### 2021-06-10 floa_doc.Rmd
Initial commit of floa documentation/overview file

---
title: "floa_doc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-06-10 floa_doc.Rmd
Initial commit

@@
 ---
-title: "floa_doc"
-output: html_document
+title: "Documentation/overview Functional Limits of Agreement"
+author:
+  - name: Daniel Koska
+    email: daniel.koska@hsw.tu-chemnitz.de
+    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
+    footnote: Corresponding Author
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: github_document
+# output:
+#   html_document: 
+#     keep_md: yes
+#     theme: united
+#     toc: yes
+#     toc_float: yes
 ---

### 2021-06-10 floa_doc.Rmd
Change the boilerplate text

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+## Projektbescheibung
 
-## R Markdown
-
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
-
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
-
-```{r cars}
-summary(cars)
-```
-
-## Including Plots
-
-You can also embed plots, for example:
-
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
-
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+Test

### 2021-06-10 floa_doc.Rmd
Added project description

@@
 
-## Projektbescheibung
+## Kurze Zusammenfassung
 
-Test
+Aufgabenstellung: Vergleich zweier Messsysteme mit kontinuierlichem Output
+
+Standardvorgehen: Bland & Altman Plots (diskrete Parameter)
+
+Erweiterung für funktionale Daten: Roislien et al. (2012) Functional Limits of Agreement
+
+#### Functional Limits of Agreement nach Roislien et al.
+
++ Behandlung von Zeitreihen als funktionale Objekte (Ramsay & Silverman)
+
++ Auswahl einer Zeitreihe je Proband (bei n=7! im Beispiel) und Schätzung der LoA durch Bootstrapping
+
+#### Koska & Maiwald (2020, submitted to Gait & Posture) ... Vergleich (coverage) 3er Verfahren:
+
++ Pointwise (continuous) LoA (FLoA~Point~)
+
++ Functional Limits of Agreement nach Roislien et al. (FLoA~2SD~)
+
++ Functional limits of agreement using a randomized cluster bootstrap (FLoA~RCB~, Koska & Maiwald)
+
+Illustriert wurden die Verfahren am Beispiel eines Datensatzes zur Validierung einer IMU (MoCap vs. IMU)
+
+#### Übergeordnete Fragestellung
+
+"What is the coverage performance of functional prediction bands when analyzing the agreement between two sets of biomechanical curves?"
+
+#### Methodikbeschreibung FLoA~RCB~ (kopiert aus dem Paper)
+
+- The agreement between the two measurement systems was analyzed based on time-normalized difference curves
+
+- Initially, time-discrete curves 𝑑⃗(𝑡) were converted to functional objects 𝑥(𝑡) using a linear combination of 𝐾 (Fourier series) functions with weight coefficients 𝑐
+  + 𝐾 was set relatively high (𝐾=50) to avoid smoothing artifacts when fitting the function. The quality of the fit was checked visually using plots.
+
+This model includes intra-individual variance from repeated measurements.
+To account for the cluster structure generated by similar curves of the same subject (Fig.
+1), we implemented the randomized cluster bootstrap described in Davison & Hinkley [20].
+In the first stage, 𝑛 clusters were drawn with replacement from the original sample.
+A single cluster corresponds to all curves of a subject.
+In the second stage, observations within the newly generated sample were randomly permuted (drawing without replacement).
+The process was repeated 𝐵 times and mean curves were determined during every iteration.
+
+From the resulting empirical distribution of mean curves 𝑋𝐵∗𝑛∗, pointwise quantiles (Q2.5, Q97.5) were determined as upper and lower interval limits: 𝐹𝐿𝑜𝐴𝑅𝐶𝐵= ± 𝑄𝛼/2(𝑋𝑖 𝐵∗𝑛 ∗(𝑡))
+
+## Flowchart FLoA~RCB~
+
+## Reviewerkommentare (sortiert nach Wichtigkeit)
+
+#### Vorab etwas Positivität
+
++ Reviewer #1: I would like to congratulate the authors for a complex, well written and meaningful manuscript. [...] I have only minor comments: ...
+
++ Reviewer #2: The general idea behind the manuscript, to evaluate agreement between biomechanical curves and to construct prediction bands are highly relevant and interesting.
+
++ Reviewer #3: The idea that the authors are presenting is quite interesting, but theanalyis and evaluation of the various mathods contains come serious errors.
+
+#### Major comments
+
++ Evaluierung anhand weiterer (simulierter) Datensätze
+
++ In looking at limits of agreement, the authors are considering the extent of within pair variation.
+However, in looking at the point wise limits of agreement, these are defined to incorporate a between pair standard deviation.
+Even if it was appropriate to utilise between pair variability, the authors should note that it is variances that are additive and not SDs.
+Perhaps the authors lack of understanding comes from their erroneous assumption (iii).
+
+(iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject.
+If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
+The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
+
++ [...] The authors are producing the arithmetic mean curves, the variability of which will reflect variation between means.
+It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of 11.
+
++ As I understand the description of the paper, the coverage is estimated for the curves that are used to estimate the prediction bands
+
++ Detaillierte Beschreibung der Methode erforderlich
+
++ Mathematische Notation
+
+#### Was meint der Reviewer?
+
+- Although the pointwise approach ignores the temporal dependence, it is not an issue when it comes to estimation. For inference, on the other hand, it should be taken into account
+
+#### Irrtümer der Reviewer!?
+
++ Fourier Series sind overfitted: "The effect of overfitting will be that the fitted curves wll reflect to a large extent the measurement errors associated with each individual point."
+
+#### Mögliche Lösungsansätze
+
++ Herausnehmen der punktweisen LoA: Die Idee funktionaler Methode könnte man als etabliert betrachten.
+Beim Lesen der Reviewerkommentare drängt sich auch der Verdacht auf, dass die relativ hohe Komplexität des Papers sich negativ auf das Verständnis (der Reviewer) auswirkt.
+
+Die FLoA~2SD~ (Roislien et al.) würde ich hingegen erstmal drin lassen, da a.
+unsere Methode darauf aufbaut und b.
+die Vorteile unserer Methode so besser veranschaulicht werden können.
+Zudem kann ich mir vorstellen, dass die Forderung nach einer Vergleichsmethode beim Einreichen eines neuen Papers zwangsläufig irgendwann kommt.
+
++ Datensätze mit ähnlichen Eigenschaften wie typische biomechanische Signale simulieren (z. B. Beschleunigungssignale, EMG, EEG)
+
+... evtl. kommen die (vermuteten) Vorteile der FLoA~RCB~ bei Signalen mit höherem Rauschanteil besser zum Tragen.

### 2021-06-10 floa_doc.Rmd
Minor changes documentation files
Minor changes to documentation file

@@
 
-## Kurze Zusammenfassung
+## Zusammenfassung
 
-Aufgabenstellung: Vergleich zweier Messsysteme mit kontinuierlichem Output
+#### Übergeordnete Fragestellung
+
+"What is the coverage performance of functional prediction bands when analyzing the agreement between two sets of biomechanical curves?"
+
+Frage ergab sich ursprünglich im Rahmen der Validierung einer hauseigenen IMU gegenüber einem kamerabasierten Messsystem (Vicon).
 
@@
 Illustriert wurden die Verfahren am Beispiel eines Datensatzes zur Validierung einer IMU (MoCap vs. IMU)
-
-#### Übergeordnete Fragestellung
-
-"What is the coverage performance of functional prediction bands when analyzing the agreement between two sets of biomechanical curves?"
 
@@
 Even if it was appropriate to utilise between pair variability, the authors should note that it is variances that are additive and not SDs.
-Perhaps the authors lack of understanding comes from their erroneous assumption (iii).
+Perhaps the authors lack of understanding comes from their *erroneous assumption* (iii).
 
-(iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject.
-If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
-The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
+*iii Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject.*
+
+*If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
+The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.*
+
+Die erwähnte assumption bezieht sich auf eine Grundannahme die im Roislien Paper getroffen wurde und in unserem Paper kritisch aufgegriffen wurde.
 
@@
 
-#### Was meint der Reviewer?
-
-- Although the pointwise approach ignores the temporal dependence, it is not an issue when it comes to estimation. For inference, on the other hand, it should be taken into account
-
-#### Irrtümer der Reviewer!?
+#### Fragwürdige Einschätzungen seitens der Reviewer
 
@@
 
-#### Mögliche Lösungsansätze
+### Mögliche Lösungsansätze
 
@@
 
-Die FLoA~2SD~ (Roislien et al.) würde ich hingegen erstmal drin lassen, da a.
+...
+siehe dieser Kommentar der Reviewer: Although the pointwise approach ignores the temporal dependence, it is not an issue when it comes to estimation.
+For inference, on the other hand, it should be taken into account
+
++ Die FLoA~2SD~ (Roislien et al.) würde ich hingegen erstmal drin lassen, da a.
 unsere Methode darauf aufbaut und b.
@@
 ... evtl. kommen die (vermuteten) Vorteile der FLoA~RCB~ bei Signalen mit höherem Rauschanteil besser zum Tragen.
+
+## Sonstiges
+
++ FLoA~RCB~ als R package
+
++ Github repository: https://github.com/koda86/floa

### 2021-06-10 floa_doc.Rmd
Flowchart old algorithm added

@@
 
+#### Eingereichte Version (problematisch)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+library(DiagrammeR)
+
+# (Nested) Randomized Cluster Bootstrap
+#
+# In the first stage (draw_clusters()), all curves of a single (random) subject
+# are drawn. From this set, the (functional) mean is calculated. The process is
+# repeated n.boot times (n.boot = number of bootstrap iterations)
+
+grViz("digraph flowchart {
+
+      # node definitions with substituted label text
+      node [fontname = Helvetica, shape = rectangle]        
+      tab1 [label = '@@1']
+      tab2 [label = '@@2']
+      tab3 [label = '@@3']
+      tab4 [label = '@@4']
+      tab5 [label = '@@5']
+      tab6 [label = '@@6']
+      tab7 [label = '@@7']
+      tab8 [label = '@@8']
+
+      # edge definitions with the node IDs
+      tab1 -> tab2
+      tab2 -> tab3
+      tab3 -> tab4
+      tab3 -> tab5
+      tab3 -> tab6
+      tab4 -> tab7
+      tab4 -> tab8
+      }
+
+      [1]: 'Time normalize curves of IMU - MC'
+      [2]: 'Create difference curves IMU - MC'
+      [3]: 'Convert difference curves to function data objects using Fourier series'
+      [4]: 'floa_rcb '
+      [5]: 'FLOAboot_2SD '
+      [6]: 'floa_point'
+      [7]: 'First stage'
+      [8]: 'Second stage'
+      ")
+
+```
+
+#### Vorschlag für verbesserte Variante
+
 ## Reviewerkommentare (sortiert nach Wichtigkeit)

### 2021-06-10 floa_doc.Rmd
Flowchart for old version completed

--- 2021-06-10 floa_doc.Rmd
@@
 
+#### Eingereichte Version (problematisch)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+library(DiagrammeR)
+
+# (Nested) Randomized Cluster Bootstrap
+#
+# In the first stage (draw_clusters()), all curves of a single (random) subject
+# are drawn. From this set, the (functional) mean is calculated. The process is
+# repeated n.boot times (n.boot = number of bootstrap iterations)
+
+grViz("digraph flowchart {
+
+      # node definitions with substituted label text
+      node [fontname = Helvetica, shape = rectangle]
+
+      tab1 [label = '@@1']
+      tab2 [label = '@@2']
+      tab3 [label = '@@3']
+      tab4 [label = '@@4']
+      tab5 [label = '@@5']
+      tab6 [label = '@@6']
+      tab7 [label = '@@7']
+      tab8 [label = '@@8']
+      tab9 [label = '@@9']
+      tab10 [label = '@@10']
+
+      # edge definitions with the node IDs
+      tab1 -> tab2
+      tab2 -> tab3
+      tab3 -> tab4
+      tab3 -> tab5
+      tab3 -> tab6
+      tab4 -> tab7
+      tab7 -> tab8
+      tab8 -> tab9
+      tab9 -> tab10
+      }
+
+      [1]: 'Time normalize curves of IMU - MC'
+      [2]: 'Create difference curves IMU - MC'
+      [3]: 'Convert difference curves to function data objects using Fourier series'
+      [4]: 'floa_rcb '
+      [5]: 'FLOAboot_2SD '
+      [6]: 'floa_point'
+      [7]: paste0('First stage', ':\\n ', 'draw_clusters', '\\n ', 'All strides from 11 subjects (randomly permuted)', '\\n ', 'are drawn with replacement')
+      [8]: 'Functional mean for each cluster'
+      [9]: paste0('Second stage', ':\\n ', 'Repeat the first stage 1000 times')
+      [10]: 'Quantile'
+      ")
+
+```
+
+Asymmetrische Limits of Agreement aus einem hierarchischen (zweistufigen) Bootstrap in Anlehnung an Davison & Hinkley (1997)
+
+#### Vorschlag für verbesserte Variante
+
 ## Reviewerkommentare (sortiert nach Wichtigkeit)

### 2021-08-05 R/main.R
Refine raw data plots

@@ -60,20 +60,28 @@ source("plot_cov_ver.R")
 #
 # (* Empirical validation data: "imu_mc")
 # * Smooth, wave data (normal error, constant variance, no trend): "smooth"
-# * Smooth wave data with nonlinear trend (no bias, constant variance): "smooth_trend"
+# * Smooth wave data with nonlinear trend (constant variance): "smooth_trend"
 # * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
 # * Data with shock peaks (no bias, no trend): "shock"
 
 data <- example_data(dat = "non_gaussian", dir.data)
 
-# Plot data of a single subject
-subject <- 1
-data.single.mc <- subset(data, subjectID == subject & device == "MC")
-data.single.imu <- subset(data, subjectID == subject & device == "IMU")
+# Plot data
+# data$subjectID <- as.factor(data$subjectID)
 
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"))
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
+  geom_line(alpha = 0.7) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
+  labs(x = "Time-normalized signal [%]", y = "Difference") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 20),
+        axis.title.x = element_text(size = 22),
+        axis.text.y = element_text(size = 20),
+        axis.title.y = element_text(size = 22),
+        legend.position = "none")
 
 PLOT
 
@@ -148,7 +156,7 @@ print(coverage)
 # Select a version:
 # v1  : n=length(subjects) random strides from all strides
 # v2  : Functional data version of v1
-# v3  : Fetch a single stride only form all strides
+# v3  : Fetch a single stride only from all strides
 #
 # Output:()
 #   * Coverage levels [%] across n=length(subjectID) iterations

### 2021-08-05 paper.Rmd
Add Rmd file
First rough draft of the scientific paper

---
title: "Asymmetric functional prediction bands for assessing the validity of biomechanical curve data"
author:
  - name: Daniel Koska
    email: daniel.koska@hsw.tu-chemnitz.de
    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
    footnote: Corresponding Author
  - name: Doris Oriwol
    email: doris.oriwol@partner.kit.edu
    affiliation: Karlsruhe Institute of Technology
  - name: Christian Maiwald
    email: christian.maiwal@hsw.tu-chemnitz.de
    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
# address:
#   - code: Chemnitz University of Technology
#     address:  Thüringer Weg 11, 09126 Chemnitz
#   - code: Another University
#     address: Department, Street, City, State, Zip
abstract: |
  This is the abstract.

date: "`r format(Sys.time(), '%d %B, %Y')`"
journal: "An awesome journal"
# bibliography: sendaFAB.bib
# output: rticles::elsevier_article
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

### TODO
- bar

```{r echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)

dir.script <- "C:/Users/Daniel/Desktop/tmp/floa/R"
dir.data <- "C:/Users/Daniel/Desktop/tmp/floa/R/examples"

# setwd(dir.script)
source(paste0(dir.script, "/example_data.R"))
source(paste0(dir.script, "/draw_clusters.R"))
source(paste0(dir.script, "/floa_rcb.R"))
source(paste0(dir.script, "/floa_point.R"))
source(paste0(dir.script, "/plot_loa.R"))
source(paste0(dir.script, "/get_coverage.R"))
source(paste0(dir.script, "/crossval_coverage.R"))
source(paste0(dir.script, "/plot_cov_ver.R"))
# source("fdaDelta.R")
```

# Introduction

# Methods

## Limits of agreement methods

FLoA derived by different methods are compared

* Randomized Cluster Bootstrap (floa.rcb)
  + Different strategies

* Point-by-point Gaussian intervals (floa.point)

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Wrapper function for example data sets. Function arguments:
#
# (* Empirical validation data: "imu_mc")
# * Smooth, wave data (normal error, constant variance, no trend): "smooth"
# * Smooth wave data with nonlinear trend (constant variance): "smooth_trend"
# * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
# * Data with shock peaks (no bias, no trend): "shock"

data <- example_data(dat = "smooth", dir.data)

n.boot <- 100

floa <- floa_rcb(data, n.boot, ver = "v2")

# Mean and SD are calculated across all strides (and subjects).
# No bootstrap or other resampling strategies are applied.
floa.point <- floa_point(data)
```

## Validation

### Synthetic data sets

```{r echo = FALSE, warning = FALSE, message = FALSE}

```

### Smooth, wave data (normal error, constant variance, no trend)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "smooth", dir.data)

data.single.mc <- subset(data, device == "MC")
data.single.imu <- subset(data, device == "IMU")

PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
  geom_line(alpha = 0.7) +
  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
  labs(x = "Time-normalized signal [%]", y = "Difference") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 22),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        legend.position = "none")

PLOT
```

### Smooth wave data with nonlinear trend

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "smooth_trend", dir.data)

data.single.mc <- subset(data, device == "MC")
data.single.imu <- subset(data, device == "IMU")

PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
  geom_line(alpha = 0.7) +
  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
  labs(x = "Time-normalized signal [%]", y = "Difference") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 22),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        legend.position = "none")

PLOT
```

#### Data with non-gaussian (Weibull distributed) error (no trend)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "non_gaussian", dir.data)

data.single.mc <- subset(data, device == "MC")
data.single.imu <- subset(data, device == "IMU")

PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
  geom_line(alpha = 0.7) +
  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
  labs(x = "Time-normalized signal [%]", y = "Difference") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20),
        axis.title.x = element_text(size = 22),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 22),
        legend.position = "none")

PLOT
```

### Cross validation

Leave-one (subject) out method to estimate the achieved coverage

Versions:

* v1   : n = length(subjects) random strides from all strides

* v1.1 : Functional data version of v1

* v2   : One random stride per subject

* v3   : Fetch a SINGLE random stride from all strides

<!-- $$ -->
<!-- RMSE_{delta} = \sqrt{\frac{1}{N}(target_1 - pedal_1)^2 + (target_2 - pedal_2)^2 + ... + (target_N - pedal_N)^2} -->
<!-- $$ -->

<!-- $$ -->
<!-- RMSE_{pedal} = \sqrt{\frac{1}{n}[(pedal_1 - \mu)^2 + (m_2 - \mu)^2 + ... + (pedal_N - \mu)^2]}, where \mu = \frac{1}{N}(pedal_1 + ... + pedal_N) -->
<!-- $$ -->

<!-- $$ -->
<!-- Z = \{z_1, z_2, z_3,...,z_i\} -->
<!-- $$ -->

# Results

### Cross validation

#### v1 (n = length(subjects) random strides from all strides)

```{r echo = FALSE, warning = FALSE, message = FALSE}
data <- example_data(dat = "smooth", dir.data)
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
cover.cross.v1 <- crossval_coverage(data, n.boot, method = "all", ver = "v1")

plot_cov_ver(cover.cross.v1)
```

### V2 (One random stride per subject)

```{r echo = FALSE, warning = FALSE, message = FALSE}
cover.cross.v2 <- crossval_coverage(data, n.boot, method = "all", ver = "v2")

plot_cov_ver(cover.cross.v2)
```

### v3 (Fetch a SINGLE random stride from all strides)

```{r echo = FALSE, warning = FALSE, message = FALSE}
cover.cross.v3 <- crossval_coverage(data, n.boot, method = "all", ver = "v3")

plot_cov_ver(cover.cross.v3)
```

# Discussion

# Conclusion

### 2021-08-05 R/crossval_coverage.R

@@ -21,7 +21,7 @@ crossval_coverage <- function (data, n.boot, method, ver) {
                             get_coverage(data, t(floa.rcb)) # Select floa method
                             )
 
-      print(paste("floa.rcb", i))
+      # print(paste("floa.rcb", i))
 
       # FLoA Point
       # --------------------------------------------------------------------
@@ -43,7 +43,7 @@ crossval_coverage <- function (data, n.boot, method, ver) {
                         get_coverage(data, t(floa.rcb)) # Select floa method
       )
 
-      print(paste("floa.rcb", i))
+      # print(paste("floa.rcb", i))
 
     } else if (method == "floa.point" | method == "all") {
 
@@ -54,7 +54,7 @@ crossval_coverage <- function (data, n.boot, method, ver) {
                         get_coverage(data, t(floa.point)) # Select floa method
       )
 
-      print(paste("floa.point", i))
+      # print(paste("floa.point", i))
 
     } else {
 

### 2021-08-05 paper.Rmd
Refine paper draft Rmd

@@
 ---
-title: "Asymmetric functional prediction bands for assessing the validity of biomechanical curve data"
+title: "Asymmetric functional prediction bands for assessing the validity of spatiotemporally smooth biomechanical curve data"
 author:
@@
 
+Knowing the size of the measurement error is a basic requirement for evaluating the results of empirical studies.
+Uncertainty about the error increases the risk of false positive (i.e.
+errors are incorrectly interpreted as effects) or false negative (i.e.
+effect and noise are inseperable) results and may negatively impact the reproducibility of study outcomes.
+
+The classic linear error model describes measurement error as the sum of systematic and random deviations from the true value.
+The full characterization of measurement error, however, is considerably more complex, as different sources of variance (within, between) and noise must be considered.
+In particular, the output of many biomechanical measurement systems is available in the form of spatiotemporally smooth curves within which the error can vary over time.
+
+However, to apply statistical standard models, in most cases the dimensionality of curves is reduced to just discrete features, such as local extrema, rates of change, or ranges of motion.
+A number of problems arise from this: (i) non-stationary error processes are not considered (ii) most traditional point statistics are either not suitable (e.g.
+statistical tests of mean differences), or are only suitable to a limited extent (e.g.
+product-moment correlation) to
+quantify the agreement between two measurement systems (Bland and Altman, 1986) (iii) the validity of statements derived from discrete variables is limited by domain alignment (i.
+e.
+phase shift) and the risk of examining points that have little relevance for the system under investigation [Donoghue et al.
+(2008), Pataky et al.
+(2008), Richter et al.
+(2014), Park et al.
+(2017)].
+
+Therefore, when analyzing the validity of continuous data, the entire length of the curve should be accounted for. This can be accomplished by either breaking down the curve into a single
+value using e. g. distance measures such as the root mean square error, or by applying statistics (e. g. standard deviation) separately to all points of the curve [Pini et al. (2019)].
+
+The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points of biomechanical curves are locally correlated [Deluzio et al.
+(2007), Pataky].
+From a statistical point of view, this implies that random measurement error is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for [Lenhoff et al., 1999].
+
+The principal object of this paper was to demonstrate a method for curve data that addresses the aforementioned issues (functional data representation, linear model assumptions, repeated measures), and thus enables the adequate characterization of the agreement between measurement curves on the basis of easy to interpret prediction bands (i.e.
+LoA in the actual measurement unit).
+
 # Methods
@@
 
-n.boot <- 100
-
-floa <- floa_rcb(data, n.boot, ver = "v2")
-
 # Mean and SD are calculated across all strides (and subjects).
@@
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
+FÜr die Validierung der Methode wurden Zeitreihen simuliert, die gängigen biomechanischen Messsignalen qualitativ ähneln und typische Fehlercharakteristika abbilden.
+Die verschiedenen Zeitreihen wurden als Fourierreihen aus additiv überlagerten Sinusfunktionen modelliert.
+Jeder Datensatz enthält n.ts = 100 Kurven von n.subj = 11 fiktiven Probanden.
+Jede Kurve besteht aus t = 101 Datenpunkten.
 
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+n.subj <- 11
+n.ts <- 100
+
+t <- seq(0, 100)
+
+for (subject.idx in 1:n.subj) {
+
+  # Subjectwise parameters
+  offset.mean <- runif(1, min = -0.5, max = 0.5)
+
+  a.sd <- runif(1, min = 0.05, max = 0.15)
+  b.sd <- runif(1, min = 0.0001, max = 0.002)
+
+  for (ts.idx in 1:(n.ts)) {
+
+    a1.device1 <- rnorm(1, mean = 3, sd = a.sd)
+    a1.device2 <- rnorm(1, mean = 3, sd = a.sd)
+    a2.device1 <- 0.08
+    a2.device2 <- 0.08
+    b1.device1 <- rnorm(1, mean = 0.06, sd = b.sd)
+    b1.device2 <- rnorm(1, mean = 0.06, sd = b.sd)
+    b2.device1 <- rnorm(1, mean = 0.58, sd = b.sd)
+    b2.device2 <- rnorm(1, mean = 0.58, sd = b.sd)
+    c <- 2
+
+    sine.1 <- a1.device1 * sin(b1.device1 * t) ^ (c + 3)
+    sine.2 <- a2.device1 * sin(b2.device1 * t)
+    sine.3 <- a1.device2 * sin(b1.device2 * t) ^ (c + 3)
+    sine.4 <- a2.device2 * sin(b2.device1 * t)
+
+    offset <- rnorm(1, offset.mean, 0.05)
+
+    # Model (i). (ii), or (iii)
+    # ...
 ```
 
-### Smooth, wave data (normal error, constant variance, no trend)
+* (i) Smooth curves with normally distributed errors, constant variance, and no trend.
+Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
+The data resembles e.
+g.
+kinematic joint angle curves captured by camera-based measurement systems or goniometers.
+
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+# ...
+
+# Model (i)
+device.1 <- sine.1 + sine.2
+device.2 <- offset + sine.3 + sine.4
+```
+
+* (ii) Smooth curves created using the same model and parameters as in (i), but with a nonlinear trend underlying one of the two signals.
+A similar trend (drift) behavior can be observed in e.
+g.
+when the same joint angle curves are calculated by integrating the signal of gyropscopes applied to measure the movement.
+
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+# ...
+trend <- (1 / 100000) * seq(0.5, 50.5, 0.5)^3
+
+# Model (ii)
+device.1 <- sine.1 + sine.2
+device.1 <- offset + sine.3 + sine.4 + trend
+```
+
+* (iii) Smooth curves created using the same model and parameters as in (i), but Weibulll distributed random curve (amplitude) parameters.
+This introduces non-gaussian measurement error (asymmetric/skewed distribution of differences around the mean difference).
+
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+# ...
+a1.device2 <- rnorm(n = 1,
+                    mean = rweibull(1, shape = 1.5, scale=1) - factorial(1/1.5), # factorial() used to center around 0
+                    sd = a.sd)
+# ...
+sine.3 <- a1.device2 * sin(b1.device2 * t) ^ (c + 3)
+# ...
+
+# Model (iii)
+device.1 <- sine.1 + sine.2
+device.2 <- offset + sine.3 + sine.4
+```
+
+### Smooth curves (normal error, constant variance, no trend)
 
@@
 
-### Smooth wave data with nonlinear trend
+### Smooth curves nonlinear error trend
 
@@
 
-#### Data with non-gaussian (Weibull distributed) error (no trend)
+#### Smooth curves with non-gaussian (Weibull) error (no trend)
 
@@
 
+### Functional limits of agreement
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth", dir.data)
+n.boot <- 100
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+
+data <- example_data(dat = "smooth_trend", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+
+data <- example_data(dat = "non_gaussian", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
 ### Cross validation
@@
 
-### V2 (One random stride per subject)
+#### V2 (One random stride per subject)
 
@@
 
-### v3 (Fetch a SINGLE random stride from all strides)
+#### v3 (Fetch a SINGLE random stride from all strides)
 

### 2021-08-05 R/simulate_data.R

@@ -174,20 +174,20 @@ for (subj.idx in 1:n.subj) {
 
   a.sd <- runif(1, min = 0.05, max = 0.15) # .1
 
-  b.sd <- runif(1, min = 0.0001, max = 0.002) # rweibull(1, shape = 1.5, scale=1) - factorial(1/1.5)
+  b.sd <- runif(1, min = 0.0001, max = 0.002)
 
   for (stride.idx in 1:(n.strides)) {
 
-    a1.1 <- rnorm(1, mean = 3, sd = a.sd) # .1
-    a1.2 <- rnorm(1, mean = rweibull(1, shape = 1.5, scale=1) -
-                    factorial(1/1.5), # center the error around 0
-                  sd = a.sd) # .1
+    a1.1 <- rnorm(1, mean = 3, sd = a.sd)
+    a1.2 <- rnorm(n = 1,
+                  mean = rweibull(1, shape = 1.5, scale=1) - factorial(1/1.5), # factorial() used to center around 0
+                  sd = a.sd)
     a2.1 <- 0.08
     a2.2 <- 0.08
-    b1.1 <- rnorm(1, mean = 0.06, sd = b.sd) # 0.001
-    b1.2 <- rnorm(1, mean = 0.06, sd = b.sd) # 0.001
-    b2.1 <- rnorm(1, mean = 0.58, sd = b.sd) # 0.001
-    b2.2 <- rnorm(1, mean = 0.58, sd = b.sd) # 0.001
+    b1.1 <- rnorm(1, mean = 0.06, sd = b.sd)
+    b1.2 <- rnorm(1, mean = 0.06, sd = b.sd)
+    b2.1 <- rnorm(1, mean = 0.58, sd = b.sd)
+    b2.2 <- rnorm(1, mean = 0.58, sd = b.sd)
     c <- 2
 
     sine.1 <- a1.1 * sin(b1.1 * t) ^ (c + 3)

### 2021-08-05 R/floa_rcb.R
Minor style formatting

@@ -44,6 +44,7 @@ floa_rcb <- function(data, n.boot, ver) { # , fd.basis
   # Add (pointwise) mean
   floa.boot.mean <- colMeans(clust.agg.intrp, dims = 1)
   floa <- rbind(floa.boot.percentiles.split, floa.boot.mean)
+
   row.names(floa) <- c("lower", "median", "upper", "mean")
 
   return(floa)

### 2021-08-05 R/main.R

@@ -1,12 +1,9 @@
-################################################################################
 # Main script Functional Limits of Agreement (FLoA)
 #
 # FLoA derived by different methods are compared
 #   * Randomized Cluster Bootstrap      (floa.rcb)
 #     * Different strategies
 #   * Point-by-point Gaussian intervals (floa.point)
-#   (* Method Lenhoff et al. (1999)      (curently not implemented))
-#   (* Method Roislien et al. (2012)     (curently not implemented))
 #
 # Thus far, a transformation of time series data to functional data (Fourier,
 # Splines etc.) is not implemented.
@@ -15,9 +12,7 @@
 # (see subsection data sets)
 #
 # TODO:
-#   + Paper als *.RMD
 #   + Implementieren FDA
-#   + Datensätze: 1. "smooth", 2. "smooth+trend" 3. "smooth + non-gaussian error (asymmetric) # Chi-Quadrat"
 #   + Quantile: Über die gesamte Verteilung oder die "Ausreisser-Quantile" einzelner/extremer Probanden
 #     + quantile() function: Bias correction useful/necessary?
 #   + Konvergenzanalyse --> ja, aber "nur" intern
@@ -28,6 +23,7 @@
 #   + https://cran.r-project.org/web/packages/smooth/vignettes/simulate.html
 #
 # TODO STYLE:
+#   + Make variable names more universal (e.g. "device.1" instead "mc")
 #   + Add namespaces (package names ::)
 #   + No line between header and body in loops
 ################################################################################
@@ -64,7 +60,7 @@ source("plot_cov_ver.R")
 # * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
 # * Data with shock peaks (no bias, no trend): "shock"
 
-data <- example_data(dat = "non_gaussian", dir.data)
+data <- example_data(dat = "smooth", dir.data)
 
 # Plot data
 # data$subjectID <- as.factor(data$subjectID)
@@ -115,28 +111,21 @@ floa <- floa_rcb(data, n.boot, ver = "v2")
 # No bootstrap or other resampling strategies are applied.
 floa.point <- floa_point(data)
 
-# # Method Lenhoff et al. (1999)
-# # ------------------------------------------------------------------------------
-
-# # Method Roislien et al. (2012)
-# # ------------------------------------------------------------------------------
-
 
 ################################### Plot data ##################################
 
 # Select limits of agreement method and central tendency parameter
 plot_loa(data, floa, central.tendency = "mean")
 
[…]

### 2021-08-05 paper.Rmd
Correct minor typo

@@
 
-FÜr die Validierung der Methode wurden Zeitreihen simuliert, die gängigen biomechanischen Messsignalen qualitativ ähneln und typische Fehlercharakteristika abbilden.
+Für die Validierung der Methode wurden Zeitreihen simuliert, die gängigen biomechanischen Messsignalen qualitativ ähneln und typische Fehlercharakteristika abbilden.
 Die verschiedenen Zeitreihen wurden als Fourierreihen aus additiv überlagerten Sinusfunktionen modelliert.

### 2021-08-05 paper.Rmd
Bugfix in paper.Rmd

@@
   - name: Christian Maiwald
-    email: christian.maiwal@hsw.tu-chemnitz.de
+    email: christian.maiwald@hsw.tu-chemnitz.de
     affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-cover.cross.v1 <- crossval_coverage(data, n.boot, method = "all", ver = "v1")
+cover.cross.v1 <- crossval_coverage(data, floa, floa.point, method = "all", ver = "v1")
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-cover.cross.v2 <- crossval_coverage(data, n.boot, method = "all", ver = "v2")
+cover.cross.v2 <- crossval_coverage(data, floa, floa.point, method = "all", ver = "v2")
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-cover.cross.v3 <- crossval_coverage(data, n.boot, method = "all", ver = "v3")
+cover.cross.v3 <- crossval_coverage(data, floa, floa.point, method = "all", ver = "v3")
 

### 2021-08-05 R/crossval_coverage.R

@@ -1,4 +1,4 @@
-crossval_coverage <- function (data, floa, floa.point, method, ver) {
+crossval_coverage <- function (data, floa, floa.point, method) {
 
   n.subj <- unique(data$subjectID)
 

### 2021-08-06 paper.Rmd
Get paper.Rmd to compile

@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-cover.cross.v1 <- crossval_coverage(data, floa, floa.point, method = "all", ver = "v1")
+floa <- floa_rcb(data, n.boot, ver = "v1")
+floa.point <- floa_point(data)
+
+cover.cross.v1 <- crossval_coverage(data, floa, floa.point, method = "all")
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-cover.cross.v2 <- crossval_coverage(data, floa, floa.point, method = "all", ver = "v2")
+floa <- floa_rcb(data, n.boot, ver = "v2")
+floa.point <- floa_point(data)
+
+cover.cross.v2 <- crossval_coverage(data, floa, floa.point, method = "all")
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-cover.cross.v3 <- crossval_coverage(data, floa, floa.point, method = "all", ver = "v3")
+floa <- floa_rcb(data, n.boot, ver = "v3")
+floa.point <- floa_point(data)
+
+cover.cross.v3 <- crossval_coverage(data, floa, floa.point, method = "all")
 

### 2021-08-06 paper.Rmd
Comment out fda functions for the moment
Continue to work on the Rmd file (introduction section)

@@
 ### TODO
-- bar
+- Simulierte Daten durch Funktionen ersetzen?
+Eigentlich aktuell eher unnötig, da die Zeitreihen ja eigentlich bereits mithilfe von Funktionen modelliert wurden.
+Zudem wird nicht interpoliert, abgeleitet o.
+Ä.
+...
+letztlich für die aktuelle Idee des Papers eher ungeeignet (plus man umgeht Probleme wie mögliche Glättungsartefakte, oder die Definition eines functional mean).
+- Falls doch FDA, dann evtl. als eigenständiges Skript forken
+- .bib file anlegen
+- Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert)
+- Ratkowsky besorgen
+
+### TODO read (ee 'Literature' folder on Nextcloud)
+- Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem
+- Generell die Kooperationspaper von Hörmann und Kokoszka zu FDA Statistik,insbesondere 'Inference for Functional Data with Applications' (2012)
+- Ratkowsky 1990 Handbook of nonlinear regression models
 
@@
 
-Knowing the size of the measurement error is a basic requirement for evaluating the results of empirical studies.
-Uncertainty about the error increases the risk of false positive (i.e.
-errors are incorrectly interpreted as effects) or false negative (i.e.
-effect and noise are inseperable) results and may negatively impact the reproducibility of study outcomes.
+Having a good estimate of the amount of measurement error is a basic requirement when analyzing movement data.
+The classic linear error model describes measurement error as the sum of systematic and random deviations from the true value.
+A complete characterization of the measurement error, however, is complex, as different sources of variance (within, between) and noise must be considered.
+The analysis of sensor data is further complicated by the fact that the output of many measurement systems in biomechanics - such as e.
+g.
+joint angles, ground reaction forces, or center of mass trajectories - is available in the form of rather smooth curves within which the error can vary over time.
 
-The classic linear error model describes measurement error as the sum of systematic and random deviations from the true value.
-The full characterization of measurement error, however, is considerably more complex, as different sources of variance (within, between) and noise must be considered.
-In particular, the output of many biomechanical measurement systems is available in the form of spatiotemporally smooth curves within which the error can vary over time.
+Therefore, when analyzing the validity of continuous data, the entire length of the curve should be accounted for.
+This can be accomplished e.
+g.
+by applying statistics (e.
+g.
+standard deviation) separately to all points of the curve [Pini et al.
+(2019)].
+The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points of biomechanical curves are locally correlated [Deluzio & Astephen, 2007; Pataky, 2010].
+From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for [Lenhoff et al., 1999].
+In addition, most traditional point statistics are either not suitable (e.g.
+statistical tests of mean differences), or are only suitable to a limited extent (e.g.
+product-moment correlation) to quantify the agreement between two measurement systems (Bland and Altman, 1986).
 
-However, to apply statistical standard models, in most cases the dimensionality of curves is reduced to just discrete features, such as local extrema, rates of change, or ranges of motion.
-A number of problems arise from this: (i) non-stationary error processes are not considered (ii) most traditional point statistics are either not suitable (e.g.
-statistical tests of mean differences), or are only suitable to a limited extent (e.g.
-product-moment correlation) to
-quantify the agreement between two measurement systems (Bland and Altman, 1986) (iii) the validity of statements derived from discrete variables is limited by domain alignment (i.
-e.
-phase shift) and the risk of examining points that have little relevance for the system under investigation [Donoghue et al.
-(2008), Pataky et al.
-(2008), Richter et al.
-(2014), Park et al.
-(2017)].
+One possibility to address te problem of a pointwise analysis is to treat curves as single objects rather than as a series of unconnected points [Ramsay & Silverman, 2005].
+Røislien et al.
+[2012] presented a method that extends the Limits of Agreement (LoA) concept [11] to functional data.
+This is achieved by approximating curves using Fourier series and resampling the thus obtained functions.
+The approach accounts for the autocorrelation characteristic in the signal and at the same time allows for an intuitive description of the systematic and random error terms in the original sampling space (e.g.
+actual degrees).
+Røislien et al.
+[10], however, make a number of strong assumptions, which are often not or only partially fulfilled in empirical data sets:
 
-Therefore, when analyzing the validity of continuous data, the entire length of the curve should be accounted for. This can be accomplished by either breaking down the curve into a single
-value using e. g. distance measures such as the root mean square error, or by applying statistics (e. g. standard deviation) separately to all points of the curve [Pini et al. (2019)].
+(i) Time-continuous differences of two measurement systems and the corresponding averages must be uncorrelated.
+This assumption is violated considerably in many validation data sets [12].
+E.g., Atallah et al.
+[13] found that the agreement between an ear-worn accelerometer and a force-plate instrumented treadmill was associated with gait speed when estimating gait cycle duration.
+When validating a foot-worn gyroscope for treadmill running, Koska & Maiwald [14] observed a linear error trend that was associated with foot strike behavior.
 
-The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points of biomechanical curves are locally correlated [Deluzio et al.
-(2007), Pataky].
-From a statistical point of view, this implies that random measurement error is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for [Lenhoff et al., 1999].
+(ii) The error distribution is approximately normal.
+The construction of symmetric prediction bands as described in Røislien et al.
+[10] relies on normally distributed measurement data.
+Non-normality, however, is often present when e.g.
+sensor values are truncated because of limited range or resolution, or because values drift over time.
 
-The principal object of this paper was to demonstrate a method for curve data that addresses the aforementioned issues (functional data representation, linear model assumptions, repeated measures), and thus enables the adequate characterization of the agreement between measurement curves on the basis of easy to interpret prediction bands (i.e.
+(iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject.
+If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
+The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
+
+Olsen et al.
+[15] examined the validity of an inertial measurement unit (IMU) for analyzing the spatial displacement of the distal limbs of horses using functional LoA (FLoA).
+Their approach differs from Røislien et al.
+[10] in that initially linear mixed models were fitted to the repeated measurements of different horses.
+Functional data estimates of the mean trends in the strides were then determined from the model fits and FLoA were calculated.
+This procedure accounts for repeated measurements and eliminates the need to emulate empirical curves with Fourier series or splines.
+The FLoA computed this way performed well in comparison with traditional Bland & Altman LoA.
+The procedure, however, is subject to the same restrictions as any conventional linear regression model and is therefore limited when model assumptions are violated.
+
+The principal object of this paper was to demonstrate a method for curve data that addresses the aforementioned issues (analyzing the whole curve, linear model assumptions, repeated measures), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret prediction bands (i.e.
 LoA in the actual measurement unit).
+
+We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters.
 

### 2021-08-06 R/fdaDelta.R

@@ -2,7 +2,7 @@ fdaDelta <- function(data, fd.basis){
 
   # Approximates time series using Fourier functions ---------------------------
 
-  library(Data2fd)
+  library(fda)
 
   # First device
   # --------------------------
@@ -10,8 +10,8 @@ fdaDelta <- function(data, fd.basis){
 
   dev1.wide <- matrix(dev1, ncol = length(dev1) / 101)
 
-  # Functional data object
-  dev1.wide.fd <- Data2fd(argvals = dev1.wide, basisobj = fd.basis)
+  # Create functional data object
+  dev1.wide.fd <- fda::Data2fd(argvals = dev1.wide, basisobj = fd.basis)
 
   # Second device
   # --------------------------

### 2021-08-06 R/main.R

@@ -82,11 +82,14 @@ PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID
 PLOT
 
 
-# ************* Approximate time series using Fourier functions ****************
-
-fd.basis <- 1 # Number of basic (Fourier) functions
-
-tmp <- fdaDelta(data, fd.basis)
+# # ****** Approximate time series (differences) using Fourier functions *********
+#
+# # Number of basic (Fourier) functions
+# # ... appears to plateau around 50 basis vectors
+# fd.basis <- fda::create.fourier.basis(nbasis = 50)
+#
+# # Returns functional data representation of difference curves
+# data.diff.fd <- fdaDelta(data, fd.basis)
 
 
 # ****************************** Calculate FLoA ********************************
@@ -112,8 +115,7 @@ floa <- floa_rcb(data, n.boot, ver = "v2")
 
 # Pointwise LoA ----------------------------------------------------------------
 
-# Mean and SD are calculated across all strides (and subjects).
-# No bootstrap or other resampling strategies are applied.
+# Mean + SD are calculated across all strides/subjects using linear mixed models
 floa.point <- floa_point(data)
 
 

### 2021-08-10 paper.Rmd
Tons of stuff

@@
 source(paste0(dir.script, "/example_data.R"))
+source(paste0(dir.script, "/pick_subwise_curves.R"))
 source(paste0(dir.script, "/draw_clusters.R"))
@@
 source(paste0(dir.script, "/floa_point.R"))
+source(paste0(dir.script, "/floa_roislien.R"))
 source(paste0(dir.script, "/plot_loa.R"))
@@
 source(paste0(dir.script, "/plot_cov_ver.R"))
-# source("fdaDelta.R")
 ```
@@
 This is achieved by approximating curves using Fourier series and resampling the thus obtained functions.
+
+<!-- Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
+werden 95% Konfidenzintervalle gebootstrapped.
+The authors are producing the arithmetic mean curves, the variability of which will reflect variation between means.
+It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of n.subj.
+-->
+
 The approach accounts for the autocorrelation characteristic in the signal and at the same time allows for an intuitive description of the systematic and random error terms in the original sampling space (e.g.
@@
 
-* v1.1 : Functional data version of v1
-
 * v2   : One random stride per subject
@@
 * v3   : Fetch a SINGLE random stride from all strides
+
+* v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
+       strap the resulting sample (of length (n=length(subjects))
 
@@
 
-### Cross validation
-
-#### v1 (n = length(subjects) random strides from all strides)
+### Leave-one-out cross validation
 
@@
 data <- example_data(dat = "smooth", dir.data)
+
+n.boot <- 100
+cover.cross <- crossval_coverage(data, n.boot)
+
+plot_cov_ver(cover.cross)
 ```
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-floa <- floa_rcb(data, n.boot, ver = "v1")
-floa.point <- floa_point(data)
+<!-- #### v1 (n = length(subjects) random strides from all strides) -->
 
-cover.cross.v1 <- crossval_coverage(data, floa, floa.point, method = "all")
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- data <- example_data(dat = "smooth", dir.data) -->
+<!-- ``` -->
 
-plot_cov_ver(cover.cross.v1)
-```
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- floa <- floa_rcb(data, n.boot, ver = "v1") -->
+<!-- floa.point <- floa_point(data) -->
 
-#### V2 (One random stride per subject)
+<!-- cover.cross.v1 <- crossval_coverage(data, floa, floa.point, method = "all") -->
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-floa <- floa_rcb(data, n.boot, ver = "v2")
-floa.point <- floa_point(data)
+<!-- plot_cov_ver(cover.cross.v1) -->
+<!-- ``` -->
 
-cover.cross.v2 <- crossval_coverage(data, floa, floa.point, method = "all")
+<!-- #### V2 (One random stride per subject) -->
 
-plot_cov_ver(cover.cross.v2)
-```
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
+<!-- floa.point <- floa_point(data) -->
 
-#### v3 (Fetch a SINGLE random stride from all strides)
+<!-- cover.cross.v2 <- crossval_coverage(data, floa, floa.point, method = "all") -->
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-floa <- floa_rcb(data, n.boot, ver = "v3")
-floa.point <- floa_point(data)
+<!-- plot_cov_ver(cover.cross.v2) -->
+<!-- ``` -->
 
-cover.cross.v3 <- crossval_coverage(data, floa, floa.point, method = "all")
+<!-- #### v3 (Fetch a SINGLE random stride from all strides) -->
 
-plot_cov_ver(cover.cross.v3)
-```
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- floa <- floa_rcb(data, n.boot, ver = "v3") -->
+<!-- floa.point <- floa_point(data) -->
+
+<!-- cover.cross.v3 <- crossval_coverage(data, floa, floa.point, method = "all") -->
+
+<!-- plot_cov_ver(cover.cross.v3) -->
+<!-- ``` -->
 
@@
 
+Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
+werden 95% Konfidenzintervalle gebootstrapped.
+The authors are producing the arithmetic mean curves, the variability of which will reflect variation between means.
+It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of n.subj.
+
 # Conclusion

### 2021-08-10 R/boot_mean_sd.R

@@ -1,0 +1,17 @@
+boot_mean_sd <- function (data, n.boot) {
+
+  mean.boot <- vector(mode = "list", length = n.boot)
+  sd.boot <- vector(mode = "list", length = n.boot)
+
+  for (boot.idx in 1:n.boot) {
+
+    data.subset <- pick_subwise_curves(data)
+
+    mean.boot[[boot.idx]] <- functional_mean(data.subset)
+  }
+
+  func.mean.boot <- unlist(mean.boot)
+  func.mean.boot <- matrix(func.mean.boot, ncol = n.boot)
+
+  return(func.mean.boot)
+}

### 2021-08-10 R/crossval_coverage.R

@@ -20,22 +20,28 @@ crossval_coverage <- function (data, n.boot) {
 
   n.subj <- unique(data$subjectID)
 
-  floa.point  <- floa_point(data)
-  floa.v1     <- floa_rcb(data, n.boot, ver = "v1")
-  floa.v2     <- floa_rcb(data, n.boot, ver = "v2")
-  floa.v3     <- floa_rcb(data, n.boot, ver = "v3")
-  floa.roislien     <- floa_roislien(data, n.boot)
-
-  cover.cross  <- c()
-  cover.cross.rcb  <- c()
-  cover.cross.point  <- c()
+  cover.cross.rcb.v1   <- c()
+  cover.cross.rcb.v2   <- c()
+  cover.cross.rcb.v3   <- c()
+  cover.cross.point    <- c()
+  cover.cross.roislien <- c()
 
   for (i in n.subj) {
 
-    data.subset <- subset(data, subjectID != i) # Leave subject "i" out
-
-    # FLoA
+    # Calculate FLoA with one subject left out
     # --------------------------------------------------------------------
+    data.one.out <- subset(data, subjectID != i) # Leave subject "i" out
+
+    floa.point    <- floa_point(data.one.out)
+    floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
+    floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
+    floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
+    floa.roislien <- floa_roislien(data.one.out, n.boot)
+
+    # Get coverage of unseen (excluded) subject vs. the FLoA
+    # --------------------------------------------------------------------
+    data.subset <- subset(data, subjectID == i) # Only one subject "i"
+
     cover.cross.rcb.v1 <- c(cover.cross.rcb.v1,
                              get_coverage(data.subset, floa.v1)
                              )

### 2021-08-10 R/draw_clusters.R

@@ -8,12 +8,13 @@ draw_clusters <- function(data, ver) { # fd.basis
   # v1   : n = length(subjects) random strides from all strides
   # v1.1 : Functional data version of v1
   # v2   : One stride per subject
-  # v3   : Fetch a single stride only form all strides
+  # v3   : Fetch a single stride only from all available strides
   # ----------------------------------------------------------------------------
+
+  source("pick_subwise_curves.R")
 
   if (ver == "v1") {
 
-    #### v1 ####
     # Strides are selected from the entire set of curves (NOT! one curve per subject)
     # ----------------------------------------------------------------------------
     # Pick n=length(subjects) random strides
@@ -56,39 +57,13 @@ draw_clusters <- function(data, ver) { # fd.basis
 
   if (ver == "v2") {
 
-    #### v2 ####
     # One stride per subject
     # ----------------------------------------------------------------------------
-    n.subj <- length(unique(data$subjectID)) # Select one random curve per subject
-
-    curve.idx <- c()
-
-    # Index starts at two because during the cross validation one subject is always removed (leave one out approach)
-    for (subj.idx in 1:n.subj) {
-
-      if (subj.idx %in% unique(data$subjectID)) {
-
-        tmp <- data[data$subjectID == subj.idx, ]
-
-        curve.idx.subj <- as.numeric(sample(tmp$strideID, size = 1))
-        curve.idx <- c(curve.idx, curve.idx.subj)
-      }
-    }
-
-    curve <- data[data$strideID %in% curve.idx, ]
-
-    curve0 <- subset(curve, device  == "IMU")$value
-    curve0 <- matrix(curve0, ncol = length(curve0) / 100)
-
-    curve1 <- subset(curve, device  == "MC")$value
-    curve1 <- matrix(curve1, ncol = length(curve1) / 100)
-
-    diff.curves <- curve0 - curve1
+    diff.curves <- pick_subwise_curves(data)
   }
 
   if (ver == "v3") {
 
-    # v3
     # Fetch a single stride only (from the entire set of curves)
     # ----------------------------------------------------------------------------
     curve.idx <- sample(unique(data$strideID), size = 1)

### 2021-08-10 R/floa_point.R

@@ -6,10 +6,12 @@ floa_point <- function(data) {
 
   library(lme4)
 
-  mean.diff <- list()
-  median.diff <- list()
-  sd.between <- list()
-  sd.within <- list()
+  n.frames <- length(unique(data$frame))
+
+  mean.diff <- vector(mode = "list", length = n.frames)
+  median.diff <- vector(mode = "list", length = n.frames)
+  sd.between <- vector(mode = "list", length = n.frames)
+  sd.within <- vector(mode = "list", length = n.frames)
 
   for (frame.idx in 0:100){
 

### 2021-08-10 R/floa_rcb.R

@@ -17,7 +17,7 @@ floa_rcb <- function(data, n.boot, ver) { # , fd.basis
 
   # (Row-wise) Arrange difference curves to facilitate computing percentiles
   clust.agg.intrp <- matrix(unlist(clust.boot.agg),
-                      ncol  = 101, # length(fd.basis$names)
+                      ncol  = length(unique(data$frame)), # length(fd.basis$names)
                       byrow = TRUE)
 
   # clust.agg.intrp <- t(sapply(apply(t(clust.agg), 2, approx, n = 101), "[[", "y"))

### 2021-08-10 R/floa_roislien.R

@@ -1,26 +1,33 @@
-floa_roislien <- function(data, n.boot, ver) { # , fd.basis
+floa_roislien <- function(data, n.boot) {
 
-  # ############################################################################
+  # ****************************************************************************
   # Functional limits of agreement according to Roislien et al. (2012)
-  # ############################################################################
+  #
+  # In Roislien et al., FLoA are calculated as 95% CI.
+  # ****************************************************************************
 
   # Get one random stride from each subject ONCE and bootstrap the resulting
   # sample (of length (n=length(subjects))
   # ----------------------------------------------------------------------------
-  curve.idx <- sample(unique(data$strideID), size = 1)
 
-  curve <- data[data$strideID %in% curve.idx, ]
+  source("pick_subwise_curves.R")
+  source("functional_mean.R")
+  source("functional_sd.R")
+  source("boot_mean_sd.R")
 
-  curve0 <- subset(curve, device  == "IMU")$value
-  curve0 <- matrix(curve0, ncol = length(curve0) / 100)
-
-  curve1 <- subset(curve, device  == "MC")$value
-  curve1 <- matrix(curve1, ncol = length(curve1) / 100)
-
-  diff.curves <- curve0 - curve1
+  # Bootstrap to get 95% CI FLoA -----------------------------------------------
+  func.boot <- boot_mean_sd(data, n.boot) # data need to have dimension [101, x]
 
   # Calculate Limits of Agreement ----------------------------------------------
+  func.mean.boot <- functional_mean(func.boot) # mean of bootstrapped distribution
+  func.sd.boot <- functional_sd(func.boot)
 
+  # Get the same structure as returned by the other methods (i. e. floa_rcb)
+  floa.roislien <- rbind(func.mean.boot + 1.96 * func.sd.boot,
+                         func.mean.boot,
+                         func.mean.boot - 1.96 * func.sd.boot)
 
-  return(floa)
+  row.names(floa.roislien) <- c("upper", "mean", "lower")
+
+  return(floa.roislien)
 }

### 2021-08-10 R/functional_mean.R

@@ -1,0 +1,13 @@
+functional_mean <- function (data) {
+
+  # data need to have dimension [101, x]
+  n.frames <- dim(data)[1]
+
+  func.mean <- vector(mode = "numeric", length = n.frames)
+
+  for (i in 1:n.frames) {
+    func.mean[i] <- mean(data[i, ])
+  }
+
+  return(func.mean)
+}

### 2021-08-10 R/functional_sd.R

@@ -1,0 +1,13 @@
+functional_sd <- function (data) {
+
+  # data need to have dimension [101, x]
+  n.frames <- dim(data)[1]
+
+  func.sd <- vector(mode = "numeric", length = n.frames)
+
+  for (i in 1:n.frames) {
+    func.sd[i] <- sd(data[i, ])
+  }
+
+  return(func.sd)
+}

### 2021-08-10 R/get_coverage.R

@@ -24,7 +24,10 @@ get_coverage <- function (data, floa) {
   n.strides <- length(unique(device.diff$strideID))
   outside <- 0
 
-  for (stride.idx in 1:n.strides){
+  # Account for stride indices not starting at 1
+  stride.indices <- unique(data$strideID)
+
+  for (stride.idx in stride.indices){
 
     tmp <- subset(device.diff, strideID == stride.idx)
 

### 2021-08-10 R/main.R

@@ -10,9 +10,11 @@
 # (see subsection data sets)
 #
 # TODO:
+#   + Check leave-one-out implementation
 #   + Finish floa_roislien()
 #   + Reimplement crossval_coverage.R
 #   + Implementieren FDA
+#   + Preallocation in floa_rcb()
 #   + Quantile: Über die gesamte Verteilung oder die "Ausreisser-Quantile" einzelner/extremer Probanden
 #     + quantile() function: Bias correction useful/necessary?
 #   + Konvergenzanalyse --> ja, aber "nur" intern
@@ -41,6 +43,7 @@ dir.data <- "C:/Users/Daniel/Desktop/tmp/floa/R/examples"
 setwd(dir.script)
 source("example_data.R")
 source("fdaDelta.R")
+source("pick_subwise_curves.R")
 source("draw_clusters.R")
 source("floa_rcb.R")
 source("floa_point.R")
@@ -60,6 +63,7 @@ source("plot_cov_ver.R")
 # * Smooth wave data with nonlinear trend (constant variance): "smooth_trend"
 # * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
 # * Data with shock peaks (no bias, no trend): "shock"
+
 
 data <- example_data(dat = "smooth", dir.data)
 
@@ -159,6 +163,6 @@ print(coverage)
 
 cover.cross <- crossval_coverage(data, n.boot)
 
-plot_cov_ver(cover.cross.v1)
+plot_cov_ver(cover.cross)
 
 

### 2021-08-10 R/pick_subwise_curves.R

@@ -1,0 +1,29 @@
+pick_subwise_curves <- function (data) {
+
+  n.subj <- length(unique(data$subjectID))
+
+  curve.idx <- vector(mode = "numeric", length = n.subj)
+
+  for (subj.idx in 1:n.subj) {
+
+    if (subj.idx %in% unique(data$subjectID)) {
+
+      tmp <- data[data$subjectID == subj.idx, ]
+
+      curve.idx.subj <- as.numeric(sample(tmp$strideID, size = 1))
+      curve.idx <- c(curve.idx, curve.idx.subj)
+    }
+  }
+
+  curve <- data[data$strideID %in% curve.idx, ]
+
+  curve0 <- subset(curve, device  == "IMU")$value
+  curve0 <- matrix(curve0, ncol = length(curve0) / 100)
+
+  curve1 <- subset(curve, device  == "MC")$value
+  curve1 <- matrix(curve1, ncol = length(curve1) / 100)
+
+  diff.curves <- curve0 - curve1
+
+  return(diff.curves)
+}

### 2021-08-10 R/plot_cov_ver.R

@@ -8,6 +8,10 @@ plot_cov_ver <- function (cover.cross) {
   names(cover.cross.long)[names(cover.cross.long) != 'time'] <- 'value'
   names(cover.cross.long)[names(cover.cross.long) == 'time'] <- 'method'
 
-  cover.PLOT <- ggplot(cover.cross.long, aes(x = method, y = value)) + geom_boxplot()
+  cover.PLOT <- ggplot(cover.cross.long, aes(x = method, y = value)) +
+    geom_boxplot() +
+    ylim(0, 1) +
+    ylab("coverage")
+
   cover.PLOT
 }

### 2021-08-11 paper.Rmd
Adapt coverage functions

@@
 abstract: |
-  This is the abstract.
+  Paper als short communication / technical note?
 
@@
 
-# setwd(dir.script)
 source(paste0(dir.script, "/example_data.R"))
@@
 source(paste0(dir.script, "/draw_clusters.R"))
+source(paste0(dir.script, "/functional_mean.R"))
+source(paste0(dir.script, "/functional_sd.R"))
+source(paste0(dir.script, "/boot_mean_sd.R"))
 source(paste0(dir.script, "/floa_rcb.R"))
@@
 source(paste0(dir.script, "/get_coverage.R"))
+source(paste0(dir.script, "/get_coverage_fraction.R"))
 source(paste0(dir.script, "/crossval_coverage.R"))
+source(paste0(dir.script, "/crossval_coverage_fraction.R"))
 source(paste0(dir.script, "/plot_cov_ver.R"))
@@
 
-Leave-one (subject) out method to estimate the achieved coverage
+Leave-one out method to estimate the achieved coverage
 
-Versions:
+Prediction intervals are calculated from size.dataset = number.subjects - 1 and then validated against the left out subjects curves for five different methods:
 
-* v1   : n = length(subjects) random strides from all strides
+* FLoA~RCB_v1~   : n = length(subjects) random strides from all strides
 
-* v2   : One random stride per subject
+* FLoA~RCB_v2~   : One random stride per subject
 
-* v3   : Fetch a SINGLE random stride from all strides
+* FLoA~RCB_v3~   : Fetch a SINGLE random stride from all strides
 
-* v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
+* FLoA~RCB_Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
        strap the resulting sample (of length (n=length(subjects))
+
+* FLoA~RCB_Point~ : Pointwise B & A Limits of Agreement
 
@@
 data <- example_data(dat = "smooth", dir.data)
-n.boot <- 100
+n.boot <- 400
 floa <- floa_rcb(data, n.boot, ver = "v2")
@@
 
+Displayed are boxes for five different methods:
+
+* 1. FLoA~RCB_v1~ : n = length(subjects) random strides from all strides
+
+* 2. FLoA~RCB_v2~ : One random stride per subject
+
+* 3. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides
+
+* 4. FLoA~RCB_Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
+       strap the resulting sample (of length (n=length(subjects))
+
+* 5. FLoA~RCB_Point~ : Pointwise B & A Limits of Agreement
+
+Coverages are displayed for two different scenarios:
+
+* 1: Coverage of the entire curve
+
+* 2: Percentage of curve points within the respective FLoA
+
+#### Smooth, wave data (normal error, constant variance, no trend)
+
+Coverage of the entire curve
+
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-n.boot <- 100
 cover.cross <- crossval_coverage(data, n.boot)
+# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
 
 plot_cov_ver(cover.cross)
+
+# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
 ```
 
-<!-- #### v1 (n = length(subjects) random strides from all strides) -->
+Percentage of curve points within the respective FLoA
 
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- data <- example_data(dat = "smooth", dir.data) -->
-<!-- ``` -->
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Display cross validation coverages (fraction of the curve that is covered) across methods
+cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
 
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v1") -->
-<!-- floa.point <- floa_point(data) -->
+plot_cov_ver(cover.cross.fraction)
+```
 
-<!-- cover.cross.v1 <- crossval_coverage(data, floa, floa.point, method = "all") -->
+#### Smooth wave data with nonlinear trend (constant variance)
 
-<!-- plot_cov_ver(cover.cross.v1) -->
-<!-- ``` -->
+Coverage of the entire curve
 
-<!-- #### V2 (One random stride per subject) -->
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_trend", dir.data)
 
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
+cover.cross <- crossval_coverage(data, n.boot)
+# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
 
-<!-- cover.cross.v2 <- crossval_coverage(data, floa, floa.point, method = "all") -->
+plot_cov_ver(cover.cross)
 
-<!-- plot_cov_ver(cover.cross.v2) -->
-<!-- ``` -->
+# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
+```
 
-<!-- #### v3 (Fetch a SINGLE random stride from all strides) -->
+Percentage of curve points within the respective FLoA
 
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v3") -->
-<!-- floa.point <- floa_point(data) -->
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Display cross validation coverages (fraction of the curve that is covered) across methods
+cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
 
-<!-- cover.cross.v3 <- crossval_coverage(data, floa, floa.point, method = "all") -->
+plot_cov_ver(cover.cross.fraction)
+```
 
-<!-- plot_cov_ver(cover.cross.v3) -->
-<!-- ``` -->
+#### Data with non-gaussian (Weibull distributed) error (no trend)
+
+Coverage of the entire curve
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "non_gaussian", dir.data)
+
+cover.cross <- crossval_coverage(data, n.boot)
+# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+plot_cov_ver(cover.cross)
+
+# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
+```
+
+Percentage of curve points within the respective FLoA
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Display cross validation coverages (fraction of the curve that is covered) across methods
+cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+
+plot_cov_ver(cover.cross.fraction)
+```
 
@@
 
+It might be beneficial to work with functional data as suggested by Lenhoff et al.
+(1999) or Roislien et al.
+(2012) when e.
+g.
+interpolating data to a standardized length.
+This approach, however, wasn't necessary in this paper, since the synthetic data were created from functions and were designed to all have equal length (101 data points).
+
 # Conclusion

### 2021-08-11 R/crossval_coverage_fraction.R

@@ -1,0 +1,83 @@
+crossval_coverage_fraction <- function (data, n.boot) {
+
+  # ####################################################################
+  # Leave-one (subject) out method to estimate the achieved coverage
+  # See e. g. Lenhoff et al. (1999)
+  # ####################################################################
+  #
+  # Currently, different versions of the sampling process in draw_clusters()
+  # (nested in floa_rcb()) are implemented:
+  #
+  # v1 : n = length(subjects) random strides from all strides
+  # v2 : One random stride per subject
+  # v3 : Fetch a SINGLE random stride from all strides
+  # v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
+  #      strap the resulting sample (of length (n=length(subjects))
+  #
+  # Output:
+  #   * Coverage levels [%] across n=length(subjectID) iterations
+  # ####################################################################
+
+  n.subj <- unique(data$subjectID)
+
+  cover.cross.rcb.v1   <- c()
+  cover.cross.rcb.v2   <- c()
+  cover.cross.rcb.v3   <- c()
+  cover.cross.point    <- c()
+  cover.cross.roislien <- c()
+
+  for (i in n.subj) {
+
+    # Calculate FLoA with one subject left out
+    # --------------------------------------------------------------------
+    data.one.out <- subset(data, subjectID != i) # Leave subject "i" out
+
+    floa.point    <- floa_point(data.one.out)
+    floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
+    floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
+    floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
+    floa.roislien <- floa_roislien(data.one.out, n.boot)
+
+    # Get coverage of the left out (unseen) curves of subject "i"
+    # --------------------------------------------------------------------
+    data.subset <- subset(data, subjectID == i)
+
+    cover.cross.rcb.v1 <- c(cover.cross.rcb.v1,
+                            get_coverage_fraction(data.subset, floa.v1)
+    )
+
+    # plot(floa.v1["mean", ], type = "l", col = "red", ylim = c(-1, 1))
+    # lines(floa.v1["upper", ], col = "red")
+    # lines(floa.v1["lower", ], col = "red")
+    # lines(tmp$value)
+
+    cover.cross.rcb.v2 <- c(cover.cross.rcb.v2,
[…]

### 2021-08-11 R/draw_clusters.R

@@ -11,7 +11,7 @@ draw_clusters <- function(data, ver) { # fd.basis
   # v3   : Fetch a single stride only from all available strides
   # ----------------------------------------------------------------------------
 
-  source("pick_subwise_curves.R")
+  # source("pick_subwise_curves.R")
 
   if (ver == "v1") {
 

### 2021-08-11 R/floa_roislien.R

@@ -10,10 +10,10 @@ floa_roislien <- function(data, n.boot) {
   # sample (of length (n=length(subjects))
   # ----------------------------------------------------------------------------
 
-  source("pick_subwise_curves.R")
-  source("functional_mean.R")
-  source("functional_sd.R")
-  source("boot_mean_sd.R")
+  # source("pick_subwise_curves.R")
+  # source("functional_mean.R")
+  # source("functional_sd.R")
+  # source("boot_mean_sd.R")
 
   # Bootstrap to get 95% CI FLoA -----------------------------------------------
   func.boot <- boot_mean_sd(data, n.boot) # data need to have dimension [101, x]

### 2021-08-11 R/get_coverage_fraction.R

@@ -1,0 +1,49 @@
+get_coverage_fraction <- function (data, floa) {
+
+  # Get difference curves
+  # ----------------------------------------------------------------------------
+  device1 <- data.frame(
+    subset(data, device == "IMU")
+  )
+
+  device2 <- data.frame(
+    subset(data, device == "MC")
+  )
+
+  device.diff <- subset(device1,
+                        select = c(subjectID, strideID, frame))
+
+  device.diff$value <- device1$value - device2$value
+
+
+  # Calculate coverage (entire curves within the percentile boundaries)
+  # ----------------------------------------------------------------------------
+  lwr.bnd <- floa["lower", ]
+  upr.bnd <- floa["upper", ]
+
+  n.strides <- length(unique(device.diff$strideID))
+  inside.thresh.perc <- c()
+
+  # Account for stride indices not starting at 1
+  stride.indices <- unique(data$strideID)
+
+  for (stride.idx in stride.indices){
+
+    tmp <- subset(device.diff, strideID == stride.idx)
+
+    # Compare difference curves with upper and lower boundaries
+    below.thresh <- tmp$value < lwr.bnd
+    above.thresh <- tmp$value > upr.bnd
+
+    # Percentage of points outside limits
+    inside <- below.thresh == above.thresh
+    inside.thresh.perc <- c(inside.thresh.perc,
+                             round(mean(sum(inside[TRUE]) / length(tmp$value)), 2))
+
+    # percent.out <- round(mean(outside.thresh.perc), 2) # Percentage of points outside limits
+  }
+
+  coverage <- round(mean(inside.thresh.perc), 2)
+
+  return(coverage)
+}

### 2021-08-11 R/main.R

@@ -46,12 +46,17 @@ source("example_data.R")
 source("fdaDelta.R")
 source("pick_subwise_curves.R")
 source("draw_clusters.R")
+source("functional_mean.R")
+source("functional_sd.R")
+source("boot_mean_sd.R")
 source("floa_rcb.R")
 source("floa_point.R")
 source("floa_roislien.R")
 source("plot_loa.R")
 source("get_coverage.R")
+source("get_coverage_fraction.R")
 source("crossval_coverage.R")
+source("crossval_coverage_fraction.R")
 source("plot_cov_ver.R")
 
 
@@ -161,7 +166,11 @@ print(coverage)
 # ------------------------------------------------------------------------------
 cover.cross <- crossval_coverage(data, n.boot)
 
-# Display cross validation coverages as boxplots across methods
+# Display cross validation coverages (entire curve covered) across methods
 plot_cov_ver(cover.cross)
 
+# Display cross validation coverages (fraction of the curve that is covered) across methods
+cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+
+plot_cov_ver(cover.cross.fraction)
 

### 2021-08-11 R/plot_cov_ver.R

@@ -9,9 +9,12 @@ plot_cov_ver <- function (cover.cross) {
   names(cover.cross.long)[names(cover.cross.long) == 'time'] <- 'method'
 
   cover.PLOT <- ggplot(cover.cross.long, aes(x = method, y = value)) +
-    geom_boxplot() +
-    ylim(0, 1) +
+    geom_boxplot(width = .5) +
+    # geom_jitter(size = 3, width = .15, colour = "red") +
+    # geom_violin() +
+    ylim(-0.1, 1.1) +
     ylab("coverage")
 
   cover.PLOT
-}
+
+  }

### 2021-08-11 R/plot_loa.R

@@ -29,7 +29,6 @@ plot_loa <- function (data, method.loa, central.tendency) {
 
 
   # Actually plot the prepared data --------------------------------------------
-
   if (central.tendency == "median") {
 
     # For line graphs, the data points must be grouped so that it knows which points to connect.

### 2021-08-12 R/get_coverage.R
Fix bug in pointwise coverage

@@ -29,11 +29,11 @@ get_coverage <- function (data, floa) {
 
   for (stride.idx in stride.indices){
 
-    tmp <- subset(device.diff, strideID == stride.idx)
+    curve <- subset(device.diff, strideID == stride.idx)
 
     # Compare difference curves with upper and lower boundaries
-    below.thresh <- tmp$value < lwr.bnd
-    above.thresh <- tmp$value > upr.bnd
+    below.thresh <- curve$value < lwr.bnd
+    above.thresh <- curve$value > upr.bnd
 
     points.outside <- sum(above.thresh) + sum(below.thresh)
 

### 2021-08-12 R/get_coverage_fraction.R

@@ -16,34 +16,31 @@ get_coverage_fraction <- function (data, floa) {
   device.diff$value <- device1$value - device2$value
 
 
-  # Calculate coverage (entire curves within the percentile boundaries)
+  # Calculate coverage (points within the limits)
   # ----------------------------------------------------------------------------
   lwr.bnd <- floa["lower", ]
   upr.bnd <- floa["upper", ]
 
   n.strides <- length(unique(device.diff$strideID))
   inside.thresh.perc <- c()
-
   # Account for stride indices not starting at 1
   stride.indices <- unique(data$strideID)
 
   for (stride.idx in stride.indices){
 
-    tmp <- subset(device.diff, strideID == stride.idx)
+    curve <- subset(device.diff, strideID == stride.idx)
 
     # Compare difference curves with upper and lower boundaries
-    below.thresh <- tmp$value < lwr.bnd
-    above.thresh <- tmp$value > upr.bnd
+    below.thresh <- curve$value < lwr.bnd
+    above.thresh <- curve$value > upr.bnd
 
-    # Percentage of points outside limits
+    # Mean percentage of points outside the limits
     inside <- below.thresh == above.thresh
     inside.thresh.perc <- c(inside.thresh.perc,
-                             round(mean(sum(inside[TRUE]) / length(tmp$value)), 2))
-
-    # percent.out <- round(mean(outside.thresh.perc), 2) # Percentage of points outside limits
+                            sum(inside[TRUE]) / length(curve$value))
   }
 
-  coverage <- round(mean(inside.thresh.perc), 2)
+  coverage <- round(mean(inside.thresh.perc), 3)
 
   return(coverage)
 }

### 2021-08-12 R/main.R

@@ -169,7 +169,7 @@ cover.cross <- crossval_coverage(data, n.boot)
 # Display cross validation coverages (entire curve covered) across methods
 plot_cov_ver(cover.cross)
 
-# Display cross validation coverages (fraction of the curve that is covered) across methods
+# Mean fraction of the curve that is covered) across methods
 cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
 
 plot_cov_ver(cover.cross.fraction)

### 2021-08-12 paper.Rmd
Rewrite introduction

@@
 ---
-title: "Asymmetric functional prediction bands for assessing the validity of spatiotemporally smooth biomechanical curve data"
+title: "Asymmetric continuous prediction bands for assessing the validity of spatiotemporally smooth biomechanical curve data"
 author:
@@
 
-### TODO read (ee 'Literature' folder on Nextcloud)
+### TODO read
 - Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem
@@
 g.
-joint angles, ground reaction forces, or center of mass trajectories - is available in the form of rather smooth curves within which the error can vary over time.
+joint angles, ground reaction forces, or center of mass trajectories - is available in the form of continuous signals (curve data) within which the error can vary over time.
 
-Therefore, when analyzing the validity of continuous data, the entire length of the curve should be accounted for.
+Therefore, when analyzing the validity of curve data, the entire length of the curve should be accounted for.
 This can be accomplished e.
@@
 (2019)].
-The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points of biomechanical curves are locally correlated [Deluzio & Astephen, 2007; Pataky, 2010].
-From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for [Lenhoff et al., 1999].
+The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points of biomechanical curves are locally correlated (Deluzio & Astephen, 2007; Pataky, 2010).
+From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
 In addition, most traditional point statistics are either not suitable (e.g.
@@
 
-One possibility to address te problem of a pointwise analysis is to treat curves as single objects rather than as a series of unconnected points [Ramsay & Silverman, 2005].
+One possibility to address the problem of pointwise analysis is to treat entire curves as (functional) objects rather than as a series of unconnected points [Ramsay & Silverman, 2005].
 Røislien et al.
-[2012] presented a method that extends the Limits of Agreement (LoA) concept [11] to functional data.
-This is achieved by approximating curves using Fourier series and resampling the thus obtained functions.
+(2012) presented a method that extends the Limits of Agreement (LoA) suggested by Bland & Altman to functional data by approximating measurement signals using Fourier series.
+These function are used to calculated functional difference curves which are then bootstrapped to obtain 95% confidence intervals.
+The approach accounts for the autocorrelation characteristic in the signal and at the same time allows for an intuitive description of the systematic and random error terms in the original sampling space (e.g.
+actual degrees).
+Røislien et al., however, make a number of strong assumptions, which are often not or only partially fulfilled in empirical data sets:
+
+(i) The error distribution is approximately normal.
+
+The construction of symmetric prediction bands as described in Røislien et al.
+[10] relies on normally distributed measurement data.
+Non-normality, however, is often present when e.g.
+sensor values are truncated because of limited range or resolution, or because values drift over time.
+
+(ii) Time-continuous differences of two measurement systems and the corresponding averages must be uncorrelated.
+
+This assumption is violated considerably in many validation data sets [12].
+E.g., Atallah et al.
+[13] found that the agreement between an ear-worn accelerometer and a force-plate instrumented treadmill was associated with gait speed when estimating gait cycle duration.
+When validating a foot-worn gyroscope for treadmill running, Koska & Maiwald [14] observed a linear error trend that was associated with foot strike behavior.
+
+(iii) Aus einem Datensatz bestehend aus einer Kurve eines Probanden werden 95% Konfidenzintervalle gebootrapped, die ein geeignetes Maß für die estimation uncertainty darstellen.
+
+In jeder Iteration des Bootstrap wird eine arithmetische mean curve gebildet. ... the limits will be too narrow by a factor of approximately the square root of n.subj.
+
+<!-- Prädiktionsbänder sind weiter als Konfidenzbänder ... und eigentlich will man ja Prädiktionsbänder, oder!?! -->
 
@@
 
-The approach accounts for the autocorrelation characteristic in the signal and at the same time allows for an intuitive description of the systematic and random error terms in the original sampling space (e.g.
-actual degrees).
-Røislien et al.
-[10], however, make a number of strong assumptions, which are often not or only partially fulfilled in empirical data sets:
+<!-- Formulierung aus dem alten Skript -->
+<!-- (iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject. -->
+<!-- If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
+The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
+-->
 
-(i) Time-continuous differences of two measurement systems and the corresponding averages must be uncorrelated.
-This assumption is violated considerably in many validation data sets [12].
-E.g., Atallah et al.
-[13] found that the agreement between an ear-worn accelerometer and a force-plate instrumented treadmill was associated with gait speed when estimating gait cycle duration.
-When validating a foot-worn gyroscope for treadmill running, Koska & Maiwald [14] observed a linear error trend that was associated with foot strike behavior.
-
-(ii) The error distribution is approximately normal.
-The construction of symmetric prediction bands as described in Røislien et al.
-[10] relies on normally distributed measurement data.
-Non-normality, however, is often present when e.g.
-sensor values are truncated because of limited range or resolution, or because values drift over time.
-
-(iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject.
-If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
-The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
+<!-- Weiterer ("zugehöriger) Reviewerkommentar: -->
+<!-- In looking at limits of agreement, the authors are considering the extent of within pair variation.
+However, in looking at the point wise limits of agreement, these are defined to incorporate a between pair standard deviation.
+Even if it was appropriate to utilise between pair variability, the authors should note that it is variances that are additive and not SDs.
+Perhaps the authors' lack of understanding comes from their erroneous assumption (iii).
+-->
 
@@
 
-The principal object of this paper was to demonstrate a method for curve data that addresses the aforementioned issues (analyzing the whole curve, linear model assumptions, repeated measures), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret prediction bands (i.e.
+The principal object of this paper was to demonstrate a method for constructing continuous prediction intervals from curve data that addresses the aforementioned issues (analyzing the whole curve, including repeated measures, (parametric) model assumptions), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret error bands (i.e.
 LoA in the actual measurement unit).
 
-We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters.
+We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters and LoA constructed similar to the method described in Roislien et al.
+(2012).
+For model validation, the respective LoA are cross validated using a leave-one-out approach.
 
@@
 
-* Randomized Cluster Bootstrap (floa.rcb)
-  + Different strategies
+* Randomized Cluster Bootstrap (FLoA~RCB~)
+  + n = length(subjects) random strides from all strides (FLoA~RCB_v1~)
+  + One random stride per subject (FLoA~RCB_v2~)
+  + Fetch a SINGLE random stride from all strides (FLoA~RCB_v3~)
 
-* Point-by-point Gaussian intervals (floa.point)
+* Pointwise Gaussian intervals (FLoA~Point~)
+  + FLoA~Point~ are calculated using mean and standard deviations (derived from linear mixed effects models)
+
+* Roislien-like intervals (FLoA~Roslien~)
+  + (Get one random stride from each subject ONCE and boot-
+       strap the resulting sample (of length (n=length(subjects))
 
@@
 Die verschiedenen Zeitreihen wurden als Fourierreihen aus additiv überlagerten Sinusfunktionen modelliert.
-Jeder Datensatz enthält n.ts = 100 Kurven von n.subj = 11 fiktiven Probanden.
+Jeder Datensatz enthält n.strides = 10 Kurven von n.subj = 11 fiktiven Probanden.
 Jede Kurve besteht aus t = 101 Datenpunkten.
@@
 
-  for (ts.idx in 1:(n.ts)) {
+  for (stride.idx in 1:(n.strides)) {
 
@@
 
-* FLoA~RCB_Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
+* FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
        strap the resulting sample (of length (n=length(subjects))
 
-* FLoA~RCB_Point~ : Pointwise B & A Limits of Agreement
+* FLoA~Point~ : Pointwise B & A Limits of Agreement
 
@@
 
-* 4. FLoA~RCB_Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
+* 4. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
        strap the resulting sample (of length (n=length(subjects))
 
-* 5. FLoA~RCB_Point~ : Pointwise B & A Limits of Agreement
+* 5. FLoA~Point~ : Pointwise B & A Limits of Agreement
 
@@
 
-* 1: Coverage of the entire curve
+* a: Percentage of entirely covered curves (all points within the limits): Each box consists of n~ = 11 values (one percentage for each subject).
 
-* 2: Percentage of curve points within the respective FLoA
+* b: Mean percentage of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
 
@@
 
-Coverage of the entire curve
+a. (coverage of the entire curve)
+
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+lwr.bnd <- floa["lower", ]
+upr.bnd <- floa["upper", ]
+
+# ...
+
+for (stride.idx in stride.indices){
+    curve <- subset(device.diff, strideID == stride.idx)
+
+    # Compare difference curves with upper and lower boundaries
+    below.thresh <- curve$value < lwr.bnd
+    above.thresh <- curve$value > upr.bnd
+
+    points.outside <- sum(above.thresh) + sum(below.thresh)
+
+    # Check if the entire curve is within the limits
+    if (points.outside > 0) {
+      outside <- outside + 1
+    }
+  }
+
+coverage <- 1 - (outside / n.strides)
+```
 
@@
 
-Percentage of curve points within the respective FLoA
+b. (mean percentage of covered curve points)
+
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+lwr.bnd <- floa["lower", ]
+upr.bnd <- floa["upper", ]
+
+# ...
+
+for (stride.idx in stride.indices){
+
+  curve <- subset(device.diff, strideID == stride.idx)
+
+  # Compare difference curves with upper and lower boundaries
+  below.thresh <- curve$value < lwr.bnd
+  above.thresh <- curve$value > upr.bnd
+
+  # Mean percentage of points outside the limits
+  inside <- below.thresh == above.thresh
+  inside.thresh.perc <- c(inside.thresh.perc,
+                          mean(sum(inside[TRUE]) / length(curve$value)))
+}
+```
 
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Display cross validation coverages (fraction of the curve that is covered) across methods
+# Display cross validation coverages (mean fraction of the curve that is covered) across methods
 cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
@@
 
-Coverage of the entire curve
+a. (coverage of the entire curve)
 
@@
 
-Percentage of curve points within the respective FLoA
+b. (mean percentage of covered curve points)
 
@@
 
-Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
+<!-- Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
 werden 95% Konfidenzintervalle gebootstrapped.
@@
 It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of n.subj.
+-->
 

### 2021-08-22 paper.Rmd
Add ideas to the paper draft

@@
 ### TODO
+- Evtl. simulierte Datensätze um Beispiele aus https://mjskay.github.io/ggdist/articles/lineribbon.html erweitern (insbesondere das erste Beispiel mit x-versetzten Gausskurven)!
 - Simulierte Daten durch Funktionen ersetzen?
@@
 - Falls doch FDA, dann evtl. als eigenständiges Skript forken
+- Bergründung warum Prädiktionsintervalle anstelle von Konfidenzintervallen
+- Lineribbon “gradients”
+<!-- - Umbennung entire curve in "All-or-nothing" (entlehnt von Juul et al.) -->
 - .bib file anlegen
@@
 - Ratkowsky besorgen
+- Begründung warum Prädiktionsintervalle und nicht Konfidenzintervalle (theoretisch könnte man beides machen; Prädiktionsintervalle scheinen mir aber eher das zu sein, was man will)
 
-### TODO read
+### TODO read (again)
+- Olshen et al. (1989) Gait analysis and the bootstrap
+- Curvewise point and interval summaries for tidy data frames of draws from distributions https://mjskay.github.io/ggdist/reference/curve_interval.html
 - Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem
@@
 g.
-by applying statistics (e.
-g.
-standard deviation) separately to all points of the curve [Pini et al.
-(2019)].
-The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points of biomechanical curves are locally correlated (Deluzio & Astephen, 2007; Pataky, 2010).
+by applying statistics separately to all points of the curve [Pini et al.
+(2019).
+The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points within a curve are locally correlated (Deluzio & Astephen, 2007; Pataky, 2010).
+From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
+
+... and likely results in a systematic underrepresentation of ... values (Juul et al., 2020; Mirzagar et al. (2014).
+
+<!-- https://mjskay.github.io/ggdist/reference/curve_interval.html -->
+<!-- Hier referenzierte Paper u. a.  -->
+<!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
+(2020).
+"Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles".
+arXiv e-print.
+arXiv:2007.05035 -->
+<!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
+(2014).
+"Curve Boxplot: Generalization of Boxplot for Ensembles of Curves".
+IEEE Transactions on Visualization and Computer Graphics.
+20(12): 2654-2663.
+doi: 10.1109/TVCG.2014.2346455 -->
+
 From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
@@
 
-One possibility to address the problem of pointwise analysis is to treat entire curves as (functional) objects rather than as a series of unconnected points [Ramsay & Silverman, 2005].
+One possibility to address the problem of pointwise analysis is to treat entire curves as (functional) objects rather than as a series of unconnected points.
+
+<!-- Es gibt eine Reihe von Ansätzen aus anderen Bereichen (z. B. Covid Prädiktion), bei denen sog. Function oder Curve Boxplots eingesetzt werden. -->
+<!-- https://mjskay.github.io/ggdist/reference/curve_interval.html -->
+<!-- Hier referenzierte Paper u. a.  -->
+<!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
+(2020).
+"Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles".
+arXiv e-print.
+arXiv:2007.05035 -->
+<!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
+(2014).
+"Curve Boxplot: Generalization of Boxplot for Ensembles of Curves".
+IEEE Transactions on Visualization and Computer Graphics.
+20(12): 2654-2663.
+doi: 10.1109/TVCG.2014.2346455 -->
+
+Sutherland et al.
+(1988) introduced a method in which continuous signals are reconstructed using Fourier series and bootstrapped to estimate of the variation of the sample mean and thus estimate confidence intervals.
+Lenhoff et al.
+(1999) illustrated their method on a generic gait analysis data set (comprised of joint angle curves) and compared it to pointwise Gaussian intervals.
+The authors found that Sutherland et al.
+bootstrap bands provide appropriate coverage for continuous curve gait data (86% coverage for a targeted coverage of 90%) while pointwise Gaussian bands are shown to provide inadequate coverage (54% for a targeted coverage of 90%).
+
 Røislien et al.
-(2012) presented a method that extends the Limits of Agreement (LoA) suggested by Bland & Altman to functional data by approximating measurement signals using Fourier series.
-These function are used to calculated functional difference curves which are then bootstrapped to obtain 95% confidence intervals.
+(2012) presented a similar approach to calculate a functional version of the the Limits of Agreement (LoA) method suggested by Bland & Altman.
+They first calculate functional difference curves which are then bootstrapped to obtain 95% confidence intervals.
 The approach accounts for the autocorrelation characteristic in the signal and at the same time allows for an intuitive description of the systematic and random error terms in the original sampling space (e.g.
 actual degrees).
-Røislien et al., however, make a number of strong assumptions, which are often not or only partially fulfilled in empirical data sets:
+Røislien et al., however, rely on a number of strong model assumptions, which are often not or only partially fulfilled in empirical data sets:
 
@@
 The construction of symmetric prediction bands as described in Røislien et al.
-[10] relies on normally distributed measurement data.
+[10] relies on Gaussian assumptions.
 Non-normality, however, is often present when e.g.
@@
 
-(iii) Aus einem Datensatz bestehend aus einer Kurve eines Probanden werden 95% Konfidenzintervalle gebootrapped, die ein geeignetes Maß für die estimation uncertainty darstellen.
+(iii) Aus einem Datensatz bestehend aus einer Kurve eines Probanden lassen sich durch Bootstrapping von mittleren Kurven Unsicherheitsintervalle bilden, die ein geeignetes Maß für die estimation uncertainty darstellen.
 
-In jeder Iteration des Bootstrap wird eine arithmetische mean curve gebildet. ... the limits will be too narrow by a factor of approximately the square root of n.subj.
+Hier werden arithmetische mean curves verwendet, the variability of which will reflect variation between means.
+In many cases, however, it is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of the number of curves within the sample.
 
-<!-- Prädiktionsbänder sind weiter als Konfidenzbänder ... und eigentlich will man ja Prädiktionsbänder, oder!?! -->
+Ein weiteres Problem ist, dass die Methode keine messwiederholten Daten berücksichtigt.
+Da sich messwiederholte Kurven i.
+d.
+R.
+ähnlicher sind als unabhängige Kurven, besteht die Gefahr die Streuung des Messsystems zu überschätzen.
 
@@
 
-Olsen et al.
+<!-- Olsen et al.
 [15] examined the validity of an inertial measurement unit (IMU) for analyzing the spatial displacement of the distal limbs of horses using functional LoA (FLoA).
@@
 The procedure, however, is subject to the same restrictions as any conventional linear regression model and is therefore limited when model assumptions are violated.
+-->
 
-The principal object of this paper was to demonstrate a method for constructing continuous prediction intervals from curve data that addresses the aforementioned issues (analyzing the whole curve, including repeated measures, (parametric) model assumptions), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret error bands (i.e.
+<!-- Warum eigentlich Prädiktionsbänder und keine Konfidenzbänder? (Siehe Einleitung Lenhoff) -->
+<!-- Aktuell würde ich es so formulieren: In der Mehrheit der Fälle ist man an der nächsten (Einzel)Observation interessiert, nicht unbedingt an Mittelwerten (obwohl es durchaus Szenarien gibt, wo das akzeptabel sein kann) + im Zweifel generell lieber breitere Fehlerintervalle -->
+
+The principal object of this paper was to demonstrate a method for constructing continuous prediction intervals from curve data that addresses the aforementioned issues (analyzing the whole curve, including repeated measures, model assumptions (normal distribution, equal variances)), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret error bands (i.e.
 LoA in the actual measurement unit).
@@
 
-## Validation
+s## Validation
 
@@
 
+Wir empfehlen die Daten zu plotten.
+Am Ende ist jeder Datensatz bzw.
+jede Fragestellung unterschiedlich und ein pauschaler coverage value wird der Qualität des Messgeräts (in Bezug auf die jeweilige Fragestellung möglicherweise nicht gerecht).
+
+<!-- https://mjskay.github.io/ggdist/articles/lineribbon.html -->
+<!-- In general I have found that there is no one method that consistently works well on all datasets.
+No matter the method, intervals often become problematic above 50%, hence the default .width for curve_interval() is 0.5 (unlike the default for point_interval(), which is 0.95).
+In any case, caution when using these intervals is advised.
+-->
+
+<!-- Welcher coverage Strategie ist besser (Entire curve (all-or-nothing) vs. percent-of-curve)? -->
+<!-- Denkansatz aus Juul et al. (2020) -->
+<!-- At the same time, however, it is clear that there is still important research to be done. We have suggested different ways of ranking centrality of ensemble curves. Each rank- -->
+<!-- ing has its merits: For example, the all-or-nothing ranking penalizes curves which are shaped differently than other ensemble curves and a weighted ranking can reward curves that are central in the very near future.
+A thorough investigation of these merits and trade-offs is needed.
+Until this has been investigated, we encourage researchers to be creative and mindful about the problems that each statistical method could have.
+And to communicate this uncertainty openly to decision makers.
+-->
+
+<!-- Für die hier untersuchten biomechanischen Signale nicht so relevant, aber an sich ein sehr interessantes Feature ist das hier: -->
+<!-- Leicht abgewandelt nach Juul et al. (2020): -->
+<!-- In addition to the ranking methods (for finding quantiles/percentiles) mentioned above, one can rank the curves according to some feature of interest.
+In Fig.
+2E we show the curve box plots obtained when we rank curves according to their projected maximum values of newly hospitalized cases in a single day; in other words, the median projected peak value received -->
+<!-- the highest centrality. -->
+
+<!-- Interessant beim Juul Paper ist auch der Umgang mit ("abgeschnittenen") Extremwerten: -->
+<!-- Here we have argued that computing confidence intervals using fixed-time descriptive systematically suppresses trajectory extremes. This is natural. -->
+<!-- Fixed-time descriptive statistics are designed to show the least extreme predictions on a given date, not to take entire curves into account. -->
+
 # Conclusion
+
+<!-- Aus Juul et al. (2020) -->
+<!-- In summary, when making forecasts of epidemic trajectories, it is important to represent the resulting curve ensembles in a way that captures the quantities of interest -->
+<!-- in an intuitive way. -->

### 2021-08-23 paper.Rmd
Add idea to paper introduction
Continue developing the paper

@@
 ---
-title: "Asymmetric continuous prediction bands for assessing the validity of spatiotemporally smooth biomechanical curve data"
+title: "Asymmetric continuous prediction bands for assessing the validity of smooth biomechanical curve data"
 author:
@@
 ### TODO
+- Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen
+- Weitere synthetische Daten siehe Paper Robinson et al. (2021)
 - Evtl. simulierte Datensätze um Beispiele aus https://mjskay.github.io/ggdist/articles/lineribbon.html erweitern (insbesondere das erste Beispiel mit x-versetzten Gausskurven)!
@@
 - Falls doch FDA, dann evtl. als eigenständiges Skript forken
-- Bergründung warum Prädiktionsintervalle anstelle von Konfidenzintervallen
-- Lineribbon “gradients”
+- Begründung warum Prädiktionsintervalle anstelle von Konfidenzintervallen
 <!-- - Umbennung entire curve in "All-or-nothing" (entlehnt von Juul et al.) -->
@@
 ### TODO read (again)
+- Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis
 - Olshen et al. (1989) Gait analysis and the bootstrap
@@
 
+Hier ein anderes Intro schreiben ... Ansatz: Neben der Validierung von Messgeräten spielt auch die Validierung von simulierten Modellergebnissen eine Rolle.
+
 Therefore, when analyzing the validity of curve data, the entire length of the curve should be accounted for.
@@
 
-... and likely results in a systematic underrepresentation of ... values (Juul et al., 2020; Mirzagar et al. (2014).
+<!-- ... and likely results in a systematic underrepresentation of ... values (Juul et al., 2020; Mirzagar et al. (2014). -->
 
@@
 
+<!-- Es gibt bereits Ansätze für funktionale Boxplots (inkl. Vergleich mit punktweisen Intervallen), das Ganze wurde aber noch nie für Differenzenkurven untersucht. -->
+
 We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters and LoA constructed similar to the method described in Roislien et al.
@@
 
+### FLoA~RCB~ construction
+
+<!-- See Mirzargar et al. (2014) or Juul et al. (2020) for an accessible introduction to data depth and curve boxplots / functional boxplots. -->
+
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-s## Validation
+## Validation
 
@@
 
-Für die Validierung der Methode wurden Zeitreihen simuliert, die gängigen biomechanischen Messsignalen qualitativ ähneln und typische Fehlercharakteristika abbilden.
-Die verschiedenen Zeitreihen wurden als Fourierreihen aus additiv überlagerten Sinusfunktionen modelliert.
-Jeder Datensatz enthält n.strides = 10 Kurven von n.subj = 11 fiktiven Probanden.
-Jede Kurve besteht aus t = 101 Datenpunkten.
+To compare different methods for constructing continuous LoA, smooth curve data were simulated across a range of common biomechanical signals and error characteristics.
+
+The curves were modeled as Fourier series from additively superimposed sine functions.
+Each data set contains n.strides = 10 curves from n.subj = 11 fictitious test subjects.
+Each curve consists of t = 101 data points.
+Note that the curves do not necessarily correspond to real-world signals as we were more concerned with illustrating the underlying principle.
+
+<!-- Aus Robinson et al. (2021) -->
+<!-- Exemplar 1D eects from the biomechanics literature across a range of signal types were collated and reported (Bakke & Besier, 2020, Barrios & Willson, 2017, Bovi, Rabuetti, Mazzoleni, & Ferrarin, 2011, Phan et al., 2017, Gomes, Ackermann, Ferreira, Orselli, & Sacco, 2017, Robinson et al., 2014).
+-->
+
+<!-- the code to reproduce all 1D eects is provided at https://github.com/m-a-robinson/sample-size -->
 
@@
 
-* (i) Smooth curves with normally distributed errors, constant variance, and no trend.
+* (1) Homoscedastic (normal) error.
 Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
-The data resembles e.
+The data could be generated e.
 g.
-kinematic joint angle curves captured by camera-based measurement systems or goniometers.
+from simultaneous joint angle measurements with two kinematic measurement systems (e.
+g.
+a camera-based measurement systems and a goniometer).
 
@@
 
-* (ii) Smooth curves created using the same model and parameters as in (i), but with a nonlinear trend underlying one of the two signals.
-A similar trend (drift) behavior can be observed in e.
+* (2) Same model as in (i), but with a nonlinear trend underlying one of the two signals.
+Similar errors can be observed e.
 g.
-when the same joint angle curves are calculated by integrating the signal of gyropscopes applied to measure the movement.
+when joint angles are calculated (numerically integrated) from gyroscope signals.
 
@@
 
-* (iii) Smooth curves created using the same model and parameters as in (i), but Weibulll distributed random curve (amplitude) parameters.
+* (3) Heteroscedasticity across the signal.
+The example consists of knee momentcurves calculated using two different methods (direct and inverse kinematic approach) (Robinson et al., 2014).
+The curves were generated using the example code in Robinson et al.
+(2021).
+
+<!-- Robinson et al. (2014) Impact of Knee Modeling Approach on Indicators and Classification of Anterior Cruciate Ligament Injury Risk -->
+
+<!-- Smooth curves created using the same model and parameters as in (i), but Weibulll distributed random curve (amplitude) parameters.
 This introduces non-gaussian measurement error (asymmetric/skewed distribution of differences around the mean difference).
+-->
 
@@
 ```
+
+* (4) Smooth curves with peaks
+
+Ground reaction force data
 

### 2021-08-23 paper.Rmd
Continue to develop the paper

@@
 
-Having a good estimate of the amount of measurement error is a basic requirement when analyzing movement data.
+<!-- Having a good estimate of the amount of measurement error is a basic requirement when analyzing movement data.
 The classic linear error model describes measurement error as the sum of systematic and random deviations from the true value.
@@
 joint angles, ground reaction forces, or center of mass trajectories - is available in the form of continuous signals (curve data) within which the error can vary over time.
+-->
 
-Hier ein anderes Intro schreiben ... Ansatz: Neben der Validierung von Messgeräten spielt auch die Validierung von simulierten Modellergebnissen eine Rolle.
+Hier ein anderes Intro schreiben ... Ansatz: Neben der Validierung von Messgeräten spielt auch die Validierung von Modellergebnissen eine Rolle!
 
@@
 Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
-The data could be generated e.
+The data are representative of e.
 g.
-from simultaneous joint angle measurements with two kinematic measurement systems (e.
+joint angle measurements simultaneously recorded from two kinematic measurement systems (e.
 g.
@@
 
-* (2) Same model as in (i), but with a nonlinear trend underlying one of the two signals.
+* (2) Trend error: Same model as in (i), but with a nonlinear trend underlying one of the two signals.
 Similar errors can be observed e.
@@
 
-* (3) Heteroscedasticity across the signal.
-The example consists of knee momentcurves calculated using two different methods (direct and inverse kinematic approach) (Robinson et al., 2014).
-The curves were generated using the example code in Robinson et al.
+* (3) Heteroscedastic error: Heteroscedasticity across the signal length.
+The example consists of knee moment curves calculated using two different methods (direct and inverse kinematic approach) (Robinson et al., 2014).
+The curves were recreated using the example code provided in Robinson et al.
 (2021).
@@
 
-* (4) Smooth curves with peaks
+* (4) Data with (missing) peaks.
+Same model as in (i), except a couple of frames where a short, prominent peak occurs in one of the two signals.
+This typically happens in high-frequency signals such as accelerations or ground reaction forces, where peaks are underrepresented or entirely missing when e.
+g.
+sampling rates are too low or lowpass filter artefacts occur.
+Also, modelling these kind of abrupt changes in the signal (when e.
+g.
+to estimate ground reaction forces from acceleration values) is more difficult and my result in similar errors (Alcantara, 2021).
 
-Ground reaction force data
+<!-- Alcantara et al. Modelling/estimating ground reaction forces from acceleration values, z. B. https://www.biorxiv.org/content/10.1101/2021.03.17.435901v1.full -->
+<!-- Johnson et al. (2020) Multidimensional ground reaction forces and moments from wearable sensor accelerations via deep learning -->
 
-### Smooth curves (normal error, constant variance, no trend)
+### Smooth curves (homoscedasticity, normal error, no trend)
 
@@
 
-### Smooth curves nonlinear error trend
+### Smooth curves with nonlinear error trend
 
@@
 
-#### Smooth curves with non-gaussian (Weibull) error (no trend)
+#### Heteroscedasticity across the signal length (Robinson et al., 2014)
+
+<!-- Smooth curves with non-gaussian (Weibull) error (no trend) -->
 
@@
 PLOT
+```
+
+#### Smooth curves with (missing) peaks
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+
 ```

### 2021-08-23 paper.Rmd
Modify shock data

@@
 
-#### Smooth wave data with nonlinear trend (constant variance)
+#### Curves with nonlinear trend (constant variance)
 
@@
 
-#### Data with non-gaussian (Weibull distributed) error (no trend)
+#### Curves with non-gaussian (Weibull distributed) error (no trend)
 
@@
 data <- example_data(dat = "non_gaussian", dir.data)
+
+cover.cross <- crossval_coverage(data, n.boot)
+# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+plot_cov_ver(cover.cross)
+
+# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
+```
+
+b. (mean percentage of covered curve points)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Display cross validation coverages (fraction of the curve that is covered) across methods
+cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+
+plot_cov_ver(cover.cross.fraction)
+```
+
+#### Curves with (missing) peaks
+
+a. (coverage of the entire curve)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "shock", dir.data)
 

### 2021-08-23 R/main.R

@@ -67,9 +67,9 @@ source("plot_cov_ver.R")
 # (* Empirical validation data: "imu_mc")
 # * Smooth, wave data (normal error, constant variance, no trend): "smooth"
 # * Smooth wave data with nonlinear trend (constant variance): "smooth_trend"
-# * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
+# * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian" # see Robinson et al. (2021)
 # * Data with shock peaks (no bias, no trend): "shock"
-data <- example_data(dat = "smooth", dir.data)
+data <- example_data(dat = "non_gaussian", dir.data)
 
 # Plot data --------------------------------------------------------------------
 

### 2021-08-23 R/simulate_data.R

@@ -526,6 +526,7 @@ saveRDS(data, file = paste0("C:/Users/Daniel/Desktop/tmp/floa/R/examples/", "dat
 # Shock (spike) data (normal error) -----------------------------------------
 # ------------------------------------------------------------------------------
 
+# Smooth wave data (constant variance, no trend) -------------------------------
 device <- c()
 value <- c()
 subjectID <- c()
@@ -536,26 +537,45 @@ for (subj.idx in 1:n.subj) {
   device.subj <- c()
   subjectID.subj <- c()
 
-  # Sample subjectwise parameters
-  subj.sd <- 0.05 # runif(1, 0.1, 0.2)
+  t <- seq(0, 100)
+
+  # Subjectwise wave parameters
+  # offset.mean <- runif(1, min = -0.5, max = 0.5)
+
+  a.sd <- runif(1, min = 0.05, max = 0.15) # .1
+  b.sd <- runif(1, min = 0.0001, max = 0.002) # .001
 
   for (stride.idx in 1:(n.strides)) {
 
-    # Error terms
-    w <- rnorm(n.frames, mean = 0, sd = subj.sd)
-    v <- rnorm(n.frames, mean = 0, sd = subj.sd)
+    a1.1 <- rnorm(1, mean = 3, sd = a.sd) # .1
+    a1.2 <- rnorm(1, mean = 3, sd = a.sd) # .1
+    a2.1 <- 0.08
+    a2.2 <- 0.08
+    b1.1 <- rnorm(1, mean = 0.06, sd = b.sd) # 0.001
+    b1.2 <- rnorm(1, mean = 0.06, sd = b.sd) # 0.001
+    b2.1 <- rnorm(1, mean = 0.58, sd = b.sd) # 0.001
+    b2.2 <- rnorm(1, mean = 0.58, sd = b.sd) # 0.001
+    c <- 2
 
-    t <- seq(0, 2*pi, 0.0625)
+    sine.1 <- a1.1 * sin(b1.1 * t) ^ (c + 3)
+    sine.2 <- a2.1 * sin(b2.1 * t)
+    sine.3 <- a1.2 * sin(b1.2 * t) ^ (c + 3)
+    sine.4 <- a2.2 * sin(b2.1 * t)
 
-    mc <- sin(t) + rnorm(n.frames, sd = subj.sd)
-    imu <- 2 * sin(t) + rnorm(n.frames, sd = subj.sd)
+    # offset <- rnorm(1, offset.mean, 0.05)
+
[…]

### 2021-08-23 paper.Rmd
Continue to write the paper draft

@@
 
-To compare different methods for constructing continuous LoA, smooth curve data were simulated across a range of common biomechanical signals and error characteristics.
+To compare different methods for constructing continuous LoA, several sets of smooth curve data were simulated across a range of common biomechanical signals and error characteristics.
+The quality of the curves is similar, as the paper is primarily developed for a validation context in which the same movement process is analyzed using similar methods (two measurement devices or models).
+For the same reason, the curves do not necessarily correspond to real-world signals as we were more concerned with illustrating the respective methods and error characterstics.
 
-The curves were modeled as Fourier series from additively superimposed sine functions.
+All curves were modeled as Fourier series from additively superimposed sine functions.
 Each data set contains n.strides = 10 curves from n.subj = 11 fictitious test subjects.
 Each curve consists of t = 101 data points.
-Note that the curves do not necessarily correspond to real-world signals as we were more concerned with illustrating the underlying principle.
 
@@
 
-* (3) Heteroscedastic error: Heteroscedasticity across the signal length.
+* (3) Heteroscedastic (non-gaussian, Weibull distributed) error: Heteroscedasticity across the signal length.
 The example consists of knee moment curves calculated using two different methods (direct and inverse kinematic approach) (Robinson et al., 2014).

### 2021-08-24 paper.Rmd
Add missing plots for shock data

@@
 
-#### Heteroscedasticity across the signal length (Robinson et al., 2014)
+#### Heteroscedasticity across the signal length
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "shock", dir.data)
 
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
+  geom_line(alpha = 0.7) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
+  labs(x = "Time-normalized signal [%]", y = "Difference") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 20),
+        axis.title.x = element_text(size = 22),
+        axis.text.y = element_text(size = 20),
+        axis.title.y = element_text(size = 22),
+        legend.position = "none")
+
+PLOT
 ```
@@
 data <- example_data(dat = "non_gaussian", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+
+data <- example_data(dat = "shock", dir.data)
 floa <- floa_rcb(data, n.boot, ver = "v2")
@@
 
+<!-- Limitations of curvewise intervals -->
 <!-- https://mjskay.github.io/ggdist/articles/lineribbon.html -->
+<!-- One challenge with curvewise intervals is that they can tend to be very conservative, especially at moderate-to-large intervals widths. -->
+<!-- Notice how noisy the curvewise intervals are.
+In addition, because a number of curves tend to start low and end high (or vice versa), above 50%, the bands rapidly expand to cover almost all of the curves in the sample, regardless of coverage level.
+-->
 <!-- In general I have found that there is no one method that consistently works well on all datasets.

### 2021-08-25 paper.Rmd

@@
 
+#### Curves phase shifted in x-axis direction
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "shift", dir.data)
+
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
+  geom_line(alpha = 0.7) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
+  labs(x = "Time-normalized signal [%]", y = "Difference") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 20),
+        axis.title.x = element_text(size = 22),
+        axis.text.y = element_text(size = 20),
+        axis.title.y = element_text(size = 22),
+        legend.position = "none")
+
+PLOT
+```
+
 ### Cross validation
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
+n.boot <- 400
+
 data <- example_data(dat = "smooth", dir.data)
-n.boot <- 400
 floa <- floa_rcb(data, n.boot, ver = "v2")
@@
 data <- example_data(dat = "shock", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+
+data <- example_data(dat = "shift", dir.data)
 floa <- floa_rcb(data, n.boot, ver = "v2")
@@
 
+#### Curves phase shifted in x-axis direction
+
+a. (coverage of the entire curve)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "shift", dir.data)
+
+cover.cross <- crossval_coverage(data, n.boot)
+# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+plot_cov_ver(cover.cross)
+
+# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
+```
+
+b. (mean percentage of covered curve points)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Display cross validation coverages (fraction of the curve that is covered) across methods
+cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+
+plot_cov_ver(cover.cross.fraction)
+```
+
 # Discussion

### 2021-08-27 paper.Rmd
Change windows to unix path in Rmd
Correctly implement coverage for single curve cross validation´

@@
 
-dir.script <- "C:/Users/Daniel/Desktop/tmp/floa/R"
-dir.data <- "C:/Users/Daniel/Desktop/tmp/floa/R/examples"
+dir.script <- "~/floa/R"
+dir.data <- "~/floa/R/examples"
 
@@
 source(paste0(dir.script, "/get_coverage.R"))
+source(paste0(dir.script, "/get_coverage_singlecurve.R"))
 source(paste0(dir.script, "/get_coverage_fraction.R"))
+source(paste0(dir.script, "/get_coverage_singlecurve_fraction.R"))
 source(paste0(dir.script, "/crossval_coverage.R"))
+source(paste0(dir.script, "/singlecurve_coverage.R"))
 source(paste0(dir.script, "/crossval_coverage_fraction.R"))
+source(paste0(dir.script, "/singlecurve_coverage_fraction.R"))
 source(paste0(dir.script, "/plot_cov_ver.R"))
@@
 <!-- Aus Robinson et al. (2021) -->
-<!-- Exemplar 1D eects from the biomechanics literature across a range of signal types were collated and reported (Bakke & Besier, 2020, Barrios & Willson, 2017, Bovi, Rabuetti, Mazzoleni, & Ferrarin, 2011, Phan et al., 2017, Gomes, Ackermann, Ferreira, Orselli, & Sacco, 2017, Robinson et al., 2014).
+<!-- Exemplar 1D effects from the biomechanics literature across a range of signal types were collated and reported (Bakke & Besier, 2020, Barrios & Willson, 2017, Bovi, Rabuetti, Mazzoleni, & Ferrarin, 2011, Phan et al., 2017, Gomes, Ackermann, Ferreira, Orselli, & Sacco, 2017, Robinson et al., 2014).
 -->
 
-<!-- the code to reproduce all 1D eects is provided at https://github.com/m-a-robinson/sample-size -->
+<!-- the code to reproduce all 1D effects is provided at https://github.com/m-a-robinson/sample-size -->
 
@@
 
-### Leave-one-out cross validation
+### Leave-one (curve) out cross validation
 
@@
 
-* 5. FLoA~Point~ : Pointwise B & A Limits of Agreement
+* 5. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)
 
@@
 
-* a: Percentage of entirely covered curves (all points within the limits): Each box consists of n~ = 11 values (one percentage for each subject).
+* a: Percentage of entirely covered curves (all points within the limits)
 
@@
 ```{r warning = FALSE, message = FALSE, eval = FALSE}
+get_coverage_singlecurve <- function (curve, floa) {
+  # Curve is a numerical vector
+  # floa is a "matrix" "array" containing "lower" and "upper" limits
+
+  # Calculate coverage (entire curves within the percentile boundaries)
+  # ----------------------------------------------------------------------------
   lwr.bnd <- floa["lower", ]
@@
 
-# ...
+  outside <- 0
 
-for (stride.idx in stride.indices){
-    curve <- subset(device.diff, strideID == stride.idx)
+  # Compare curves with upper and lower boundaries
+  below.thresh <- curve < lwr.bnd
+  above.thresh <- curve > upr.bnd
+
+  points.outside <- sum(above.thresh) + sum(below.thresh)
+
+  coverage <- ifelse(points.outside > 0, 0, 1)
+
+  return(coverage)
+}
+```
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth", dir.data)
+
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
+
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross <- crossval_coverage(data, n.boot)
+# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# 
+# plot_cov_ver(cover.cross)
+```
+
+b. (mean percentage of covered curve points)
+
+```{r warning = FALSE, message = FALSE, eval = FALSE}
+get_coverage_singlecurve_fraction <- function (curve, floa) {
+  # Curve is a numerical vector
+  # floa is a "matrix" "array" containing "lower" and "upper" limits
+
+  # Calculate coverage (curve points within the percentile boundaries)
+  # ----------------------------------------------------------------------------
+  lwr.bnd <- floa["lower", ]
+  upr.bnd <- floa["upper", ]
+
+  outside <- 0
 
@@
 
-    # Check if the entire curve is within the limits
-    if (points.outside > 0) {
-      outside <- outside + 1
-    }
-  }
+  # Divide by 101 to adjust for 101 data points
+  coverage <- 1 - (points.outside / 101)
+  coverage <- round(coverage, digits = 2)
 
-coverage <- 1 - (outside / n.strides)
-```
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth", dir.data)
-
-cover.cross <- crossval_coverage(data, n.boot)
-# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-
-plot_cov_ver(cover.cross)
-
-# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
-```
-
-b. (mean percentage of covered curve points)
-
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-lwr.bnd <- floa["lower", ]
-upr.bnd <- floa["upper", ]
-
-# ...
-
-for (stride.idx in stride.indices){
-
-  curve <- subset(device.diff, strideID == stride.idx)
-
-  # Compare difference curves with upper and lower boundaries
-  below.thresh <- curve$value < lwr.bnd
-  above.thresh <- curve$value > upr.bnd
-
-  # Mean percentage of points outside the limits
-  inside <- below.thresh == above.thresh
-  inside.thresh.perc <- c(inside.thresh.perc,
-                          mean(sum(inside[TRUE]) / length(curve$value)))
+  return(coverage)
 }
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Display cross validation coverages (mean fraction of the curve that is covered) across methods
-cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+plot_cov_ver(cover.cross.fraction.singlecurve)
 
-plot_cov_ver(cover.cross.fraction)
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+# plot_cov_ver(cover.cross.fraction)
 ```
@@
 
-cover.cross <- crossval_coverage(data, n.boot)
-# colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
 
-plot_cov_ver(cover.cross)
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
 
-# knitr::kable(cover.cross, caption = "Percentages (entire curve covered) for every leave one out fold")
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross <- crossval_coverage(data, n.boot)
+# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# 
+# plot_cov_ver(cover.cross)
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Display cross validation coverages (fraction of the curve that is covered) across methods
-cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+plot_cov_ver(cover.cross.fraction.singlecurve)
 
-plot_cov_ver(cover.cross.fraction)
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+# plot_cov_ver(cover.cross.fraction)
 ```

### 2021-08-27 R/get_coverage_singlecurve.R

@@ -1,0 +1,21 @@
+get_coverage_singlecurve <- function (curve, floa) {
+  # Curve is a numerical vector
+  # floa is a "matrix" "array" containing "lower" and "upper" limits
+
+  # Calculate coverage (entire curves within the percentile boundaries)
+  # ----------------------------------------------------------------------------
+  lwr.bnd <- floa["lower", ]
+  upr.bnd <- floa["upper", ]
+
+  outside <- 0
+
+  # Compare curves with upper and lower boundaries
+  below.thresh <- curve < lwr.bnd
+  above.thresh <- curve > upr.bnd
+
+  points.outside <- sum(above.thresh) + sum(below.thresh)
+
+  coverage <- ifelse(points.outside > 0, 0, 1)
+
+  return(coverage)
+}

### 2021-08-27 R/get_coverage_singlecurve_fraction.R

@@ -1,0 +1,23 @@
+get_coverage_singlecurve_fraction <- function (curve, floa) {
+  # Curve is a numerical vector
+  # floa is a "matrix" "array" containing "lower" and "upper" limits
+
+  # Calculate coverage (curve points within the percentile boundaries)
+  # ----------------------------------------------------------------------------
+  lwr.bnd <- floa["lower", ]
+  upr.bnd <- floa["upper", ]
+
+  outside <- 0
+
+  # Compare difference curves with upper and lower boundaries
+  below.thresh <- curve < lwr.bnd
+  above.thresh <- curve > upr.bnd
+
+  points.outside <- sum(above.thresh) + sum(below.thresh)
+
+  # Divide by 101 to adjust for 101 data points
+  coverage <- 1 - (points.outside / 101)
+  coverage <- round(coverage, digits = 2)
+
+  return(coverage)
+}

### 2021-08-27 R/main.R

@@ -43,6 +43,7 @@ dir.script <- "~/floa/R"
 dir.data <- "~/floa/R/examples"
 
 setwd(dir.script)
+
 source("example_data.R")
 source("fdaDelta.R")
 source("pick_subwise_curves.R")
@@ -56,8 +57,12 @@ source("floa_roislien.R")
 source("plot_loa.R")
 source("get_coverage.R")
 source("get_coverage_fraction.R")
+source("get_coverage_singlecurve.R")
+source("get_coverage_singlecurve_fraction.R")
 source("crossval_coverage.R")
 source("crossval_coverage_fraction.R")
+source("singlecurve_coverage.R")
+source("singlecurve_coverage_fraction.R")
 source("plot_cov_ver.R")
 
 
@@ -169,7 +174,7 @@ print(coverage)
 #   * Coverage levels [%] across n=length(subjectID) iterations
 # ------------------------------------------------------------------------------
 
-# 1. Leave-one (subject) out
+# 1. Leave-one subject out
 # ------------------------------------------------------------------------------
 cover.cross.subject <- crossval_coverage(data, n.boot)
 
@@ -181,8 +186,16 @@ cover.cross.fraction.subject <- crossval_coverage_fraction(data, n.boot)
 
 plot_cov_ver(cover.cross.fraction.subject)
 
-# 2. Leave-one (curve) out
+# 2. Leave-one curve out
 # ------------------------------------------------------------------------------
 cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+
 cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-plot_cov_ver(cover.cross.fraction.subject)
+plot_cov_ver(cover.cross.fraction.singlecurve)

### 2021-08-27 R/plot_cov_ver.R

@@ -16,5 +16,4 @@ plot_cov_ver <- function (cover.cross) {
     ylab("coverage")
 
   cover.PLOT
-
   }

### 2021-08-27 R/singlecurve_coverage.R

@@ -1,7 +1,7 @@
 singlecurve_coverage <- function (data, n.boot) {
 
   # ####################################################################
-  # Leave-one (curve) out method to estimate the uncertainty int the
+  # Leave-one (curve) out method to estimate the uncertainty in the
   # achieved coverage (see Lenhoff et al. (1999))
   # ####################################################################
   #
@@ -13,9 +13,7 @@ singlecurve_coverage <- function (data, n.boot) {
   # v3 : Fetch a SINGLE random stride from all strides
   # v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
   #      strap the resulting sample (of length (n=length(subjects))
-  #
-  # Output:
-  #   * Coverage levels [%] across n=length(subjectID) iterations
+  # v5 : Pointwise B&A limits (SD from linear mixed effects model)
   # ####################################################################
 
   n.curves <- unique(data$strideID)
@@ -30,7 +28,6 @@ singlecurve_coverage <- function (data, n.boot) {
 
     # Calculate FLoA with one curve left out
     # --------------------------------------------------------------------
-    # Leave curve at curve.idx out
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
@@ -39,36 +36,35 @@ singlecurve_coverage <- function (data, n.boot) {
     floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out, n.boot)
 
-    # Get coverage for the left out curve
-    # --------------------------------------------------------------------
+    # Get the difference curve
+    # ----------------------------------------------------------------------------
     data.subset <- subset(data, strideID == curve.idx)
 
-    cover.cross.rcb.v1[curve.idx] <- get_coverage(data.subset, floa.v1)
+    device1 <- data.frame(subset(data.subset, device == "IMU"))
+    device2 <- data.frame(subset(data.subset, device == "MC"))
 
-    # plot(floa.v1["mean", ], type = "l", col = "red", ylim = c(-1, 1))
+    device.diff <- device1$value - device2$value
+
+    # plot(floa.v1["mean", ], type = "l", col = "red", ylim = c(-5, 5))
     # lines(floa.v1["upper", ], col = "red")
[…]

### 2021-08-27 R/singlecurve_coverage_fraction.R

@@ -1,8 +1,8 @@
-crossval_coverage_fraction <- function (data, n.boot) {
+singlecurve_coverage_fraction <- function (data, n.boot) {
 
   # ####################################################################
-  # Leave-one (subject) out method to estimate the achieved coverage
-  # See e. g. Lenhoff et al. (1999)
+  # Leave-one (curve) out method to estimate the uncertainty in the
+  # achieved coverage (see Lenhoff et al. (1999))
   # ####################################################################
   #
   # Currently, different versions of the sampling process in draw_clusters()
@@ -13,9 +13,7 @@ crossval_coverage_fraction <- function (data, n.boot) {
   # v3 : Fetch a SINGLE random stride from all strides
   # v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
   #      strap the resulting sample (of length (n=length(subjects))
-  #
-  # Output:
-  #   * Coverage levels [%] across n=length(subjectID) iterations
+  # v5 : Pointwise B&A limits (SD from linear mixed effects model)
   # ####################################################################
 
   n.curves <- unique(data$strideID)
@@ -30,7 +28,6 @@ crossval_coverage_fraction <- function (data, n.boot) {
 
     # Calculate FLoA with one curve left out
     # --------------------------------------------------------------------
-    # Leave curve at curve.idx out
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
@@ -39,36 +36,35 @@ crossval_coverage_fraction <- function (data, n.boot) {
     floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out, n.boot)
 
-    # Get coverage for the left out curve
-    # --------------------------------------------------------------------
+    # Get the difference curve
+    # ----------------------------------------------------------------------------
     data.subset <- subset(data, strideID == curve.idx)
 
-    cover.cross.rcb.v1[curve.idx] <- get_coverage_fraction(data.subset, floa.v1)
+    device1 <- data.frame(subset(data.subset, device == "IMU"))
+    device2 <- data.frame(subset(data.subset, device == "MC"))
 
-    # plot(floa.v1["mean", ], type = "l", col = "red", ylim = c(-1, 1))
[…]

### 2021-09-08 paper.Rmd
Complete overhaul

@@
 ### TODO
+- Alternativtitel: Comparing different methods for constructing continuous prediction bands
 - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen
@@
 ### TODO read (again)
+- https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
 - Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis
@@
 library(ggplot2)
+library(boot)
+# library(fda)
+# library(funData)
 
@@
 
-source(paste0(dir.script, "/example_data.R"))
-source(paste0(dir.script, "/pick_subwise_curves.R"))
-source(paste0(dir.script, "/draw_clusters.R"))
-source(paste0(dir.script, "/functional_mean.R"))
-source(paste0(dir.script, "/functional_sd.R"))
-source(paste0(dir.script, "/boot_mean_sd.R"))
-source(paste0(dir.script, "/floa_rcb.R"))
-source(paste0(dir.script, "/floa_point.R"))
-source(paste0(dir.script, "/floa_roislien.R"))
-source(paste0(dir.script, "/plot_loa.R"))
-source(paste0(dir.script, "/get_coverage.R"))
-source(paste0(dir.script, "/get_coverage_singlecurve.R"))
-source(paste0(dir.script, "/get_coverage_fraction.R"))
-source(paste0(dir.script, "/get_coverage_singlecurve_fraction.R"))
-source(paste0(dir.script, "/crossval_coverage.R"))
-source(paste0(dir.script, "/singlecurve_coverage.R"))
-source(paste0(dir.script, "/crossval_coverage_fraction.R"))
-source(paste0(dir.script, "/singlecurve_coverage_fraction.R"))
-source(paste0(dir.script, "/plot_cov_ver.R"))
+setwd(dir.script)
+
+source("example_data.R")
+# source("fdaDelta.R")
+source("pick_subwise_curves.R")
+source("draw_clusters.R")
+source("floa_rcb.R")
+source("floa_point.R")
+source("floa_roislien.R")
+source("plot_loa.R")
+source("get_coverage.R")
+source("get_coverage_fraction.R")
+source("get_coverage_singlecurve.R")
+source("get_coverage_singlecurve_fraction.R")
+source("crossval_coverage.R")
+source("crossval_coverage_fraction.R")
+source("singlecurve_coverage.R")
+source("singlecurve_coverage_fraction.R")
+source("plot_cov_ver.R")
 ```
@@
 
+<!-- Aus Bland and Altman (2007): If each pair of X and Y measurements is treated as if from a different individual the structure of the data is ignored and incorrect estimates are  -->
+<!-- Ähnliche Formulierung aus https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x : In the repeated measures case, applying the standard limits of agreement to the data will result in limits that are too narrow because they do not take into account the reduction in variability that arises when working with averages of readings.
+-->
+
 # Methods
@@
 # * Data with shock peaks (no bias, no trend): "shock"
-
 data <- example_data(dat = "smooth", dir.data)

### 2021-09-08 R/boot_mean_sd.R

@@ -1,17 +1,0 @@
-boot_mean_sd <- function (data, n.boot) {
-
-  mean.boot <- vector(mode = "list", length = n.boot)
-  sd.boot <- vector(mode = "list", length = n.boot)
-
-  for (boot.idx in 1:n.boot) {
-
-    data.subset <- pick_subwise_curves(data)
-
-    mean.boot[[boot.idx]] <- functional_mean(data.subset)
-  }
-
-  func.mean.boot <- unlist(mean.boot)
-  func.mean.boot <- matrix(func.mean.boot, ncol = n.boot)
-
-  return(func.mean.boot)
-}

### 2021-09-08 R/draw_clusters.R

@@ -10,9 +10,6 @@ draw_clusters <- function(data, ver) { # fd.basis
   # v2   : One stride per subject
   # v3   : Fetch a single stride only from all available strides
   # ----------------------------------------------------------------------------
-
-  # source("pick_subwise_curves.R")
-
   if (ver == "v1") {
 
     # Strides are selected from the entire set of curves (NOT! one curve per subject)
@@ -79,5 +76,5 @@ draw_clusters <- function(data, ver) { # fd.basis
     diff.curves <- curve0 - curve1
   }
 
-  return(diff.curves) # diff.curve.samp # cluster.fdata
+  return(diff.curves)
 }

### 2021-09-08 R/example_data.R

@@ -34,6 +34,14 @@ example_data <- function(dat, dir.data) {
     data <- readRDS(paste0(dir.data, "/", "shift.rds"))
   }
 
+  # Convert continuous strideID to identical strideID's for repeated measures across subjects
+  # Needed for floa_point()
+  n.subj <- length(unique(data$subjectID))
+  n.devices <- length(unique(data$device))
+  strides.per.subject <- length(unique(data$strideID)) / length(unique(data$subjectID))
+  n.frames <- length(unique(data$frame))
+  data$strideID.rep <- rep(1:strides.per.subject, each = n.frames * n.devices, times = n.subj)
+
   return(data)
 
 }

### 2021-09-08 R/floa_point.R

@@ -3,45 +3,50 @@ floa_point <- function(data) {
   # ----------------------------------------------------------------------------
   #  Pointwise FLoA (FLoA_Point) calculated using linear mixed effects models
   # ----------------------------------------------------------------------------
-
+  # Implemetation inspired by:
+  # https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
   library(lme4)
 
   n.frames <- length(unique(data$frame))
 
   mean.diff <- vector(mode = "list", length = n.frames)
-  median.diff <- vector(mode = "list", length = n.frames)
-  sd.between <- vector(mode = "list", length = n.frames)
-  sd.within <- vector(mode = "list", length = n.frames)
-
+  total.sd <- vector(mode = "list", length = n.frames)
+  # median.diff <- vector(mode = "list", length = n.frames)
   for (frame.idx in 0:100){
-
     data.by.frame <- subset(data, frame == frame.idx)
 
     # Get within and between subject standard deviation from framewise linear
     # mixed effects models
-    LMEM <- lmer(value ~ device + (1|subjectID),
-                 data = data.by.frame,
-                 REML = FALSE) # method=REML to get unbiased estimates of the variance
+    LMEM <- lmer(value ~ device + (subjectID|strideID.rep), #  (1|subjectID) + (1|strideID.rep),
+                 data = data.by.frame)
+                 # REML = FALSE) # method=REML to get unbiased estimates of the variance
 
     mean.diff[[frame.idx + 1]] <- mean(data.by.frame$value[data.by.frame$device == "IMU"] - data.by.frame$value[data.by.frame$device == "MC"])
-    median.diff[[frame.idx + 1]] <- mean(data.by.frame$value[data.by.frame$device == "IMU"] - data.by.frame$value[data.by.frame$device == "MC"])
 
-    sd.between[[frame.idx + 1]] <- summary(LMEM)$coefficients[2, 2] # standard error of the difference MC - IMU
-    sd.within[[frame.idx + 1]] <- data.frame(summary(LMEM)$varcor)$sdcor[1]
+    # # Code from https://static-content.springer.com/esm/art%3A10.1186%2Fs12874-020-01022-x/MediaObjects/12874_2020_1022_MOESM1_ESM.pdf
+    # total.sd <- sqrt(as.numeric(summary(LMEM)$varcor[1]) +
+    #                  as.numeric(summary(LMEM)$varcor[2]) +
[…]

### 2021-09-08 R/floa_rcb.R

@@ -3,11 +3,9 @@ floa_rcb <- function(data, n.boot, ver) { # , fd.basis
   # ############################################################################
   # Randomized Cluster Bootstrap
   # ############################################################################
-
   clust.boot.agg <- c()
 
   for (boot.idx in 1:n.boot) {
-
     # draw_clusters returns difference curves (device1 - device2)
     #
     # Currently, different versions of the sampling process in draw_clusters() are
@@ -17,17 +15,13 @@ floa_rcb <- function(data, n.boot, ver) { # , fd.basis
 
   # (Row-wise) Arrange difference curves to facilitate computing percentiles
   clust.agg.intrp <- matrix(unlist(clust.boot.agg),
-                      ncol  = length(unique(data$frame)), # length(fd.basis$names)
-                      byrow = TRUE)
-
-  # clust.agg.intrp <- t(sapply(apply(t(clust.agg), 2, approx, n = 101), "[[", "y"))
+                            ncol  = length(unique(data$frame)), # length(fd.basis$names)
+                            byrow = TRUE)
 
   # Calculate percentiles ------------------------------------------------------
   floa.boot.percentiles <- c()
 
   for (i in 1:ncol(clust.agg.intrp)) {
-
-    # TODO: Bias correction useful/necessary?
     floa.boot.percentiles <- c(floa.boot.percentiles, quantile(clust.agg.intrp[, i], probs = c(0.025, 0.5, 0.975)))
   }
 
@@ -39,7 +33,7 @@ floa_rcb <- function(data, n.boot, ver) { # , fd.basis
   floa.boot.percentiles.split <- rbind(approx(perc2.5, n = 101)$y,
                                        approx(perc50, n = 101)$y,
                                        approx(perc97.5, n = 101)$y
-                                       )
+  )
 
   # Add (pointwise) mean
   floa.boot.mean <- colMeans(clust.agg.intrp, dims = 1)

### 2021-09-08 R/floa_roislien.R

@@ -1,33 +1,28 @@
-floa_roislien <- function(data, n.boot) {
+floa_roislien <- function(data) {
 
   # ****************************************************************************
   # Functional limits of agreement according to Roislien et al. (2012)
   #
   # In Roislien et al., FLoA are calculated as 95% CI.
   # ****************************************************************************
-
-  # Get one random stride from each subject ONCE and bootstrap the resulting
-  # sample (of length (n=length(subjects))
-  # ----------------------------------------------------------------------------
-
-  # source("pick_subwise_curves.R")
-  # source("functional_mean.R")
-  # source("functional_sd.R")
-  # source("boot_mean_sd.R")
-
-  # Bootstrap to get 95% CI FLoA -----------------------------------------------
-  func.boot <- boot_mean_sd(data, n.boot) # data need to have dimension [101, x]
+  diff.curves <- pick_subwise_curves(data)
 
   # Calculate Limits of Agreement ----------------------------------------------
-  func.mean.boot <- functional_mean(func.boot) # mean of bootstrapped distribution
-  func.sd.boot <- functional_sd(func.boot)
+  func.mean <- apply(diff.curves, 1, mean)
+  func.sd <- apply(diff.curves, 1, sd)
 
   # Get the same structure as returned by the other methods (i. e. floa_rcb)
-  floa.roislien <- rbind(func.mean.boot + 1.96 * func.sd.boot,
-                         func.mean.boot,
-                         func.mean.boot - 1.96 * func.sd.boot)
+  floa.roislien <- rbind(func.mean + 1.96 * func.sd,
+                         func.mean,
+                         func.mean - 1.96 * func.sd)
 
-  row.names(floa.roislien) <- c("upper", "mean", "lower")
+  rownames(floa.roislien) <- c("upper", "mean", "lower")
+
+  # # Plausibility check
+  # plot(diff.curves[, 1], type="l", ylim = c(-5, 5))
+  # apply(diff.curves, 2, lines, type = "l")
+  # lines(floa.roislien["upper", ], col = "red", size = 3)
+  # lines(floa.roislien["lower", ], col = "red", size = 3)
 
   return(floa.roislien)
 }

### 2021-09-08 R/functional_mean.R

@@ -1,13 +1,0 @@
-functional_mean <- function (data) {
-
-  # data need to have dimension [101, x]
-  n.frames <- dim(data)[1]
-
-  func.mean <- vector(mode = "numeric", length = n.frames)
-
-  for (i in 1:n.frames) {
-    func.mean[i] <- mean(data[i, ])
-  }
-
-  return(func.mean)
-}

### 2021-09-08 R/functional_sd.R

@@ -1,13 +1,0 @@
-functional_sd <- function (data) {
-
-  # data need to have dimension [101, x]
-  n.frames <- dim(data)[1]
-
-  func.sd <- vector(mode = "numeric", length = n.frames)
-
-  for (i in 1:n.frames) {
-    func.sd[i] <- sd(data[i, ])
-  }
-
-  return(func.sd)
-}

### 2021-09-08 R/main.R

@@ -1,4 +1,5 @@
 # Main script Functional Limits of Agreement (FLoA)
+# ------------------------------------------------------------------------------
 #
 # FLoA derived by different methods are compared
 #   * Randomized Cluster Bootstrap      (floa.rcb)
@@ -10,34 +11,33 @@
 # (see subsection data sets)
 #
 # TODO:
-#   + Leave-one-curve out implementieren (zusätzlich zu leave one subject out)
-#   + Beispieldaten Robinson et al. (2021) hinzufuegen
-#   + Rename n.strides to n.curves()
-#   + Zeitlichen Versatz (x-Achse) in example_data() einbringen (siehe Grafiken Lenhoff)
-#   + Check leave-one-out implementation
-#   + Programming challenge: Vectorize (or at least preallocate as much as possible)
-#   + Implementieren FDA
+#   + Pointwise LoA korrigieren. (ez)ANOVA Ansatz (B & A, 2007)
+#   + 95% coverage implemetieren
+#   + Reduce number of methods to 3 or 4
+#   + (Implementieren FDA)
 #   + Preallocation in floa_rcb()
-#   + Quantile: Über die gesamte Verteilung oder die "Ausreisser-Quantile" einzelner/extremer Probanden
-#     + quantile() function: Bias correction useful/necessary?
 #   + Implement balanced data in floa_rcb.R
 #
 # TODO READ:
+#   + B&A, 2007
+#   + https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
 #   + Ratkowsky - Handbook of nonlinear regression models
+#   + Robinson et al. (2021)
 #   + https://cran.r-project.org/web/packages/smooth/vignettes/simulate.html
 #
 # TODO STYLE:
-#   + Make variable names more universal (e.g. "device.1" instead "mc")
 #   + Add namespaces (package names ::)
+#   + Vectorize wherever possible
+#   + Make variable names more universal (e.g. "device.1" instead "mc")
+#   + Rename n.strides to n.curves()
 #   + No line between header and body in loops
 # ******************************************************************************
-
-# library(fda)
-# library(funData)
-
 rm(list = ls())
 
 library(ggplot2)
+library(boot)
+# library(fda)
+# library(funData)
 
 dir.script <- "~/floa/R"
 dir.data <- "~/floa/R/examples"
[…]

### 2021-09-08 R/pick_subwise_curves.R

@@ -2,12 +2,9 @@ pick_subwise_curves <- function (data) {
 
   n.subj <- length(unique(data$subjectID))
 
-  curve.idx <- vector(mode = "numeric", length = n.subj)
-
+  curve.idx <- c()
   for (subj.idx in 1:n.subj) {
-
     if (subj.idx %in% unique(data$subjectID)) {
-
       tmp <- data[data$subjectID == subj.idx, ]
 
       curve.idx.subj <- as.numeric(sample(tmp$strideID, size = 1))
@@ -16,10 +13,8 @@ pick_subwise_curves <- function (data) {
   }
 
   curve <- data[data$strideID %in% curve.idx, ]
-
   curve0 <- subset(curve, device  == "IMU")$value
   curve0 <- matrix(curve0, ncol = length(curve0) / 100)
-
   curve1 <- subset(curve, device  == "MC")$value
   curve1 <- matrix(curve1, ncol = length(curve1) / 100)
 

### 2021-09-08 R/plot_loa.R

@@ -1,17 +1,22 @@
-plot_loa <- function (data, method.loa, central.tendency) {
+plot_loa <- function (data, floa, central.tendency) {
+  # plot_loa() function arguments:
+  # 1. floa: Limits of agreement
+  # + a. floa (rcb method)
+  # + b. floa.point # requires: data.frame(t(floa.point))
+  # 2. central.tendency: Central tendency parameter
+  # + a. mean
+  # + b. median
 
   # floa variables need to be arranged in columns
   if (dim(floa)[1] < dim(floa)[2]) {
     floa <- t(floa)
   }
 
-  # floa need to be a data.frame
   if (!is.data.frame(floa)) {
     floa <- as.data.frame(floa)
   }
 
   # Prepare data for ggploting -------------------------------------------------
-
   device1 <- data.frame(subset(data, device == "IMU")$value)
   device2 <- data.frame(subset(data, device == "MC")$value)
 

### 2021-09-08 R/singlecurve_coverage.R

@@ -1,9 +1,8 @@
 singlecurve_coverage <- function (data, n.boot) {
-
   # ####################################################################
   # Leave-one (curve) out method to estimate the uncertainty in the
   # achieved coverage (see Lenhoff et al. (1999))
-  # ####################################################################
+  # --------------------------------------------------------------------
   #
   # Currently, different versions of the sampling process in draw_clusters()
   # (nested in floa_rcb()) are implemented:
@@ -11,11 +10,10 @@ singlecurve_coverage <- function (data, n.boot) {
   # v1 : n = length(subjects) random strides from all strides
   # v2 : One random stride per subject
   # v3 : Fetch a SINGLE random stride from all strides
-  # v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
-  #      strap the resulting sample (of length (n=length(subjects))
-  # v5 : Pointwise B&A limits (SD from linear mixed effects model)
+  # v4 : Roislien approach (Get one random stride from each subject ONCE
+  #      and bootstrap the resulting sample
+  # v5 : Pointwise B&A limits
   # ####################################################################
-
   n.curves <- unique(data$strideID)
 
   cover.cross.rcb.v1   <- vector(mode = "list", length = length(n.curves))
@@ -23,35 +21,71 @@ singlecurve_coverage <- function (data, n.boot) {
   cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
-
   for (curve.idx in n.curves) {
-
-    # Calculate FLoA with one curve left out
-    # --------------------------------------------------------------------
+    # Calculate FLoA with one curve left out -----------------------------
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
     floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
     floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
     floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
-    floa.roislien <- floa_roislien(data.one.out, n.boot)
+    floa.roislien <- floa_roislien(data.one.out)
 
[…]

### 2021-09-08 R/singlecurve_coverage_fraction.R

@@ -16,6 +16,14 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
   # v5 : Pointwise B&A limits (SD from linear mixed effects model)
   # ####################################################################
 
+  # Convert continuous strideID to identical strideID's for repeated measures across subjects
+  # Needed for floa_point()
+  n.subj <- length(unique(data$subjectID))
+  n.devices <- length(unique(data$device))
+  strides.per.subject <- length(unique(data$strideID)) / length(unique(data$subjectID))
+  n.frames <- length(unique(data$frame))
+  data$strideID.rep <- rep(1:strides.per.subject, each = n.frames * n.devices, times = n.subj)
+
   n.curves <- unique(data$strideID)
 
   cover.cross.rcb.v1   <- vector(mode = "list", length = length(n.curves))
@@ -23,32 +31,22 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
   cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
-
   for (curve.idx in n.curves) {
-
-    # Calculate FLoA with one curve left out
-    # --------------------------------------------------------------------
+    # Calculate FLoA with one curve left out -----------------------------
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
     floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
     floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
     floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
-    floa.roislien <- floa_roislien(data.one.out, n.boot)
+    floa.roislien <- floa_roislien(data.one.out)
 
-    # Get the difference curve
-    # ----------------------------------------------------------------------------
+    # Plot left out curve vs. various FLoA methods -----------------------
     data.subset <- subset(data, strideID == curve.idx)
 
     device1 <- data.frame(subset(data.subset, device == "IMU"))
     device2 <- data.frame(subset(data.subset, device == "MC"))
-
     device.diff <- device1$value - device2$value
-
-    # plot(floa.v1["mean", ], type = "l", col = "red", ylim = c(-5, 5))
-    # lines(floa.v1["upper", ], col = "red")
-    # lines(floa.v1["lower", ], col = "red")
[…]

### 2021-09-12 paper.Rmd
Edit uncertainty plots

@@
 ### TODO
-- Alternativtitel: Comparing different methods for constructing continuous prediction bands
-- Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen
+- Punktweise LoA wasserdicht machen (Parker et al., 2020)
+<!-- - Alternativtitel: Comparing different methods for constructing continuous prediction bands -->
+<!-- - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen -->
 - Weitere synthetische Daten siehe Paper Robinson et al. (2021)
@@
 ### TODO read (again)
-- https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
+- Parker et al. (2021) https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
 - Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis
@@
 source("plot_cov_ver.R")
+source("estimate_uncertainty_loa.R")
 ```
@@
 * Randomized Cluster Bootstrap (FLoA~RCB~)
-  + n = length(subjects) random strides from all strides (FLoA~RCB_v1~)
   + One random stride per subject (FLoA~RCB_v2~)
@@
 
+* Roislien-like intervals (FLoA~Roslien~)
+  + Select one random stride from each subject ONCE and bootstrap the resulting sample
+
+Pointwise continuous LoA are constructed using mean +- 1.96 * SD
+
 * Pointwise Gaussian intervals (FLoA~Point~)
-  + FLoA~Point~ are calculated using mean and standard deviations (derived from linear mixed effects models)
+  + FLoA~Point~ are calculated using using mean +- 1.96 * SD (standard deviations derived from linear mixed effects models)
 
-* Roislien-like intervals (FLoA~Roslien~)
-  + (Get one random stride from each subject ONCE and boot-
-       strap the resulting sample (of length (n=length(subjects))
+Linear mixed effects models (lme4 package (Bates, Maechler, Bolker, & Walker, 2015)) were used because of the advantages over fixed-effects models (e.
+g.
+ANOVA as suggested by Bland & Altman, 1999), such as e.
+g.
+(i) missing or unbalanced data are less problematic and (ii) inference can be made based on a wider population of patients (Parker et al., 2020).
+Implementation details are provided in the R code within the supplementary material.
 
@@
 
-#### Smooth curves with (missing) peaks
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "shock", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
-  geom_line(alpha = 0.7) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
-  labs(x = "Time-normalized signal [%]", y = "Difference") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 22),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 22),
-        legend.position = "none")
-
-PLOT
-```
-
 #### Curves phase shifted in x-axis direction
@@
 
-Leave-one out method to estimate the achieved coverage
+Leave-one (curve) out method to estimate the achieved coverage
 
-Prediction intervals are calculated from size.dataset = number.subjects - 1 and then validated against the left out subjects curves for five different methods:
-
-* FLoA~RCB_v1~   : n = length(subjects) random strides from all strides
+Prediction intervals are calculated from size.dataset = number.subjects - 1 and then validated against the left out curve for four different methods:
 
@@
 
-* FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
-       strap the resulting sample (of length (n=length(subjects))
+* FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and bootstrap the resulting sample)
 
 * FLoA~Point~ : Pointwise B & A Limits of Agreement
-
-<!-- $$ -->
-<!-- RMSE_{delta} = \sqrt{\frac{1}{N}(target_1 - pedal_1)^2 + (target_2 - pedal_2)^2 + ... + (target_N - pedal_N)^2} -->
-<!-- $$ -->
-
-<!-- $$ -->
-<!-- RMSE_{pedal} = \sqrt{\frac{1}{n}[(pedal_1 - \mu)^2 + (m_2 - \mu)^2 + ... + (pedal_N - \mu)^2]}, where \mu = \frac{1}{N}(pedal_1 + ... + pedal_N) -->
-<!-- $$ -->
-
-<!-- $$ -->
-<!-- Z = \{z_1, z_2, z_3,...,z_i\} -->
-<!-- $$ -->
 
@@
 
-data <- example_data(dat = "smooth", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
-
-data <- example_data(dat = "smooth_trend", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
-
-data <- example_data(dat = "non_gaussian", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
-
-data <- example_data(dat = "shock", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
-
-data <- example_data(dat = "shift", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+# data <- example_data(dat = "smooth", dir.data)
+# floa <- floa_rcb(data, n.boot, ver = "v2")
+# plot_loa(data, floa, central.tendency = "mean")
+# 
+# data <- example_data(dat = "smooth_trend", dir.data)
+# floa <- floa_rcb(data, n.boot, ver = "v2")
+# plot_loa(data, floa, central.tendency = "mean")
+# 
+# data <- example_data(dat = "non_gaussian", dir.data)
+# floa <- floa_rcb(data, n.boot, ver = "v2")
+# plot_loa(data, floa, central.tendency = "mean")
+# 
+# data <- example_data(dat = "shift", dir.data)
+# floa <- floa_rcb(data, n.boot, ver = "v2")
+# plot_loa(data, floa, central.tendency = "mean")
 ```
@@
 
-* 1. FLoA~RCB_v1~ : n = length(subjects) random strides from all strides
+* 1. FLoA~RCB_v2~ : One random stride per subject
 
-* 2. FLoA~RCB_v2~ : One random stride per subject
+* 2. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides
 
-* 3. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides
-
-* 4. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
+* 3. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
        strap the resulting sample (of length (n=length(subjects))
 
-* 5. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)
+* 4. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)
 
@@
 
-* b: Mean percentage of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
+* b: Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
+
+* c: Uncertainty in LoA (95% percentile intervals)
 
@@
 
-a. (coverage of the entire curve)
-
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-get_coverage_singlecurve <- function (curve, floa) {
-  # Curve is a numerical vector
-  # floa is a "matrix" "array" containing "lower" and "upper" limits
-
-  # Calculate coverage (entire curves within the percentile boundaries)
-  # ----------------------------------------------------------------------------
-  lwr.bnd <- floa["lower", ]
-  upr.bnd <- floa["upper", ]
-
-  outside <- 0
-
-  # Compare curves with upper and lower boundaries
-  below.thresh <- curve < lwr.bnd
-  above.thresh <- curve > upr.bnd
-
-  points.outside <- sum(above.thresh) + sum(below.thresh)
-
-  coverage <- ifelse(points.outside > 0, 0, 1)
-
-  return(coverage)
-}
-```
+a. Coverage of the entire curve
 
@@
 covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
 
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
 
@@
 
-b. (mean percentage of covered curve points)
-
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-get_coverage_singlecurve_fraction <- function (curve, floa) {
-  # Curve is a numerical vector
-  # floa is a "matrix" "array" containing "lower" and "upper" limits
-
-  # Calculate coverage (curve points within the percentile boundaries)
-  # ----------------------------------------------------------------------------
-  lwr.bnd <- floa["lower", ]
-  upr.bnd <- floa["upper", ]
-
-  outside <- 0
-
-  # Compare difference curves with upper and lower boundaries
-  below.thresh <- curve$value < lwr.bnd
-  above.thresh <- curve$value > upr.bnd
-
-  points.outside <- sum(above.thresh) + sum(below.thresh)
-
-  # Divide by 101 to adjust for 101 data points
-  coverage <- 1 - (points.outside / 101)
-  coverage <- round(coverage, digits = 2)
-
-  return(coverage)
-}
-```
+b. Mean (SD) of covered curve points
 
@@
 cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-plot_cov_ver(cover.cross.fraction.singlecurve)
 
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
-# plot_cov_ver(cover.cross.fraction)
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+
+# # Display as boxplots across methods
+# plot_cov_ver(cover.cross.fraction.singlecurve)
+```
+
+c. Uncertainty in LoA (95% percentile intervals)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+estimate_uncertainty_loa(data, n.boot)
 ```
@@
 
-a. (coverage of the entire curve)
+a. Coverage of the entire curve
 
@@
 covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
 
 knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
 ```
 
-b. (mean percentage of covered curve points)
+b. Mean (SD) of covered curve points
 
@@
 cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-plot_cov_ver(cover.cross.fraction.singlecurve)
 
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
-# plot_cov_ver(cover.cross.fraction)
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+```
+
+c. Uncertainty in LoA (95% percentile intervals)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+estimate_uncertainty_loa(data, n.boot)
 ```
@@
 covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
 
@@
 
-b. (mean percentage of covered curve points)
+b. Mean (SD) of covered curve points
 
@@
 cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-plot_cov_ver(cover.cross.fraction.singlecurve)
+
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points")
+
+# # Display as boxplots acroos methods
+# plot_cov_ver(cover.cross.fraction.singlecurve)
 
@@
 
-#### Curves with (missing) peaks
-
-a. (coverage of the entire curve)
+c. Uncertainty in LoA (95% percentile intervals)
 
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "shock", dir.data)
-
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
-```
-
-b. (mean percentage of covered curve points)
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-plot_cov_ver(cover.cross.fraction.singlecurve)
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
-# plot_cov_ver(cover.cross.fraction)
+estimate_uncertainty_loa(data, n.boot)
 ```
@@
 
+c. Uncertainty in LoA (95% percentile intervals)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+estimate_uncertainty_loa(data, n.boot)
+```
+
 # Discussion
+
+Punktweise FLoA unterrepräsentieren den Messfehler
+
+FLoA Roislien tendieren dazu zu weit auszufallen (den Messfehler zu überrepräsentieren)
 

### 2021-09-12 R/estimate_uncertainty_loa.R

@@ -1,0 +1,112 @@
+# Estimate uncertainty in different LoA approaches (95% percentile intervals for the LoA)
+estimate_uncertainty_loa <- function (data, n.boot) {
+
+  floa.v2.upper <- vector(mode = "list", length = n.boot)
+  floa.v2.lower <- vector(mode = "list", length = n.boot)
+  floa.roislien.upper <- vector(mode = "list", length = n.boot)
+  floa.roislien.lower <- vector(mode = "list", length = n.boot)
+  floa.point.upper <- vector(mode = "list", length = n.boot)
+  floa.point.lower <- vector(mode = "list", length = n.boot)
+  for (i in 1:n.boot) {
+    floa.point    <- floa_point(data)
+    floa.v1       <- floa_rcb(data, n.boot, ver = "v1")
+    floa.v2       <- floa_rcb(data, n.boot, ver = "v2")
+    floa.v3       <- floa_rcb(data, n.boot, ver = "v3")
+    floa.roislien <- floa_roislien(data)
+
+    floa.v2.upper[[i]] <- floa.v2["upper", ]
+    floa.v2.lower[[i]] <- floa.v2["lower", ]
+    floa.roislien.upper[[i]] <- floa.roislien["upper", ]
+    floa.roislien.lower[[i]] <- floa.roislien["lower", ]
+    # floa.point.upper[[i]] <- floa.roislien["upper", ]
+    # floa.point.lower[[i]] <- floa.roislien["lower", ]
+  }
+
+  floa.v2.upper.unlist <- matrix(unlist(floa.v2.upper), 101)
+  floa.v2.lower.unlist <- matrix(unlist(floa.v2.lower), 101)
+  floa.roislien.upper.unlist <- matrix(unlist(floa.roislien.upper), 101)
+  floa.roislien.lower.unlist <- matrix(unlist(floa.roislien.lower), 101)
+  # floa.point.upper.unlist <- matrix(unlist(floa.point.upper), 101)
+  # floa.point.lower.unlist <- matrix(unlist(floa.point.lower), 101)
+
+  # Pointwise 95% percentile intervals for the LoA
+  # floa.v1.pi <- quantile(floa.v1.upper.unlist, probs = c(0.025, 0.975))
+  # floa.roislien.pi <- quantile(floa.roislien.upper.unlist, probs = c(0.025, 0.975))
+  floa.v2.upper.pi <- apply(floa.v2.upper.unlist, 1, quantile, probs = c(0.025, 0.975))
[…]

### 2021-09-16 paper.Rmd
Modify plots

@@
 
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + # , colour = subjectID
-  geom_line(alpha = 0.7) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, col = "red"), alpha = 0.3) + #, colour = subjectID
-  labs(x = "Time-normalized signal [%]", y = "Difference") +
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Signal value") +
   theme_minimal() +
   theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 22),
+        axis.title.x = element_text(size = 18),
         axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 22),
+        axis.title.y = element_text(size = 18),
         legend.position = "none")

### 2021-09-20 paper.Rmd
Implement new data sets

@@
 
-### Synthetic data sets
-
+### Data generating process
 To compare different methods for constructing continuous LoA, several sets of smooth curve data were simulated across a range of common biomechanical signals and error characteristics.
@@
 
+### Realistic smooth wave data (constant variance, no trend)
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_realistic", dir.data)
+
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Signal value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 20),
+        axis.title.x = element_text(size = 18),
+        axis.text.y = element_text(size = 20),
+        axis.title.y = element_text(size = 18),
+        legend.position = "none")
+
+PLOT
+```
+
 ### Smooth curves (homoscedasticity, normal error, no trend)
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-### Smooth curves with nonlinear error trend
+<!-- ### Smooth curves with nonlinear error trend -->
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- data <- example_data(dat = "smooth_trend", dir.data) -->
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_trend", dir.data)
+<!-- data.single.mc <- subset(data, device == "MC") -->
+<!-- data.single.imu <- subset(data, device == "IMU") -->
 
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
+<!-- PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) + -->
+<!--   geom_line() + -->
+<!--   scale_color_grey(start = 0.6, end = 0.9) + -->
+<!--   geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) + -->
+<!--   labs(x = "Time-normalized signal [%]", y = "Signal value") + -->
+<!--   theme_minimal() + -->
+<!--   theme(axis.text.x = element_text(size = 20), -->
+<!--         axis.title.x = element_text(size = 18), -->
+<!--         axis.text.y = element_text(size = 20), -->
+<!--         axis.title.y = element_text(size = 18), -->
+<!--         legend.position = "none") -->
 
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
+<!-- PLOT -->
+<!-- ``` -->
 
 #### Heteroscedasticity across the signal length
-
 <!-- Smooth curves with non-gaussian (Weibull) error (no trend) -->
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 #### Curves phase shifted in x-axis direction
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-### Cross validation
+#### Real-world validation data
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "imu_mc", dir.data)
 
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Signal value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 20),
+        axis.title.x = element_text(size = 18),
+        axis.text.y = element_text(size = 20),
+        axis.title.y = element_text(size = 18),
+        legend.position = "none")
+
+PLOT
+```
+
+### Leave-one (curve) out cross validation
 Leave-one (curve) out method to estimate the achieved coverage
@@
 
+#### Coverages
+Coverages are displayed for two different scenarios:
+
+* a: Percentage of entirely covered curves (all points within the limits)
+
+* b: Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
+
+### LoA uncertainty
+In order to quantify the uncertainty of the calculated LoA across different methods, LoA were calculated n.boot times and compared using uncertainty bands.
+
 # Results
@@
 ### Functional limits of agreement
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-n.boot <- 400
+n.boot <- 1000
 
@@
 ### Leave-one (curve) out cross validation
+Displayed are boxes for different calculation methods (FLoA~RCB_v2~, FLoA~RCB_v3~, FLoA~Roislien~, FLoA~Point~)
 
-Displayed are boxes for five different methods:
+#### Realistic smooth wave data (constant variance, no trend)
+a. Coverage of the entire curve
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_realistic", dir.data)
 
-* 1. FLoA~RCB_v2~ : One random stride per subject
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
 
-* 2. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
 
-* 3. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
-       strap the resulting sample (of length (n=length(subjects))
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
 
-* 4. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)
+knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
 
-Coverages are displayed for two different scenarios:
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross <- crossval_coverage(data, n.boot)
+# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# 
+# plot_cov_ver(cover.cross)
+```
 
-* a: Percentage of entirely covered curves (all points within the limits)
+b. Mean (SD) of covered curve points
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
 
-* b: Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
 
-* c: Uncertainty in LoA (95% percentile intervals)
+# # Display as boxplots across methods
+# plot_cov_ver(cover.cross.fraction.singlecurve)
+```
+
+c. Uncertainty in LoA (95% percentile intervals)
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+estimate_uncertainty_loa(data, n.boot)
+```
 
 #### Smooth, wave data (normal error, constant variance, no trend)
-
 a. Coverage of the entire curve
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 b. Mean (SD) of covered curve points
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 c. Uncertainty in LoA (95% percentile intervals)
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-#### Curves with nonlinear trend (constant variance)
+<!-- #### Curves with nonlinear trend (constant variance) -->
 
-a. Coverage of the entire curve
+<!-- a. Coverage of the entire curve -->
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_trend", dir.data)
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- data <- example_data(dat = "smooth_trend", dir.data) -->
 
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+<!-- # Leave-one curve out approach -->
+<!-- # ------------------------------------------------------------------------------ -->
+<!-- cover.cross.singlecurve <- singlecurve_coverage(data, n.boot) -->
 
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
+<!-- # Calculate percent from counts -->
+<!-- calculate_percent <- function(data) { -->
+<!--   percentage <- (sum(data) / length(data)) -->
+<!--   percentage <- round(percentage, digits = 2) -->
+<!-- } -->
 
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
+<!-- covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent) -->
+<!-- covered.curves.percent <- data.frame(t(covered.curves.percent)) -->
+<!-- colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5") -->
 
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-```
+<!-- knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves") -->
+<!-- ``` -->
 
-b. Mean (SD) of covered curve points
+<!-- b. Mean (SD) of covered curve points -->
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- # Leave-one curve out approach -->
+<!-- # ------------------------------------------------------------------------------ -->
+<!-- cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot) -->
 
-cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-)
-knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-```
+<!-- cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2), -->
+<!--              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2) -->
+<!-- ) -->
+<!-- knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve") -->
+<!-- ``` -->
 
-c. Uncertainty in LoA (95% percentile intervals)
+<!-- c. Uncertainty in LoA (95% percentile intervals) -->
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-estimate_uncertainty_loa(data, n.boot)
-```
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- estimate_uncertainty_loa(data, n.boot) -->
+<!-- ``` -->
 
@@
 #### Curves phase shifted in x-axis direction
-
 a. Coverage of the entire curve
@@
 data <- example_data(dat = "shift", dir.data)
+
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
+
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross <- crossval_coverage(data, n.boot)
+# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# 
+# plot_cov_ver(cover.cross)
+```
+
+b. Mean (SD) of covered curve points
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points")
+
+# # Display as boxplots acroos methods
+# plot_cov_ver(cover.cross.fraction.singlecurve)
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+# plot_cov_ver(cover.cross.fraction)
+```
+
+c. Uncertainty in LoA (95% percentile intervals)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+estimate_uncertainty_loa(data, n.boot)
+```
+
+#### Real-world validation data
+a. Coverage of the entire curve
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "imu_mc", dir.data)
 

### 2021-09-20 R/example_data.R

@@ -5,13 +5,17 @@ example_data <- function(dat, dir.data) {
   if (dat == "imu_mc") {
 
     # IMU vs. MC validation data ---------------------------------------------------
-    data <- readRDS(paste0(dir.data, "/", "data.rds"))
+    data <- readRDS(paste0(dir.data, "/", "imu_mc.rds")) # data.rds contains the unbalanced data
 
     # # Intermediate step: Fit functions to discrete time series data ----------------
     # # Thus far, a transformation of time series data to functional data (Fourier,
     # # Splines etc.) is not implemented.
     # fd.basis <- create.fourier.basis(nbasis=50) # Plateau around 50 basis vectors
     # fda.delta <- fdaDelta(data, fd.basis) # Fit Fourier, returns delta curves (mean, sd)
+
+  } else if (dat == "smooth_realistic") {
+
+    data <- readRDS(paste0(dir.data, "/", "smooth_realistic.rds"))
 
   } else if (dat == "smooth") {
 
@@ -20,6 +24,10 @@ example_data <- function(dat, dir.data) {
   } else if (dat == "smooth_trend") {
 
     data <- readRDS(paste0(dir.data, "/", "smooth_trend.rds"))
+
+  } else if (dat == "smooth_biased") {
+
+    data <- readRDS(paste0(dir.data, "/", "smooth_biased.rds"))
 
   } else if (dat == "non_gaussian") {
 

### 2021-09-20 R/main.R

@@ -68,13 +68,15 @@ source("estimate_uncertainty_loa.R")
 # ********************************* Data sets **********************************
 # Wrapper function for example data sets. Function arguments:
 #
-# (* Empirical validation data: "imu_mc")
-# * Smooth, wave data (normal error, constant variance, no trend): "smooth"
+# * Real world empirical validation data: "imu_mc"
+# * Realistic looking, smooth, wave data (normal error, no trend): "smooth_realistic"
+# * Smooth, wave data (normal error, no trend): "smooth"
 # * Smooth wave data with nonlinear trend (constant variance): "smooth_trend"
+# * Heteroscedastic errors (no trend): "heteroscedastic"
 # * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
 # * Data with shock peaks (no bias, no trend): "shock"
 # * Phase shifted data (x-axis direction): "shift"
-data <- example_data(dat = "smooth", dir.data)
+data <- example_data(dat = "imu_mc", dir.data)
 
 # Plot data --------------------------------------------------------------------
 # uncommment when subject differences need to be plotted

### 2021-09-20 R/simulate_data.R

@@ -8,6 +8,7 @@
 #
 # Simulated data sets:
 #
+# 0. Real-world validation data: "imu_mc"
 # 1. Smooth, wave data (normal error, constant variance, no trend): "smooth"
 # 2. Non-stationary data (trend, no bias) data:"non_stationary"
 # 3. Non-gaussian (Weibull distributed) error (no trend)
@@ -15,46 +16,128 @@
 # 5. Shift data (data phase shifted (x-axis direction))
 ################################################################################
 
+# Real-world validation data: "imu_mc" -----------------------------------------
+# Prepare (balance) data for the main script
+imu.mc <- readRDS(paste0(dir.data, "/", "data.rds"))
+
+# Pick 10 random strides from each of the 11 subjects
+stride.idx <- imu.mc %>% dplyr:::group_by(subjectID)  %>% dplyr::summarize(sample_n(data.frame(strideID), 10, replace = FALSE)) # sample(strideID, size = 10, replace = FALSE)
+stride.idx <- stride.idx$strideID
+stride.idx <- as.numeric(stride.idx)
+
+double.idx <- which(duplicated(stride.idx))
+stride.idx[double.idx] <- double.idx + 1
+imu.mc.balanced <- imu.mc[imu.mc$strideID %in% stride.idx, ]
+
+if (sum(duplicated(stride.idx))) {
+  saveRDS(imu.mc.balanced, file = paste0("~/floa/R/examples/", "imu_mc.rds"))
+}
+
+
+
 n.subj <- 11
 n.strides <- 10
 n.devices <- 2
 n.frames <- 101
 
-# Smooth wave data (constant variance, no trend) -------------------------------
+# Realistic smooth wave data (constant variance, no trend) ---------------------
 device <- c()
 value <- c()
 subjectID <- c()
 
 for (subj.idx in 1:n.subj) {
-
   value.subj <- c()
   device.subj <- c()
   subjectID.subj <- c()
 
   t <- seq(0, 100)
 
-  # Subjectwise wave parameters
+  # Subjectwise parameters
[…]

### 2021-09-26 paper.Rmd
Implement new data sets and Lenhoff method

--- 2021-09-16 paper.Rmd
@@
 ---
-title: "Asymmetric continuous prediction bands for assessing the validity of smooth biomechanical curve data"
+title: "Analysis of different methods for constructing continuous limits of agreement from smooth biomechanical data"
 author:
@@
 ### TODO
-- Punktweise LoA wasserdicht machen (Parker et al., 2020)
-<!-- - Alternativtitel: Comparing different methods for constructing continuous prediction bands -->
 <!-- - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen -->
@@
 source("plot_cov_ver.R")
+source("distance_2_floa.R")
 source("estimate_uncertainty_loa.R")
@@
 
+<!-- # Alternatives Intro -->
+
+<!-- SHORT NOTE!? -->
+
+<!-- Ansatz: Neben der Validierung von Messgeräten spielt auch die Validierung von Modellergebnissen eine Rolle! -->
+
+<!-- 1.
+Es gibt einen Unterschied zwischen Konfidenz- und Prädiktionsintervallen: Prediction intervals are a prediction of a future observation, while confidence intervals represent a predicted mean response.
+-->
+
+<!-- <!-- https://stats.stackexchange.com/questions/16493/difference-between-confidence-intervals-and-prediction-intervals --> -->
+<!-- The difference between a prediction interval and a confidence interval is the standard error. -->
+<!-- The standard error for a confidence interval on the mean takes into account the uncertainty due to sampling.
+The line you computed from your sample will be different from the line that would have been computed if you had the entire population, the standard error takes this uncertainty into account.
+-->
+<!-- The standard error for a prediction interval on an individual observation takes into account the uncertainty due to sampling like above, but also takes into account the variability of the individuals around the predicted mean.
+The standard error for the prediction interval will be wider than for the confidence interval and hence the prediction interval will be wider than the confidence interval.
+-->
+
+<!-- Für die Validierung von Messinstrumenten sind aus unserer Sicht eher Prädiktionsintervalle interessant! Daher beziehen sich alle folgenden Aussagen ausschließlich auf Prädiktionsintervalle. -->
+
+<!-- 2.
+Punktweise Prädiktionsbänder (und Konfidenzbänder) sind gegenüber simultatnen zu eng (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
+Daher wurden in der Vergangenheit Korrekturverfahren (e.
+.g.
+Lenhoff, Bonferroni, Šidák, projection, sup-t bands) vorgeschlagen -->
+
+<!-- <!-- See Mirzargar et al. (2014) or Juul et al. (2020) for an accessible introduction to data depth and curve boxplots / functional boxplots. --> -->
+
 # Introduction
 
-<!-- Having a good estimate of the amount of measurement error is a basic requirement when analyzing movement data.
+Having a good estimate of the amount of measurement error is a basic requirement when analyzing movement data.
 The classic linear error model describes measurement error as the sum of systematic and random deviations from the true value.
@@
 joint angles, ground reaction forces, or center of mass trajectories - is available in the form of continuous signals (curve data) within which the error can vary over time.
--->
-
-Hier ein anderes Intro schreiben ... Ansatz: Neben der Validierung von Messgeräten spielt auch die Validierung von Modellergebnissen eine Rolle!
 
@@
 
-<!-- ... and likely results in a systematic underrepresentation of ... values (Juul et al., 2020; Mirzagar et al. (2014). -->
+Es gibt verschiedene Ansätze dieses Problem zu adressieren.
+Was bislang fehlt ist eine systematische Validierung und Gegenüberstellung verschiedener Methoden.
+Im Paper sollen daher wichtige Methoden zur Konstruktion kontinuierlicher Übereinstimmnugsbänder anhand repräsentativer Datensätze (synthetisch und real-world) verglichen werden und deren Vor- und Nachteile gegenübergestellt werden.
+<!-- Zudem sollen relevante methodische Aspekte diskutiert (z. B. Prädiktions- vs. Konfidenzbänder für die Validierung von Messsystemen) -->
 
-<!-- https://mjskay.github.io/ggdist/reference/curve_interval.html -->
-<!-- Hier referenzierte Paper u. a.  -->
-<!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
+Bei allen etablierten Vorzuügen funktionaler/kontinuerlicher Verfahren ist aktuell zu konstatieren, dass diese Methoden in der angewandten Forschung im Bereich Biomechanik/Bewegungswissenschafte kaum Anwendung finden.
+Wir spekulieren, dass für die praktische Umsetzung Fähigkeiten in den Bereichen Statistik und Programmierung erforderlich sind, die bei vielen "Praktikern" nicht ausreichend vorhanden sind.
+Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die im Paper analysierten Methoden als frei verfügbaren R Code zur Verfügung.
+
+<!-- <!-- ... and likely results in a systematic underrepresentation of ... values (Juul et al., 2020; Mirzagar et al. (2014). --> -->
+
+<!-- <!-- https://mjskay.github.io/ggdist/reference/curve_interval.html --> -->
+<!-- <!-- Hier referenzierte Paper u. a.  --> -->
+<!-- <!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
 (2020).
@@
 arXiv e-print.
-arXiv:2007.05035 -->
-<!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
+arXiv:2007.05035 --> -->
+<!-- <!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
 (2014).
@@
 20(12): 2654-2663.
-doi: 10.1109/TVCG.2014.2346455 -->
+doi: 10.1109/TVCG.2014.2346455 --> -->
 
-From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
+<!-- From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
 In addition, most traditional point statistics are either not suitable (e.g.
@@
 product-moment correlation) to quantify the agreement between two measurement systems (Bland and Altman, 1986).
+-->
 
-One possibility to address the problem of pointwise analysis is to treat entire curves as (functional) objects rather than as a series of unconnected points.
+<!-- One possibility to address the problem of pointwise analysis is to treat entire curves as (functional) objects rather than as a series of unconnected points. -->
 
-<!-- Es gibt eine Reihe von Ansätzen aus anderen Bereichen (z. B. Covid Prädiktion), bei denen sog. Function oder Curve Boxplots eingesetzt werden. -->
-<!-- https://mjskay.github.io/ggdist/reference/curve_interval.html -->
-<!-- Hier referenzierte Paper u. a.  -->
-<!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
+<!-- <!-- Es gibt eine Reihe von Ansätzen aus anderen Bereichen (z. B. Covid Prädiktion), bei denen sog. Function oder Curve Boxplots eingesetzt werden. --> -->
+<!-- <!-- https://mjskay.github.io/ggdist/reference/curve_interval.html --> -->
+<!-- <!-- Hier referenzierte Paper u. a.  --> -->
+<!-- <!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
 (2020).
@@
 20(12): 2654-2663.
-doi: 10.1109/TVCG.2014.2346455 -->
+doi: 10.1109/TVCG.2014.2346455 --> -->
 
-Sutherland et al.
+<!-- Sutherland et al.
 (1988) introduced a method in which continuous signals are reconstructed using Fourier series and bootstrapped to estimate of the variation of the sample mean and thus estimate confidence intervals.
@@
 bootstrap bands provide appropriate coverage for continuous curve gait data (86% coverage for a targeted coverage of 90%) while pointwise Gaussian bands are shown to provide inadequate coverage (54% for a targeted coverage of 90%).
+-->
 
-Røislien et al.
+<!-- <!-- In other words, a single 0.95-CI for each one of the quantities of interest is computed and, in this sense, they are called pointwise CIs.
+However, if the target is the estimation of a vector of values (such as the CDF or the return level function evaluated in a grid of points, for instance), more accurate than computing the corresponding pointwise CIs for each one of the quantities of interest is to construct simultaneous CIs (also called a confidence band).
+Simultaneous CIs are a group of CIs (a CI for each one of the quantities of interest) designed to jointly contain this vector of unknown values with a prescribed high probability (95%, for instance).
+In this way, it would be possible to make a joint probabilistic assessment for some quantiles of the corresponding CDF.
+Because simultaneous CIs include information on model reliability, they are are much more informative than pointwise CIs [14].
+--> -->
+
+<!-- Røislien et al.
 (2012) presented a similar approach to calculate a functional version of the the Limits of Agreement (LoA) method suggested by Bland & Altman.
@@
 actual degrees).
-Røislien et al., however, rely on a number of strong model assumptions, which are often not or only partially fulfilled in empirical data sets:
+Røislien et al., however, rely on a number of strong model assumptions, which are often not or only partially fulfilled in empirical data sets: -->
 
-(i) The error distribution is approximately normal.
+<!-- (i) The error distribution is approximately normal. -->
 
-The construction of symmetric prediction bands as described in Røislien et al.
+<!-- The construction of symmetric prediction bands as described in Røislien et al.
 [10] relies on Gaussian assumptions.
@@
 sensor values are truncated because of limited range or resolution, or because values drift over time.
+-->
 
-(ii) Time-continuous differences of two measurement systems and the corresponding averages must be uncorrelated.
+<!-- (ii) Time-continuous differences of two measurement systems and the corresponding averages must be uncorrelated. -->
 
-This assumption is violated considerably in many validation data sets [12].
+<!-- This assumption is violated considerably in many validation data sets [12].
 E.g., Atallah et al.
@@
 When validating a foot-worn gyroscope for treadmill running, Koska & Maiwald [14] observed a linear error trend that was associated with foot strike behavior.
+-->
 
-(iii) Aus einem Datensatz bestehend aus einer Kurve eines Probanden lassen sich durch Bootstrapping von mittleren Kurven Unsicherheitsintervalle bilden, die ein geeignetes Maß für die estimation uncertainty darstellen.
+<!-- (iii) Aus einem Datensatz bestehend aus einer Kurve eines Probanden lassen sich durch Bootstrapping von mittleren Kurven Unsicherheitsintervalle bilden, die ein geeignetes Maß für die estimation uncertainty darstellen.
+-->
 
-Hier werden arithmetische mean curves verwendet, the variability of which will reflect variation between means.
+<!-- Hier werden arithmetische mean curves verwendet, the variability of which will reflect variation between means.
 In many cases, however, it is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of the number of curves within the sample.
+-->
 
-Ein weiteres Problem ist, dass die Methode keine messwiederholten Daten berücksichtigt.
+<!-- Ein weiteres Problem ist, dass die Methode keine messwiederholten Daten berücksichtigt.
 Da sich messwiederholte Kurven i.
@@
 ähnlicher sind als unabhängige Kurven, besteht die Gefahr die Streuung des Messsystems zu überschätzen.
+-->
 
-<!-- Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
+<!-- <!-- Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
 werden 95% Konfidenzintervalle gebootstrapped.
@@
 It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of n.subj.
--->
+--> -->
 
-<!-- Formulierung aus dem alten Skript -->
-<!-- (iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject. -->
-<!-- If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
+<!-- <!-- Formulierung aus dem alten Skript --> -->
+<!-- <!-- (iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject. --> -->
+<!-- <!-- If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
 The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
--->
+--> -->
 
-<!-- Weiterer ("zugehöriger) Reviewerkommentar: -->
-<!-- In looking at limits of agreement, the authors are considering the extent of within pair variation.
+<!-- <!-- Weiterer ("zugehöriger) Reviewerkommentar: --> -->
+<!-- <!-- In looking at limits of agreement, the authors are considering the extent of within pair variation.
 However, in looking at the point wise limits of agreement, these are defined to incorporate a between pair standard deviation.
@@
 Perhaps the authors' lack of understanding comes from their erroneous assumption (iii).
--->
+--> -->
 
-<!-- Olsen et al.
+<!-- <!-- Olsen et al.
 [15] examined the validity of an inertial measurement unit (IMU) for analyzing the spatial displacement of the distal limbs of horses using functional LoA (FLoA).
@@
 The procedure, however, is subject to the same restrictions as any conventional linear regression model and is therefore limited when model assumptions are violated.
+--> -->
+
+<!-- <!-- Warum eigentlich Prädiktionsbänder und keine Konfidenzbänder? (Siehe Einleitung Lenhoff) --> -->
+<!-- <!-- Aktuell würde ich es so formulieren: In der Mehrheit der Fälle ist man an der nächsten (Einzel)Observation interessiert, nicht unbedingt an Mittelwerten (obwohl es durchaus Szenarien gibt, wo das akzeptabel sein kann) + im Zweifel generell lieber breitere Fehlerintervalle --> -->
+
+<!-- The principal object of this paper was to demonstrate a method for constructing continuous prediction intervals from curve data that addresses the aforementioned issues (analyzing the whole curve, including repeated measures, model assumptions (normal distribution, equal variances)), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret error bands (i.e.
+LoA in the actual measurement unit).
 -->
 
-<!-- Warum eigentlich Prädiktionsbänder und keine Konfidenzbänder? (Siehe Einleitung Lenhoff) -->
-<!-- Aktuell würde ich es so formulieren: In der Mehrheit der Fälle ist man an der nächsten (Einzel)Observation interessiert, nicht unbedingt an Mittelwerten (obwohl es durchaus Szenarien gibt, wo das akzeptabel sein kann) + im Zweifel generell lieber breitere Fehlerintervalle -->
+<!-- <!-- Es gibt bereits Ansätze für funktionale Boxplots (inkl. Vergleich mit punktweisen Intervallen), das Ganze wurde aber noch nie für Differenzenkurven untersucht. --> -->
 
-The principal object of this paper was to demonstrate a method for constructing continuous prediction intervals from curve data that addresses the aforementioned issues (analyzing the whole curve, including repeated measures, model assumptions (normal distribution, equal variances)), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret error bands (i.e.
-LoA in the actual measurement unit).
-
-<!-- Es gibt bereits Ansätze für funktionale Boxplots (inkl. Vergleich mit punktweisen Intervallen), das Ganze wurde aber noch nie für Differenzenkurven untersucht. -->
-
-We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters and LoA constructed similar to the method described in Roislien et al.
+<!-- We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters and LoA constructed similar to the method described in Roislien et al.
 (2012).
 For model validation, the respective LoA are cross validated using a leave-one-out approach.
+-->
 
-<!-- Aus Bland and Altman (2007): If each pair of X and Y measurements is treated as if from a different individual the structure of the data is ignored and incorrect estimates are  -->
-<!-- Ähnliche Formulierung aus https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x : In the repeated measures case, applying the standard limits of agreement to the data will result in limits that are too narrow because they do not take into account the reduction in variability that arises when working with averages of readings.
--->
+<!-- <!-- Aus Bland and Altman (2007): If each pair of X and Y measurements is treated as if from a different individual the structure of the data is ignored and incorrect estimates are  --> -->
+<!-- <!-- Ähnliche Formulierung aus https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x : In the repeated measures case, applying the standard limits of agreement to the data will result in limits that are too narrow because they do not take into account the reduction in variability that arises when working with averages of readings.
+--> -->
 
@@
 
-### Synthetic data sets
-
+#### Data generating process
 To compare different methods for constructing continuous LoA, several sets of smooth curve data were simulated across a range of common biomechanical signals and error characteristics.
@@
 All curves were modeled as Fourier series from additively superimposed sine functions.
-Each data set contains n.strides = 10 curves from n.subj = 11 fictitious test subjects.
+Each data set contains n.strides = 10 curves from n.subj = 11 simulated subjects.
 Each curve consists of t = 101 data points.
@@
 
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-n.subj <- 11
-n.ts <- 100
-
-t <- seq(0, 100)
-
-for (subject.idx in 1:n.subj) {
-
-  # Subjectwise parameters
-  offset.mean <- runif(1, min = -0.5, max = 0.5)
-
-  a.sd <- runif(1, min = 0.05, max = 0.15)
-  b.sd <- runif(1, min = 0.0001, max = 0.002)
-
-  for (stride.idx in 1:(n.strides)) {
-
-    a1.device1 <- rnorm(1, mean = 3, sd = a.sd)
-    a1.device2 <- rnorm(1, mean = 3, sd = a.sd)
-    a2.device1 <- 0.08
-    a2.device2 <- 0.08
-    b1.device1 <- rnorm(1, mean = 0.06, sd = b.sd)
-    b1.device2 <- rnorm(1, mean = 0.06, sd = b.sd)
-    b2.device1 <- rnorm(1, mean = 0.58, sd = b.sd)
-    b2.device2 <- rnorm(1, mean = 0.58, sd = b.sd)
-    c <- 2
-
-    sine.1 <- a1.device1 * sin(b1.device1 * t) ^ (c + 3)
-    sine.2 <- a2.device1 * sin(b2.device1 * t)
-    sine.3 <- a1.device2 * sin(b1.device2 * t) ^ (c + 3)
-    sine.4 <- a2.device2 * sin(b2.device1 * t)
-
-    offset <- rnorm(1, offset.mean, 0.05)
-
-    # Model (i). (ii), or (iii)
-    # ...
-```
-
-* (1) Homoscedastic (normal) error.
-Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
-The data are representative of e.
-g.
-joint angle measurements simultaneously recorded from two kinematic measurement systems (e.
-g.
-a camera-based measurement systems and a goniometer).
-
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-# ...
-
-# Model (i)
-device.1 <- sine.1 + sine.2
-device.2 <- offset + sine.3 + sine.4
-```
-
-* (2) Trend error: Same model as in (i), but with a nonlinear trend underlying one of the two signals.
-Similar errors can be observed e.
-g.
-when joint angles are calculated (numerically integrated) from gyroscope signals.
-
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-# ...
-trend <- (1 / 100000) * seq(0.5, 50.5, 0.5)^3
-
-# Model (ii)
-device.1 <- sine.1 + sine.2
-device.1 <- offset + sine.3 + sine.4 + trend
-```
-
-* (3) Heteroscedastic (non-gaussian, Weibull distributed) error: Heteroscedasticity across the signal length.
-The example consists of knee moment curves calculated using two different methods (direct and inverse kinematic approach) (Robinson et al., 2014).
-The curves were recreated using the example code provided in Robinson et al.
-(2021).
-
 <!-- Robinson et al. (2014) Impact of Knee Modeling Approach on Indicators and Classification of Anterior Cruciate Ligament Injury Risk -->
 
-<!-- Smooth curves created using the same model and parameters as in (i), but Weibulll distributed random curve (amplitude) parameters.
-This introduces non-gaussian measurement error (asymmetric/skewed distribution of differences around the mean difference).
--->
-
-```{r warning = FALSE, message = FALSE, eval = FALSE}
-# ...
-a1.device2 <- rnorm(n = 1,
-                    mean = rweibull(1, shape = 1.5, scale=1) - factorial(1/1.5), # factorial() used to center around 0
-                    sd = a.sd)
-# ...
-sine.3 <- a1.device2 * sin(b1.device2 * t) ^ (c + 3)
-# ...
-
-# Model (iii)
-device.1 <- sine.1 + sine.2
-device.2 <- offset + sine.3 + sine.4
-```
-
-* (4) Data with (missing) peaks.
+<!-- * (4) Data with (missing) peaks.
 Same model as in (i), except a couple of frames where a short, prominent peak occurs in one of the two signals.
@@
 to estimate ground reaction forces from acceleration values) is more difficult and my result in similar errors (Alcantara, 2021).
+-->
+<!-- <!-- Alcantara et al. Modelling/estimating ground reaction forces from acceleration values, z. B. https://www.biorxiv.org/content/10.1101/2021.03.17.435901v1.full --> -->
+<!-- <!-- Johnson et al. (2020) Multidimensional ground reaction forces and moments from wearable sensor accelerations via deep learning --> -->
 
-<!-- Alcantara et al. Modelling/estimating ground reaction forces from acceleration values, z. B. https://www.biorxiv.org/content/10.1101/2021.03.17.435901v1.full -->
-<!-- Johnson et al. (2020) Multidimensional ground reaction forces and moments from wearable sensor accelerations via deep learning -->
+#### Smooth wave data (constant variance, no trend) with subjectwise bias
+Homoscedastic (normal) error.
+Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
+The data are representative of e.
+g.
+joint angle measurements simultaneously recorded from two kinematic measurement systems (e.
+g.
+a camera-based measurement systems and a goniometer).
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_realistic", dir.data)
 
-### Smooth curves (homoscedasticity, normal error, no trend)
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
 
+PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Signal value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 20),
+        axis.title.x = element_text(size = 18),
+        axis.text.y = element_text(size = 20),
+        axis.title.y = element_text(size = 18),
+        legend.position = "none")
+
+PLOT
+```
+
+#### Smooth curves (homoscedasticity, normal error, no trend)
+Homoscedastic (normal) error.
+Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
+The data are representative of e.
+g.
+joint angle measurements simultaneously recorded from two kinematic measurement systems (e.
+g.
+a camera-based measurement systems and a goniometer).
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-### Smooth curves with nonlinear error trend
-
+#### Smooth curves with nonlinear error trend
+Same model as in (i), but with a nonlinear trend underlying one of the two signals.
+Similar errors can be observed e.
+g.
+when joint angles are calculated (numerically integrated) from gyroscope signals.
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-#### Heteroscedasticity across the signal length
-
+#### Heteroscedastic measurement error (no trend)
+Smooth curves created using the same model and parameters as in (i), but Weibulll distributed random curve (amplitude) parameters.
+This introduces non-gaussian measurement error (asymmetric/skewed distribution of differences around the mean difference).
 <!-- Smooth curves with non-gaussian (Weibull) error (no trend) -->
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-# Results
-
-### Functional limits of agreement
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-n.boot <- 400
-
-# data <- example_data(dat = "smooth", dir.data)
-# floa <- floa_rcb(data, n.boot, ver = "v2")
-# plot_loa(data, floa, central.tendency = "mean")
-# 
-# data <- example_data(dat = "smooth_trend", dir.data)
-# floa <- floa_rcb(data, n.boot, ver = "v2")
-# plot_loa(data, floa, central.tendency = "mean")
-# 
-# data <- example_data(dat = "non_gaussian", dir.data)
-# floa <- floa_rcb(data, n.boot, ver = "v2")
-# plot_loa(data, floa, central.tendency = "mean")
-# 
-# data <- example_data(dat = "shift", dir.data)
-# floa <- floa_rcb(data, n.boot, ver = "v2")
-# plot_loa(data, floa, central.tendency = "mean")
-```
-
-### Leave-one (curve) out cross validation
-
-Displayed are boxes for five different methods:
-
-* 1. FLoA~RCB_v2~ : One random stride per subject
-
-* 2. FLoA~RCB_v3~ : Fetch a SINGLE random stride from all strides
-
-* 3. FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and boot-
-       strap the resulting sample (of length (n=length(subjects))
-
-* 4. FLoA~Point~ : Pointwise B & A Limits of Agreement (SD derived from linear mixed effects model)
-
+#### Coverages
 Coverages are displayed for two different scenarios:
@@
 
-* c: Uncertainty in LoA (95% percentile intervals)
+### LoA uncertainty
+In order to quantify the uncertainty of the calculated LoA across different methods, LoA were calculated n.boot times and compared using uncertainty bands.
+
+# Results
+
+### Functional limits of agreement
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+n.boot <- 100
+```
+
+#### Smooth wave data (constant variance, no trend) with subjectwise bias
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_realistic", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
+#### Smooth curves (homoscedasticity, normal error, no trend)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
+#### Smooth curves with nonlinear error trend
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_trend", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
+#### Heteroscedastic measurement error (no trend)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "non_gaussian", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
+#### Curves phase shifted in x-axis direction
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "shift", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
+#### Real-world validation data
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "imu_mc", dir.data)
+floa <- floa_rcb(data, n.boot, ver = "v2")
+plot_loa(data, floa, central.tendency = "mean")
+```
+
+### Leave-one (curve) out cross validation
+Displayed are boxes for different calculation methods (FLoA~RCB_v2~, FLoA~RCB_v3~, FLoA~Roislien~, FLoA~Point~)
+
+#### Realistic smooth wave data (constant variance, no trend)
+a. Coverage of the entire curve
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_realistic", dir.data)
+
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
+
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross <- crossval_coverage(data, n.boot)
+# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# 
+# plot_cov_ver(cover.cross)
+```
+
+b. Mean (SD) of covered curve points
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+
+# # Display as boxplots across methods
+# plot_cov_ver(cover.cross.fraction.singlecurve)
+```
+
+c. Average distance to the outer point of the curve set
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+floa <- floa_rcb(data, n.boot, ver = "v2")
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+
+dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
+                       distance_2_floa(data, floa)$lwr.dist,
+                       distance_2_floa(data, floa.point)$upr.dist,
+                       distance_2_floa(data, floa.point)$lwr.dist,
+                       distance_2_floa(data, floa.roislien)$upr.dist,
+                       distance_2_floa(data, floa.roislien)$lwr.dist
+                       )
+
+colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw")
+
+boxplot(dist.all, ylab = "floa boundary - curve maximum")
+```
+
+d. Uncertainty in LoA (95% percentile intervals)
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+estimate_uncertainty_loa(data, n.boot)
+```
 
 #### Smooth, wave data (normal error, constant variance, no trend)
-
 a. Coverage of the entire curve
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-c. Uncertainty in LoA (95% percentile intervals)
+c. Average distance to the outer point of the curve set
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+floa <- floa_rcb(data, n.boot, ver = "v2")
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
 
+dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
+                       distance_2_floa(data, floa)$lwr.dist,
+                       distance_2_floa(data, floa.point)$upr.dist,
+                       distance_2_floa(data, floa.point)$lwr.dist,
+                       distance_2_floa(data, floa.roislien)$upr.dist,
+                       distance_2_floa(data, floa.roislien)$lwr.dist
+                       )
+
+colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw")
+
+boxplot(dist.all, ylab = "floa boundary - curve maximum")
+```
+
+d. Uncertainty in LoA (95% percentile intervals)
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-estimate_uncertainty_loa(data, n.boot)
+# estimate_uncertainty_loa(data, n.boot)
 ```
@@
 a. Coverage of the entire curve
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 #### Curves with non-gaussian (Weibull distributed) error (no trend)
-
 a. Coverage of the entire curve
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 #### Curves phase shifted in x-axis direction
-
 a. Coverage of the entire curve
-
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 
-c. Uncertainty in LoA (95% percentile intervals)
+c. Average distance to the outer point of the curve set
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+floa <- floa_rcb(data, n.boot, ver = "v2")
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
 
+dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
+                       distance_2_floa(data, floa)$lwr.dist,
+                       distance_2_floa(data, floa.point)$upr.dist,
+                       distance_2_floa(data, floa.point)$lwr.dist,
+                       distance_2_floa(data, floa.roislien)$upr.dist,
+                       distance_2_floa(data, floa.roislien)$lwr.dist
+                       )
+
+colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw")
+
+boxplot(dist.all, ylab = "floa boundary - curve maximum")
+```
+
+d. Uncertainty in LoA (95% percentile intervals)
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-estimate_uncertainty_loa(data, n.boot)
+# estimate_uncertainty_loa(data, n.boot)
+```
+
+#### Real-world validation data
+a. Coverage of the entire curve
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "imu_mc", dir.data)
+
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
+
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
+
+knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross <- crossval_coverage(data, n.boot)
+# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
+# 
+# plot_cov_ver(cover.cross)
+```
+
+b. Mean (SD) of covered curve points
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Leave-one curve out approach
+# ------------------------------------------------------------------------------
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+
+cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+)
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points")
+
+# # Display as boxplots acroos methods
+# plot_cov_ver(cover.cross.fraction.singlecurve)
+
+# # Leave-one subject out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
+# plot_cov_ver(cover.cross.fraction)
+```
+
+c. Average distance to the outer point of the curve set
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+floa <- floa_rcb(data, n.boot, ver = "v2")
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+
+dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
+                       distance_2_floa(data, floa)$lwr.dist,
+                       distance_2_floa(data, floa.point)$upr.dist,
+                       distance_2_floa(data, floa.point)$lwr.dist,
+                       distance_2_floa(data, floa.roislien)$upr.dist,
+                       distance_2_floa(data, floa.roislien)$lwr.dist
+                       )
+
+colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw")
+
+boxplot(dist.all, ylab = "floa boundary - curve maximum")
+```
+
+d. Uncertainty in LoA (95% percentile intervals)
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# estimate_uncertainty_loa(data, n.boot)
 ```
@@
 -->
-
+d
 It might be beneficial to work with functional data as suggested by Lenhoff et al.

### 2021-09-26 R/distance_2_floa.R

@@ -1,0 +1,51 @@
+# Average distance to the outer point of the curve set
+distance_2_floa <- function (data, floa) {
+
+  # floa variables need to be arranged in columns
+  if (dim(floa)[1] < dim(floa)[2]) {
+    floa <- t(floa)
+  }
+
+  if (!is.data.frame(floa)) {
+    floa <- as.data.frame(floa)
+  }
+
+  device1 <- data.frame(subset(data, device == "IMU")$value)
+  device2 <- data.frame(subset(data, device == "MC")$value)
+  device.diff <- device1 - device2
+  colnames(device.diff)[1] <- "value"
+
+  device.diff$frame <- seq(0, 100)
+  n.frames <- length(unique(data$frame))
+  n.strides <- length(unique(data$strideID))
+  n.subjects <- length(unique(data$subjectID))
+  strides.per.subject <- length(unique(data$strideID)) / n.subjects
+  device.diff$strideID <- as.factor(rep(1:n.strides, each = 101))
+  device.diff$subjectID <- as.factor(rep(1:n.subjects, each = strides.per.subject * n.frames))
+
+  upr.dist <- vector(mode = "list", length = n.frames)
+  lwr.dist <- vector(mode = "list", length = n.frames)
+  for (frame.idx in seq(0, 100)) {
+    data.per.frame <- subset(device.diff, subset = frame == frame.idx)
+
+    curve.extrema <- range(data.per.frame$value)
+    curve.max <- max(curve.extrema)
+    curve.min <- min(curve.extrema)
+
+    # Account for the loop index starting at 0
+    frame.idx.correct <- ifelse(frame.idx == 0, frame.idx + 1, frame.idx)
+
+    floa.upr <- floa$upper[frame.idx + 1]
+    floa.lwr <- floa$lower[frame.idx + 1]
+
+    upr.dist[[frame.idx.correct]] <- floa.upr - curve.max
+    lwr.dist[[frame.idx.correct]] <- floa.lwr - curve.min
+  }
+
+  distance <- data.frame(unlist(upr.dist),
+                         unlist(lwr.dist))
+
+  colnames(distance) <- c("upr.dist", "lwr.dist")
+
+  return(distance)
+}

### 2021-09-26 R/estimate_uncertainty_loa.R

@@ -100,13 +100,13 @@ estimate_uncertainty_loa <- function (data, n.boot) {
   device.diff$floa.roislien.lower.2.5 <- rep(pi95$floa.roislien.lower.2.5, length.out = nrow(device.diff))
   device.diff$floa.roislien.lower.97.5 <- rep(pi95$floa.roislien.lower.97.5, length.out = nrow(device.diff))
 
-  # For line graphs, the data points must be grouped so that it knows which points to connect.
+  # For line graphs, the data points must be grouped so that it knows whic h points to connect.
   # In this case, it is simple -- all points should be connected, so group=1.
   # When more variables are used and multiple lines are drawn, the grouping for lines is usually done by variable.
   PLOT.DIFF <- ggplot(data = device.diff, aes(x = frame, y = value)) +
-    geom_line(data = device.diff,
-              aes(group = strideID),
-              alpha = 0.25) +
+    # geom_line(data = device.diff,
+    #           aes(group = strideID),
+    #           alpha = 0.25) +
     geom_ribbon(aes(ymin = floa.roislien.upper.2.5,
                     ymax = floa.roislien.upper.97.5),
                 fill = "burlywood",

### 2021-09-26 R/fdaDelta.R

@@ -1,33 +1,31 @@
 fdaDelta <- function(data, fd.basis){
 
-  # Approximates time series using Fourier functions ---------------------------
-
-  library(fda)
+  # Approximates time series using Fourier functions
+  # ----------------------------------------------------------------------------
 
   # First device
-  # --------------------------
+  # -------------------------------------------------------
   dev1 <- subset(data, device  == "IMU")$value
-
   dev1.wide <- matrix(dev1, ncol = length(dev1) / 101)
 
   # Create functional data object
   dev1.wide.fd <- fda::Data2fd(argvals = dev1.wide, basisobj = fd.basis)
 
   # Second device
-  # --------------------------
+  # -------------------------------------------------------
   dev2 <- subset(data, device == "MC")$value
-
   dev2.wide <- matrix(dev2, ncol = length(dev2) / 101)
 
-  dev2.wide.fd <- Data2fd(argvals = dev2.wide, basisobj = fd.basis)
+  dev2.wide.fd <- fda::Data2fd(argvals = dev2.wide, basisobj = fd.basis)
 
   # Differences
-  # --------------------------
+  # -------------------------------------------------------
   fd.diff <- dev1.wide.fd - dev2.wide.fd
 
-  # point-wise statistics
-  mean.diff.fd <- mean.fd(fd.diff)
-  sd.diff.fd <- std.fd(fd.diff)
+  # Point-wise statistics
+  # -------------------------------------------------------
+  mean.diff.fd <- fda::mean.fd(fd.diff)
+  sd.diff.fd <- fda::std.fd(fd.diff)
 
   fda <- list()
 

### 2021-09-26 R/floa_lenhoff.R

@@ -1,0 +1,139 @@
+floa_lenhoff <- function(data, n.boot, cp.begin, alpha) {
+
+  # Limit of Agreement calculation adapted from Lenhoff et al. (1999)
+  # ----------------------------------------------------------------------------
+  #
+  # Default values:
+  # Bootstrap iterations (n.boot)     : 100
+  # Anfangswert Quantil (cp_begin)    : 0
+  # Irrtumswahrscheinlichkeit (alpha) : 0.05
+  # ----------------------------------------------------------------------------
+
+  # Dimensions
+  n.frames <- length(unique(data$frame))
+  n.curves <- length(unique(data$strideID))
+
+  # ----------------------------------------------------------------------------
+  # Fourier Transformation
+  #
+  # Approximate time series (differences) using Fourier functions
+  # ----------------------------------------------------------------------------
+
+  # Number of basic (Fourier) functions
+  # ... appears to plateau around 50 basis vectors
+  k <- 50
+
+  fd.basis <- fda::create.fourier.basis(nbasis = k)
+
+  # Returns functional data representation of difference curves
+  data.diff.fd <- fdaDelta(data, fd.basis)
+
+  original.sample.fd <- data.diff.fd[[1]]
+  original.sample.mean.fd <- data.diff.fd[[2]]
+  original.sample.sd.fd <- data.diff.fd[[3]]
+
+  # ----------------------------------------------------------------------------
+  # Bootstrap
+  # ----------------------------------------------------------------------------
+
+  # for i=1:n_repeat
+  # % neues Bootstrap Sample bilden
+  # for k=1:size(daten,2)
+  # bootstrap.zz(k,i)=floor(unifrnd(1,size(daten,2)+1));
+  # bootstrap.pseudo_koeffi(:,k,i) = fourier.koeffi(:,bootstrap.zz(k,i));
+  # bootstrap.real(:,k,i)=fourier.s*bootstrap.pseudo_koeffi(:,k,i);
+  # end
+
+  for (b in 1:n.boot) {
+    # Extract information from class fd object
+    # https://rdrr.io/cran/fda/src/R/plot.fd.R
[…]

### 2021-09-26 R/floa_rcb.R

@@ -17,18 +17,31 @@ floa_rcb <- function(data, n.boot, ver) { # , fd.basis
                             byrow = TRUE)
 
   # Calculate percentiles ------------------------------------------------------
+  alpha <- 0.5
+
+  # Alpha correction (Bonferroni)
+  m <- length(unique(data$frame))
+  alpha <- alpha/m
+  # Round is necessary because quantile() can only take two decimal places in the "probs" argument
+  upr <- round(1 - alpha/2)
+  lwr <- round(alpha/2, 2)
+
   floa.boot.percentiles <- c()
   for (i in 1:ncol(clust.agg.intrp)) {
     floa.boot.percentiles <- c(floa.boot.percentiles,
                                quantile(clust.agg.intrp[, i],
-                                        probs = c(0.025, 0.5, 0.975)
+                                        probs = c(lwr, 0.5, upr)
                                         )
                                )
   }
 
-  perc2.5 <- floa.boot.percentiles[which(names(floa.boot.percentiles) == "2.5%")]
-  perc50 <- floa.boot.percentiles[which(names(floa.boot.percentiles) == "50%")]
-  perc97.5 <- floa.boot.percentiles[which(names(floa.boot.percentiles) == "97.5%")]
+  name.lower.percentile <- unique(names(floa.boot.percentiles))[1]
+  name.mid.percentile <- unique(names(floa.boot.percentiles))[2]
+  name.upper.percentile <- unique(names(floa.boot.percentiles))[3]
+
+  perc2.5 <- floa.boot.percentiles[which(names(floa.boot.percentiles) == name.lower.percentile)]
+  perc50 <- floa.boot.percentiles[which(names(floa.boot.percentiles) == name.mid.percentile)]
+  perc97.5 <- floa.boot.percentiles[which(names(floa.boot.percentiles) == name.upper.percentile)]
 
   # Split the long percentile vector into 101 data points each -----------------
   floa.boot.percentiles.split <- rbind(approx(perc2.5, n = 101)$y,

### 2021-09-26 R/get_coverage.R

@@ -21,21 +21,29 @@ get_coverage <- function (data, floa) {
   lwr.bnd <- floa["lower", ]
   upr.bnd <- floa["upper", ]
 
+  # # Plausibility check
+  # plot(lwr.bnd, type = "l", ylim = c(-5, 5), col = "red", lwd = 4)
+  # lines(upr.bnd, col = "red", lwd = 4)
+
   n.strides <- length(unique(device.diff$strideID))
   outside <- 0
 
   # Account for stride indices not starting at 1
   stride.indices <- unique(data$strideID)
 
-  for (stride.idx in stride.indices){
-
+  for (stride.idx in 1:20){ # stride.indices
     curve <- subset(device.diff, strideID == stride.idx)
+
+    # # Plausibility check
+    # lines(curve$value)
 
     # Compare difference curves with upper and lower boundaries
     below.thresh <- curve$value < lwr.bnd
     above.thresh <- curve$value > upr.bnd
 
     points.outside <- sum(above.thresh) + sum(below.thresh)
+
+    outside
 
     # Check if the entire curve is within the limits
     if (points.outside > 0) {

### 2021-09-26 R/lenhoff_doris.R

@@ -1,0 +1,169 @@
+floa_lenhoff <- function(data, K, n.boot, cp.begin, alpha) {
+
+  # Limit of Agreement calculation adapted from Lenhoff et al. (1999)
+  # ----------------------------------------------------------------------------
+  #
+  # Function arguments:
+  # n.boot   : Number of bootstrap iterations
+  # k_reihe  : number of coefficients
+  # cp.begin : Anfangswert Quantil
+  # alpha    : Irrtumswahrscheinlichkeit
+  # ----------------------------------------------------------------------------
+
+  # ----------------------------------------------------------------------------
+  # Get difference curves
+  # ----------------------------------------------------------------------------
+
+  devices <- unique(data$device)
+
+  device.1 <- subset(data, device  == devices[1])$value
+  device.1.wide <- matrix(device.1, ncol = length(device.1) / 101)
+  device.2 <- subset(data, device == devices[2])$value
+  device.2.wide <- matrix(device.2, ncol = length(device.2) / 101)
+
+  data.diff <- device.1.wide - device.2.wide
+
+  # Dimensions
+  n_time    <- dim(data.diff)[1]
+  n_kurven  <- dim(data.diff)[2]
+  time <- seq(0, (n_time-1)) # 0:(n_time-1)
+
+  # ----------------------------------------------------------------------------
+  # Fourier Transformation
+  #
+  # Approximate time series (differences) using Fourier functions
+  # ----------------------------------------------------------------------------
+
+  fourier.koeffi    = matlab::zeros(c(k_reihe*2 + 1, n_kurven))
+  fourier.real      = matlab::zeros(n_time, n_kurven)
+  fourier.mean      = matlab::zeros(k_reihe*2 + 1)
+  fourier.real_mw   = matlab::zeros(n_time, 1)
+  fourier.std1      = matlab::zeros(k_reihe*2 + 1, k_reihe*2 + 1, n_kurven)
+  fourier.kovarianz = matlab::zeros(k_reihe*2 + 1, k_reihe*2 + 1)
+  fourier.std_all   = matlab::zeros(n_time, n_time)
[…]

### 2021-09-26 R/main.R

@@ -38,8 +38,6 @@
 rm(list = ls())
 
 library(ggplot2)
-# library(fda)
-# library(funData)
 
 dir.script <- "~/floa/R"
 dir.data <- "~/floa/R/examples"
@@ -47,7 +45,7 @@ dir.data <- "~/floa/R/examples"
 setwd(dir.script)
 
 source("example_data.R")
-# source("fdaDelta.R")
+source("fdaDelta.R")
 source("pick_subwise_curves.R")
 source("draw_clusters.R")
 source("floa_rcb.R")
@@ -63,6 +61,7 @@ source("crossval_coverage_fraction.R")
 source("singlecurve_coverage.R")
 source("singlecurve_coverage_fraction.R")
 source("plot_cov_ver.R")
+source("distance_2_floa.R")
 source("estimate_uncertainty_loa.R")
 
 # ********************************* Data sets **********************************
@@ -76,7 +75,7 @@ source("estimate_uncertainty_loa.R")
 # * Data with non-gaussian (Weibull distributed) error (no trend): "non_gaussian"
 # * Data with shock peaks (no bias, no trend): "shock"
 # * Phase shifted data (x-axis direction): "shift"
-data <- example_data(dat = "imu_mc", dir.data)
+data <- example_data(dat = "smooth_realistic", dir.data)
 
 # Plot data --------------------------------------------------------------------
 # uncommment when subject differences need to be plotted
@@ -96,15 +95,6 @@ PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID
         legend.position = "none")
 
 PLOT
-
-# # ****** Approximate time series (differences) using Fourier functions *********
-#
-# # Number of basic (Fourier) functions
-# # ... appears to plateau around 50 basis vectors
-# fd.basis <- fda::create.fourier.basis(nbasis = 50)
-#
-# # Returns functional data representation of difference curves
-# data.diff.fd <- fdaDelta(data, fd.basis)
 
 
 # ****************************** Calculate FLoA ********************************
@@ -186,6 +176,21 @@ plot_cov_ver(cover.cross.fraction.singlecurve)
 
 
 
-
-
+# Estimate the uncertainty across several iterations
+# Returns results for all implemented methods
 estimate_uncertainty_loa(data, n.boot)
+
+
+# Average distance to the outer point of the curve set
+dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
[…]

### 2021-09-26 R/plot_loa.R

@@ -29,7 +29,7 @@ plot_loa <- function (data, floa, central.tendency) {
   n.strides <- length(unique(data$strideID))
   n.subjects <- length(unique(data$subjectID))
   strides.per.subject <- length(unique(data$strideID)) / n.subjects
-  device.diff$strideID <- as.factor(rep(1:n.strides, each = 101)) # as.factor(seq(0, n.strides, by = n.frames)) # as.factor(rep(seq(1, strides.per.subject), each = 101))
+  device.diff$strideID <- as.factor(rep(1:n.strides, each = 101))
   device.diff$subjectID <- as.factor(rep(1:n.subjects, each = strides.per.subject * n.frames))
 
   floa.point <- data.frame(t(floa_point(data)))
@@ -78,30 +78,41 @@ plot_loa <- function (data, floa, central.tendency) {
     # For line graphs, the data points must be grouped so that it knows which points to connect.
     # In this case, it is simple -- all points should be connected, so group=1.
     # When more variables are used and multiple lines are drawn, the grouping for lines is usually done by variable.
-    PLOT.DIFF <- ggplot(data = device.diff, aes(x = frame, y = value)) + # , color = subjectID, group = strideID
-      geom_line() +
+    PLOT.DIFF <- ggplot(data = device.diff, aes(x = frame, y = value)) +
+      geom_line(data = device.diff,
+                aes(group = strideID),
+                alpha = 0.25) +
       scale_color_grey(start = 0.8, end = 0.2) +
       # Plot lower limit of agreement
       geom_line(data = floa,
                 aes(x = seq(0, 100), y = upper, col = "red", group = 1),
                 linetype = "solid",
-                size = 3,
+                size = 2,
                 colour = "red") +
-      # Plot mean difference
-      geom_line(data = floa,
-                aes(x = seq(0, 100), y = mean, col = "red", group = 1),
-                linetype = "dotted",
-                size = 3,
-                colour = "red") +
+      # # Plot mean difference
+      # geom_line(data = floa,
+      #           aes(x = seq(0, 100), y = mean, col = "red", group = 1),
+      #           linetype = "dotted",
+      #           size = 2,
+      #           colour = "red") +
       # Plot upper limits of agreement
       geom_line(data = floa,
                 aes(x = seq(0, 100), y = lower, col = "red", group = 1),
                 linetype = "solid",
-                size = 3,
+                size = 2,
                 colour = "red") +
[…]

### 2021-09-26 R/simulate_data.R

@@ -9,14 +9,13 @@
 # Simulated data sets:
 #
 # 0. Real-world validation data: "imu_mc"
-# 1. Realistic smooth wave data (constant variance, no trend)
-# 2. Smooth, wave data (normal error, constant variance, no trend): "smooth"
-# 3. Homoscedastic error
-# 4. Heteroscedastic error
-# 5. Non-stationary data (trend, no bias) data:"non_stationary"
-# 6. Non-gaussian (Weibull distributed) error (no trend)
-# 7. Data with shock peaks (no bias, no trend): "shock"
-# 8. Shift data (data phase shifted (x-axis direction))
+# 1. Realistic smooth wave data (normal error, constant variance, no trend)
+# 2. Same model as in (1), but almost no variation in one of the devices: "smooth"
+# 3. Heteroscedastic error
+# 4. Non-stationary data (trend, no bias) data:"non_stationary"
+# 5. Non-gaussian (Weibull distributed) error (no trend)
+# 6. Data with shock peaks (no bias, no trend): "shock"
+# 7. Shift data (data phase shifted (x-axis direction))
 ################################################################################
 
 # 0. Real-world validation data: "imu_mc" --------------------------------------
@@ -44,6 +43,8 @@ n.devices <- 2
 n.frames <- 101
 
 # 1. Realistic smooth wave data (constant variance, no trend) ------------------
+set.seed(3)
+
 device <- c()
 value <- c()
 subjectID <- c()
@@ -108,6 +109,9 @@ saveRDS(data, file = paste0("~/floa/R/examples/", "smooth_realistic.rds"))
 
 
 # 2. Smooth wave data (constant variance, no trend) with subjectwise bias ------
+# Same model as in (1), but almost no variation in one of the devices
+set.seed(3)
+
 device <- c()
 value <- c()
 subjectID <- c()
@@ -174,79 +178,9 @@ saveRDS(data, file = paste0("~/floa/R/examples/", "smooth.rds"))
 
 
 
[…]

### 2021-09-28 paper.Rmd
Adjust rmd file

@@
 library(boot)
-# library(fda)
-# library(funData)
 
@@
 source("example_data.R")
-# source("fdaDelta.R")
 source("pick_subwise_curves.R")
@@
 source("floa_rcb.R")
+source("lenhoff_doris.R")
 source("floa_point.R")
@@
 source("plot_cov_ver.R")
-source("distance_2_floa.R")
+# source("distance_2_floa.R")
 source("estimate_uncertainty_loa.R")
@@
 data <- example_data(dat = "smooth_realistic", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 ```
@@
 data <- example_data(dat = "smooth", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 ```
@@
 data <- example_data(dat = "smooth_trend", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 ```
@@
 data <- example_data(dat = "non_gaussian", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 ```
@@
 data <- example_data(dat = "shift", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 ```
@@
 data <- example_data(dat = "imu_mc", dir.data)
-floa <- floa_rcb(data, n.boot, ver = "v2")
-plot_loa(data, floa, central.tendency = "mean")
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 ```
@@
 covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("RCB_2", "RCB_3", "Roislien_4", "Point_5")
+colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
 
@@
 )
+colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
+
 knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
@@
 
-c. Average distance to the outer point of the curve set
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-floa <- floa_rcb(data, n.boot, ver = "v2")
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-
-dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
-                       distance_2_floa(data, floa)$lwr.dist,
-                       distance_2_floa(data, floa.point)$upr.dist,
-                       distance_2_floa(data, floa.point)$lwr.dist,
-                       distance_2_floa(data, floa.roislien)$upr.dist,
-                       distance_2_floa(data, floa.roislien)$lwr.dist
-                       )
-
-colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw")
-
-boxplot(dist.all, ylab = "floa boundary - curve maximum")
-```
-
-d. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-estimate_uncertainty_loa(data, n.boot)
+<!-- c. Average distance to the outer point of the curve set -->
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
+<!-- floa.point <- floa_point(data) -->
+<!-- floa.roislien <- floa_roislien(data) -->
+
+<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
+<!--                        distance_2_floa(data, floa)$lwr.dist, -->
+<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
+<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
+<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
+<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
+<!--                        ) -->
+
+<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
+
+<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
+<!-- ``` -->
+
+c. Uncertainty in LoA (95% percentile intervals)
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# estimate_uncertainty_loa(data, n.boot)
 ```
@@
 
-c. Average distance to the outer point of the curve set
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-floa <- floa_rcb(data, n.boot, ver = "v2")
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-
-dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist,
-                       distance_2_floa(data, floa)$lwr.dist,
-                       distance_2_floa(data, floa.point)$upr.dist,
-                       distance_2_floa(data, floa.point)$lwr.dist,
-                       distance_2_floa(data, floa.roislien)$upr.dist,
-                       distance_2_floa(data, floa.roislien)$lwr.dist
-                       )
-
-colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw")
-
-boxplot(dist.all, ylab = "floa boundary - curve maximum")
-```
-
-d. Uncertainty in LoA (95% percentile intervals)
+<!-- c. Average distance to the outer point of the curve set -->
+<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
+<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
+<!-- floa.point <- floa_point(data) -->
+<!-- floa.roislien <- floa_roislien(data) -->
+
+<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
+<!--                        distance_2_floa(data, floa)$lwr.dist, -->
+<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
+<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
+<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
+<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
+<!--                        ) -->
+
+<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
+
+<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
+<!-- ``` -->
+
+c. Uncertainty in LoA (95% percentile intervals)
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
@@
 )
-knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points")
+colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
+
+knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
 

### 2021-09-28 R/lenhoff_doris.R

@@ -1,4 +1,4 @@
-floa_lenhoff <- function(data, K, n.boot, cp.begin, alpha) {
+floa_lenhoff <- function(data, k_reihe, n.boot, cp.begin, alpha) {
 
   # Limit of Agreement calculation adapted from Lenhoff et al. (1999)
   # ----------------------------------------------------------------------------
@@ -54,7 +54,7 @@ floa_lenhoff <- function(data, K, n.boot, cp.begin, alpha) {
   for (i in 1:n_time) {
     # Least squares Regression
     # pracma::mldivide emulates the Matlab backslash operator “\” (Matrix left division)
-    fourier.koeffi[, i] = pracma::mldivide(fourier.s, data.diff[, 1]) # fourier.s\daten[, i]
+    fourier.koeffi[, i] = pracma::mldivide(fourier.s, data.diff[, i]) # fourier.s\daten[, i]
     # Fourierkurve
     fourier.real[, i] = fourier.s %*% fourier.koeffi[, i]
   }
@@ -66,11 +66,11 @@ floa_lenhoff <- function(data, K, n.boot, cp.begin, alpha) {
   # Standardabweichung der Fourierkurve
   for (i in 1:n_kurven) {
     # Lenhoff, Appendix A, Eq. (0.5)
-    fourier.std1[, , i] <- t((fourier.koeffi[, i] - fourier.mean[, 1])) * (fourier.koeffi[, i] - fourier.mean[, 1])
+    fourier.std1[, , i] <- (fourier.koeffi[, i] - fourier.mean[, 1]) %*% t(fourier.koeffi[, i] - fourier.mean[, 1])
   }
 
   fourier.kovarianz <- apply(fourier.std1, c(1, 2), mean) # nanmean(fourier.std1, 3)
-  fourier.std_all <- sqrt(fourier.s * fourier.kovarianz * t(fourier.s))
+  fourier.std_all <- sqrt(fourier.s %*% fourier.kovarianz %*% t(fourier.s))
 
   for (i in 1:n_time) {
     # Werte liegen auf der Diagonalen der quadratischen Matrix fourier.std_all
@@ -81,8 +81,8 @@ floa_lenhoff <- function(data, K, n.boot, cp.begin, alpha) {
   # fourier.std <- apply(fourier.real, 1, sd)
   # # plot(fourier.std)
 
[…]

### 2021-09-28 R/main.R

@@ -49,6 +49,7 @@ source("fdaDelta.R")
 source("pick_subwise_curves.R")
 source("draw_clusters.R")
 source("floa_rcb.R")
+source("lenhoff_doris.R")
 source("floa_point.R")
 source("floa_roislien.R")
 source("plot_loa.R")
@@ -114,20 +115,18 @@ n.boot <- 400
 # v3 : Fetch a SINGLE random stride from all strides
 # v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
 #      strap the resulting sample (of length (n=length(subjects))
-floa <- floa_rcb(data, n.boot, ver = "v2")
+# floa <- floa_rcb(data, n.boot, ver = "v2")
 
-# Pointwise LoA ----------------------------------------------------------------
-# Mean + SD are calculated across all strides/subjects using linear mixed models
 floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
 
-
-# ********************************* Plot data **********************************
-plot_loa(data, floa = floa, central.tendency = "mean")
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
 
 
 # ********************************* Coverage ***********************************
 # Calculate coverage (entire curves within the limits of agreement)
-coverage <- get_coverage(data, floa.point) # Select floa method: floa or floa.point
+coverage <- get_coverage(data, floa.roislien) # Select floa method: floa or floa.point
 print(coverage)
 
 

### 2021-09-28 R/plot_loa.R

@@ -1,4 +1,4 @@
-plot_loa <- function (data, floa, central.tendency) {
+plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff) {
   # plot_loa() function arguments:
   # 1. floa: Limits of agreement
   # + a. floa (rcb method)
@@ -7,14 +7,18 @@ plot_loa <- function (data, floa, central.tendency) {
   # + a. mean
   # + b. median
 
-  # floa variables need to be arranged in columns
-  if (dim(floa)[1] < dim(floa)[2]) {
-    floa <- t(floa)
-  }
+  # # floa variables need to be arranged in columns
+  # if (dim(floa)[1] < dim(floa)[2]) {
+  #   floa <- t(floa)
+  # }
+  #
+  # if (!is.data.frame(floa)) {
+  #   floa <- as.data.frame(floa)
+  # }
 
-  if (!is.data.frame(floa)) {
-    floa <- as.data.frame(floa)
-  }
+  floa.point <- data.frame(t(floa.point))
+  floa.roislien <- data.frame(t(floa.roislien))
+  floa.lenhoff <- data.frame(t(floa.lenhoff))
 
   # Prepare data for ggploting -------------------------------------------------
   device1 <- data.frame(subset(data, device == "IMU")$value)
@@ -32,96 +36,55 @@ plot_loa <- function (data, floa, central.tendency) {
   device.diff$strideID <- as.factor(rep(1:n.strides, each = 101))
   device.diff$subjectID <- as.factor(rep(1:n.subjects, each = strides.per.subject * n.frames))
 
-  floa.point <- data.frame(t(floa_point(data)))
 
+  # For line graphs, the data points must be grouped so that it knows which points to connect.
+  # In this case, it is simple -- all points should be connected, so group=1.
+  # When more variables are used and multiple lines are drawn, the grouping for lines is usually done by variable.
+  PLOT.DIFF <- ggplot(data = device.diff, aes(x = frame, y = value)) +
+    geom_line(data = device.diff,
+              aes(group = strideID),
+              alpha = 0.25) +
+    scale_color_grey(start = 0.8, end = 0.2) +
+    # Plot lower limit of agreement
+    geom_line(data = floa.lenhoff,
+              aes(x = seq(0, 100), y = upper, col = "red", group = 1),
+              linetype = "solid",
+              size = 1.5,
+              colour = "#D55E00") +
+    geom_line(data = floa.lenhoff,
[…]

### 2021-09-28 R/singlecurve_coverage.R

@@ -17,19 +17,22 @@ singlecurve_coverage <- function (data, n.boot) {
   n.curves <- unique(data$strideID)
 
   # cover.cross.rcb.v1   <- vector(mode = "list", length = length(n.curves))
-  cover.cross.rcb.v2   <- vector(mode = "list", length = length(n.curves))
-  cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
+  # cover.cross.rcb.v2   <- vector(mode = "list", length = length(n.curves))
+  # cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
+  cover.cross.lenhoff <- vector(mode = "list", length = length(n.curves))
+
   for (curve.idx in n.curves) {
     # Calculate FLoA with one curve left out -----------------------------
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
     # floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
-    floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
-    floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
+    # floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
+    # floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out)
+    floa.lenhoff <- floa_lenhoff(data.one.out, k_reihe = 30, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
 
     # Plot left out curve vs. various FLoA methods -----------------------
     data.subset <- subset(data, strideID == curve.idx)
@@ -38,67 +41,22 @@ singlecurve_coverage <- function (data, n.boot) {
     device2 <- data.frame(subset(data.subset, device == "MC"))
     device.diff <- device1$value - device2$value
 
-    floa.all <- data.frame(device.diff)
-    floa.all$frame <- seq(0, 100)
-    # floa.all$floa.v1.upper <- floa.v1["upper", ]
-    # floa.all$floa.v1.lower <- floa.v1["lower", ]
-    floa.all$floa.v2.upper <- floa.v2["upper", ]
[…]

### 2021-09-28 R/singlecurve_coverage_fraction.R

@@ -27,19 +27,22 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
   n.curves <- unique(data$strideID)
 
   # cover.cross.rcb.v1   <- vector(mode = "list", length = length(n.curves))
-  cover.cross.rcb.v2   <- vector(mode = "list", length = length(n.curves))
-  cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
+  # cover.cross.rcb.v2   <- vector(mode = "list", length = length(n.curves))
+  # cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
+  cover.cross.lenhoff <- vector(mode = "list", length = length(n.curves))
+
   for (curve.idx in n.curves) {
     # Calculate FLoA with one curve left out -----------------------------
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
     # floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
-    floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
-    floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
+    # floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
+    # floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out)
+    floa.lenhoff <- floa_lenhoff(data.one.out, k_reihe = 30, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
 
     # Plot left out curve vs. various FLoA methods -----------------------
     data.subset <- subset(data, strideID == curve.idx)
@@ -51,18 +54,20 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
     # Get coverage for the left out (difference) curve
     # --------------------------------------------------------------------
     # cover.cross.rcb.v1[curve.idx]   <- get_coverage_singlecurve_fraction(device.diff, floa.v1)
-    cover.cross.rcb.v2[curve.idx]   <- get_coverage_singlecurve_fraction(device.diff, floa.v2)
-    cover.cross.rcb.v3[curve.idx]   <- get_coverage_singlecurve_fraction(device.diff, floa.v3)
[…]

### 2021-09-28 paper.Rmd
Add y-axis range as function argument to plot_loa()

@@
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
 ```
@@
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-10, 10))
 ```
@@
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-15, 15))
 ```
@@
 ### Leave-one (curve) out cross validation
-Displayed are boxes for different calculation methods (FLoA~RCB_v2~, FLoA~RCB_v3~, FLoA~Roislien~, FLoA~Point~)
 

### 2021-09-28 R/main.R

@@ -121,7 +121,7 @@ floa.point <- floa_point(data)
 floa.roislien <- floa_roislien(data)
 floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff)
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
 
 
 # ********************************* Coverage ***********************************

### 2021-09-28 R/plot_loa.R

@@ -1,31 +1,25 @@
-plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff) {
-  # plot_loa() function arguments:
-  # 1. floa: Limits of agreement
-  # + a. floa (rcb method)
-  # + b. floa.point # requires: data.frame(t(floa.point))
-  # 2. central.tendency: Central tendency parameter
-  # + a. mean
-  # + b. median
+plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
 
-  # # floa variables need to be arranged in columns
-  # if (dim(floa)[1] < dim(floa)[2]) {
-  #   floa <- t(floa)
-  # }
+  # Plots the limits across the respective methods against all difference curves
+  # in the data set
+  # ----------------------------------------------------------------------------
   #
-  # if (!is.data.frame(floa)) {
-  #   floa <- as.data.frame(floa)
-  # }
+  # Arguments:
+  # floa.point: point-wise limits of agreement (Bland & Altman, 2007)
+  # floa.roislien: limits of agreement as in Roislien et al. (2012)
+  # floa.lenhoff: limits of agreement as in Lenhoff et al. (1999)
+  # ylim: y-axis limits (2 element numeric object)
+  # ----------------------------------------------------------------------------
 
   floa.point <- data.frame(t(floa.point))
   floa.roislien <- data.frame(t(floa.roislien))
   floa.lenhoff <- data.frame(t(floa.lenhoff))
 
-  # Prepare data for ggploting -------------------------------------------------
+  # Difference curves -------------------------------------------------
   device1 <- data.frame(subset(data, device == "IMU")$value)
   device2 <- data.frame(subset(data, device == "MC")$value)
 
   device.diff <- device1 - device2
-
   colnames(device.diff)[1] <- "value"
 
   device.diff$frame <- seq(0, 100)
@@ -76,7 +70,7 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff) {
               linetype = "dotted",
               size = 1.5,
               colour = "#009E73") +
-    scale_y_continuous(limits = c(-5, 5)) +
+    scale_y_continuous(limits = c(ylim[1], ylim[2])) +
     labs(x = "Time-normalized signal [%]", y = "Difference") +
     theme_minimal() +
     theme(axis.text.x = element_text(size = 20),

### 2021-10-14 paper.Rmd
Continued paper draft

@@
 ---
-title: "Analysis of different methods for constructing continuous limits of agreement from smooth biomechanical data"
+title: "Comparison of continuous prediction bands for the validation of smooth 1D biomechanical time series data"
+
 author:
@@
 <!-- - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen -->
-- Weitere synthetische Daten siehe Paper Robinson et al. (2021)
-- Evtl. simulierte Datensätze um Beispiele aus https://mjskay.github.io/ggdist/articles/lineribbon.html erweitern (insbesondere das erste Beispiel mit x-versetzten Gausskurven)!
-- Simulierte Daten durch Funktionen ersetzen?
-Eigentlich aktuell eher unnötig, da die Zeitreihen ja eigentlich bereits mithilfe von Funktionen modelliert wurden.
-Zudem wird nicht interpoliert, abgeleitet o.
-Ä.
-...
-letztlich für die aktuelle Idee des Papers eher ungeeignet (plus man umgeht Probleme wie mögliche Glättungsartefakte, oder die Definition eines functional mean).
-- Falls doch FDA, dann evtl. als eigenständiges Skript forken
-- Begründung warum Prädiktionsintervalle anstelle von Konfidenzintervallen
 <!-- - Umbennung entire curve in "All-or-nothing" (entlehnt von Juul et al.) -->
+<!-- - Toleranzintervalle mit reinnehmen? -->
+<!-- - ROISLIEN durch Toleranzintervalle ersetzen! -->
 - .bib file anlegen
 - Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert)
-- Ratkowsky besorgen
-- Begründung warum Prädiktionsintervalle und nicht Konfidenzintervalle (theoretisch könnte man beides machen; Prädiktionsintervalle scheinen mir aber eher das zu sein, was man will)
-
-### TODO read (again)
-- Parker et al. (2021) https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
-- Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis
-- Olshen et al. (1989) Gait analysis and the bootstrap
-- Curvewise point and interval summaries for tidy data frames of draws from distributions https://mjskay.github.io/ggdist/reference/curve_interval.html
-- Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem
-- Generell die Kooperationspaper von Hörmann und Kokoszka zu FDA Statistik,insbesondere 'Inference for Functional Data with Applications' (2012)
-- Ratkowsky 1990 Handbook of nonlinear regression models
+
+<!-- ### TODO read (again) -->
+<!-- - White Basics of Estimating Measurement Uncertainty -->
+<!-- - Francq B, Govaerts B. How to regress and predict in a Bland-Altman plot? Review and contribution based on tolerance intervals and -->
+<!-- correlated-errors-in-variables models. Statist Med. 2016;35:2328-2358. -->
+<!-- - Francq Confidence, prediction, and tolerance in linear mixed models -->
+<!-- - Vock (2016) -->
+<!-- - Parker et al. (2021) https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x -->
+<!-- - Robinson, Vanrenterghem, Pataky (2021) Sample size estimation for biomechanical waveforms: Current practice, recommendations and a comparison to discrete power analysis -->
+<!-- - Olshen et al. (1989) Gait analysis and the bootstrap -->
+<!-- - Curvewise point and interval summaries for tidy data frames of draws from distributions https://mjskay.github.io/ggdist/reference/curve_interval.html -->
+<!-- - Horvath, Kokoszka & Reeder (2013) Estimation of the mean of functional time series and a two-sample problem -->
+<!-- - Generell die Kooperationspaper von Hörmann und Kokoszka zu FDA Statistik,insbesondere 'Inference for Functional Data with Applications' (2012) -->
+<!-- - Ratkowsky 1990 Handbook of nonlinear regression models -->
 
@@
 library(ggplot2)
-library(boot)
+# library(boot)
 
@@
 
-<!-- # Alternatives Intro -->
-
-<!-- SHORT NOTE!? -->
-
-<!-- Ansatz: Neben der Validierung von Messgeräten spielt auch die Validierung von Modellergebnissen eine Rolle! -->
-
-<!-- 1.
-Es gibt einen Unterschied zwischen Konfidenz- und Prädiktionsintervallen: Prediction intervals are a prediction of a future observation, while confidence intervals represent a predicted mean response.
--->
-
-<!-- <!-- https://stats.stackexchange.com/questions/16493/difference-between-confidence-intervals-and-prediction-intervals --> -->
-<!-- The difference between a prediction interval and a confidence interval is the standard error. -->
-<!-- The standard error for a confidence interval on the mean takes into account the uncertainty due to sampling.
-The line you computed from your sample will be different from the line that would have been computed if you had the entire population, the standard error takes this uncertainty into account.
--->
-<!-- The standard error for a prediction interval on an individual observation takes into account the uncertainty due to sampling like above, but also takes into account the variability of the individuals around the predicted mean.
-The standard error for the prediction interval will be wider than for the confidence interval and hence the prediction interval will be wider than the confidence interval.
--->
-
-<!-- Für die Validierung von Messinstrumenten sind aus unserer Sicht eher Prädiktionsintervalle interessant! Daher beziehen sich alle folgenden Aussagen ausschließlich auf Prädiktionsintervalle. -->
-
-<!-- 2.
-Punktweise Prädiktionsbänder (und Konfidenzbänder) sind gegenüber simultatnen zu eng (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
-Daher wurden in der Vergangenheit Korrekturverfahren (e.
-.g.
-Lenhoff, Bonferroni, Šidák, projection, sup-t bands) vorgeschlagen -->
-
-<!-- <!-- See Mirzargar et al. (2014) or Juul et al. (2020) for an accessible introduction to data depth and curve boxplots / functional boxplots. --> -->
-
 # Introduction
 
-Having a good estimate of the amount of measurement error is a basic requirement when analyzing movement data.
-The classic linear error model describes measurement error as the sum of systematic and random deviations from the true value.
-A complete characterization of the measurement error, however, is complex, as different sources of variance (within, between) and noise must be considered.
-The analysis of sensor data is further complicated by the fact that the output of many measurement systems in biomechanics - such as e.
-g.
-joint angles, ground reaction forces, or center of mass trajectories - is available in the form of continuous signals (curve data) within which the error can vary over time.
-
-Therefore, when analyzing the validity of curve data, the entire length of the curve should be accounted for.
-This can be accomplished e.
-g.
-by applying statistics separately to all points of the curve [Pini et al.
-(2019).
-The pointwise approach allows for an error representation over the entire domain, but ignores the fact, that the sampling points within a curve are locally correlated (Deluzio & Astephen, 2007; Pataky, 2010).
-From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
-
-Es gibt verschiedene Ansätze dieses Problem zu adressieren.
-Was bislang fehlt ist eine systematische Validierung und Gegenüberstellung verschiedener Methoden.
-Im Paper sollen daher wichtige Methoden zur Konstruktion kontinuierlicher Übereinstimmnugsbänder anhand repräsentativer Datensätze (synthetisch und real-world) verglichen werden und deren Vor- und Nachteile gegenübergestellt werden.
-<!-- Zudem sollen relevante methodische Aspekte diskutiert (z. B. Prädiktions- vs. Konfidenzbänder für die Validierung von Messsystemen) -->
-
-Bei allen etablierten Vorzuügen funktionaler/kontinuerlicher Verfahren ist aktuell zu konstatieren, dass diese Methoden in der angewandten Forschung im Bereich Biomechanik/Bewegungswissenschafte kaum Anwendung finden.
-Wir spekulieren, dass für die praktische Umsetzung Fähigkeiten in den Bereichen Statistik und Programmierung erforderlich sind, die bei vielen "Praktikern" nicht ausreichend vorhanden sind.
-Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die im Paper analysierten Methoden als frei verfügbaren R Code zur Verfügung.
-
-<!-- <!-- ... and likely results in a systematic underrepresentation of ... values (Juul et al., 2020; Mirzagar et al. (2014). --> -->
-
-<!-- <!-- https://mjskay.github.io/ggdist/reference/curve_interval.html --> -->
-<!-- <!-- Hier referenzierte Paper u. a.  --> -->
-<!-- <!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
-(2020).
-"Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles".
-arXiv e-print.
-arXiv:2007.05035 --> -->
-<!-- <!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
-(2014).
-"Curve Boxplot: Generalization of Boxplot for Ensembles of Curves".
-IEEE Transactions on Visualization and Computer Graphics.
-20(12): 2654-2663.
-doi: 10.1109/TVCG.2014.2346455 --> -->
-
-<!-- From a statistical point of view, this implies that the random error component is likely underrepresented when the (nonzero) covariance term of adjacent sampling points is not accounted for (Lenhoff et al., 1999).
-In addition, most traditional point statistics are either not suitable (e.g.
-statistical tests of mean differences), or are only suitable to a limited extent (e.g.
-product-moment correlation) to quantify the agreement between two measurement systems (Bland and Altman, 1986).
--->
-
-<!-- One possibility to address the problem of pointwise analysis is to treat entire curves as (functional) objects rather than as a series of unconnected points. -->
-
-<!-- <!-- Es gibt eine Reihe von Ansätzen aus anderen Bereichen (z. B. Covid Prädiktion), bei denen sog. Function oder Curve Boxplots eingesetzt werden. --> -->
-<!-- <!-- https://mjskay.github.io/ggdist/reference/curve_interval.html --> -->
-<!-- <!-- Hier referenzierte Paper u. a.  --> -->
-<!-- <!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
-(2020).
-"Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles".
-arXiv e-print.
-arXiv:2007.05035 --> -->
-<!-- <!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
-(2014).
-"Curve Boxplot: Generalization of Boxplot for Ensembles of Curves".
-IEEE Transactions on Visualization and Computer Graphics.
-20(12): 2654-2663.
-doi: 10.1109/TVCG.2014.2346455 --> -->
-
-<!-- Sutherland et al.
-(1988) introduced a method in which continuous signals are reconstructed using Fourier series and bootstrapped to estimate of the variation of the sample mean and thus estimate confidence intervals.
+Die Beurteilung der Validität biomechanischer Messverfahren erfolgt üblicherweise über den Vergleich eines 'neuen' Messsystems mit einem etablierten.
+Dabei werden Bewegungen mit zwei Messsystemen simultan aufgezeichnet und Messwerte miteinander verglichen, die theoretisch den gleichen Wert aufweisen sollten.
+Die statistische Aufbereitung der zwangsläufig zu beobachtenden Abweichungen ist dabei aus verschiedenen Gründen nicht trivial.
+Eine besondere Herausforderung stellt die Tatsache dar, dass der Output der meisten Messsysteme in Form höherdimensionaler Daten (z.
+B.
+1D Kurvendaten wie EMG-Signale oder COP Verläufe, 2D plantare Druckmatrizen oder 3D bone strain fields) vorliegt, innerhalb derer der Messfehler variiert.
+Der Fokus dieses Papers liegt auf 1D Kurvendaten, die den häufigsten Anwendungsfall in biomechanischen Bewegungsanalysen darstellen.
+
+In den allermeisten Validierungsstudien wird der Zeitreihencharakter zugunsten einfacher(er) Modelle vernachlässigt und Kurvendaten auf einzelne diskrete (0D)Punkte oder skalare Features (z.
+B.
+local extrema, rates of change, or ranges of motion) reduziert.
+Daraus ergeben sich eine Reihe von Problemen für die Bewertung des Messfehlers: (i) non-stationary error processes are not considered (ii) the validity of statements derived from discrete variables is limited by domain alignment (i.
+e.
+phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) biomechanische Daten, die die Bewegung von Körpersegmenten direkt messen, sind durch die fundamentalen Viskositätseigenschaften biologischer Gewebe (Fung, 1981) und Freiheitsgrade menschlicher Gelenkbewegungen charakterisisiert.
+Beispielsweise sind im Verlauf einer Gelenkwinkelkurve keine Sprünge zu erwarten.
+Aus statistischer Sicht bedeutet das, dass die Samplingpunkte innerhalb einer Kurve keine lose Ansammlung von Einzelpunkten darstellen, sondern lokal korreliert sind (Deluzio & Astephen, 2007; Pataky, 2010).
+Bei der einfachen Übertragung punktweiser Statistiken auf alle Samplingpunkte der Kurve ist damit zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird, da der (nonzero) covariance term of adjacent sampling points nicht berücksichtigt wird (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
+Die üblicherweise zur Korrektur beim multiplen Testen eingesetzten Korrekturverfahren zur Anpassung des alpha levels (Hochberg & , 1987) führen aufgrund der hohen Anzahl an Datenpunkten pro Kurve ($t \le 100$) zu sehr konservativen Schätzungen weshalb dieses Vorgehen hier nicht zielführend ist.
+Ein besserer Ansatz ist es Kurven direkt als zusammenhängende Objekte zu betrachten.
+
+Sutherland, Olshen, Biden & Wyatt (1988, 1989) (BOOT) haben für Gelenkwinkeldaten im Kontext von Ganganalysen Ende der 1980er Jahre eine Methode vorgestellt, bei der Kurvendaten mithilfe eines random (Fourier) coefficient Regressionsmodells approximiert werden und aus den so gebildeten funktionalen Objekte anschließend durch Bootstrapping Prädiktionsbänder für normales Gehen gebildet werden.
 Lenhoff et al.
-(1999) illustrated their method on a generic gait analysis data set (comprised of joint angle curves) and compared it to pointwise Gaussian intervals.
-The authors found that Sutherland et al.
-bootstrap bands provide appropriate coverage for continuous curve gait data (86% coverage for a targeted coverage of 90%) while pointwise Gaussian bands are shown to provide inadequate coverage (54% for a targeted coverage of 90%).
--->
-
-<!-- <!-- In other words, a single 0.95-CI for each one of the quantities of interest is computed and, in this sense, they are called pointwise CIs.
-However, if the target is the estimation of a vector of values (such as the CDF or the return level function evaluated in a grid of points, for instance), more accurate than computing the corresponding pointwise CIs for each one of the quantities of interest is to construct simultaneous CIs (also called a confidence band).
-Simultaneous CIs are a group of CIs (a CI for each one of the quantities of interest) designed to jointly contain this vector of unknown values with a prescribed high probability (95%, for instance).
-In this way, it would be possible to make a joint probabilistic assessment for some quantiles of the corresponding CDF.
-Because simultaneous CIs include information on model reliability, they are are much more informative than pointwise CIs [14].
---> -->
-
-<!-- Røislien et al.
-(2012) presented a similar approach to calculate a functional version of the the Limits of Agreement (LoA) method suggested by Bland & Altman.
-They first calculate functional difference curves which are then bootstrapped to obtain 95% confidence intervals.
-The approach accounts for the autocorrelation characteristic in the signal and at the same time allows for an intuitive description of the systematic and random error terms in the original sampling space (e.g.
-actual degrees).
-Røislien et al., however, rely on a number of strong model assumptions, which are often not or only partially fulfilled in empirical data sets: -->
-
-<!-- (i) The error distribution is approximately normal. -->
-
-<!-- The construction of symmetric prediction bands as described in Røislien et al.
-[10] relies on Gaussian assumptions.
-Non-normality, however, is often present when e.g.
-sensor values are truncated because of limited range or resolution, or because values drift over time.
--->
-
-<!-- (ii) Time-continuous differences of two measurement systems and the corresponding averages must be uncorrelated. -->
-
-<!-- This assumption is violated considerably in many validation data sets [12].
-E.g., Atallah et al.
-[13] found that the agreement between an ear-worn accelerometer and a force-plate instrumented treadmill was associated with gait speed when estimating gait cycle duration.
-When validating a foot-worn gyroscope for treadmill running, Koska & Maiwald [14] observed a linear error trend that was associated with foot strike behavior.
--->
-
-<!-- (iii) Aus einem Datensatz bestehend aus einer Kurve eines Probanden lassen sich durch Bootstrapping von mittleren Kurven Unsicherheitsintervalle bilden, die ein geeignetes Maß für die estimation uncertainty darstellen.
--->
-
-<!-- Hier werden arithmetische mean curves verwendet, the variability of which will reflect variation between means.
-In many cases, however, it is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of the number of curves within the sample.
--->
-
-<!-- Ein weiteres Problem ist, dass die Methode keine messwiederholten Daten berücksichtigt.
-Da sich messwiederholte Kurven i.
-d.
-R.
-ähnlicher sind als unabhängige Kurven, besteht die Gefahr die Streuung des Messsystems zu überschätzen.
--->
-
-<!-- <!-- Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
-werden 95% Konfidenzintervalle gebootstrapped.
-The authors are producing the arithmetic mean curves, the variability of which will reflect variation between means.
-It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of n.subj.
---> -->
-
-<!-- <!-- Formulierung aus dem alten Skript --> -->
-<!-- <!-- (iii) Measurement uncertainty can be estimated adequately by bootstrapping the error variance from a sample that consists of a single curve per subject. --> -->
-<!-- <!-- If no repeated measurements are included in the sample, an essential error component (intra-individual variance) is ignored.
-The bootstrap – like any other statistical procedure – cannot solve this problem, as it can only estimate proportions of variance that are contained in the sample.
---> -->
-
-<!-- <!-- Weiterer ("zugehöriger) Reviewerkommentar: --> -->
-<!-- <!-- In looking at limits of agreement, the authors are considering the extent of within pair variation.
-However, in looking at the point wise limits of agreement, these are defined to incorporate a between pair standard deviation.
-Even if it was appropriate to utilise between pair variability, the authors should note that it is variances that are additive and not SDs.
-Perhaps the authors' lack of understanding comes from their erroneous assumption (iii).
---> -->
-
-<!-- <!-- Olsen et al.
-[15] examined the validity of an inertial measurement unit (IMU) for analyzing the spatial displacement of the distal limbs of horses using functional LoA (FLoA).
-Their approach differs from Røislien et al.
-[10] in that initially linear mixed models were fitted to the repeated measurements of different horses.
-Functional data estimates of the mean trends in the strides were then determined from the model fits and FLoA were calculated.
-This procedure accounts for repeated measurements and eliminates the need to emulate empirical curves with Fourier series or splines.
-The FLoA computed this way performed well in comparison with traditional Bland & Altman LoA.
-The procedure, however, is subject to the same restrictions as any conventional linear regression model and is therefore limited when model assumptions are violated.
---> -->
-
-<!-- <!-- Warum eigentlich Prädiktionsbänder und keine Konfidenzbänder? (Siehe Einleitung Lenhoff) --> -->
-<!-- <!-- Aktuell würde ich es so formulieren: In der Mehrheit der Fälle ist man an der nächsten (Einzel)Observation interessiert, nicht unbedingt an Mittelwerten (obwohl es durchaus Szenarien gibt, wo das akzeptabel sein kann) + im Zweifel generell lieber breitere Fehlerintervalle --> -->
-
-<!-- The principal object of this paper was to demonstrate a method for constructing continuous prediction intervals from curve data that addresses the aforementioned issues (analyzing the whole curve, including repeated measures, model assumptions (normal distribution, equal variances)), and thus enables a more appropriate characterization of the agreement between measurement curves on the basis of easy to interpret error bands (i.e.
-LoA in the actual measurement unit).
--->
-
-<!-- <!-- Es gibt bereits Ansätze für funktionale Boxplots (inkl. Vergleich mit punktweisen Intervallen), das Ganze wurde aber noch nie für Differenzenkurven untersucht. --> -->
-
-<!-- We use synthetic time series data sets containing typical curves of biomechanical measurement systems to illustrate the method and compare the resulting prediction bands to pointwise Bland & Altman parameters and LoA constructed similar to the method described in Roislien et al.
-(2012).
-For model validation, the respective LoA are cross validated using a leave-one-out approach.
--->
-
-<!-- <!-- Aus Bland and Altman (2007): If each pair of X and Y measurements is treated as if from a different individual the structure of the data is ignored and incorrect estimates are  --> -->
-<!-- <!-- Ähnliche Formulierung aus https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x : In the repeated measures case, applying the standard limits of agreement to the data will result in limits that are too narrow because they do not take into account the reduction in variability that arises when working with averages of readings.
---> -->
+(1999) illustrated the same method on a generic gait analysis data set (comprised of joint angle curves) and compared it to pointwise Gaussian intervals.
+By means of a leave-one out cross validation they found that the Sutherland prediction bands provide appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands are shown to provide inadequate coverage (54%).
+Røislien et al.
+(2012) haben einen Ansatz präsentiert, mit dem die Methode von Bland & Altman (1983) für kontinuierliche Daten erweitert wird.
+Der Ansatz von Bland & Altman ist die am weitesten verbreitete Methode in Methodenvergleichen (Ludbrook, 2010) und verwendet die Differenzen zwischen zwei Messmethoden für die Konstruktion von Limits Agreement (LoA).
+
+Das vorliegende Paper strebt eine systematische Validierung verschiedener Methoden zur Konstruktion kontinuierlicher Fehlerintervalle bei Methodenvergleichen an.
+Dazu sollen drei verschiedene Methoden anhand simulierter und realer Datensätze charakteristischer glatter Kurven biomechanischer Messysteme (Gelenkwinkelverläufe von zwei Messsystemen) verglichen werden: (1) punktweise Bland and Altman LoA, (2) LoA nach Roislien et al.
+und (3) gebootstrappte funktionale Prädiktionsbänder (Sutherland, Olshen, Biden & Wyatt (1988, 1989)).
+Da sowohl die Intervalle von Sutherland als auch Roislien im Wesentlichen Prädiktionsintervalle repräsentieren, soll der Methodenvergleich exemplarisch am Beispiel kontinuierlicher Prädiktionsbänder vollzogen werden.
+Die Methode lässt sich jedoch prinzipiell auch auf andere Intervallmethoden wie Konfidenz- oder Toleranzintervalle übertragen.
+Zur Beurteilung der Intervallgüte werden die qualitative Verläufe der Intervallgrenzen der jeweiligen Verfahren gegen die tatsächlichen Differenzen geplottet und die entsprechenden coverage properties berechnet.
+Für die Beurteilung der Intervallgüte wird im Rahmen einer leave-one out cross validation die Unsicherheit in den berechneten LoA angegeben.
+
+Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
+In diesem Sinne stellen wir auch den zugehörigen R Code im Anhang zur freien Verfügung.
 
@@
 
-## Limits of agreement methods
-
-FLoA derived by different methods are compared
-
-* Randomized Cluster Bootstrap (FLoA~RCB~)
-  + One random stride per subject (FLoA~RCB_v2~)
-  + Fetch a SINGLE random stride from all strides (FLoA~RCB_v3~)
-
-* Roislien-like intervals (FLoA~Roslien~)
-  + Select one random stride from each subject ONCE and bootstrap the resulting sample
-
-Pointwise continuous LoA are constructed using mean +- 1.96 * SD
-
-* Pointwise Gaussian intervals (FLoA~Point~)
-  + FLoA~Point~ are calculated using using mean +- 1.96 * SD (standard deviations derived from linear mixed effects models)
-
-Linear mixed effects models (lme4 package (Bates, Maechler, Bolker, & Walker, 2015)) were used because of the advantages over fixed-effects models (e.
-g.
-ANOVA as suggested by Bland & Altman, 1999), such as e.
-g.
-(i) missing or unbalanced data are less problematic and (ii) inference can be made based on a wider population of patients (Parker et al., 2020).
-Implementation details are provided in the R code within the supplementary material.
-
-### FLoA~RCB~ construction
-
-<!-- See Mirzargar et al. (2014) or Juul et al. (2020) for an accessible introduction to data depth and curve boxplots / functional boxplots. -->
+### Data sets
+
+Für den Vergleich der Methoden wurden vier (simulierte und reale) Datensätze verwendet.
+Jeder der Datensätze enthält jeweils $N = 220$ Kurven der Länge $T = 101$ Datenpunkte aus $M = 2$ Messsystemen $(m=1,2)$.
+Das entspricht in jedem Datensatz $D = 110$ Differenzkurven, die durch Subtraktion gebildet werden:
+
+$\sum D_{ij} = \sum_{i=1}^{I} \sum_{j=1}^{J} Y_{ij} - X_{ij} = 110$
+
+X und Y sind (I x J) Matrizen der beiden Messsysteme, mit $I = 11$ Probanden und $J = 10$ wiederholten Kurven pro Proband.
+Die Größe der Datensätze entspricht in etwa dem typischen Stichprobenumfang in Validierungsstudien.
+
+<!-- Die $nN$ Kurven entsprechen $J = 10$ repeated measures von $I = 11$ Probanden und $m = 2$ Messsystemen: -->
+<!-- $n = m * \sum_{i=1}^{11} \sum_{j=1}^{10} n_{ij} = 220$ -->
+
+Alle simulierten Datensätze wurden als glatte Kurven, ähnlich Gelenkwinkelverläufen, mit dem gleichen Grundmodell generiert.
+Die Kurven korrespondieren nicht mit realen Signalen und wurden teilweise bewusst überspitzt konstruiert um den Einfluss bestimmter Differenzcharakteristika auf das Verhalten der Bänder möglichst deutlich darzustellen.
+Die Kurven wurden als trigonometrische Funktionen durch Addition zweier Sinusfunktionen konstruiert:
+
+$$
+X_{ij}(t) = (offset_{m}) + f_{1} + f_{2}, Y_{ij}(t) = (offset_{m}) + f_{3} + f_{4}
+$$
+mit
+
+$$
+f_{1, 3} = Asin(Bt)^{c + 3}
+$$
+und
+
+$$
+f_{2, 4} = Asin(Bt)
+$$
+$A$, $B$ sind Zufallsvariablen mit unterschiedlichen Verteilungseigenschaften (normal, uniform, Weibull) über deren Parameter systematische und zufällige Unterschiede zwischen Probanden (between subject variation), Schritten (within subject variation) und Messystemen modelliert werden.
+$c$ ist eine Konstante ($c = 2$).
+Zudem wurden - je nach Datensatz - unterschiedliche Offsets als normalverteilte Zufallsvariablen mit variierenden $\mu$ und $\sigma$ simuliert:
+
+$$
+offset_{m} = \sim N(\mu, \sigma_{i})
+$$
+Im Sinne der Reproduzierbarkeit wurden alle Zufallsvariablen mit einem random seed initialisiert. Der Code für die Generierung der Kurven kann im Anhang im Detail nachvollzogen werden.
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+data <- example_data(dat = "smooth_realistic", dir.data)
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT.1 <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 14),
+        axis.title.x = element_text(size = 14),
+        axis.text.y = element_text(size = 14),
+        axis.title.y = element_text(size = 14),
+        legend.position = "none")
+
+data <- example_data(dat = "non_gaussian", dir.data)
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT.2 <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 14),
+        axis.title.x = element_text(size = 14),
+        axis.text.y = element_text(size = 14),
+        axis.title.y = element_text(size = 14),
+        legend.position = "none")
+
+data <- example_data(dat = "shift", dir.data)
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT.3 <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 14),
+        axis.title.x = element_text(size = 14),
+        axis.text.y = element_text(size = 14),
+        axis.title.y = element_text(size = 14),
+        legend.position = "none")
+
+data <- example_data(dat = "imu_mc", dir.data)
+data.single.mc <- subset(data, device == "MC")
+data.single.imu <- subset(data, device == "IMU")
+
+PLOT.4 <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
+  geom_line() +
+  scale_color_grey(start = 0.6, end = 0.9) +
+  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
+  labs(x = "Time-normalized signal [%]", y = "Value") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(size = 14),
+        axis.title.x = element_text(size = 14),
+        axis.text.y = element_text(size = 14),
+        axis.title.y = element_text(size = 14),
+        legend.position = "none")
+```
+
+```{r echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Figure 1: Kurvenverläufe in den jeweiligen Datensätzen (A - D). Schwarze Kurven im Hintergrund sind die Werte des Referenzsystems."}
+PLOT <- ggpubr::ggarrange(PLOT.1, PLOT.2, PLOT.3, PLOT.4,
+                          labels = c("A", "B", "C", "D"),
+                          ncol = 2,
+                          nrow = 2)
+PLOT
+```
+
+Fig. 1 stellt die originalen Kurven beider Messsysteme in den vier verwendeten Datensätzen (A - D) dar:
+
+* Datensatz A : Simulated curves with Gaussian error model ($\epsilon_{i} \sim N (0, \sigma_{i}^2)$).
+
+Die Daten repräsentieren den Fall, dass die Kurven zweier Messsysteme qualitativ ähnlich sind und kein relevanter Bias vorliegt.
+
+* Datensatz B : Simulated curves with non-Gaussian erros and heteroskedasticity
+
+Der Skalierungsparameter des zweiten Messsystems ist als normalverteilte Zufallsvariable mit Weibull-verteiltem Mittelwert modelliert ($\epsilon_{i} \sim N(\mu \sim Wei(\gamma = 1.5,\alpha  = 1), \sigma_{i})$).
+Wie in (A) liegt kein Bias vor.
+
+Solche Daten liegen vor, wenn mindestens eines der Messsysteme Kurven produziert, die qualitativ und quantitativ deutlich vom wahren Wert abweichen (hier: Messsystem 2, grau).
+In diesem Fall sind elementare Modellvoraussetzungen für parametrische Verfahren wie Bland and Altman erheblich verletzt.
+
+* Datensatz C : Simulated curves curves with Gaussian error model and phase shift in x-axis direction.
+
+Die Kurven wurden mit dem selben Modell wie in (A) erstellt, aber die Kurven von Messsystem 2 (grau) sind auf der x-Achse verschoben.
+Auch hier sind Modellvoraussetzungen erheblich verletzt, die Differenzen weisen aber eine andere Verteilung auf als in Modell (B).
+In der Praxis treten vergleichbare Kurvenverläufe auf, wenn die Messung eines Systems in ihrem zeitlichen Verlauf in irgendeiner Weise systematisch beeinflusst wird, z.
+B.
+wenn die Messsysteme infolge der Platzierung am Körper unterschiedliche Bewegungen aufzeichnen, oder durch Filterartefakte.
+
+* Datensatz D : Reale Messdaten aus der Validierung einer inertialen Messeinheit (IMU)
+
+Real-world hip joint angle curves from healthy subjects walking at 6 km/h on a treadmill without gradient recorded simultaneously using an optical motion capture system (MC) and IMU.
+Die Kurven unterscheiden sich qualitativ und in der absoluten Ausprägung der Differenzen.
+
+### Calculation methods
+
+Um die untersuchten Methoden im gleichen Wahrscheinlichkeitsraum (probability space, function space?) zu analysieren, wurden Differenzkurven zwischen den beiden Systemen als Berechnungsgrundlage gebildet.
+Im Kontext von Messsystemvalidierungen bietet sich die Verwendung von Differenzkurven zudem an, weil mit diesen Kurven unmittelbar die (von Bland & Altman formulierte) Fragestellung beantwortet wird, inwiefern die verglichenen Methoden in der Praxis austauschbar eingesetzt werden können?
+
+Auf Grundlage der Differenzkurven wurden Prädiktionsintervalle mit drei unterschiedlichen Methoden berechnet:
+
+* 1. POINT : Pointwise continuous LoA according to Bland & Altman (1999, 2007)
+
+Für alle $t = 101$ Punkte wurden jeweils über alle $d$ Kurven Bland & Altman LoA berechnet.
+Wenn diese Differenzen normalverteilt (homoskedastisch, unabhängig) sind, kann erwartet werden, dass 95% der Differenzen im Intervall mean difference $\pm$ 1.96 $*$ standard deviation of the differences liegen.
+Dieses einfache Modell kann erweitert werden wenn die Modellvoraussetzungen nicht ausreichend erfüllt sind (z.
+B.
+nichtnormalverteilte Differenzen (logarithmische Transformation), Abhängigkeit der Differenzen vom zugehörigen Mittelwert (Regressionsansatz), Abhängigkeit der Kurven (messwiederholte Daten)).
+
+Wir verwenden in der vorliegenden Arbeit ein Modell für wiederholte Messungen, in dem sich der wahre Wert über jedes Differenzpaar ändert.
+Das Modell beschreibt die Differenz zwischen zwei Messsystemen als Summe der mittleren Differenz ($\overline{d}$) und zweier Varianzkomponenten (between (\sigma^2_{db}), within (\sigma^2_{dw})):
+<!-- Auch wenn hier jeweils die gleiche Anzahl Kurven für jedes Messsystem vorliegt, erlaubt die verwendete Formel die Übertragung auf den unbalancierten Fall. -->
+
+$$
+LoA = D_{ij} = \overline{d} + \sqrt{\sigma^2_{db} + \sigma^2_{dw}}
+$$
+Die beiden Varianzkomponenten können mithilfe einer einfaktoriellen Varianzanalyse bestimmt werden.
+\sigma^2_{dw} wird dabei als residual mean square ($MS_{w}$) direkt aus dem ANOVA table abgelesen.
+\sigma^2_{db} berechnet sich aus der Differenz aus between-subject mean squares und $MS_{w}$, geteilt durch einen Term, der die Anzahl der $i$ Probanden und $j$ Kurven berücksichtigt.
+Wenn für alle Probanden die gleiche Anzahl an Datenpunkten vorliegt (wie im vorliegenden Fall), reduziert sich dieser Term auf die Anzahl der Datenpunkte (Kurven) pro Proband ($j = 10$):
+
+$$
+\sigma^2_{db} = \frac{(MS_{b} - MS_{w})}{j}
+$$
+
+Für weiterführende Informationen zur Berechnung (z. B. bei ungleicher Anzahl Observationen pro Proband), siehe Bland & Altman (1999) sowie die zugehörige Implementierung als R Code im Anhang.
+
+* 2. ROISLIEN : Functional limits of agreement (FLoA) according to Roislien et al. (2012)
+
+Die Methode von Røislien et al.
+(2012) wurde als Erweiterung der Methode von Bland & Altman (1983) für kontinuierliche Differenzen ($d(t)$) entwickelt.
+Roislien et al.
+approximieren die Kurven im Datensatz daher zunächst mithilfe elementarer trigonometrischer Funktionen.
+Die Berechnung der Modellparameter (mean, standard deviation) erfolgt im Anschluss jedoch - genau wie in POINT - separat für jeden Punkt der Kurve.
+Daher macht es in der praktischen Umsetzung de facto keinen Unterschied ob die originalen (diskreten) oder funktionalen Kurven verwendet werden.
+Wir verzichten daher im Rahmen dieses Papers auf die Repräsentation der Kurven als Funktionen.
+Damit ergeben sich gegenüber POINT zwei Unterschiede:
+
+(i) Da die Unabhängigkeit der Kurven vorausgesetzt wird, werden keine wiederholten Kurven verwendet, sondern je eine Kurve pro Proband ausgewählt.
+Es wird also nicht wie bei POINT mit allen Kurven gerechnet, sondern mit einer reduzierten Version des Datensatzes.
+
+(ii) Die (punktweise) Berechnung der LoA erfolgt für diese $j = i$ Kurven mithilfe des minimalen Bland & Altman Modells als $FLoA = d(t) \pm 1.96SD(t)$ berechnet, wobei $\overline{d}(t)$ die kontinuierliche mean difference bezeichnet.
+
+* 3. BOOT : Boostrapped functional prediction bands according Sutherland et al. (1988), Olshen et al. (1989) and Lenhoff et al. (1999)
+
+Die Methode
+
+Das Bootstrapping der Prädiktionsintervalle wird nicht mit diskreten Werten gemessener Kurven durchgeführt, sondern mit Funktionen der Kurven.
+Die Funktionen werden durch Kurvenanpassung mit Fourier-Reihen erzeugt.
+Kurvenanpassung ist notwendig, da diskrete Messwerte eines Kurvenverlaufs keine unabhängig voneinander auftretenden Ereignisse sind, die einer Normalverteilung unterliegen.
+
+Im ersten Schritt des Verfahrens werden alle gemessenen Kurven (n) mithilfe von Fourier-Reihen als Funktionen modelliert:
+
+$$
+f(t) = \mu_{i} + \sum_{k=1}^K(\alpha_{i, k}cos(\frac{2\pi kt}{T}) + \beta_{i, k} sin(\frac{2\pi kt}{T}))
+$$
+Die Freiheitsgrade $K$ sind dabei so zu wählen, dass die Kurven ausreichend gut angepasst werden.
+In der vorliegenden Arbeit wurden die Kurven großzügig mit K = 50 approximiert.
+Die Koeffizienten der Kurven werden durch die Methode der kleinsten Quadrate angenähert.
+Für das weitere Vorgehen ist dabei wichtig, dass $\mu$ für alle Kurven 0 sein muss, da die Standardabweichungskurve der Kurvenschar sonst einen Offset aufweist.
+Durch die Kurvenanpassung wird für jede Kurve ein Vektor mit $2k+1$ Koeffizenten erzeugt:
+
+$$
+W_{i} = (\mu_{i}, \alpha_{i, 1}, \beta_{i, k}, ..., \beta_{i, K})
+$$
+Die Vektoren werden in einer Matrix M mit den Dimensionen n (Zeilen) und 2k+1 (Spalten) zusammengefasst.
+Durch Mittelwertbildung und Berechnung der Varianz über die Spalten können jetzt Vektoren der Mittelwert- (W) und der Varianzkoeffizienten (Σw) berechnet werden.
+Durch Multiplikation von W (Formel 18) und Σw (Formel 19) mit
+
+$$
+l(t) = (1, cos(\frac{2\pi 1t}{T}, sin(\frac{2\pi 1t}{T}, ..., cos(\frac{2\pi Kt}{T}, sin(\frac{2\pi Kt}{T},))
+$$
+lassen sich die Mittelwertkurve
+
+$$
+\hat{f}(t) = \overline{W}^T l (t)
+$$
+und der Verlauf der Standardabweichung
+
+$$
+\hat{\sigma}_{f(t)} = \sqrt(l(T)^T \sum_{w} l(t))
+$$
+der Kurvenschar anpassen.
+
+Anschließend werden aus der Matrix M die Koeffizienten von n Kurven im
+Bootstrappingverfahren b mal gezogen (Ziehen mit Zurücklegen).
+Jede Bootstrap-Wiederholung erzeugt ein 'bootstrap sample' (Bootstrap-Ziehung) entsprechend der ursprünglichen Kurvenanzahl n.
+Aus den Koeffizienten der n Kurven jeder Bootstrap-Ziehung werden mit
+
+$$
+\hat{f}
+$$
+
+und
+
+$$
+
+$$
+
+$$
+P \{\max_{t} (\frac{|\hat{f}_{n+1}(t) - \hat{f}(t)|}{\hat{\sigma}_{f(t)}} \le C_{p}\} = 1-\alpha
+$$
+
+$$
+\max_{\{P^{\circ}_t(j)\}}
+$$
+
+1. Fourier
+2. Bootstrap
+3. Prädiktionsbänder
+
+Let fn + 1(t) denote a future draw from the population mean curve that is defined by Eq.
+(0.1) and let Yn + 1(t) be an observed curve generated by Eqs. (0.2) and (0.3). Given a desired confidence level 100(1− a)%, we chose the constant Cp so that
+
+95% prediction bands
+
+The bootstrap method does not rely on any parametric assumption concerning the distribution of the data at each point of the curve.
+Bootstrap prediction and confidence intervals can be theoretically justified asymptotically although their performance in small-sample settings has to be studied on a case-by-case basis.
+
+400 bootstrap iterations
+
+Für weitere Details zur Methode siehe Lenhoff et al. (1999), Appendix A, sowie die zugehörige Implementierung als R Code im Anhang.
 
@@
 
-## Validation
-
-#### Data generating process
-To compare different methods for constructing continuous LoA, several sets of smooth curve data were simulated across a range of common biomechanical signals and error characteristics.
-The quality of the curves is similar, as the paper is primarily developed for a validation context in which the same movement process is analyzed using similar methods (two measurement devices or models).
-For the same reason, the curves do not necessarily correspond to real-world signals as we were more concerned with illustrating the respective methods and error characterstics.
-
-All curves were modeled as Fourier series from additively superimposed sine functions.
-Each data set contains n.strides = 10 curves from n.subj = 11 simulated subjects.
-Each curve consists of t = 101 data points.
-
-<!-- Aus Robinson et al. (2021) -->
-<!-- Exemplar 1D effects from the biomechanics literature across a range of signal types were collated and reported (Bakke & Besier, 2020, Barrios & Willson, 2017, Bovi, Rabuetti, Mazzoleni, & Ferrarin, 2011, Phan et al., 2017, Gomes, Ackermann, Ferreira, Orselli, & Sacco, 2017, Robinson et al., 2014).
--->
-
-<!-- the code to reproduce all 1D effects is provided at https://github.com/m-a-robinson/sample-size -->
-
-<!-- Robinson et al. (2014) Impact of Knee Modeling Approach on Indicators and Classification of Anterior Cruciate Ligament Injury Risk -->
-
-<!-- * (4) Data with (missing) peaks.
-Same model as in (i), except a couple of frames where a short, prominent peak occurs in one of the two signals.
-This typically happens in high-frequency signals such as accelerations or ground reaction forces, where peaks are underrepresented or entirely missing when e.
-g.
-sampling rates are too low or lowpass filter artefacts occur.
-Also, modelling these kind of abrupt changes in the signal (when e.
-g.
-to estimate ground reaction forces from acceleration values) is more difficult and my result in similar errors (Alcantara, 2021).
--->
-<!-- <!-- Alcantara et al. Modelling/estimating ground reaction forces from acceleration values, z. B. https://www.biorxiv.org/content/10.1101/2021.03.17.435901v1.full --> -->
-<!-- <!-- Johnson et al. (2020) Multidimensional ground reaction forces and moments from wearable sensor accelerations via deep learning --> -->
-
-#### Smooth wave data (constant variance, no trend) with subjectwise bias
-Homoscedastic (normal) error.
-Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
-The data are representative of e.
-g.
-joint angle measurements simultaneously recorded from two kinematic measurement systems (e.
-g.
-a camera-based measurement systems and a goniometer).
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_realistic", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
-
-#### Smooth curves (homoscedasticity, normal error, no trend)
-Homoscedastic (normal) error.
-Within and between subject differences between curves were modeled using random systematic offset (bias) and random variation of curve parameters.
-The data are representative of e.
-g.
-joint angle measurements simultaneously recorded from two kinematic measurement systems (e.
-g.
-a camera-based measurement systems and a goniometer).
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
-
-#### Smooth curves with nonlinear error trend
-Same model as in (i), but with a nonlinear trend underlying one of the two signals.
-Similar errors can be observed e.
-g.
-when joint angles are calculated (numerically integrated) from gyroscope signals.
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_trend", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
-
-#### Heteroscedastic measurement error (no trend)
-Smooth curves created using the same model and parameters as in (i), but Weibulll distributed random curve (amplitude) parameters.
-This introduces non-gaussian measurement error (asymmetric/skewed distribution of differences around the mean difference).
-<!-- Smooth curves with non-gaussian (Weibull) error (no trend) -->
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "non_gaussian", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
-
-#### Curves phase shifted in x-axis direction
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "shift", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
-
-#### Real-world validation data
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "imu_mc", dir.data)
-
-data.single.mc <- subset(data, device == "MC")
-data.single.imu <- subset(data, device == "IMU")
-
-PLOT <- ggplot(data = data.single.mc, aes(x = frame, y = value, group = strideID)) +
-  geom_line() +
-  scale_color_grey(start = 0.6, end = 0.9) +
-  geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Signal value") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(size = 20),
-        axis.title.x = element_text(size = 18),
-        axis.text.y = element_text(size = 20),
-        axis.title.y = element_text(size = 18),
-        legend.position = "none")
-
-PLOT
-```
-
-### Leave-one (curve) out cross validation
-Leave-one (curve) out method to estimate the achieved coverage
-
-Prediction intervals are calculated from size.dataset = number.subjects - 1 and then validated against the left out curve for four different methods:
-
-* FLoA~RCB_v2~   : One random stride per subject
-
-* FLoA~RCB_v3~   : Fetch a SINGLE random stride from all strides
-
-* FLoA~Roislien~ : Roislien approach (Get one random stride from each subject ONCE and bootstrap the resulting sample)
-
-* FLoA~Point~ : Pointwise B & A Limits of Agreement
-
-#### Coverages
-Coverages are displayed for two different scenarios:
-
-* a: Percentage of entirely covered curves (all points within the limits)
-
-* b: Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
-
-### LoA uncertainty
-In order to quantify the uncertainty of the calculated LoA across different methods, LoA were calculated n.boot times and compared using uncertainty bands.
+### Leave-one out cross validation
+
+Um die coverage Performance der untersuchten Methoden zu bewerten, wurden die jeweiligen Intervalle im leave-one-out Verfahren kreuzvalidiert.
+Dabei wird eine Kurve vom Datensatz entfernt und die Intervalle anhand der übrigen Kurven berechnet.
+Dieser Vorgang wird für jede der $n$ Kurven eines Datensatzes wiederholt.
+Der Anteil der vorgehaltenen Kurven, die in den Bändern enthalten sind, wird mit zwei verschiedenen Kennwerten beschrieben:
+
+* Percentage of entirely covered curves (all points within the limits)
+
+* Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
+
+Für die punktweisen wurde
+
+### LoA uncertainty estimation
+
+Um die Schwankung der Intervallgrenzen über wiederholte Berechnungen hinweg zu veranschaulichen, wurden für die BOOT und ROISLIEN Intervalle Unsicherheitsbänder gebootstrapped (b = 400 Wiederholungen).
+Die Unsicherheit in diesen Verfahren entsteht durch das wiederholte Ziehen mit Zurücklegen im Rahmen des Bootstrapverfahrens (BOOT) bzw.
+die jeweils unterschiedlichen gezogenen Einzelkurven (ROISLIEN).
+
+Im Falle der POINT Intervalle werden alle Kurven eines Datensatzes verwendet, weshalb keine Schwankung der Intervallgrenzen auftritt.
+Verschiedene Autoren schlagen jedoch Konfidenzintervalle als Erweiterung ihres ursprünglichen Ansatzes vor, um den sampling error insbesondere kleinere Stichproben (Bland & Altman, 1986; Hamilton & Stamey, 2007; Giavarina, 2015) zu berücksichtigen.
+Bland & Altman (1986) schätzen den Standardfehler der Differenzen auf ungefähr $\sqrt{3*s² / c}$.
+
+mit $s$ ... Standardabweichung
+und $c$ ... Anzahl der Kurven im Datensatz
+
+Für die Berechnung von Konfidenzintervallen wird noch das geeignete Quantil der t-Verteilung mit n − 1 Freiheitsgraden benötigt.
+Das Konfidenzintervall reicht dann vom oberen/unteren Limit minus t$*$Standardfehler bis zum oberen/unteren Limit plus t$*$Standardfehler.
+Im vorliegenden Paper wurden 95% Konfidenzintervalle gerechnet.
+
+Die vollständige von Bland & Altman vorgeschlagene Formel (95% CI für 95% LoA) lautet damit:
+
+$(\overline{d} \pm z_{0.975}*s) \pm t_{0.975, n-1}*s*\sqrt{3*s² / c}$, mit $z_{0.975} = 1.96$
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-n.boot <- 100
-```
-
-#### Smooth wave data (constant variance, no trend) with subjectwise bias
+n.boot <- 30
+```
+
+#### Simulated curves with Gaussian error model (A)
 
@@
 floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
-
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
-```
-
-#### Smooth curves (homoscedasticity, normal error, no trend)
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth", dir.data)
-
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
-
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
-```
-
-#### Smooth curves with nonlinear error trend
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_trend", dir.data)
-
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
-
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
-```
-
-#### Heteroscedastic measurement error (no trend)
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "non_gaussian", dir.data)
-
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
-
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-10, 10))
-```
-
-#### Curves phase shifted in x-axis direction
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "shift", dir.data)
-
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
-
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
-```
-
-#### Real-world validation data
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "imu_mc", dir.data)
-
-floa.point <- floa_point(data)
-floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
-
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-15, 15))
+floa.lenhoff.pred  <- floa_lenhoff(data,
+                                   k_reihe = 50,
+                                   n.boot = n.boot,
+                                   band = "prediction",
+                                   cp.begin = 0,
+                                   alpha = 0.05)
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-5, 5))
+```
+
+#### Simulated curves with non-normal, heteroskedastic differences (B)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# data <- example_data(dat = "non_gaussian", dir.data)
+# 
+# floa.point <- floa_point(data)
+# floa.roislien <- floa_roislien(data)
+# floa.lenhoff.pred  <- floa_lenhoff(data,
+#                                    k_reihe = 50,
+#                                    n.boot = n.boot,
+#                                    band = "prediction",
+#                                    cp.begin = 0,
+#                                    alpha = 0.05)
+# 
+# plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-10, 10))
+```
+
+#### Simulated curves curves with Gaussian error model created and phase shift in x-axis direction (C)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# data <- example_data(dat = "shift", dir.data)
+# 
+# floa.point <- floa_point(data)
+# floa.roislien <- floa_roislien(data)
+# floa.lenhoff.pred  <- floa_lenhoff(data,
+#                                    k_reihe = 50,
+#                                    n.boot = n.boot,
+#                                    band = "prediction",
+#                                    cp.begin = 0,
+#                                    alpha = 0.05)
+# 
+# plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-5, 5))
+```
+
+#### Real-world validation data (hip joint angles)
+
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# data <- example_data(dat = "imu_mc", dir.data)
+# 
+# floa.point <- floa_point(data)
+# floa.roislien <- floa_roislien(data)
+# floa.lenhoff.pred  <- floa_lenhoff(data,
+#                                    k_reihe = 50,
+#                                    n.boot = n.boot,
+#                                    band = "prediction",
+#                                    cp.begin = 0,
+#                                    alpha = 0.05)
+# 
+# plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-15, 15))
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_realistic", dir.data)
-
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
+# data <- example_data(dat = "smooth_realistic", dir.data)
+# 
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+# 
+# # Calculate percent from counts
+# calculate_percent <- function(data) {
+#   percentage <- (sum(data) / length(data))
+#   percentage <- round(percentage, digits = 2)
+# }
+# 
+# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+# covered.curves.percent <- data.frame(t(covered.curves.percent))
+# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-
-cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-)
-colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-
-# # Display as boxplots across methods
-# plot_cov_ver(cover.cross.fraction.singlecurve)
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+# 
+# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+# )
+# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+# 
+# # # Display as boxplots across methods
+# # plot_cov_ver(cover.cross.fraction.singlecurve)
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth", dir.data)
-
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
+# data <- example_data(dat = "smooth", dir.data)
+# 
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+# 
+# # Calculate percent from counts
+# calculate_percent <- function(data) {
+#   percentage <- (sum(data) / length(data))
+#   percentage <- round(percentage, digits = 2)
+# }
+# 
+# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+# covered.curves.percent <- data.frame(t(covered.curves.percent))
+# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "smooth_trend", dir.data)
-
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+# data <- example_data(dat = "smooth_trend", dir.data)
+# 
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+# 
+# # Calculate percent from counts
+# calculate_percent <- function(data) {
+#   percentage <- (sum(data) / length(data))
+#   percentage <- round(percentage, digits = 2)
+# }
+# 
+# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+# covered.curves.percent <- data.frame(t(covered.curves.percent))
+# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-
-cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-)
-colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+# 
+# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+# )
+# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "non_gaussian", dir.data)
-
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
+# data <- example_data(dat = "non_gaussian", dir.data)
+# 
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+# 
+# # Calculate percent from counts
+# calculate_percent <- function(data) {
+#   percentage <- (sum(data) / length(data))
+#   percentage <- round(percentage, digits = 2)
+# }
+# 
+# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+# covered.curves.percent <- data.frame(t(covered.curves.percent))
+# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-
-cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-)
-colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-
-# # Display as boxplots acroos methods
-# plot_cov_ver(cover.cross.fraction.singlecurve)
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
-# plot_cov_ver(cover.cross.fraction)
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+# 
+# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+# )
+# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+# 
+# # # Display as boxplots acroos methods
+# # plot_cov_ver(cover.cross.fraction.singlecurve)
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "shift", dir.data)
-
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
+# data <- example_data(dat = "shift", dir.data)
+# 
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+# 
+# # Calculate percent from counts
+# calculate_percent <- function(data) {
+#   percentage <- (sum(data) / length(data))
+#   percentage <- round(percentage, digits = 2)
+# }
+# 
+# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+# covered.curves.percent <- data.frame(t(covered.curves.percent))
+# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
 ```
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-data <- example_data(dat = "imu_mc", dir.data)
-
+# data <- example_data(dat = "imu_mc", dir.data)
+# 
+# # Leave-one curve out approach
+# # ------------------------------------------------------------------------------
+# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+# 
+# # Calculate percent from counts
+# calculate_percent <- function(data) {
+#   percentage <- (sum(data) / length(data))
+#   percentage <- round(percentage, digits = 2)
+# }
+# 
+# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+# covered.curves.percent <- data.frame(t(covered.curves.percent))
+# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
+```
+
+b. Mean (SD) of covered curve points
+```{r echo = FALSE, warning = FALSE, message = FALSE}
 # Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-
-# Calculate percent from counts
-calculate_percent <- function(data) {
-  percentage <- (sum(data) / length(data))
-  percentage <- round(percentage, digits = 2)
-}
-
-covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-covered.curves.percent <- data.frame(t(covered.curves.percent))
-colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross <- crossval_coverage(data, n.boot)
-# # colnames(cover.cross) <- c("RCB_1", "RCB_2", "RCB_3", "Roislien_4", "Point_5")
-# 
-# plot_cov_ver(cover.cross)
-```
-
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# ------------------------------------------------------------------------------
-cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-
-cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-             round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-)
-colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-
-knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-
-# # Display as boxplots acroos methods
-# plot_cov_ver(cover.cross.fraction.singlecurve)
-
-# # Leave-one subject out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction <- crossval_coverage_fraction(data, n.boot)
-# plot_cov_ver(cover.cross.fraction)
+# # ------------------------------------------------------------------------------
+# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+# 
+# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
+#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
+# )
+# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
+# 
+# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
+# 
+# # # Display as boxplots acroos methods
+# # plot_cov_ver(cover.cross.fraction.singlecurve)
 ```
@@
 
-Punktweise FLoA unterrepräsentieren den Messfehler
+Punktweise FLoA unterrepräsentieren den Messfehler sind aber ganz ok.
+Allerdings gilt es zu konstatieren, dass die korrekte Berechnung (ANOVA-Varianzanteile, Konfidenzintervalle) auch nicht unbedingt trivial ausfällt.
 
@@
 
-<!-- Anwendung der Argumentation der Reviewer* auf die "Roislien_FLoA": In Roislien et al.
-werden 95% Konfidenzintervalle gebootstrapped.
-The authors are producing the arithmetic mean curves, the variability of which will reflect variation between means.
-It is the variability of individual values that is relevant, and the limits will be too narrow by a factor of approximately the square root of n.subj.
+Im Allgemeinen sind Prädiktionsintervalle überall dort von Interesse, wo neue Messwerte gegen normative Daten (z.
+B.
+aus einer Referenzdatenbank) verglichen werden.
+Wenn die Abweichung ein bestimmtes Maß überschreitet, d.
+h.
+der Proband nicht zur Referenzpopulation gehört, dann werden - wie bei Olshen et al.
+(1989) - bestimmte pathologische Muster angenommen und eine - wie auch immer geartete - Therapie vorgeschlagen.
+Entsprechend wichtig ist es diese Intervalle mit der richtigen Breite zu konstruieren, andernfalls gehen etwaige Effekte im Rauschen unter (zu weit) oder es werden Effekte angenommen, die so gar nicht auftreten (zu eng).
+
+Die Bland & Altman Methode setzt normalverteilte, homoskedastische und unabhängige Differenzen voraus.
+Diese Annahmen sind allerdings beim Vergleich verschiedener Messmethoden häufig verletzt.
+Die Verteilung der Differenzen zwischen zwei Systemen wird z.
+B.
+erheblich verzerrt, wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder der Auflösung abgeschnitten werden, oder wegen (nichtlinearem) Drift.
+Die Unabhängigkeitsannahme wird dann verletzt, wenn messwiederholte Daten vorliegen.
+Meswiederholte Werte sind sich i.
+d.
+R.
+ähnlicher (d.
+h.
+sie streuen weniger) als die Kurven verschiedener Probanden.
+Entsprechend besteht die Gefahr, die Streuung der Differenzen zwischen zwei Messmethoden zu unterschätzen.
+Um dem entgegenzuwirken, verwenden Roislien nur jeweils eine Kurve je Proband, was wiederum zu Lasten einer erheblichen Unsicherheit bezüglich der Streuung der berechneten LoA erkauft wird.
+
+KONFIDENZINTERVALLE VS. PRÄDIKTIONSINTERVALLE
+<!-- Prädiktionsbänder sind kontinuierliche Intervalle, in denen eine zukünftige Beobachtung einer Zufallsvariablen (Messkurve) mit einem gegebenen Konfidenzniveau enthalten ist.
+Anhand dieser Prädiktionsbänder werden Entscheidungen über die Form und den Effekt von Behandlungen für einzelne Probanden getroffen.
+Zu enge Prädiktionsbänder implizieren im beschriebenen Anwendungsfall, dass der Gang eines Probanden ein von der Norm abweichendes Muster aufweist, dass behandelt werden muss.
 -->
-d
-It might be beneficial to work with functional data as suggested by Lenhoff et al.
-(1999) or Roislien et al.
-(2012) when e.
-g.
-interpolating data to a standardized length.
-This approach, however, wasn't necessary in this paper, since the synthetic data were created from functions and were designed to all have equal length (101 data points).
-
-Wir empfehlen die Daten zu plotten.
+Der hier beschriebene Anwendungsfall (Prädiktionsbänder) deckt natürlich nur ein gewisses Spektrum relevanter Fragestellungen ab.
+In vielen weit verbreiteten Studiendesigns, z.
+B.
+Interventionsstudien, ist mit den breiten Prädiktionsbändern oder den noch breiteren Toleranzbändern nicht viel zu holen.
+Mit den im Paper kontruierten Bändern würde nahezu jeder Effekt im Messfehlintervall untergehen.
+In solchen Fällen Studien sind aber ohnehin eher Mittelwerte von Interesse, die in Signifikanztests auf Unterschiede untersucht werden.
+In diesem Fall sind die schmaleren Konfidenzintervalle eher geeignet um die Abweichung/Übereinstimmung zwischen den gemessenen Mittelwertkurven zweier Messsysteme zu beschreiben.
+<!-- Für Details zu Unterschieden in den coverage properties von LoA und herkömmlichen Prädiktionsintervallen sei an dieser Stelle auf Francq et al. (2020) verwiesen. -->
+
+<!-- LENHOFF -->
+<!-- In some research studies it is important to determine confidence bands for the mean curve of a given population.
+For example, a researcher might want to characterize the effect upon the mean external varus moment of a group of subjects treated with a given type of knee brace.
+In this example, the researcher would wish to determine a confidence band for the mean difference in the external varus moment for subjects while wearing the brace and the external varus moment for the same subjects while not wearing the brace.
+If the confidence -->
+<!-- band about the difference of the braced and un-braced curves of each subject contains the horizontal line at value zero, then one could state that there was no significant difference between the two conditions.
+This application is illustrated in Fig.
+1.
+-->
+
+https://statmodeling.stat.columbia.edu/2010/12/05/what_do_practit/
+The key assumptions of a regression model are validity and additivity.
+Except when you’re focused on predictions, don’t spend one minute worrying about distributional issues such as normality or equal variance of the errors.
+
+Wir empfehlen grundsätzlich die Daten zu plotten.
 Am Ende ist jeder Datensatz bzw.
 jede Fragestellung unterschiedlich und ein pauschaler coverage value wird der Qualität des Messgeräts (in Bezug auf die jeweilige Fragestellung möglicherweise nicht gerecht).
+
+Die Erkenntnisse der vorliegenden Studie können nicht 1:1 auf andere Arten von Daten (z. B. Bodenreaktionskraftkurven, EMG Signale) übertragen, sondern gelten primär für glatte Kurven.
+
+SPM
+Pataky: A methodology called Statistical Parametric Mapping (SPM) (Friston et al., 2007) can partially offset these limitations by providing a framework for the continuous statistical analysis of smooth bounded nD fields.
+
+Trotz der theoretisch und praktisch demonstrierten Überlegenheit funktionaler Verfahren konnten sich deren Einsatz im Kontext der Validierung von Messinstrumenten noch nicht recht durchsetzen.
+Wir spekulieren, dass dies in direktem Zusammenhang mit der Komplexität der Verfahren steht, die eine hohe Hürde für viele Praktiker mit weniger ausgeprägten Statistik- und Programmierkenntnissen darstellt.
+Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code frei zur Verfügung (GPL).
+
+Angesichts der überragenden Bedeutung der Kenntniss des Messfehlers ist es erstaunlich wie wenig sich viele Studien mit der Frage nach der Größe des Messfehlers beschäftigen.
+Gut demonstrieren lässt sich das am Beispiel von IMU's ...
+
+Wir glauben dass die geringe Verbreitung kontinuierlicher Methoden trotz erwiesenermaßen gut funktionierender Methoden in der Praxis an der programmiertechnischen Umsetzung scheitert.
+Wir hoffen mit diesem Paper und dem Code im Anhang einen Beitrag dazu zu leisten, dass mehr Forscher diese Methoden adaptieren und veralteten/ungeeigneten Methoden (Korrelationskoeffizienten etc.) den Rücken kehren.
+
+<!-- Es gibt auch immer mehr readily available (und kostenlose) Lösungen, siehe https://mjskay.github.io/ggdist/articles/lineribbon.html -->
 
@@
 
+<!-- <!-- Es gibt eine Reihe von Ansätzen aus anderen Bereichen (z. B. Covid Prädiktion), bei denen sog. Function oder Curve Boxplots eingesetzt werden. -->
+<!-- <!-- https://mjskay.github.io/ggdist/reference/curve_interval.html -->
+<!-- <!-- Hier referenzierte Paper u. a.  -->
+<!-- <!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
+(2020).
+"Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles".
+arXiv e-print.
+arXiv:2007.05035 -->
+<!-- <!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
+(2014).
+"Curve Boxplot: Generalization of Boxplot for Ensembles of Curves".
+IEEE Transactions on Visualization and Computer Graphics.
+20(12): 2654-2663.
+doi: 10.1109/TVCG.2014.2346455 -->
+
+<!-- <!-- See Mirzargar et al. (2014) or Juul et al. (2020) for an accessible introduction to data depth and curve boxplots / functional boxplots. -->
+
 # Conclusion

### 2021-10-15 paper.Rmd
Continued paper draft

@@
 ---
-title: "Comparison of continuous prediction bands for the validation of smooth 1D biomechanical time series data"
-
+title: Comparison of continuous prediction bands for the validation of smooth 1D biomechanical
+  time series data
 author:
@@
   email: daniel.koska@hsw.tu-chemnitz.de
-    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
+  affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis
+    in Biomechanics
   footnote: Corresponding Author
@@
   email: christian.maiwald@hsw.tu-chemnitz.de
-    affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis in Biomechanics
-# address:
-#   - code: Chemnitz University of Technology
-#     address:  Thüringer Weg 11, 09126 Chemnitz
-#   - code: Another University
-#     address: Department, Street, City, State, Zip
+  affiliation: Chemnitz University of Technology, Research Methodology and Data Analysis
+    in Biomechanics
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+journal: An awesome journal
 abstract: |
@@
 
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-journal: "An awesome journal"
-# bibliography: sendaFAB.bib
-# output: rticles::elsevier_article
+header-includes:
+  - \usepackage{amsmath}
 output:
   html_document:
-    toc: true
-    toc_float: true
+    toc: yes
+    toc_float: yes
     theme: united
+  pdf_document:
+    toc: yes
 ---
@@
 
-$\sum D_{ij} = \sum_{i=1}^{I} \sum_{j=1}^{J} Y_{ij} - X_{ij} = 110$
+\begin{equation}
+\tag{1}
+\sum D_{ij} = \sum_{i=1}^{I} \sum_{j=1}^{J} Y_{ij} - X_{ij} = 110
+\label{eq:mean}
+\end{equation}
 
 X und Y sind (I x J) Matrizen der beiden Messsysteme, mit $I = 11$ Probanden und $J = 10$ wiederholten Kurven pro Proband.
-Die Größe der Datensätze entspricht in etwa dem typischen Stichprobenumfang in Validierungsstudien.
+Die Größe der Datensätze entspricht in etwa dem typischen Stichprobenumfang vieler Validierungsstudien.
 
@@
 $$
-X_{ij}(t) = (offset_{m}) + f_{1} + f_{2}, Y_{ij}(t) = (offset_{m}) + f_{3} + f_{4}
+X_{ij} = (offset_{m}) + f_{1}(t) + f_{2}(t),\\Y_{ij}(t) = (offset_{m}) + f_{3}(t) + f_{4}(t)
 $$
@@
 $$
-f_{1, 3} = Asin(Bt)^{c + 3}
+f_{1, 3}(t) = Asin(Bt)^{c + 3}
 $$
@@
 $$
-f_{2, 4} = Asin(Bt)
+f_{2, 4}(t) = Asin(Bt)
 $$
@@
 $$
-offset_{m} = \sim N(\mu, \sigma_{i})
+offset_{m} \sim N(\mu_{i}, \sigma=0.05)
 $$
@@
 
-Um die untersuchten Methoden im gleichen Wahrscheinlichkeitsraum (probability space, function space?) zu analysieren, wurden Differenzkurven zwischen den beiden Systemen als Berechnungsgrundlage gebildet.
-Im Kontext von Messsystemvalidierungen bietet sich die Verwendung von Differenzkurven zudem an, weil mit diesen Kurven unmittelbar die (von Bland & Altman formulierte) Fragestellung beantwortet wird, inwiefern die verglichenen Methoden in der Praxis austauschbar eingesetzt werden können?
-
-Auf Grundlage der Differenzkurven wurden Prädiktionsintervalle mit drei unterschiedlichen Methoden berechnet:
+Die untersuchten Prädiktionsbänder wurden ausschließlich anhand der gebildeten Differenzkurven $D$ gebildet um die Vergleichbarkeit der Methoden zu gewährleisten.
+Prädiktionsbänder wurden mit drei unterschiedlichen Verfahren berechnet:
 
@@
 
-Für alle $t = 101$ Punkte wurden jeweils über alle $d$ Kurven Bland & Altman LoA berechnet.
-Wenn diese Differenzen normalverteilt (homoskedastisch, unabhängig) sind, kann erwartet werden, dass 95% der Differenzen im Intervall mean difference $\pm$ 1.96 $*$ standard deviation of the differences liegen.
-Dieses einfache Modell kann erweitert werden wenn die Modellvoraussetzungen nicht ausreichend erfüllt sind (z.
-B.
-nichtnormalverteilte Differenzen (logarithmische Transformation), Abhängigkeit der Differenzen vom zugehörigen Mittelwert (Regressionsansatz), Abhängigkeit der Kurven (messwiederholte Daten)).
-
-Wir verwenden in der vorliegenden Arbeit ein Modell für wiederholte Messungen, in dem sich der wahre Wert über jedes Differenzpaar ändert.
-Das Modell beschreibt die Differenz zwischen zwei Messsystemen als Summe der mittleren Differenz ($\overline{d}$) und zweier Varianzkomponenten (between (\sigma^2_{db}), within (\sigma^2_{dw})):
+Für alle $t = 101$ Punkte wurden jeweils über alle $D$ Kurven LoA berechnet.
+Bei erfüllten Modellvoraussetzungen (unabhängige Kurven, normalverteilte Differenzen, Homoskedastizität), kann erwartet werden, dass 95% der Differenzen im Intervall mean difference $\pm$ 1.96 $*$ standard deviation of the differences liegen.
+Wir verwenden in der vorliegenden Arbeit ein komplexeres Modell für wiederholte Messungen, in dem sich der wahre Wert über jedes Differenzpaar ändert (Bland & Altman, 1999).
+Das Modell beschreibt die Differenz zwischen zwei Messsystemen als Summe der mittleren Differenz ($\overline{d}$) und zweier Varianzkomponenten (between ($\sigma^2_{db}$), within ($\sigma^2_{dw}$)):
 <!-- Auch wenn hier jeweils die gleiche Anzahl Kurven für jedes Messsystem vorliegt, erlaubt die verwendete Formel die Übertragung auf den unbalancierten Fall. -->
@@
 Die beiden Varianzkomponenten können mithilfe einer einfaktoriellen Varianzanalyse bestimmt werden.
-\sigma^2_{dw} wird dabei als residual mean square ($MS_{w}$) direkt aus dem ANOVA table abgelesen.
+\sigma^2_{dw} kann als residual mean square ($MS_{w}$) direkt aus dem ANOVA table abgelesen werden.
 \sigma^2_{db} berechnet sich aus der Differenz aus between-subject mean squares und $MS_{w}$, geteilt durch einen Term, der die Anzahl der $i$ Probanden und $j$ Kurven berücksichtigt.
@@
 $$
-
-Für weiterführende Informationen zur Berechnung (z. B. bei ungleicher Anzahl Observationen pro Proband), siehe Bland & Altman (1999) sowie die zugehörige Implementierung als R Code im Anhang.
+Für weiterführende Informationen zur Berechnung (z. B. bei ungleicher Anzahl Observationen pro Proband) siehe Bland & Altman (1999, 2007) sowie die zugehörige Implementierung als R Code im Anhang.
 
@@
 approximieren die Kurven im Datensatz daher zunächst mithilfe elementarer trigonometrischer Funktionen.
-Die Berechnung der Modellparameter (mean, standard deviation) erfolgt im Anschluss jedoch - genau wie in POINT - separat für jeden Punkt der Kurve.
+Die Berechnung der Modellparameter ($\overline{d}$, $\sigma$) erfolgt im Anschluss allerdings - genau wie in POINT - separat für jeden Punkt der Kurve.
 Daher macht es in der praktischen Umsetzung de facto keinen Unterschied ob die originalen (diskreten) oder funktionalen Kurven verwendet werden.
 Wir verzichten daher im Rahmen dieses Papers auf die Repräsentation der Kurven als Funktionen.
-Damit ergeben sich gegenüber POINT zwei Unterschiede:
-
-(i) Da die Unabhängigkeit der Kurven vorausgesetzt wird, werden keine wiederholten Kurven verwendet, sondern je eine Kurve pro Proband ausgewählt.
-Es wird also nicht wie bei POINT mit allen Kurven gerechnet, sondern mit einer reduzierten Version des Datensatzes.
-
-(ii) Die (punktweise) Berechnung der LoA erfolgt für diese $j = i$ Kurven mithilfe des minimalen Bland & Altman Modells als $FLoA = d(t) \pm 1.96SD(t)$ berechnet, wobei $\overline{d}(t)$ die kontinuierliche mean difference bezeichnet.
+Das Modell ist dennoch interessant weil gegenüber der Variante in POINT (i) keine wiederholten Kurven verwendet werden, sondern nur je eine Kurve pro Proband um die Modellvoraussetzung der Unabhängigkeit der Kurven nicht zu verletzen und (ii) die (punktweise) Berechnung der LoA für diese $j = i$ Kurven mithilfe des minimalen Bland & Altman Modells erfolgt ($LoA = d(t) \pm 1.96SD(t)$), wobei $\overline{d}(t)$ die kontinuierliche mean difference bezeichnet.
 
@@
 
-Die Methode
-
-Das Bootstrapping der Prädiktionsintervalle wird nicht mit diskreten Werten gemessener Kurven durchgeführt, sondern mit Funktionen der Kurven.
-Die Funktionen werden durch Kurvenanpassung mit Fourier-Reihen erzeugt.
-Kurvenanpassung ist notwendig, da diskrete Messwerte eines Kurvenverlaufs keine unabhängig voneinander auftretenden Ereignisse sind, die einer Normalverteilung unterliegen.
-
-Im ersten Schritt des Verfahrens werden alle gemessenen Kurven (n) mithilfe von Fourier-Reihen als Funktionen modelliert:
-
-$$
-f(t) = \mu_{i} + \sum_{k=1}^K(\alpha_{i, k}cos(\frac{2\pi kt}{T}) + \beta_{i, k} sin(\frac{2\pi kt}{T}))
-$$
-Die Freiheitsgrade $K$ sind dabei so zu wählen, dass die Kurven ausreichend gut angepasst werden.
-In der vorliegenden Arbeit wurden die Kurven großzügig mit K = 50 approximiert.
-Die Koeffizienten der Kurven werden durch die Methode der kleinsten Quadrate angenähert.
-Für das weitere Vorgehen ist dabei wichtig, dass $\mu$ für alle Kurven 0 sein muss, da die Standardabweichungskurve der Kurvenschar sonst einen Offset aufweist.
-Durch die Kurvenanpassung wird für jede Kurve ein Vektor mit $2k+1$ Koeffizenten erzeugt:
-
-$$
-W_{i} = (\mu_{i}, \alpha_{i, 1}, \beta_{i, k}, ..., \beta_{i, K})
-$$
-Die Vektoren werden in einer Matrix M mit den Dimensionen n (Zeilen) und 2k+1 (Spalten) zusammengefasst.
-Durch Mittelwertbildung und Berechnung der Varianz über die Spalten können jetzt Vektoren der Mittelwert- (W) und der Varianzkoeffizienten (Σw) berechnet werden.
-Durch Multiplikation von W (Formel 18) und Σw (Formel 19) mit
-
-$$
-l(t) = (1, cos(\frac{2\pi 1t}{T}, sin(\frac{2\pi 1t}{T}, ..., cos(\frac{2\pi Kt}{T}, sin(\frac{2\pi Kt}{T},))
-$$
-lassen sich die Mittelwertkurve
-
-$$
-\hat{f}(t) = \overline{W}^T l (t)
-$$
-und der Verlauf der Standardabweichung
-
-$$
-\hat{\sigma}_{f(t)} = \sqrt(l(T)^T \sum_{w} l(t))
-$$
-der Kurvenschar anpassen.
-
-Anschließend werden aus der Matrix M die Koeffizienten von n Kurven im
-Bootstrappingverfahren b mal gezogen (Ziehen mit Zurücklegen).
-Jede Bootstrap-Wiederholung erzeugt ein 'bootstrap sample' (Bootstrap-Ziehung) entsprechend der ursprünglichen Kurvenanzahl n.
-Aus den Koeffizienten der n Kurven jeder Bootstrap-Ziehung werden mit
-
-$$
-\hat{f}
-$$
-
-und
-
-$$
-
-$$
-
-$$
-P \{\max_{t} (\frac{|\hat{f}_{n+1}(t) - \hat{f}(t)|}{\hat{\sigma}_{f(t)}} \le C_{p}\} = 1-\alpha
-$$
-
-$$
-\max_{\{P^{\circ}_t(j)\}}
-$$
-
-1. Fourier
-2. Bootstrap
-3. Prädiktionsbänder
-
-Let fn + 1(t) denote a future draw from the population mean curve that is defined by Eq.
-(0.1) and let Yn + 1(t) be an observed curve generated by Eqs. (0.2) and (0.3). Given a desired confidence level 100(1− a)%, we chose the constant Cp so that
-
-95% prediction bands
-
-The bootstrap method does not rely on any parametric assumption concerning the distribution of the data at each point of the curve.
-Bootstrap prediction and confidence intervals can be theoretically justified asymptotically although their performance in small-sample settings has to be studied on a case-by-case basis.
-
-400 bootstrap iterations
-
-Für weitere Details zur Methode siehe Lenhoff et al. (1999), Appendix A, sowie die zugehörige Implementierung als R Code im Anhang.
+Die Methode ist die einzige der vorgestellten Methoden, die vollständig funktional arbeitet, d.
+h.
+die Konstruktion der Prädiktionsbänder findet nicht punktweise, sondern auf Grundlage ganzer Kurven statt.
+
+Bootstrapping ist ein nicht-parametrisches Verfahren und kommt daher ohne Annahmen zur Verteilung der Differenzen aus.
+
+Im ersten Schritt des Verfahrens werden alle $N$ gemessenen Kurven mithilfe von Fourier-Reihen als Funktionen approximiert:
+
+$$
+f_{n}(t) = \mu_{n} + \sum_{k=1}^K(\alpha_{nk}cos(\frac{2\pi kt}{T}) + \beta_{nk} sin(\frac{2\pi kt}{T}))
+$$
+Die Anzahl der Fourier-Basisfunktionen $K$ ist dabei so zu wählen, dass die Kurven bestmöglich angenähert werden und keine ungewollte Glättungseffekte entstehen.
+In der vorliegenden Arbeit wurde großzügig $K = 50$ gewählt.
+Die Koeffizienten ($\mu_{n}, \alpha_{n1}, \beta_{n1}, ..., \alpha_{nK}, \beta_{nK}$) werden durch die Methode der kleinsten Quadrate bestimmt.
+
+asserts that the curve starts and ends at the same height.
+<!-- Für das weitere Vorgehen ist dabei wichtig, dass $\mu$ für alle Kurven 0 sein muss, da die Standardabweichungskurve der Kurvenschar sonst einen Offset aufweist. -->
+
+Die Koeffizienten werden in einer Matrix $M$ mit den Dimensionen $N$ (Zeilen) und $2K+1$ zusammengefasst.
+Aus dieser Matrix werden durch Bootstrapping $B = 400$ mal $N$ Kurven aus $M$ mit Zurücklegen gezogen, für die jeweils die mittlere Kurve $\hat{f^b}(t)$ und Kovarianzmatrix $\hat{\sigma_{\hat{f^b}(t)}}$ bestimmt werden.
+
+Im Mittel über alle Bootstrapiterationen lässt sich daraus der Anteil der Kurven bestimmen, deren maximale standardisierte Abweichung vom Bootstrap-Mittel ($\hat{f^b}(t)$) kleiner gleich $C_{p}$ beträgt.
+
+$$
+\frac{1}{B} \sum_{b=1}^B [\frac{1}{N} \sum_{n=1}^N I(\max_{t} (\frac{|\hat{f}_{n}(t) - \hat{f^b}(t)|}{\hat{\sigma}_{f(t)}} \le C_{p})]
+$$
+wobei $I(E)$ eine Matrix ist die Werte von entweder 0 ($E$ tritt nicht auf) oder 1 ($E$ tritt auf) annimmt.
+
+$C_{p}$ muss dabei so optimiert werden, dass es der gewünschten coverage probability des Prädiktionsbandes $P = 1- \alpha$ (hier: $\alpha=0.05$) entspricht.
+Damit ist berechnet sich das Prädiktionsband als:
+
+$$
+\hat{f}(t) \pm C_{p} * \hat{\sigma}_{f(t)}
+$$
 
@@
 
+Stichprobengröße: Es kann natürlich gut sein, dass die punnktweisen Verfahren bei höheren Stichprobenumfängen besser performen.
+
+Die Berechnung der LoA ist mit bestimmten (parametrischen) Modellvoraussetzungen verbunden, dass ist aber am Ende eigentlich egal, weil: Was mache ich denn wenn die verletzt sind?
+
+Wir können natürlich nicht mit vier Datensätzen die gesamte Bandbreite abdecken, obwohl damit schon sehr viele Anwendungsfälle abgedeckt werden.
+Weiterhin ist der Anwendungsbereich auf eine spezielle Art von Kurven (glatte Kurve) beschränkt.
+Kurven mit vergleichbaren Eigenschaften werden von vielen biomechanischen Messsystemen erzeugt:  Bestimmte Bodenreaktionskräfte (horizontal), tiefpassgefilterte Signale, COP Verläufe etc..
+Es gibt aber auch Kurven die ganz andere Charkteristiken wie z.
+B.
+Sprünge/Unstetigkeiten im Verlauf (vertikale Bodenreaktionskurven), hohe Rauschanteile (z.
+B.
+EMG, ungefilterte Beschleunigungssignale), nichtstationäres Verhalten (z.
+B.
+Drift bei intergrierten Gyrosignalen) aufweisen.
+Es ist nicht klar inwiefern die untersuchten Methoden auch bei diesen Signalen brauchbare Ergebnisse liefern.
+Gerade bei Signalen mit sprunghaften Änderungen ist denkbar dass punktweise Intervalle zu plausibleren Ergebnissen führen können.
+<!-- Lenhoff geht z. B. nicht für nichtstionäre Signale! -->
+
 KONFIDENZINTERVALLE VS. PRÄDIKTIONSINTERVALLE
@@
 Gut demonstrieren lässt sich das am Beispiel von IMU's ...
+
+Als Ansatz für die Konstruktion besserer Bänder:
+Nichtparametrischer Ansatz mit Perzentilen (asymmetrische Bänder)
 

### 2021-10-15 R/estimate_uncertainty_loa.R
Update

@@ -1,81 +1,89 @@
 # Estimate uncertainty in different LoA approaches (95% percentile intervals for the LoA)
 estimate_uncertainty_loa <- function (data, n.boot) {
 
-  floa.v2.upper <- vector(mode = "list", length = n.boot)
-  floa.v2.lower <- vector(mode = "list", length = n.boot)
-  floa.v3.upper <- vector(mode = "list", length = n.boot)
-  floa.v3.lower <- vector(mode = "list", length = n.boot)
+  # floa.v2.upper <- vector(mode = "list", length = n.boot)
+  # floa.v2.lower <- vector(mode = "list", length = n.boot)
+  # floa.v3.upper <- vector(mode = "list", length = n.boot)
+  # floa.v3.lower <- vector(mode = "list", length = n.boot)
   floa.roislien.upper <- vector(mode = "list", length = n.boot)
   floa.roislien.lower <- vector(mode = "list", length = n.boot)
-  for (i in 1:n.boot) {
-    floa.v2       <- floa_rcb(data, n.boot, ver = "v2")
-    floa.v3       <- floa_rcb(data, n.boot, ver = "v3")
-    floa.roislien <- floa_roislien(data)
+  floa.lenhoff.upper <- vector(mode = "list", length = n.boot)
+  floa.lenhoff.lower <- vector(mode = "list", length = n.boot)
 
-    floa.v2.upper[[i]] <- floa.v2["upper", ]
-    floa.v2.lower[[i]] <- floa.v2["lower", ]
-    floa.v3.upper[[i]] <- floa.v3["upper", ]
-    floa.v3.lower[[i]] <- floa.v3["lower", ]
+  for (i in 1:n.boot) {
+    # floa.v2       <- floa_rcb(data, n.boot, ver = "v2")
+    # floa.v3       <- floa_rcb(data, n.boot, ver = "v3")
+    floa.roislien <- floa_roislien(data)
+    floa.lenhoff  <- floa_lenhoff(data,
+                                  k_reihe = 50,
+                                  n.boot = n.boot,
+                                  band = "prediction",
+                                  cp.begin = 0,
+                                  alpha = 0.05)
+
+    # floa.v2.upper[[i]] <- floa.v2["upper", ]
+    # floa.v2.lower[[i]] <- floa.v2["lower", ]
[…]

### 2021-10-15 R/floa_point.R

@@ -1,22 +1,16 @@
 floa_point <- function(data) {
 
   # ****************************************************************************
-  #  Pointwise FLoA (FLoA_Point) calculated using linear mixed effects models
+  #  Pointwise FLoA (FLoA_Point)
   # ----------------------------------------------------------------------------
   #
-  # Implemetation inspired by Parker et al. (2020) Using multiple agreement methods for
-  # continuous repeated measures data: a tutorial for practitioners
-  # https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
-  #
-  # [...] We focus specifically on the linear mixed effects model implementation
-  # of the methods [...].  The justification of this focus is because mixed
-  # effects modelling is increasingly used in clinical research and has advantages
-  # over fixed effects methods (e.g. Analysis Of Variance (ANOVA)) for several
-  # reasons outlined in Brown (2015). In particular, (i) missing or unbalanced
-  # data poses fewer problems for analysis, and (ii) inference can be made based
-  # on a wider population of patients
+  # For implementation details, see Bland & Altman, 2007
+  # Design currently implemented for balanced data
   # ****************************************************************************
   n.frames <- length(unique(data$frame))
+  n.subjects <- length(unique(data$subjectID))
+  n.strides <- length(unique(data$strideID))
+  n.curves.per.subject <- n.strides / n.subjects
 
   mean.diff <- vector(mode = "list", length = n.frames)
   total.sd <- vector(mode = "list", length = n.frames)
@@ -24,6 +18,7 @@ floa_point <- function(data) {
     data.by.frame <- subset(data, frame == frame.idx)
 
     # Get variance from ANOVA (see Bland & Altman, 2007)
+
     # Prepare data for ANOVA
     data.by.frame.wide <- reshape2::dcast(data = data.by.frame, subjectID + strideID ~ device, value.var = "value")
     data.by.frame.wide$device.diff <- data.by.frame.wide$IMU - data.by.frame.wide$MC
@@ -33,7 +28,10 @@ floa_point <- function(data) {
     S <- summary(LMEM.2)
[…]

### 2021-10-15 R/floa_roislien.R

@@ -18,11 +18,5 @@ floa_roislien <- function(data) {
 
   rownames(floa.roislien) <- c("upper", "mean", "lower")
 
-  # # Plausibility check
-  # plot(diff.curves[, 1], type="l", ylim = c(-5, 5))
-  # apply(diff.curves, 2, lines, type = "l")
-  # lines(floa.roislien["upper", ], col = "red", size = 3)
-  # lines(floa.roislien["lower", ], col = "red", size = 3)
-
   return(floa.roislien)
 }

### 2021-10-15 R/lenhoff_doris.R

@@ -30,8 +30,6 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
   time <- seq(0, (n_time-1)) # 0:(n_time-1)
 
   # ----------------------------------------------------------------------------
-  # Fourier Transformation
-  #
   # Approximate time series (differences) using Fourier functions
   # ----------------------------------------------------------------------------
 
@@ -46,7 +44,7 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
 
   # Fourierreihe aufstellen
   # General: f(t) = mu + sum(alpha cos(2pi*k*t/T) + beta sin(2pi*k*t/T))
-  fourier.s = rep(1, times = n_time) # ones(n_time, 1)
+  fourier.s = rep(1, times = n_time)
   for (k in seq(1, k_reihe*2, 2)) {
     fourier.s <- cbind(fourier.s, cos(2*pi*(k/2)*time/(n_time-1)))
     fourier.s <- cbind(fourier.s, sin(2*pi*(k/2)*time/(n_time-1)))
@@ -54,7 +52,7 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
 
   for (i in 1:n_time) {
     # Least squares Regression
-    fourier.koeffi[, i] = pracma::mldivide(fourier.s, data.diff[, i]) # fourier.s\daten[, i]
+    fourier.koeffi[, i] = pracma::mldivide(fourier.s, data.diff[, i])
     # Fourier curve
     fourier.real[, i] = fourier.s %*% fourier.koeffi[, i]
   }
@@ -69,7 +67,7 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
     fourier.std1[, , i] <- (fourier.koeffi[, i] - fourier.mean[, 1]) %*% t(fourier.koeffi[, i] - fourier.mean[, 1])
   }
 
-  fourier.kovarianz <- apply(fourier.std1, c(1, 2), mean) # nanmean(fourier.std1, 3)
+  fourier.kovarianz <- apply(fourier.std1, c(1, 2), mean)
   # Lenhoff, Appendix A, Eq. (0.5)
[…]

### 2021-10-15 R/main.R

@@ -11,18 +11,10 @@
 # (see subsection data sets)
 #
 # TODO:
-#   + Wie lassen sich zu weite LoA effektiv beschreiben / in einen Parameter giessen? Abstand zur aeussersten Kurve (als MSE)?
-#   + Weitere edge case Datensaetze a la shift
-#   + range-Angabe unertainty intervals!
-#   + Wie lassen sich (potenziell) zu weite LoA quantifizieren (Roislien!?)? Estimation uncertainty in Roislien is probably a lot higher
-#   + mean or median for floa.rcb?
-#   + 95% coverage implementieren
-#   + (Implementieren FDA)
 #   + Preallocation in floa_rcb()
 #   + Implement balanced data in floa_rcb.R
 #
 # TODO READ:
-#   + B&A, 2007
 #   + https://bmcmedresmethodol.biomedcentral.com/articles/10.1186/s12874-020-01022-x
 #   + Ratkowsky - Handbook of nonlinear regression models
 #   + Robinson et al. (2021)
@@ -117,11 +109,25 @@ n.boot <- 400
 #      strap the resulting sample (of length (n=length(subjects))
 # floa <- floa_rcb(data, n.boot, ver = "v2")
 
-floa.point <- floa_point(data)
+floa.point    <- floa_point(data)
 floa.roislien <- floa_roislien(data)
-floa.lenhoff <- floa_lenhoff(data, k_reihe = 50, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+floa.lenhoff.pred  <- floa_lenhoff(data,
+                                   k_reihe = 50,
+                                   n.boot = n.boot,
+                                   band = "prediction",
+                                   cp.begin = 0,
+                                   alpha = 0.05)
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff, ylim = c(-5, 5))
+floa.lenhoff.conf  <- floa_lenhoff(data,
+                                   k_reihe = 50,
+                                   n.boot = n.boot,
+                                   band = "confidence",
+                                   cp.begin = 0,
+                                   alpha = 0.05)
+
+
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-15, 15))
+plot_loa(data, floa.point, floa.roislien, floa.lenhoff.conf, ylim = c(-15, 15))
 
 
 # ********************************* Coverage ***********************************

### 2021-10-15 R/pick_subwise_curves.R

@@ -1,4 +1,5 @@
 pick_subwise_curves <- function (data) {
+
   # Select a single random stride from every subject in data
 
   n.subj <- length(unique(data$subjectID))

### 2021-10-15 R/plot_loa.R

@@ -16,8 +16,8 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
   floa.lenhoff <- data.frame(t(floa.lenhoff))
 
   # Difference curves -------------------------------------------------
-  device1 <- data.frame(subset(data, device == "IMU")$value)
-  device2 <- data.frame(subset(data, device == "MC")$value)
+  device1 <- subset(data, device == "IMU", select = value)
+  device2 <- subset(data, device == "MC", select = value)
 
   device.diff <- device1 - device2
   colnames(device.diff)[1] <- "value"
@@ -30,14 +30,12 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
   device.diff$strideID <- as.factor(rep(1:n.strides, each = 101))
   device.diff$subjectID <- as.factor(rep(1:n.subjects, each = strides.per.subject * n.frames))
 
-
   # For line graphs, the data points must be grouped so that it knows which points to connect.
   # In this case, it is simple -- all points should be connected, so group=1.
   # When more variables are used and multiple lines are drawn, the grouping for lines is usually done by variable.
   PLOT.DIFF <- ggplot(data = device.diff, aes(x = frame, y = value)) +
-    geom_line(data = device.diff,
-              aes(group = strideID),
-              alpha = 0.25) +
+    geom_line(aes(group = strideID),
+              alpha = 0.15) +
     scale_color_grey(start = 0.8, end = 0.2) +
     # Plot lower limit of agreement
     geom_line(data = floa.lenhoff,

### 2021-10-15 R/singlecurve_coverage.R

@@ -32,7 +32,12 @@ singlecurve_coverage <- function (data, n.boot) {
     # floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
     # floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out)
-    floa.lenhoff <- floa_lenhoff(data.one.out, k_reihe = 30, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+    floa.lenhoff  <- floa_lenhoff(data,
+                                  k_reihe = 50,
+                                  n.boot = n.boot,
+                                  band = "prediction",
+                                  cp.begin = 0,
+                                  alpha = 0.05)
 
     # Plot left out curve vs. various FLoA methods -----------------------
     data.subset <- subset(data, strideID == curve.idx)

### 2021-10-15 R/singlecurve_coverage_fraction.R

@@ -42,7 +42,12 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
     # floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
     # floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out)
-    floa.lenhoff <- floa_lenhoff(data.one.out, k_reihe = 30, n.boot = n.boot, cp.begin = 0, alpha = 0.05)
+    floa.lenhoff  <- floa_lenhoff(data,
+                                  k_reihe = 50,
+                                  n.boot = n.boot,
+                                  band = "prediction",
+                                  cp.begin = 0,
+                                  alpha = 0.05)
 
     # Plot left out curve vs. various FLoA methods -----------------------
     data.subset <- subset(data, strideID == curve.idx)

### 2021-10-15 R/floa_point.R
Change formula for confidence bands

@@ -54,7 +54,8 @@ floa_point <- function(data) {
   # Confidence intervals for the 95% limits of agreement (Bland & Altman, 1986)
   # The standard error of the limits is approximately sqrt(3*s^2 / n)
   n.curves <- length(unique(data$strideID))
-  standard.error.d2s <- sqrt(3*floa.point$total.sd^2 / n.curves)
+  # standard.error.d2s <- sqrt(3*floa.point$total.sd^2 / n.curves)
+  standard.error.d2s <- sqrt(1/n.curves + (z0.975^2 / (2*(n.curves-1)))) * floa.point$total.sd
 
   # t-score of the 95th quantile of the Student t distribution with df = (n.curves - 1)
   t <- qt(.975, df = n.curves - 1)

### 2021-10-15 paper.Rmd
Update draft

@@
 Um die coverage Performance der untersuchten Methoden zu bewerten, wurden die jeweiligen Intervalle im leave-one-out Verfahren kreuzvalidiert.
-Dabei wird eine Kurve vom Datensatz entfernt und die Intervalle anhand der übrigen Kurven berechnet.
-Dieser Vorgang wird für jede der $n$ Kurven eines Datensatzes wiederholt.
+Dazu wird jeweils eine Kurve vom Datensatz entfernt und die Intervalle anhand der übrigen Kurven berechnet.
+Dieser Vorgang wird für jede der $N$ Kurven eines Datensatzes wiederholt.
 Der Anteil der vorgehaltenen Kurven, die in den Bändern enthalten sind, wird mit zwei verschiedenen Kennwerten beschrieben:
@@
 
-Für die punktweisen wurde
-
 ### LoA uncertainty estimation
 
-Um die Schwankung der Intervallgrenzen über wiederholte Berechnungen hinweg zu veranschaulichen, wurden für die BOOT und ROISLIEN Intervalle Unsicherheitsbänder gebootstrapped (b = 400 Wiederholungen).
-Die Unsicherheit in diesen Verfahren entsteht durch das wiederholte Ziehen mit Zurücklegen im Rahmen des Bootstrapverfahrens (BOOT) bzw.
+Um die Schwankung der Intervallgrenzen über wiederholte Berechnungen hinweg zu veranschaulichen, wurden für alle drei Verfahren Unsicherheitsbänder durch gebootstrapped ($B = 400$).
+Die Unsicherheit entsteht durch wiederholtes Ziehen mit Zurücklegen im Rahmen des Bootstrapverfahrens (BOOT) bzw.
 die jeweils unterschiedlichen gezogenen Einzelkurven (ROISLIEN).
@@
 Im Falle der POINT Intervalle werden alle Kurven eines Datensatzes verwendet, weshalb keine Schwankung der Intervallgrenzen auftritt.
-Verschiedene Autoren schlagen jedoch Konfidenzintervalle als Erweiterung ihres ursprünglichen Ansatzes vor, um den sampling error insbesondere kleinere Stichproben (Bland & Altman, 1986; Hamilton & Stamey, 2007; Giavarina, 2015) zu berücksichtigen.
-Bland & Altman (1986) schätzen den Standardfehler der Differenzen auf ungefähr $\sqrt{3*s² / c}$.
-
-mit $s$ ... Standardabweichung
-und $c$ ... Anzahl der Kurven im Datensatz
+Verschiedene Autoren schlagen jedoch Konfidenzintervalle (KI) als Erweiterung ihres ursprünglichen Ansatzes vor, um den sampling error insbesondere kleiner Stichproben (Bland & Altman, 1986; Hamilton & Stamey, 2007; Giavarina, 2015) zu berücksichtigen.
+Bland & Altman (1986) schlagen die folgende Formel vor (95% CI für 95% LoA):
+
+$$
+(\overline{d} \pm z_{0.975}SD) \pm t_{0.975, n-1}SD\sqrt{\frac{1}{N} + \frac{z^2_{0.975}}{2(N-1)}}
+$$
+Quelle? Wo steht die Formel drin?
+
+mit $z_{0.975} = 1.96$.
+
+<!-- schätzen den Standardfehler der Differenzen auf ungefähr $\sqrt{3*s² / c}$. -->
 
@@
 Im vorliegenden Paper wurden 95% Konfidenzintervalle gerechnet.
-
-Die vollständige von Bland & Altman vorgeschlagene Formel (95% CI für 95% LoA) lautet damit:
-
-$(\overline{d} \pm z_{0.975}*s) \pm t_{0.975, n-1}*s*\sqrt{3*s² / c}$, mit $z_{0.975} = 1.96$
 
@@
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# estimate_uncertainty_loa(data, n.boot)
+estimate_uncertainty_loa(data, n.boot)
 ```

### 2021-10-18 R/simulate_data.R
Add example data sets as txt

@@ -33,6 +33,7 @@ imu.mc.balanced <- imu.mc[imu.mc$strideID %in% stride.idx, ]
 
 if (sum(duplicated(stride.idx))) {
   saveRDS(imu.mc.balanced, file = paste0("~/floa/R/examples/", "imu_mc.rds"))
+  # write.table(imu.mc.balanced, "~/floa/R/examples/imu_mc.txt", sep="\t")
 }
 
 
@@ -105,6 +106,7 @@ frame <- rep(0:100, times = n.strides * n.devices * n.subj)
 data <- data.frame(device, subjectID, strideID, value, frame)
 
 saveRDS(data, file = paste0("~/floa/R/examples/", "smooth_realistic.rds"))
+# write.table(data, "~/floa/R/examples/smooth_realistic.txt", sep="\t")
 
 
 
@@ -175,6 +177,7 @@ frame <- rep(0:100, times = n.strides * n.devices * n.subj)
 data <- data.frame(device, subjectID, strideID, value, frame)
 
 saveRDS(data, file = paste0("~/floa/R/examples/", "smooth.rds"))
+# write.table(data, "~/floa/R/examples/smooth.txt", sep="\t")
 
 
 
@@ -244,6 +247,7 @@ frame <- rep(0:100, times = n.strides * n.devices * n.subj)
 data <- data.frame(device, subjectID, strideID, value, frame)
 
 saveRDS(data, file = paste0("~/floa/R/examples/", "heteroscedastic.rds"))
+# write.table(data, "~/floa/R/examples/heteroscedastic.txt", sep="\t")
 
 
 
@@ -313,6 +317,7 @@ frame <- rep(0:100, times = n.strides * n.devices * n.subj)
 data <- data.frame(device, subjectID, strideID, value, frame)
 
 saveRDS(data, file = paste0("~/floa/R/examples/", "smooth_trend.rds"))
+# write.table(data, "~/floa/R/examples/smooth_trend.txt", sep="\t")
 
 
 
@@ -384,6 +389,7 @@ frame <- rep(0:100, times = n.strides * n.devices * n.subj)
 data <- data.frame(device, subjectID, strideID, value, frame)
 
 saveRDS(data, file = paste0("~/floa/R/examples/", "non_gaussian.rds"))
+# write.table(data, "~/floa/R/examples/non_gaussian.txt", sep="\t")
 
 
 
@@ -459,6 +465,7 @@ frame <- rep(0:100, times = n.strides * n.devices * n.subj)
[…]

### 2021-10-18 paper.Rmd
Continue paper draft

@@
 ---
-title: Comparison of continuous prediction bands for the validation of smooth 1D biomechanical
-  time series data
+title: Comparison of continuous prediction bands for the validation of biomechanical
+  curve data
 author:
@@
 Dabei werden Bewegungen mit zwei Messsystemen simultan aufgezeichnet und Messwerte miteinander verglichen, die theoretisch den gleichen Wert aufweisen sollten.
-Die statistische Aufbereitung der zwangsläufig zu beobachtenden Abweichungen ist dabei aus verschiedenen Gründen nicht trivial.
-Eine besondere Herausforderung stellt die Tatsache dar, dass der Output der meisten Messsysteme in Form höherdimensionaler Daten (z.
+Die statistische Aufbereitung der gemessenen Abweichungen ist bereits bei 0D Punktdaten (z.
 B.
-1D Kurvendaten wie EMG-Signale oder COP Verläufe, 2D plantare Druckmatrizen oder 3D bone strain fields) vorliegt, innerhalb derer der Messfehler variiert.
-Der Fokus dieses Papers liegt auf 1D Kurvendaten, die den häufigsten Anwendungsfall in biomechanischen Bewegungsanalysen darstellen.
-
-In den allermeisten Validierungsstudien wird der Zeitreihencharakter zugunsten einfacher(er) Modelle vernachlässigt und Kurvendaten auf einzelne diskrete (0D)Punkte oder skalare Features (z.
+body mass) nicht trivial (Bland & Altman, 1983) und wird zusätzlich erschwert, wenn Messungen als höherdimensionale Daten vorliegen (z.
+B.
+1D Kurvendaten wie EMG-Signale oder COP Verläufe, 2D plantare Druckmatrizen oder 3D bone strain fields).
+Ein relevanter Anwendungsfall in biomechanischen Bewegungsanalysen sind dabei 1D Zeitreihendaten (Kurven), die zwei wesentliche Unterschiede gegenüber Punktdaten aufweisen: 1.
+Der Messfehler variiert über der Zeit.
+2.
+Die einzelnen Samplingpunkte der Kurven sind keine bloße Aneinanderreihung unabhängiger Messwerte, sondern lokal abhängige Events (Deluzio & Astephen, 2007; Pataky, 2010).
+Dieser Umstand ergibt sich aus dem anatomischen Kontext (Viskositätseigenschaften biologischer Gewebe (Fung, 1981), Freiheitsgrade menschlicher Gelenkbewegungen etc.).
+
+In den allermeisten Validierungsstudien werden diese Eigenschaften zugunsten einfacher(er) Modelle vernachlässigt und Kurvendaten auf einzelne diskrete Punkte oder skalare Features (z.
 B.
@@
 e.
-phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) biomechanische Daten, die die Bewegung von Körpersegmenten direkt messen, sind durch die fundamentalen Viskositätseigenschaften biologischer Gewebe (Fung, 1981) und Freiheitsgrade menschlicher Gelenkbewegungen charakterisisiert.
-Beispielsweise sind im Verlauf einer Gelenkwinkelkurve keine Sprünge zu erwarten.
-Aus statistischer Sicht bedeutet das, dass die Samplingpunkte innerhalb einer Kurve keine lose Ansammlung von Einzelpunkten darstellen, sondern lokal korreliert sind (Deluzio & Astephen, 2007; Pataky, 2010).
-Bei der einfachen Übertragung punktweiser Statistiken auf alle Samplingpunkte der Kurve ist damit zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird, da der (nonzero) covariance term of adjacent sampling points nicht berücksichtigt wird (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
+phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) bei der Übertragung punktweiser Statistiken auf alle Samplingpunkte lokal korrelierter Kurven ist zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird, da der (nonzero) covariance term of adjacent sampling points nicht berücksichtigt wird (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
 Die üblicherweise zur Korrektur beim multiplen Testen eingesetzten Korrekturverfahren zur Anpassung des alpha levels (Hochberg & , 1987) führen aufgrund der hohen Anzahl an Datenpunkten pro Kurve ($t \le 100$) zu sehr konservativen Schätzungen weshalb dieses Vorgehen hier nicht zielführend ist.
@@
 
-Sutherland, Olshen, Biden & Wyatt (1988, 1989) (BOOT) haben für Gelenkwinkeldaten im Kontext von Ganganalysen Ende der 1980er Jahre eine Methode vorgestellt, bei der Kurvendaten mithilfe eines random (Fourier) coefficient Regressionsmodells approximiert werden und aus den so gebildeten funktionalen Objekte anschließend durch Bootstrapping Prädiktionsbänder für normales Gehen gebildet werden.
+Sutherland, Olshen, Biden & Wyatt (1988, 1989) (BOOT) haben für Gelenkwinkeldaten im Kontext von Ganganalysen Ende der 1980er Jahre eine Methode vorgestellt, bei der Kurvendaten mithilfe eines random (Fourier) coefficient Regressionsmodells approximiert werden und aus den so gebildeten funktionalen Objekte durch Bootstrapping Prädiktionsbänder für normales Gehen gebildet werden.
 Lenhoff et al.
@@
 
-Für den Vergleich der Methoden wurden vier (simulierte und reale) Datensätze verwendet.
-Jeder der Datensätze enthält jeweils $N = 220$ Kurven der Länge $T = 101$ Datenpunkte aus $M = 2$ Messsystemen $(m=1,2)$.
-Das entspricht in jedem Datensatz $D = 110$ Differenzkurven, die durch Subtraktion gebildet werden:
+Für den Vergleich der Methoden wurden vier Datensätze (drei simulierte und ein realer) verwendet.
+Jeder dieser Datensätze enthält jeweils $N = 220$ Kurven der Länge $T = 101$ Datenpunkte aus $M = 2$ Messsystemen.
+Das entspricht in jedem Datensatz $D = 110$ Differenzkurven:
 
@@
 X und Y sind (I x J) Matrizen der beiden Messsysteme, mit $I = 11$ Probanden und $J = 10$ wiederholten Kurven pro Proband.
-Die Größe der Datensätze entspricht in etwa dem typischen Stichprobenumfang vieler Validierungsstudien.
-
-<!-- Die $nN$ Kurven entsprechen $J = 10$ repeated measures von $I = 11$ Probanden und $m = 2$ Messsystemen: -->
-<!-- $n = m * \sum_{i=1}^{11} \sum_{j=1}^{10} n_{ij} = 220$ -->
-
 Alle simulierten Datensätze wurden als glatte Kurven, ähnlich Gelenkwinkelverläufen, mit dem gleichen Grundmodell generiert.
-Die Kurven korrespondieren nicht mit realen Signalen und wurden teilweise bewusst überspitzt konstruiert um den Einfluss bestimmter Differenzcharakteristika auf das Verhalten der Bänder möglichst deutlich darzustellen.
-Die Kurven wurden als trigonometrische Funktionen durch Addition zweier Sinusfunktionen konstruiert:
+Dazu wurden für jedes Messsystem zwei ähnliche Sinusfunktionen überlagert:
 
@@
 
+Die Größe der Datensätze wurde bewusst klein gehalten, um dem typischen Stichprobenumfang vieler Validierungsstudien zu entsprechen.
+Die Kurven korrespondieren nicht mit realen Signalen und wurden teilweise bewusst überspitzt konstruiert um den Einfluss häufig auftretender Fehlercharakteristika (nicht Gauss'sche Fehler, Heteroskedastizität) auf das Verhalten der Bänder hervorzuheben.
+
+Der Code zum Erstellen der Kurven findet sich im Anhang und auf Github als R Code und txt Dateien.
+
 ### Calculation methods
@@
 $$
-\hat{f}(t) \pm C_{p} * \hat{\sigma}_{f(t)}
+\hat{f}(t) \pm C_{p} \hat{\sigma}_{f(t)}
 $$

### 2021-11-12 paper.Rmd
Update paper draft

@@
 ---
-title: Comparison of continuous prediction bands for the validation of biomechanical
-  curve data
+title: Comparison of continuous prediction bands for the validation of biomechanical curve data
 author:
@@
 ### TODO
+- Tabellen- und Abbildungsbeschriftungen
+- Erklarende Abbildung Methodenkapitel uncertainty estimation (eahnlich Diss Doris)
+- Formelnummerierung
 <!-- - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen -->
-<!-- - Umbennung entire curve in "All-or-nothing" (entlehnt von Juul et al.) -->
-<!-- - Toleranzintervalle mit reinnehmen? -->
-<!-- - ROISLIEN durch Toleranzintervalle ersetzen! -->
 - .bib file anlegen
-- Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert)
+<!-- - Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert) -->
 
@@
 # Introduction
+<!-- Wie in allen empirischen Wissenschaften basiert der Erkenntnisgewinn auf der Qualität experimenteller Daten und hängt damit direkt von der Güte der eingesetzten Messverfahren ab. -->
+Die Biomechanik ist eine empirische Wissenschaft, d.
+h.
+der Erkenntnisgewinn basiert auf der Qualität experimenteller Daten und hängt damit direkt von der Güte der eingesetzten Messverfahren ab.
+Die Beurteilung der Validität des Messverfahrens erfolgt üblicherweise über den Vergleich eines zu validierenden Messsystems mit einem bereits etablierten.
+Dabei werden Bewegungen mit beiden Messsystemen simultan aufgezeichnet und die Messwerte verglichen, wobei ein möglichst hoher Übereinstimmungsgrad vorliegen sollte.
 
-Die Beurteilung der Validität biomechanischer Messverfahren erfolgt üblicherweise über den Vergleich eines 'neuen' Messsystems mit einem etablierten.
-Dabei werden Bewegungen mit zwei Messsystemen simultan aufgezeichnet und Messwerte miteinander verglichen, die theoretisch den gleichen Wert aufweisen sollten.
 Die statistische Aufbereitung der gemessenen Abweichungen ist bereits bei 0D Punktdaten (z.
 B.
-body mass) nicht trivial (Bland & Altman, 1983) und wird zusätzlich erschwert, wenn Messungen als höherdimensionale Daten vorliegen (z.
+body mass) nicht trivial (Bland & Altman, 1983) und wird zusätzlich erschwert, wenn höherdimensionale Daten vorliegen, z.
 B.
-1D Kurvendaten wie EMG-Signale oder COP Verläufe, 2D plantare Druckmatrizen oder 3D bone strain fields).
-Ein relevanter Anwendungsfall in biomechanischen Bewegungsanalysen sind dabei 1D Zeitreihendaten (Kurven), die zwei wesentliche Unterschiede gegenüber Punktdaten aufweisen: 1.
-Der Messfehler variiert über der Zeit.
-2.
-Die einzelnen Samplingpunkte der Kurven sind keine bloße Aneinanderreihung unabhängiger Messwerte, sondern lokal abhängige Events (Deluzio & Astephen, 2007; Pataky, 2010).
-Dieser Umstand ergibt sich aus dem anatomischen Kontext (Viskositätseigenschaften biologischer Gewebe (Fung, 1981), Freiheitsgrade menschlicher Gelenkbewegungen etc.).
+in Form von 1D Kurvendaten (z.
+B.
+Gelenkwinkel, Bodenreaktionskurven, COP Trajektorien, EMG-Signale), 2D Matrizen (z.
+B.
+Druckverteilungsdaten), oder 3D Arrays (z.
+B.
+bone strain fields).
+Höherdimensionale Daten weisen gegenüber 0D Daten zwei wesentliche Unterschiede auf: 1.
+Der Messfehler variiert über der Zeit und 2.
+benachbarte Samplingpunkte sind nicht unabhängig, sondern lokal korreliert (Deluzio & Astephen, 2007; Pataky, 2010).
 
-In den allermeisten Validierungsstudien werden diese Eigenschaften zugunsten einfacher(er) Modelle vernachlässigt und Kurvendaten auf einzelne diskrete Punkte oder skalare Features (z.
+In den allermeisten Validierungsstudien werden diese Charakteristika allerdings durch Diskretisierung (Reduzierung auf Einzelpunkte oder skalare Features wie z.
 B.
-local extrema, rates of change, or ranges of motion) reduziert.
-Daraus ergeben sich eine Reihe von Problemen für die Bewertung des Messfehlers: (i) non-stationary error processes are not considered (ii) the validity of statements derived from discrete variables is limited by domain alignment (i.
+local extrema, rates of change, or ranges of motion) ignoriert um die statistische Auswertung und Interpretation der Daten zu vereinfachen.
+Daraus ergeben sich eine Reihe von Problemen hinsichtlich der Bewertung des Messfehlers: (i) non-stationary error processes are not considered (ii) the validity of statements derived from discrete variables is limited by domain alignment (i.
 e.
-phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) bei der Übertragung punktweiser Statistiken auf alle Samplingpunkte lokal korrelierter Kurven ist zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird, da der (nonzero) covariance term of adjacent sampling points nicht berücksichtigt wird (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
-Die üblicherweise zur Korrektur beim multiplen Testen eingesetzten Korrekturverfahren zur Anpassung des alpha levels (Hochberg & , 1987) führen aufgrund der hohen Anzahl an Datenpunkten pro Kurve ($t \le 100$) zu sehr konservativen Schätzungen weshalb dieses Vorgehen hier nicht zielführend ist.
-Ein besserer Ansatz ist es Kurven direkt als zusammenhängende Objekte zu betrachten.
+phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) wenn inferenzstatistische Verfahren wie Konfidenz- oder Prädiktionsintervalle simultan auf mehrere Samplingpunkte angewandt werden um kontinuierliche Fehlerbänder zu konstruieren, ist zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird weil die (nonzero) Kovarianz benachbarter Samplingpunkte nicht berücksichtigt wird (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
+Die üblicherweise eingesetzten Korrekturverfahren zur Anpassung des alpha levels (Hochberg & , 1987) führen aufgrund der hohen Anzahl an Samplingpunkten zu extrem konservativen Schätzungen des zufälligen Fehlers.
 
-Sutherland, Olshen, Biden & Wyatt (1988, 1989) (BOOT) haben für Gelenkwinkeldaten im Kontext von Ganganalysen Ende der 1980er Jahre eine Methode vorgestellt, bei der Kurvendaten mithilfe eines random (Fourier) coefficient Regressionsmodells approximiert werden und aus den so gebildeten funktionalen Objekte durch Bootstrapping Prädiktionsbänder für normales Gehen gebildet werden.
+Lenhoff et al.
+(1999) haben das Problem punktweise konstruierter Prädiktionsbänder am Beispiel von 1D Gelenkwinkelverläufen (Kurven) demonstriert.
+Dazu verglichen sie punktweise berechnete Prädiktionsbänder gegenüber einer funktionalen Methode bei der Kurvendaten zunächst mithilfe von Fourierfunktionen approximiert werden und aus den so gebildeten funktionalen Kurven durch Bootstrapping Prädiktionsbänder gebildet werden (Sutherland, Olshen, Biden & Wyatt (1988, 1989)).
+They found that the Sutherland prediction bands provided appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands were shown to provide inadequate coverage (54%).
+
+<!-- Sutherland, Olshen, Biden & Wyatt (1988, 1989) (BOOT) haben für Gelenkwinkeldaten im Kontext von Ganganalysen bereits Ende der 1980er Jahre eine Methode vorgestellt, bei der Kurvendaten mithilfe von Fourierfunktionen approximiert werden und aus den so gebildeten funktionalen Kurven durch Bootstrapping Prädiktionsbänder zur Beschreibung eines normativen Gangbildes gebildet werden.
 Lenhoff et al.
 (1999) illustrated the same method on a generic gait analysis data set (comprised of joint angle curves) and compared it to pointwise Gaussian intervals.
-By means of a leave-one out cross validation they found that the Sutherland prediction bands provide appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands are shown to provide inadequate coverage (54%).
+They found that the Sutherland prediction bands provided appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands were shown to provide inadequate coverage (54%).
+-->
+
 Røislien et al.
-(2012) haben einen Ansatz präsentiert, mit dem die Methode von Bland & Altman (1983) für kontinuierliche Daten erweitert wird.
-Der Ansatz von Bland & Altman ist die am weitesten verbreitete Methode in Methodenvergleichen (Ludbrook, 2010) und verwendet die Differenzen zwischen zwei Messmethoden für die Konstruktion von Limits Agreement (LoA).
+(2012) haben einen Ansatz präsentiert, mit dem die von Bland & Altman (1983) vorgeschlagene Methode zur Konstruktion von Übereinstimmungsintervallen, sog.
+Limits of Agreement (LoA: 95% der Differenzen zwischen zwei Messsystemen sind im Intervall mittlere Differenz $\pm$ 1.96 Standardabweichungen der Differenzen enthalten), für Kurvendaten erweitert wird.
+LoA sind das am weitesten verbreitete verfahren in Methodenvergleichen (Ludbrook, 2010) und stellen im Wesentlichen Prädiktionsintervalle dar (obwohl das tatsächliche Prädiktionslevel der LoA stets niedriger ist als das nomale Niveau und für kleine Stichproben kollabiert (Francq, 2020)).
 
-Das vorliegende Paper strebt eine systematische Validierung verschiedener Methoden zur Konstruktion kontinuierlicher Fehlerintervalle bei Methodenvergleichen an.
-Dazu sollen drei verschiedene Methoden anhand simulierter und realer Datensätze charakteristischer glatter Kurven biomechanischer Messysteme (Gelenkwinkelverläufe von zwei Messsystemen) verglichen werden: (1) punktweise Bland and Altman LoA, (2) LoA nach Roislien et al.
-und (3) gebootstrappte funktionale Prädiktionsbänder (Sutherland, Olshen, Biden & Wyatt (1988, 1989)).
-Da sowohl die Intervalle von Sutherland als auch Roislien im Wesentlichen Prädiktionsintervalle repräsentieren, soll der Methodenvergleich exemplarisch am Beispiel kontinuierlicher Prädiktionsbänder vollzogen werden.
-Die Methode lässt sich jedoch prinzipiell auch auf andere Intervallmethoden wie Konfidenz- oder Toleranzintervalle übertragen.
-Zur Beurteilung der Intervallgüte werden die qualitative Verläufe der Intervallgrenzen der jeweiligen Verfahren gegen die tatsächlichen Differenzen geplottet und die entsprechenden coverage properties berechnet.
-Für die Beurteilung der Intervallgüte wird im Rahmen einer leave-one out cross validation die Unsicherheit in den berechneten LoA angegeben.
+Ausgehend von diesen Vorarbeiten soll im vorliegenden Paper eine systematische Analyse von Verfahren zur Konstruktion kontinuierlicher Prädiktionsbändern für 1D Differenzkurven vorgenommen werden (im Kontext der Validierung von Messsystemen).
+Dazu werden drei verschiedene Verfahren anhand simulierter und realer Datensätze hinsichtlich ihrer Abdeckungseigenschaften verglichen: (1) punktweise Bland and Altman LoA (2) LoA nach Roislien et al.
+(2012) und (3) gebootstrappte funktionale Prädiktionsbänder (Sutherland, Olshen, Biden & Wyatt (1988, 1989), Lenhoff et al.
+(1999)).
 
-Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
-In diesem Sinne stellen wir auch den zugehörigen R Code im Anhang zur freien Verfügung.
+Wir hoffen mit dieser Studie einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien zu setzen und Forschern praktische Hinweise für die Auswahl und Implementierung geeigneter Verfahren zur Verfügung zu stellen.
+In diesem Sinne ist auch der zugehörige R Code im Anhang zur Verfügung gestellt.
 
@@
 Für den Vergleich der Methoden wurden vier Datensätze (drei simulierte und ein realer) verwendet.
-Jeder dieser Datensätze enthält jeweils $N = 220$ Kurven der Länge $T = 101$ Datenpunkte aus $M = 2$ Messsystemen.
-Das entspricht in jedem Datensatz $D = 110$ Differenzkurven:
+Jeder Datensatz enthält jeweils $N = 220$ Kurven der Länge $T = 101$ Datenpunkte aus $M = 2$ Messsystemen.
+Das entspricht in jedem Datensatz $D = 110$ Differenzkurven (Differenzen zwischen den Messsystemen):
 
@@
 
-X und Y sind (I x J) Matrizen der beiden Messsysteme, mit $I = 11$ Probanden und $J = 10$ wiederholten Kurven pro Proband.
+X und Y sind (I x J) Messwertmatrizen mit je $I = 11$ Probanden und $J = 10$ wiederholten Kurven pro Proband.
 Alle simulierten Datensätze wurden als glatte Kurven, ähnlich Gelenkwinkelverläufen, mit dem gleichen Grundmodell generiert.
-Dazu wurden für jedes Messsystem zwei ähnliche Sinusfunktionen überlagert:
+Dazu wurden für jedes Messsystem zwei Sinusfunktionen überlagert:
 
 $$
-X_{ij} = (offset_{m}) + f_{1}(t) + f_{2}(t),\\Y_{ij}(t) = (offset_{m}) + f_{3}(t) + f_{4}(t)
+X_{ij} = bias + A_{1}sin(B_{1}t) + A_{2}sin(B_{2}t)^{c + 3}, \\Y_{ij}(t) = bias + A_{1}sin(B_{1}t) + A_{3}sin(B_{2}t)^{c + 3}
 $$
-mit
+
+$A_{1,2,3}$ und $B_{1,2}$ sind Zufallsvariablen mit variierenden Verteilungseigenschaften (normal, uniform, Weibull), über deren Parameter Abweichungen zwischen Probanden (between subject variation), Schritten (within subject variation) und Messystemen abgebildet sind.
+$c$ ist eine Konstante ($c = 2$).
+Zudem wurden - je nach Datensatz - Offsets als normalverteilte Zufallsvariablen mit probandenweise variierenden Mittelwerten ($\mu_{i}$ modelliert:
 
 $$
-f_{1, 3}(t) = Asin(Bt)^{c + 3}
+bias \sim N(\mu_{i}, \sigma=0.05)
 $$
-und
-
-$$
-f_{2, 4}(t) = Asin(Bt)
-$$
-$A$, $B$ sind Zufallsvariablen mit unterschiedlichen Verteilungseigenschaften (normal, uniform, Weibull) über deren Parameter systematische und zufällige Unterschiede zwischen Probanden (between subject variation), Schritten (within subject variation) und Messystemen modelliert werden.
-$c$ ist eine Konstante ($c = 2$).
-Zudem wurden - je nach Datensatz - unterschiedliche Offsets als normalverteilte Zufallsvariablen mit variierenden $\mu$ und $\sigma$ simuliert:
-
-$$
-offset_{m} \sim N(\mu_{i}, \sigma=0.05)
-$$
-Im Sinne der Reproduzierbarkeit wurden alle Zufallsvariablen mit einem random seed initialisiert. Der Code für die Generierung der Kurven kann im Anhang im Detail nachvollzogen werden.
+Auf eine Modellierung systematischer Offsets zwischen beiden Messsystemen wurde verzichtet da sich diese lediglich auf die Breite der Intervalle auswirken, nicht aber auf die qualitativen Verläufe der Intervallgrenzen.
+Um die Reproduzierbarkeit der Datensätze zu gewährleisten wurden alle Zufallsvariablen mit einem random seed initialisiert.
+Der Code für die Generierung der Kurven kann im Anhang im Detail nachvollzogen werden.
 
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(size = 14),
-        axis.title.x = element_text(size = 14),
-        axis.text.y = element_text(size = 14),
-        axis.title.y = element_text(size = 14),
-        legend.position = "none")
+  theme(legend.position = "none")
 ```
@@
 
-Die Daten repräsentieren den Fall, dass die Kurven zweier Messsysteme qualitativ ähnlich sind und kein relevanter Bias vorliegt.
+Die Daten repräsentieren den Fall dass beide Messysteme qualitativ ähnliche Messkurven erzeugen.
 
-* Datensatz B : Simulated curves with non-Gaussian erros and heteroskedasticity
+* Datensatz B : Simulated curves with non-Gaussian errors and heteroskedasticity
 
-Der Skalierungsparameter des zweiten Messsystems ist als normalverteilte Zufallsvariable mit Weibull-verteiltem Mittelwert modelliert ($\epsilon_{i} \sim N(\mu \sim Wei(\gamma = 1.5,\alpha  = 1), \sigma_{i})$).
-Wie in (A) liegt kein Bias vor.
-
-Solche Daten liegen vor, wenn mindestens eines der Messsysteme Kurven produziert, die qualitativ und quantitativ deutlich vom wahren Wert abweichen (hier: Messsystem 2, grau).
-In diesem Fall sind elementare Modellvoraussetzungen für parametrische Verfahren wie Bland and Altman erheblich verletzt.
+Einer der Skalierungsparameter des zweiten Messsystems ($A_{3}$) ist als normalverteilte Zufallsvariable mit Weibull-verteiltem Mittelwert modelliert ($\epsilon_{i} \sim N(\mu \sim Wei(\gamma = 1.5,\alpha  = 1), \sigma_{i})$).
+Derartige Datensätze liegen dann vor, wenn mindestens eines der Messsysteme Kurven produziert, die in bestimmten Kurvenabschnitten deutlich vom wahren Wert abweichen (hier: Messsystem 2, grau).
+In diesem Fall sind elementare Voraussetzungen für den Einsatz parametrischer Modelle (wie z.
+B.
+Bland and Altman LoA) erheblich verletzt.
 
@@
 B.
-wenn die Messsysteme infolge der Platzierung am Körper unterschiedliche Bewegungen aufzeichnen, oder durch Filterartefakte.
+durch Filterartefakte.
 
-* Datensatz D : Reale Messdaten aus der Validierung einer inertialen Messeinheit (IMU)
+* Datensatz D : Real-world hip joint angle curves
 
-Real-world hip joint angle curves from healthy subjects walking at 6 km/h on a treadmill without gradient recorded simultaneously using an optical motion capture system (MC) and IMU.
-Die Kurven unterscheiden sich qualitativ und in der absoluten Ausprägung der Differenzen.
+Data from healthy subjects walking at 6 km/h on a treadmill without gradient recorded simultaneously recorded using an optical motion capture system (MC) and an inertial measurement unit (IMU).
+Die Kurven unterscheiden sich qualitativ und quantitativ von den übrigen Datensätzen.
+Die Streuung innerhalb und zwischen den Probanden ist stärker ausgeprägt und weniger homogen als in den simulierten Datensätzen.
 
-Die Größe der Datensätze wurde bewusst klein gehalten, um dem typischen Stichprobenumfang vieler Validierungsstudien zu entsprechen.
-Die Kurven korrespondieren nicht mit realen Signalen und wurden teilweise bewusst überspitzt konstruiert um den Einfluss häufig auftretender Fehlercharakteristika (nicht Gauss'sche Fehler, Heteroskedastizität) auf das Verhalten der Bänder hervorzuheben.
+Insgesamt wurde die Anzahl der Messkurven im Datensatz bewusst klein gehalten um dem typischen Stichprobenumfang vieler Validierungsstudien zu entsprechen.
+Die Kurven korrespondieren nicht mit realen Signalen und wurden teilweise bewusst überspitzt konstruiert um den Einfluss ausgewählter Fehlercharakteristika (nicht Gauss'sche Fehler, Heteroskedastizität, Phasenshift) auf das Verhalten der Bänder hervorzuheben.
 
@@
 
-Die untersuchten Prädiktionsbänder wurden ausschließlich anhand der gebildeten Differenzkurven $D$ gebildet um die Vergleichbarkeit der Methoden zu gewährleisten.
-Prädiktionsbänder wurden mit drei unterschiedlichen Verfahren berechnet:
+Um die Vergleichbarkeit der Ergebnisse zu gewährleisten, wurden in allen drei untersuchten Verfahren Differenzkurven als Berechnungsgrundlage der Prädiktionsbänder verwendet. Analysiert wurden:
 
-* 1. POINT : Pointwise continuous LoA according to Bland & Altman (1999, 2007)
+* 1. Pointwise continuous LoA according to Bland & Altman (1999, 2007) (POINT)
 
-Für alle $t = 101$ Punkte wurden jeweils über alle $D$ Kurven LoA berechnet.
-Bei erfüllten Modellvoraussetzungen (unabhängige Kurven, normalverteilte Differenzen, Homoskedastizität), kann erwartet werden, dass 95% der Differenzen im Intervall mean difference $\pm$ 1.96 $*$ standard deviation of the differences liegen.
-Wir verwenden in der vorliegenden Arbeit ein komplexeres Modell für wiederholte Messungen, in dem sich der wahre Wert über jedes Differenzpaar ändert (Bland & Altman, 1999).
-Das Modell beschreibt die Differenz zwischen zwei Messsystemen als Summe der mittleren Differenz ($\overline{d}$) und zweier Varianzkomponenten (between ($\sigma^2_{db}$), within ($\sigma^2_{dw}$)):
+Über alle $D$ Kurven wurden separat für alle $T = 101$ Punkte klassiche (diskrete) LoA berechnet.
+Bei erfüllten Modellvoraussetzungen (unabhängige Kurven, normalverteilte Differenzen, Homoskedastizität), kann erwartet werden, dass ungefähr 95% der Differenzen im Intervall mean difference $\pm$ 1.96standard deviation of the differences liegen.
+1.96 entspricht dem 0.975 Quantil der Standardnormalverteilung (Bland & Altman, 1986).
+In der vorliegenden Arbeit wird ein komplexeres Modell für wiederholte Messungen verwendet, in dem sich der wahre Wert über jedes Differenzpaar ändert (Bland & Altman, 1999).
+Das Modell beschreibt die Differenz zwischen zwei Messsystemen als Summe der mittleren Differenz und zweier Varianzkomponenten (between ($\sigma^2_{db}$), within ($\sigma^2_{dw}$)):
 <!-- Auch wenn hier jeweils die gleiche Anzahl Kurven für jedes Messsystem vorliegt, erlaubt die verwendete Formel die Übertragung auf den unbalancierten Fall. -->
@@
 $$
-LoA = D_{ij} = \overline{d} + \sqrt{\sigma^2_{db} + \sigma^2_{dw}}
+D_{ij} = \overline{d} + \sqrt{\sigma^2_{db} + \sigma^2_{dw}}
 $$
-Die beiden Varianzkomponenten können mithilfe einer einfaktoriellen Varianzanalyse bestimmt werden.
+Die beiden Varianzkomponenten werden mithilfe einer einfaktoriellen Varianzanalyse bestimmt.
 \sigma^2_{dw} kann als residual mean square ($MS_{w}$) direkt aus dem ANOVA table abgelesen werden.
-\sigma^2_{db} berechnet sich aus der Differenz aus between-subject mean squares und $MS_{w}$, geteilt durch einen Term, der die Anzahl der $i$ Probanden und $j$ Kurven berücksichtigt.
+\sigma^2_{db} berechnet sich aus der Differenz aus between-subject mean squares und $MS_{w}$, geteilt durch einen Term, der die Anzahl der Probanden und Kurven berücksichtigt.
 Wenn für alle Probanden die gleiche Anzahl an Datenpunkten vorliegt (wie im vorliegenden Fall), reduziert sich dieser Term auf die Anzahl der Datenpunkte (Kurven) pro Proband ($j = 10$):
@@
 
-* 2. ROISLIEN : Functional limits of agreement (FLoA) according to Roislien et al. (2012)
+* 2. Functional limits of agreement (FLoA) according to Roislien et al. (2012) (ROISLIEN)
 
 Die Methode von Røislien et al.
-(2012) wurde als Erweiterung der Methode von Bland & Altman (1983) für kontinuierliche Differenzen ($d(t)$) entwickelt.
-Roislien et al.
-approximieren die Kurven im Datensatz daher zunächst mithilfe elementarer trigonometrischer Funktionen.
-Die Berechnung der Modellparameter ($\overline{d}$, $\sigma$) erfolgt im Anschluss allerdings - genau wie in POINT - separat für jeden Punkt der Kurve.
-Daher macht es in der praktischen Umsetzung de facto keinen Unterschied ob die originalen (diskreten) oder funktionalen Kurven verwendet werden.
-Wir verzichten daher im Rahmen dieses Papers auf die Repräsentation der Kurven als Funktionen.
-Das Modell ist dennoch interessant weil gegenüber der Variante in POINT (i) keine wiederholten Kurven verwendet werden, sondern nur je eine Kurve pro Proband um die Modellvoraussetzung der Unabhängigkeit der Kurven nicht zu verletzen und (ii) die (punktweise) Berechnung der LoA für diese $j = i$ Kurven mithilfe des minimalen Bland & Altman Modells erfolgt ($LoA = d(t) \pm 1.96SD(t)$), wobei $\overline{d}(t)$ die kontinuierliche mean difference bezeichnet.
+(2012) wurde als Erweiterung der diskreten LoA für Kurvendaten vorgestellt.
+Dazu werden die Differenzkurven zunächst mithilfe von Fourier Funktionen angenähert.
+Die Berechnung der LoA erfolgt im Anschluss allerdings - genau wie in POINT - separat für jeden Punkt der Kurve.
+De facto bietet die Verwendung funktionaler Kurven mit diesem Ansatz keinen Vorteil gegenüber dem punktweisen Vorgehen, weshalb im Rahmen dieses Papers auf die Repräsentation der Kurven als Funktionen verzichtet wurde.
+Gegenüber dem Vorgehen in POINT werden zudem keine wiederholten Kurven verwendet, sondern nur je eine Kurve pro Proband um die Modellvoraussetzung der Unabhängigkeit der Kurven nicht zu verletzen.
+Die Berechnung der LoA für diesen reduzierten Datensatz ($j = i$ Kurven) erfolgt mithilfe des minimalen Bland & Altman Modells ($\overline{d} \pm 1.96SD$).
 
-* 3. BOOT : Boostrapped functional prediction bands according Sutherland et al. (1988), Olshen et al. (1989) and Lenhoff et al. (1999)
+Das ROISLIEN Verfahren wurde in die Analyse aufgenommen um den Einfluss fehlender Meswiederholunge als Repräsentation intersubjektiver Varianz zu demonstrieren.
+
+* 3. Boostrapped functional prediction bands according Sutherland et al. (1988), Olshen et al. (1989) and Lenhoff et al. (1999) (BOOT)
 
@@
 die Konstruktion der Prädiktionsbänder findet nicht punktweise, sondern auf Grundlage ganzer Kurven statt.
+Damit bleibt die lokale Abhängigkeit der Kurvenpunkte erhalten.
+Die Prädiktionsbänder werden mithilfe eines iterativ ermittelten Korrekturfaktors ($C_{p}$) bestimmt, der mittels Bootstrapping bestimmt wird.
+Bootstrapping ist ein nicht-parametrisches Verfahren und kommt daher ohne Annahmen zur Verteilung der Differenzen aus.
+Die folgende Beschreibung der Methode orientiert sich an der Beschreibung in Lenhoff et al.
+(1999).
 
-Bootstrapping ist ein nicht-parametrisches Verfahren und kommt daher ohne Annahmen zur Verteilung der Differenzen aus.
-
-Im ersten Schritt des Verfahrens werden alle $N$ gemessenen Kurven mithilfe von Fourier-Reihen als Funktionen approximiert:
+Im ersten Schritt werden alle $D$ Differenzkurven mithilfe von Fourier-Reihen als Funktionen approximiert:
 
 $$
-f_{n}(t) = \mu_{n} + \sum_{k=1}^K(\alpha_{nk}cos(\frac{2\pi kt}{T}) + \beta_{nk} sin(\frac{2\pi kt}{T}))
+f_{d}(t) = \mu_{d} + \sum_{k=1}^K(\alpha_{dk}cos(\frac{2\pi kt}{T}) + \beta_{dk} sin(\frac{2\pi kt}{T}))
 $$
-Die Anzahl der Fourier-Basisfunktionen $K$ ist dabei so zu wählen, dass die Kurven bestmöglich angenähert werden und keine ungewollte Glättungseffekte entstehen.
+
+wobei $d = 1,...,D$.
+Die Methode setzt voraus dass der Start- und Endpunkt der untersuchten Kurven auf dem gleichen Niveau liegen sollten, was bei perodischen Kurven wie Gelenkwinkeln oder Bodenreaktionsverläufen üblicherweise erfüllt ist.
+Die Anzahl der Basisfunktionen $K$ ist so zu wählen, dass die Kurven bestmöglich angenähert werden und keine ungewollten Glättungseffekte entstehen.
 In der vorliegenden Arbeit wurde großzügig $K = 50$ gewählt.
-Die Koeffizienten ($\mu_{n}, \alpha_{n1}, \beta_{n1}, ..., \alpha_{nK}, \beta_{nK}$) werden durch die Methode der kleinsten Quadrate bestimmt.
+Die Koeffizienten ($\mu_{n}, \alpha_{n1}, \beta_{n1}, ..., \alpha_{nK}, \beta_{nK}$) werden durch die Methode der kleinsten Quadrate bestimmt und ermöglichen die Schätzung der Originalkurven durch $\hat{f}_{d}(t)$ (Einzelkurven) und $\hat{f}(t)$ (mittlere Kurve).
 
-asserts that the curve starts and ends at the same height.
-<!-- Für das weitere Vorgehen ist dabei wichtig, dass $\mu$ für alle Kurven 0 sein muss, da die Standardabweichungskurve der Kurvenschar sonst einen Offset aufweist. -->
-
-Die Koeffizienten werden in einer Matrix $M$ mit den Dimensionen $N$ (Zeilen) und $2K+1$ zusammengefasst.
-Aus dieser Matrix werden durch Bootstrapping $B = 400$ mal $N$ Kurven aus $M$ mit Zurücklegen gezogen, für die jeweils die mittlere Kurve $\hat{f^b}(t)$ und Kovarianzmatrix $\hat{\sigma_{\hat{f^b}(t)}}$ bestimmt werden.
-
-Im Mittel über alle Bootstrapiterationen lässt sich daraus der Anteil der Kurven bestimmen, deren maximale standardisierte Abweichung vom Bootstrap-Mittel ($\hat{f^b}(t)$) kleiner gleich $C_{p}$ beträgt.
+Aus der zugehörigen Koeffizientenmatrix werden durch Bootstrapping $B = 400$ mal die Koeffizienten von $D$ Kurven mit Zurücklegen gezogen, für die jeweils die mittlere Kurve $\hat{f^b}(t)$ und Variabilität $\hat{\sigma}_{\hat{f}^b(t)}$ bestimmt werden.
+Im Mittel über alle Bootstrapiterationen lässt sich daraus der Anteil der Kurven bestimmen, deren maximale standardisierte Abweichung vom Bootstrap-Mittel kleiner gleich $C_{p}$ ist.
 
 $$
-\frac{1}{B} \sum_{b=1}^B [\frac{1}{N} \sum_{n=1}^N I(\max_{t} (\frac{|\hat{f}_{n}(t) - \hat{f^b}(t)|}{\hat{\sigma}_{f(t)}} \le C_{p})]
+\frac{1}{B} \sum_{b=1}^B [\frac{1}{D} \sum_{d=1}^D I(\max_{t} (\frac{|\hat{f}_{d}(t) - \hat{f^b}(t)|}{\hat{\sigma}_{f(t)}} \le C_{p})]
 $$
-wobei $I(E)$ eine Matrix ist die Werte von entweder 0 ($E$ tritt nicht auf) oder 1 ($E$ tritt auf) annimmt.
-
-$C_{p}$ muss dabei so optimiert werden, dass es der gewünschten coverage probability des Prädiktionsbandes $P = 1- \alpha$ (hier: $\alpha=0.05$) entspricht.
-Damit ist berechnet sich das Prädiktionsband als:
+wobei $I(E)$ eine logische Matrix darstellt, die Werte von 0 oder 1 annimmt, wenn die zugehörige Ungleichung zutrifft.
+$C_{p}$ ist ein zu bestimmendes Quantil das durch Bootstrapping iterativ ($B$ mal) so optimiert wird, dass es der gewünschten coverage probability des Prädiktionsbandes $P = 1- \alpha$ (hier: $\alpha=0.05$) entspricht.
+$B$ sollte dabei möglichst groß gewählt werden (hier: $B = 400$).
+Damit berechnet sich das Prädiktionsband als:
 
 $$
-\hat{f}(t) \pm C_{p} \hat{\sigma}_{f(t)}
+\hat{f}(t) \pm C_{p} \hat{\sigma}_{\hat{f}(t)}
 $$
@@
 
-Um die coverage Performance der untersuchten Methoden zu bewerten, wurden die jeweiligen Intervalle im leave-one-out Verfahren kreuzvalidiert.
-Dazu wird jeweils eine Kurve vom Datensatz entfernt und die Intervalle anhand der übrigen Kurven berechnet.
-Dieser Vorgang wird für jede der $N$ Kurven eines Datensatzes wiederholt.
-Der Anteil der vorgehaltenen Kurven, die in den Bändern enthalten sind, wird mit zwei verschiedenen Kennwerten beschrieben:
+Für die Bewertung der coverage Performance wurden die untersuchten Methoden im leave-one-out Verfahren kreuzvalidiert.
+Dazu wird der Datensatz in einen Trainings- und Testdatensatz gesplittet, wobei der Testdatensatz genau einer Differenzkurve entspricht.
+Dieser Vorgang wird $D$ mal für alle Differenzkurven des Datensatzes wiederholt und folgende coverage Parameter für die Testkurven berechnet:
 
-* Percentage of entirely covered curves (all points within the limits)
+1. Prozentualer Anteil (proportion) der im Prädiktionsband enthaltenen Punkte einer Kurve
 
-* Mean (SD) of curve points within the respective FLoA: The fraction of points along a curve within the limits is calculated and averaged across all curves of a subject.
-
-### LoA uncertainty estimation
-
-Um die Schwankung der Intervallgrenzen über wiederholte Berechnungen hinweg zu veranschaulichen, wurden für alle drei Verfahren Unsicherheitsbänder durch gebootstrapped ($B = 400$).
-Die Unsicherheit entsteht durch wiederholtes Ziehen mit Zurücklegen im Rahmen des Bootstrapverfahrens (BOOT) bzw.
-die jeweils unterschiedlichen gezogenen Einzelkurven (ROISLIEN).
-
-Im Falle der POINT Intervalle werden alle Kurven eines Datensatzes verwendet, weshalb keine Schwankung der Intervallgrenzen auftritt.
-Verschiedene Autoren schlagen jedoch Konfidenzintervalle (KI) als Erweiterung ihres ursprünglichen Ansatzes vor, um den sampling error insbesondere kleiner Stichproben (Bland & Altman, 1986; Hamilton & Stamey, 2007; Giavarina, 2015) zu berücksichtigen.
-Bland & Altman (1986) schlagen die folgende Formel vor (95% CI für 95% LoA):
+Median, Modalwert und Range (Minimum, Maximum) von $x$ über alle Kurven, wobei $x$ der prozentuale Anteil der Punkte je untersuchter Kurve ist, die innerhalb der $upper$ und $lower$ limits des Prädiktionsbandes liegen:
 
 $$
-(\overline{d} \pm z_{0.975}SD) \pm t_{0.975, n-1}SD\sqrt{\frac{1}{N} + \frac{z^2_{0.975}}{2(N-1)}}
+\{x \in D \mid  \frac{1}{T} \sum_{t=1}^T (upper \leq  x_{t} \leq lower) \cdot 100 \} \\
 $$
-Quelle? Wo steht die Formel drin?
+<!-- $$ -->
+<!-- median = -->
+<!--   \begin{cases} -->
+<!--     x_{\frac{n+1}{2}} & n\text{ odd} \\ -->
+<!--     \frac{1}{2}\bigl(x_{\frac{n}{2}}+x_{\frac{n+1}{2}}\bigr) & n\text{ even} -->
+<!--   \end{cases} -->
+<!-- $$ -->
 
-mit $z_{0.975} = 1.96$.
+2. Mittlerer prozentualer Anteil der vollständig im Prädiktionsband enthaltenen Kurven (all-or-nothing Prinzip)
 
-<!-- schätzen den Standardfehler der Differenzen auf ungefähr $\sqrt{3*s² / c}$. -->
+$$
+\frac{1}{D}  \sum_{d=1}^D H(x_{d}) \cdot 100
+$$
+wobei $H$ eine Heaviside Funktion ist
 
-Für die Berechnung von Konfidenzintervallen wird noch das geeignete Quantil der t-Verteilung mit n − 1 Freiheitsgraden benötigt.
-Das Konfidenzintervall reicht dann vom oberen/unteren Limit minus t$*$Standardfehler bis zum oberen/unteren Limit plus t$*$Standardfehler.
-Im vorliegenden Paper wurden 95% Konfidenzintervalle gerechnet.
+$$
+H(x) =
+  \begin{cases}
+    0 : x_{d} < 1 \\
+    1 : x_{d} = 1
+  \end{cases}
+$$
+
+### Uncertainty estimation
+
+Die Grenzen der Prädiktionsintervalle unterliegen einem stochastischen Prozess und variieren in Abhängigkeit vom Samplingprozess in den jeweiligen Verfahren (POINT, ROISLIEN, BOOT).
+Um diese Unsicherheit der Intervallgrenzen zu quantifizieren, wurde der Abstand zwischen einer oberen und unteren Unsicherheitsgrenze (Erläuterung und Berechnungsdetails im Folgenden) als kumulierter Flächeninhalt (Cumulated Uncertainty Measure) durch punktweises Aufsummieren der Differenzen berechnet (Oriwol, 2012):
+
+$$
+CUM = \sum_{t=1}^T|upper_{t} - lower_{t}|
+$$
+
+Im vorgestellten POINT Verfahren nach Bland & Altman ergibt sich keine Streuung der Prädiktionsintervallgrenzen, da jeweils alle Kurven in die Berechnung der LoA eingehen.
+Bland & Altman (1986) schlagen jedoch 95% Konfidenzintervalle (CI) als Erweiterung des ursprünglichen Ansatzes vor, um die - insbesondere bei kleineren Stichproben auftretenden - Schwankungen zu berücksichtigen,
+
+$$
+LoA \pm 95\% CI: (\overline{d} \pm z_{0.975}SD) \pm t_{0.975, n-1}SD\sqrt{\frac{1}{D} + \frac{z^2_{0.975}}{2(D-1)}}
+$$
+wobei $t_{0.975, n-1}$ das 97.5% Perzentil der t-Verteilung (n − 1 Freiheitsgrade) ist und $z_{0.975} = 1.96$ das entsprechende Perzentil der Standardnormalverteilung (Bland & Altman, 1999).
+Die Grenzen der Konfidenzintervalle bilden die oberen und unteren Grenzen des Streuungsintervalls der LoA und können separat für beide LoA berechnet werden.
+Als Voraussetzung für die Berechnung dieser Konfidenzintervalle gilt, dass die Differenzen annähernd normalverteilt sein müssen.
+
+Beim ROISLIEN-Verfahren entsteht durch das Ziehen nur jeweils einer zufälligen Kurve pro Proband eine zusätzliche Varianzquelle gegenüber dem POINT Ansatz.
+Im Rahmen des BOOT-Verfahren variieren die Intervallgrenzen durch wiederholtes Ziehen mit Zurücklegen.
+Die Ermittlung der Intervallgrenzenstreuung in diesen Verfahren (ROISLIEN, BOOT) erfolgt empirisch.
+Dazu werden die Intervallgrenzen $B = 400$ mal berechnet und aus der resultierenden Kurvenschar die 2.5% bzw.
+97.5% Perzentile als untere bzw.
+obere Begrenzung für die Streuung der Prädiktionsintervallgrenzen verwendet.
+
+<!-- Nach dem Gesetz der großen Zahlen kann davon ausgegangen werden, dass der Abstand zwischen den Intervallgrenzen mit steigender Iterationszahl abnimmt und sich bei einem bestimmten Wert stabilisiert.
+-->
+
+<!-- Hier noch eine Grafik der kumulierten Flächen einbringen (siehe Diss Doris und bishere uncertainty plots) -->
 
@@
 
-### Functional limits of agreement
+Figure 2 enthält die qualitative Verläufe der Prädiktionsbänder der drei Methoden gegenüber den Differenzkurven.
+Die Werte für die Abdeckung der Gesamtkurven über alle Datensätze befinden sich in Table 1.
+Die zugehörigen Parameter für den Anteil der abgedeckten Punkte pro Kurve sind in Table 2 enthalten.
+Table 3 enthält die Streuung der Prädiktionsbandgrenzen (Cumulated Uncertainty Area).
+
+Die Ergebnisse zeigen, dass die gebootstrappten Prädiktionsintervalle (BOOT) in allen untersuchten Szenarien an die nominale coverage heranreichen.
+Die BOOT Methode ist dabei robust gegenüber unterschiedlichen Fehlercharakteristika.
+Die Prädiktionsbänder der punktweisen Verfahren (POINT, ROISLIEN) fallen zum Teil deutlich schmaler aus und unterrepräsentieren den Messfehler in Bezug auf die Abdeckung ganzer Kurven.
+Es zeigt sich jedoch dass die Performance insgesamt sehr gut ist und die Differenzkurven nur an wenigen Stellen außerhalb der Prädiktionsgrenzen liegen.
+
+Die qualitativen Verläufe der Prädiktionsbänder von POINT und ROISLIEN sind recht ähnlich.
+Allerdings ist die coverage der ROISLIEN-Bänder im Mittel über alle $D = 110$ Iterationen der leave-one-out Kreuzvalidierung niedriger als die der POINT-Bänder (sowohl bei der Abdeckung ganzer Kurven als auch bei beim Anteil der abgedeckten Kurvenpunkte).
+Generell weisen die ROISLIEN-Bänder die höchste Unsicherheit bzgl.
+der Streuung der Intervallgrenzen auf, d.
+h.
+die Prädiktionsbänder nach ROISLIEN fallen häufiger zu eng oder zu weit aus.
+Dabei ist die Unsicherheit der Intervallgrenzen beim ROISLIEN-Verfahren etwas höher als beim POINT-Verfahren und nahezu 3x so hoch wie beim BOOT Verfahren.
+
+Die BOOT-Bänder weisen eine deutlich geringere Streuung der Intervallgrenzen auf als die beiden übrigen Methoden.
 
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-n.boot <- 30
-```
+n.boot <- 5
 
-#### Simulated curves with Gaussian error model (A)
-
-```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Simulated curves with Gaussian error model (A)
 data <- example_data(dat = "smooth_realistic", dir.data)
@@
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-5, 5))
+PLOT.1 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-3, 3))
+
+# Simulated curves with non-normal, heteroskedastic differences (B)
+data <- example_data(dat = "non_gaussian", dir.data)
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff.pred  <- floa_lenhoff(data,
+                                   k_reihe = 50,
+                                   n.boot = n.boot,
+                                   band = "prediction",
+                                   cp.begin = 0,
+                                   alpha = 0.05)
+
+PLOT.2 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-8, 8))
+
+# Simulated curves curves with Gaussian error model created and phase shift in x-axis direction (C)
+data <- example_data(dat = "shift", dir.data)
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff.pred  <- floa_lenhoff(data,
+                                   k_reihe = 50,
+                                   n.boot = n.boot,
+                                   band = "prediction",
+                                   cp.begin = 0,
+                                   alpha = 0.05)
+
+PLOT.3 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-5, 5))
+
+# Real-world hip joint angle curves
+data <- example_data(dat = "imu_mc", dir.data)
+
+floa.point <- floa_point(data)
+floa.roislien <- floa_roislien(data)
+floa.lenhoff.pred  <- floa_lenhoff(data,
+                                   k_reihe = 50,
+                                   n.boot = n.boot,
+                                   band = "prediction",
+                                   cp.begin = 0,
+                                   alpha = 0.05)
+
+PLOT.4 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-20, 20))
 ```
 
-#### Simulated curves with non-normal, heteroskedastic differences (B)
+```{r echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Figure 2: Verlauf der Prädiktionsbänder von POINT (grüne, gepunktete Kurve), ROISLEN (blaue Linie) und BOOT (braune Linie) in den vier Datensätzen (A - Gaussian error model, B - Non-Gaussian errors and heteroskedasticity, C - Gaussian error model and phase shift in x-axis direction, D - Real-world hip joint angle curves).
+Differenzen zwischen beiden Messsystemen sind als schwarze Kurven Linien im Hintergrund dargestellt."}
+PLOT <- ggpubr::ggarrange(PLOT.1, PLOT.2, PLOT.3, PLOT.4,
+                          labels = c("A", "B", "C", "D"),
+                          ncol = 2,
+                          nrow = 2)
+
+PLOT
+```
 
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "non_gaussian", dir.data)
-# 
-# floa.point <- floa_point(data)
-# floa.roislien <- floa_roislien(data)
-# floa.lenhoff.pred  <- floa_lenhoff(data,
-#                                    k_reihe = 50,
-#                                    n.boot = n.boot,
-#                                    band = "prediction",
-#                                    cp.begin = 0,
-#                                    alpha = 0.05)
-# 
-# plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-10, 10))
+# Realistic smooth wave data (constant variance, no trend)
+data <- example_data(dat = "smooth_realistic", dir.data)
+
+# Leave-one curve out approach
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+
+# Calculate percent from counts
+calculate_percent <- function(data) {
+  percentage <- (sum(data) / length(data))
+  percentage <- round(percentage, digits = 2)
+}
+
+covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
+covered.curves.percent <- data.frame(t(covered.curves.percent))
+colnames(covered.curves.percent) <- c("POINT", "ROISLIEN", "BOOT")
+
+# Curves with non-gaussian (Weibull distributed) error (no trend)
+data <- example_data(dat = "non_gaussian", dir.data)
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+covered.curves.percent <- rbind(covered.curves.percent, apply(cover.cross.singlecurve, 2, calculate_percent))
+
+# Curves phase shifted in x-axis direction
+data <- example_data(dat = "shift", dir.data)
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+covered.curves.percent <- rbind(covered.curves.percent, apply(cover.cross.singlecurve, 2, calculate_percent))
+
+# Real-world validation data
+data <- example_data(dat = "imu_mc", dir.data)
+cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
+covered.curves.percent <- rbind(covered.curves.percent, apply(cover.cross.singlecurve, 2, calculate_percent))
+
+rownames(covered.curves.percent) <- c("A. GAUSS", "B. NONGAUSS", "C. XSHIFT", "D. REAL")
+knitr::kable(covered.curves.percent, caption = "Table 1: Percentages of entirely covered curves for methods POINT, ROISLIEN and BOOT across data sets A - D (A - Gaussian error model, B - Non-Gaussian errors and heteroskedasticity, C - Gaussian error model and phase shift in x-axis direction, D - Real-world hip joint angle curves).
+Coverage values are presented as mean values across $D = 110$ leave-one-out test curves.")
 ```
 
-#### Simulated curves curves with Gaussian error model created and phase shift in x-axis direction (C)
+```{r echo = FALSE, warning = FALSE, message = FALSE}
+# Realistic smooth wave data (constant variance, no trend)
+data <- example_data(dat = "smooth_realistic", dir.data)
+n.strides <- length(unique(data$strideID))
+# Leave-one curve out approach
+cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
+# Create list of numeric values for histogram
+hist.list <- cover.cross.fraction.singlecurve
+cover.cross.fraction.singlecurve <- cbind(cover.cross.fraction.singlecurve, rep("A. GAUSS", n.strides))
+colnames(cover.cross.fraction.singlecurve) <- c("POINT", "ROISLIEN", "BOOT", "Data")
+
+# Curves with non-gaussian (Weibull distributed) error (no trend)
+data <- example_data(dat = "non_gaussian", dir.data)
+tmp <- singlecurve_coverage_fraction(data, n.boot)
+hist.list <- cbind(hist.list, tmp)
+cover.cross.fraction.singlecurve <- rbind(cover.cross.fraction.singlecurve, cbind(tmp, rep("B. NONGAUSS", n.strides)))
+
+hist.list <- cbind(hist.list, tmp)
+
+# Curves phase shifted in x-axis direction
+data <- example_data(dat = "shift", dir.data)
+tmp <- singlecurve_coverage_fraction(data, n.boot)
+hist.list <- cbind(hist.list, tmp)
+cover.cross.fraction.singlecurve <- rbind(cover.cross.fraction.singlecurve, cbind(tmp, rep("C. XSHIFT", n.strides)))
+
+# Real-world validation data
+data <- example_data(dat = "imu_mc", dir.data)
+tmp <- singlecurve_coverage_fraction(data, n.boot)
+hist.list <- cbind(hist.list, tmp)
+cover.cross.fraction.singlecurve <- rbind(cover.cross.fraction.singlecurve, cbind(tmp, rep("D. REAL", n.strides)))
+
+# Melt data to long format
+# Assign correct class to avoid problems when melting to long data format
+cover.cross.fraction.singlecurve <- data.frame(cover.cross.fraction.singlecurve)
+cover.cross.fraction.singlecurve$POINT <- as.numeric(cover.cross.fraction.singlecurve$POINT)
+cover.cross.fraction.singlecurve$ROISLIEN <- as.numeric(cover.cross.fraction.singlecurve$ROISLIEN)
+cover.cross.fraction.singlecurve$BOOT <- as.numeric(cover.cross.fraction.singlecurve$BOOT)
+
+cover.cross.fraction.singlecurve.melt <- reshape2::melt(cover.cross.fraction.singlecurve)
+names(cover.cross.fraction.singlecurve.melt)[2] <- "Method"
+
+# Create a table with historgrams and boxplots
+# https://vincentarelbundock.github.io/modelsummary/articles/datasummary.html#histograms-1
+library(modelsummary)
+library(kableExtra)
+
+# Create a list with individual variables
+tmp_list <- list(hist.list)
+# tmp_list <- list(sample(10), sample(10), sample(10), sample(10), sample(10), sample(10))
+
+Mode <- function(x) {
+  ux <- unique(x)
+  ux[which.max(tabulate(match(x, ux)))]
+}
+
+# emptycol = function(x) " "
+caption <- "Table 2: Proportion of covered points across the curve.
+Due to the highly skewed distribution of coverage values (most values close to 1), the distribution of coverage values across $D = 110$ leave-one-out test curves is summarized using median, mode and range variables (minimum and maximum)."
+datasummary(Data * (Method * Heading("") * value) ~ Min + Max + Median + Mode, data = cover.cross.fraction.singlecurve.melt, title = caption)
+# + Heading("Histogram") * emptycol
+# %>%
+  # column_spec(column = 7, image = spec_hist(tmp_list))
+
+```
 
 ```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "shift", dir.data)
-# 
-# floa.point <- floa_point(data)
-# floa.roislien <- floa_roislien(data)
-# floa.lenhoff.pred  <- floa_lenhoff(data,
-#                                    k_reihe = 50,
-#                                    n.boot = n.boot,
-#                                    band = "prediction",
-#                                    cp.begin = 0,
-#                                    alpha = 0.05)
-# 
-# plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-5, 5))
-```
+data <- example_data(dat = "smooth_realistic", dir.data)
+uncertainty <- estimate_uncertainty_loa(data, n.boot)
 
-#### Real-world validation data (hip joint angles)
+# Curves with non-gaussian (Weibull distributed) error (no trend)
+data <- example_data(dat = "non_gaussian", dir.data)
+NONGAUSS <- estimate_uncertainty_loa(data, n.boot)
+uncertainty <- rbind(uncertainty, NONGAUSS)
 
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "imu_mc", dir.data)
-# 
-# floa.point <- floa_point(data)
-# floa.roislien <- floa_roislien(data)
-# floa.lenhoff.pred  <- floa_lenhoff(data,
-#                                    k_reihe = 50,
-#                                    n.boot = n.boot,
-#                                    band = "prediction",
-#                                    cp.begin = 0,
-#                                    alpha = 0.05)
-# 
-# plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-15, 15))
-```
+# Curves phase shifted in x-axis direction
+data <- example_data(dat = "shift", dir.data)
+XSHIFT <- estimate_uncertainty_loa(data, n.boot)
+uncertainty <- rbind(uncertainty, XSHIFT)
 
-### Leave-one (curve) out cross validation
+# Real-world validation data
+data <- example_data(dat = "imu_mc", dir.data)
+REAL <- estimate_uncertainty_loa(data, n.boot)
+uncertainty <- rbind(uncertainty, REAL)
 
-#### Realistic smooth wave data (constant variance, no trend)
-a. Coverage of the entire curve
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "smooth_realistic", dir.data)
-# 
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-# 
-# # Calculate percent from counts
-# calculate_percent <- function(data) {
-#   percentage <- (sum(data) / length(data))
-#   percentage <- round(percentage, digits = 2)
-# }
-# 
-# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-# covered.curves.percent <- data.frame(t(covered.curves.percent))
-# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
-```
+# Adjust column order and column/row names for the use in tables
+rownames(uncertainty) <- c("GAUSS", "NONGAUSS", "XSHIFT", "REAL")
+# colnames(uncertainty) <- c("POINT", "ROISLIEN", "BOOT")
 
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-# 
-# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-# )
-# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-# 
-# # # Display as boxplots across methods
-# # plot_cov_ver(cover.cross.fraction.singlecurve)
-```
-
-<!-- c. Average distance to the outer point of the curve set -->
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
-<!-- floa.roislien <- floa_roislien(data) -->
-
-<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
-<!--                        distance_2_floa(data, floa)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
-<!--                        ) -->
-
-<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
-
-<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
-<!-- ``` -->
-
-c. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-estimate_uncertainty_loa(data, n.boot)
-```
-
-#### Smooth, wave data (normal error, constant variance, no trend)
-a. Coverage of the entire curve
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "smooth", dir.data)
-# 
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-# 
-# # Calculate percent from counts
-# calculate_percent <- function(data) {
-#   percentage <- (sum(data) / length(data))
-#   percentage <- round(percentage, digits = 2)
-# }
-# 
-# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-# covered.curves.percent <- data.frame(t(covered.curves.percent))
-# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(covered.curves.percent, caption = "Percentages of entirely covered curves")
-```
-
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-# 
-# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-# )
-# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-# 
-# # # Display as boxplots across methods
-# # plot_cov_ver(cover.cross.fraction.singlecurve)
-```
-
-<!-- c. Average distance to the outer point of the curve set -->
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
-<!-- floa.roislien <- floa_roislien(data) -->
-
-<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
-<!--                        distance_2_floa(data, floa)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
-<!--                        ) -->
-
-<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
-
-<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
-<!-- ``` -->
-
-c. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# estimate_uncertainty_loa(data, n.boot)
-```
-
-#### Curves with nonlinear trend (constant variance)
-
-a. Coverage of the entire curve
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "smooth_trend", dir.data)
-# 
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-# 
-# # Calculate percent from counts
-# calculate_percent <- function(data) {
-#   percentage <- (sum(data) / length(data))
-#   percentage <- round(percentage, digits = 2)
-# }
-# 
-# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-# covered.curves.percent <- data.frame(t(covered.curves.percent))
-# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-```
-
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-# 
-# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-# )
-# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-```
-
-<!-- c. Average distance to the outer point of the curve set -->
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
-<!-- floa.roislien <- floa_roislien(data) -->
-
-<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
-<!--                        distance_2_floa(data, floa)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
-<!--                        ) -->
-
-<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
-
-<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
-<!-- ``` -->
-
-c. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# estimate_uncertainty_loa(data, n.boot)
-```
-
-#### Curves with non-gaussian (Weibull distributed) error (no trend)
-a. Coverage of the entire curve
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "non_gaussian", dir.data)
-# 
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-# 
-# # Calculate percent from counts
-# calculate_percent <- function(data) {
-#   percentage <- (sum(data) / length(data))
-#   percentage <- round(percentage, digits = 2)
-# }
-# 
-# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-# covered.curves.percent <- data.frame(t(covered.curves.percent))
-# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-```
-
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-# 
-# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-# )
-# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-# 
-# # # Display as boxplots acroos methods
-# # plot_cov_ver(cover.cross.fraction.singlecurve)
-```
-
-<!-- c. Average distance to the outer point of the curve set -->
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
-<!-- floa.roislien <- floa_roislien(data) -->
-
-<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
-<!--                        distance_2_floa(data, floa)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
-<!--                        ) -->
-
-<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
-
-<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
-<!-- ``` -->
-
-c. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# estimate_uncertainty_loa(data, n.boot)
-```
-
-#### Curves phase shifted in x-axis direction
-a. Coverage of the entire curve
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "shift", dir.data)
-# 
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-# 
-# # Calculate percent from counts
-# calculate_percent <- function(data) {
-#   percentage <- (sum(data) / length(data))
-#   percentage <- round(percentage, digits = 2)
-# }
-# 
-# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-# covered.curves.percent <- data.frame(t(covered.curves.percent))
-# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-```
-
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-# 
-# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-# )
-# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-# 
-# # # Display as boxplots acroos methods
-# # plot_cov_ver(cover.cross.fraction.singlecurve)
-```
-
-<!-- c. Average distance to the outer point of the curve set -->
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
-<!-- floa.roislien <- floa_roislien(data) -->
-
-<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
-<!--                        distance_2_floa(data, floa)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
-<!--                        ) -->
-
-<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
-
-<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
-<!-- ``` -->
-
-c. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# estimate_uncertainty_loa(data, n.boot)
-```
-
-#### Real-world validation data
-a. Coverage of the entire curve
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# data <- example_data(dat = "imu_mc", dir.data)
-# 
-# # Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.singlecurve <- singlecurve_coverage(data, n.boot)
-# 
-# # Calculate percent from counts
-# calculate_percent <- function(data) {
-#   percentage <- (sum(data) / length(data))
-#   percentage <- round(percentage, digits = 2)
-# }
-# 
-# covered.curves.percent <- apply(cover.cross.singlecurve, 2, calculate_percent)
-# covered.curves.percent <- data.frame(t(covered.curves.percent))
-# colnames(covered.curves.percent) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(covered.curves.percent, caption = "Percentages of entire covered curves")
-```
-
-b. Mean (SD) of covered curve points
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# Leave-one curve out approach
-# # ------------------------------------------------------------------------------
-# cover.cross.fraction.singlecurve <- singlecurve_coverage_fraction(data, n.boot)
-# 
-# cover.cross.mean.sd <- rbind(round(colMeans(cover.cross.fraction.singlecurve), 2),
-#              round(apply(cover.cross.fraction.singlecurve, 2, sd), 2)
-# )
-# colnames(cover.cross.mean.sd) <- c("Point", "Roislien", "Lenhoff")
-# 
-# knitr::kable(cover.cross.mean.sd, caption = "Mean (top) and SD of covered points across the curve")
-# 
-# # # Display as boxplots acroos methods
-# # plot_cov_ver(cover.cross.fraction.singlecurve)
-```
-
-<!-- c. Average distance to the outer point of the curve set -->
-<!-- ```{r echo = FALSE, warning = FALSE, message = FALSE} -->
-<!-- floa <- floa_rcb(data, n.boot, ver = "v2") -->
-<!-- floa.point <- floa_point(data) -->
-<!-- floa.roislien <- floa_roislien(data) -->
-
-<!-- dist.all <- data.frame(distance_2_floa(data, floa)$upr.dist, -->
-<!--                        distance_2_floa(data, floa)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.point)$lwr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$upr.dist, -->
-<!--                        distance_2_floa(data, floa.roislien)$lwr.dist -->
-<!--                        ) -->
-
-<!-- colnames(dist.all) <- c("floa.up", "floa.lw", "point.up", "point.lw", "roislien.up", "roislien.lw") -->
-
-<!-- boxplot(dist.all, ylab = "floa boundary - curve maximum") -->
-<!-- ``` -->
-
-c. Uncertainty in LoA (95% percentile intervals)
-```{r echo = FALSE, warning = FALSE, message = FALSE}
-# estimate_uncertainty_loa(data, n.boot)
+knitr::kable(uncertainty, caption = "Table 3: Cumulated Uncertainty Area.
+Values are displayed separately for the upper (U) and lower (L) limits of the three methods: POINT, ROISLIEN, BOOT.
+POINT values are identical for the upper and lower limits as symmetric confidence intervals were calculated.")
 ```
@@
 # Discussion
+Die Tatsache dass ein erheblicher Teil biomechanischer Messsysteme höherdimensionale Daten liefert, spiegelt sich bislang kaum in der Methodik von Validierungsstudien wieder.
+Um die Aussagekraft von Validierungsstudien diesbezüglich zu erhöhen braucht es Fehlerintervalle die alle Punkte des Messsignals.
+Zu diesem Zweck wurden im vorliegenden Paper drei relevante Verfahren (POINT, ROISLIEN, BOOT) zur Berechnung kontinuierlicher Prädiktionsintervalle identifiziert und für ausgewählte Messfehlerszenarien am Beispiel von 1D Kurvendaten empirisch analysiert.
 
-Punktweise FLoA unterrepräsentieren den Messfehler sind aber ganz ok.
-Allerdings gilt es zu konstatieren, dass die korrekte Berechnung (ANOVA-Varianzanteile, Konfidenzintervalle) auch nicht unbedingt trivial ausfällt.
+Die Intervalle der BOOT-Methode erreichen in allen untersuchten Szenarien die nominale coverage und sind mit weniger estimation uncertainty belastet als die beiden übrigen Verfahren (POINT, ROISLIEN).
+Diese Performance ist insofern zu erwarten, dass die BOOT-Bänder anhand eines Korrekturfaktors auf das nominale Niveau hin optimiert werden.
+Dies ist auch ein wesentlicher Unterschied zu den parametrischen POINT und ROISLIEN Methoden, bei denen die Intervallgrenzen auf Grundlage der Verteilung der Differenzen in der Stichprobe geschätzt werden.
 
-FLoA Roislien tendieren dazu zu weit auszufallen (den Messfehler zu überrepräsentieren)
+Die punktweise berechneten Bänder fallen erwartungsgemäß enger aus als die funktionalen BOOT-Bänder (Figure 2, Table 1 & 2).
+Die Prädiktionsintervalle der punktweisen Verfahren performen jedoch durchaus sehr gut wenn man den Anteil der abgedeckten Punkte pro Kurve über alle $D = 110$ Iterationen der leave-one-out Validierung als Maß für die Performance verwendet.
+Die zugehörige Verteilung ist extrem linksschief, d.
+h.
+bis auf wenige Kurven liegen stets die allermeisten Kurvenpunkte im Prädiktionsintervall.
+Die ROISLIEN-Methode wird hier im Übrigen nicht als funktionales, sondern als punktweises Verfahren gezählt, da im Anschluss an die Approximation der Kurven mithilfe von Fourrierreihen eine punktweise Berechnung der Prädiktionsintervallgrenzen stattfindet.
 
+Die Verläufe der Intervallgrenzen im POINT und BOOT-Verfahren Verfahren sind qualitativ ähnlich, die coverage (Table 1 & 2) ist beim ROISLIEN-Verfahren jedoch etwas niedriger.
+Die Ursache hierfür liegt in der höheren Streuung der ROISLIEN-Intervallgrenzen über $D = 110$ Iterationen des leave-one-out Validierungsverfahrens, da hier nur eine Kurve je Proband verwendet wird und kleinere Stichproben generell einer größeren Unsicherheit unterliegen.
+Bei der POINT-Methode hingegen werden stets alle Kurven im Datensatz berücksichtigt, womit es für die Daten der Stichprobe keine Unsicherheit bezüglich der Intervallgrenzen gibt.
+Um die inhärente Unsicherheit des Samplingprozesses zu berücksichtigen schlagen Bland & Altman allerdings vor die LoA um Konfidenzbänder zu erweitern.
+Der Einsatz zusätzlicher Konfidenzbänder führt die ursprüngliche Idee einfach zu berechnender, intuitiver Unsicherheitsintervalle jedoch ein Stück weit ad absurdum (Francq et al., 2020), weshalb in dieser Arbeit auf eine Verbreiterung der LoA durch Konfidenzbänder verzichtete wurde.
+Ohnehin ist die korrekte Umsetzung der im vorliegenden Paper umgesetzten POINT-Methode nach Bland & Altman (1999) (mit Varianzanteilen aus einer ANOVA mit Messwiederholung etc.) deutlich komplexer als das im ROISLEN-Verfahren eingesetzte minimale Bland & Altman Modell ($\overline{d} \pm 1.96SD$).
+<!-- "would complicate the analysis, makes it awkward and confusing, and is useless as the exact CI with the t-distribution is available"  -->
+
+Die punktweisen Verfahren liefern zudem auch für die Datensätze plausible Verläufe, in denen erhebliche Verletzungen der Modellannahmen vorliegen (nicht-gaussverteilten Differenzen, phase shift in x-axis direction).
+Für die Praxis ist das eine wichtige Erkenntnis, da diese Annahmen beim Vergleich verschiedener Messmethoden häufig verletzt sind.
+Die Verteilung der Differenzen zwischen zwei Systemen wird z.
+B.
+erheblich verzerrt, wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder zu geringer Auflösung abgeschnitten werden, oder wegen (nichtlinearem) Drift.
+Speziell im der letzten Fall sind Alternativen zum BOOT-Verfahren von Interesse, da die Methode nur für stationäre Signale zuverlässig funktioniert (Lenhoff et al., 1999).
+
+Die BOOT-Intervalle fallen stellenweise deutlich breiter aus als die punktweisen.
+Das resultiert in hervorragenden coverage Werten, produziert aber auch stark konservative Fehlerbänder.
+Fairerweise muss man sagen das die BOOT-Intervalle nie für diesen Einsatzzweck entwickelt wurden.
 Im Allgemeinen sind Prädiktionsintervalle überall dort von Interesse, wo neue Messwerte gegen normative Daten (z.
@@
 (1989) - bestimmte pathologische Muster angenommen und eine - wie auch immer geartete - Therapie vorgeschlagen.
-Entsprechend wichtig ist es diese Intervalle mit der richtigen Breite zu konstruieren, andernfalls gehen etwaige Effekte im Rauschen unter (zu weit) oder es werden Effekte angenommen, die so gar nicht auftreten (zu eng).
+Hier ist es also wichtig dass die Bänder nicht zu eng ausfallen, weil die Konsequenzen möglicher $\beta$ Fehler in diesem Kontext schwerwiegend sind.
 
-Die Bland & Altman Methode setzt normalverteilte, homoskedastische und unabhängige Differenzen voraus.
-Diese Annahmen sind allerdings beim Vergleich verschiedener Messmethoden häufig verletzt.
-Die Verteilung der Differenzen zwischen zwei Systemen wird z.
+Generell vermitteln die punktweisen Bänder (insbesondere POINT) ein deutlich realistischeres Bild des zufälligen Messfehlers.
+Für Differenzkurven zwischen zwei biomechanischen Messsystemen braucht es eigentlich keine Prädiktionsbänder die deutlich außerhalb der äußersten Kurve liegen wie bei BOOT da hier keine große Sprünge im Kurvenverlauf oder Ausreißer zu erwarten zu erwarten sind (vorausgesetzt die Stichprobe ist ausreichend groß und repräsentativ).
+
+Apropos unrealistisch weite Intervalle: In bestimmten Fällen, wie z.
 B.
-erheblich verzerrt, wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder der Auflösung abgeschnitten werden, oder wegen (nichtlinearem) Drift.
-Die Unabhängigkeitsannahme wird dann verletzt, wenn messwiederholte Daten vorliegen.
-Meswiederholte Werte sind sich i.
+dem realweltlichen Datensatz fallen die Bänder aller drei Methoden so breit aus, dass mit dem neuen Messverfahren im Rahmen eines between-subject Designs kaum ein Effekt detektiert werden kann, der über den zufälligen Messfehler hinausgeht (Figure 2,  D).
+Anhand der Kurvenverläufe ist aber auch erkennbar, dass die Kurven geclustert auftreten weil die Streuung innerhalb der Probanden geringer ausfällt.
+Diese Eigenschaft kann ausgenutzt werden um engere Intervalle auf Grundlage der intraindividuellen Streuung zu konstruieren wenn primär die Reproduzierbarkeit der neuen Methode von Bedeutung ist.
+Das ist z.
+B.
+der Fall wenn Untersuchungen mit einem reinen within-subjects Design durchgeführt werden (z.
+B.
+bei prä-post-Vergleichen in Interventionsstudien).
+Bland & Altman (1999) schlagen hierfür vor Intervalle auf Grundlage der within-subject Standardabweichung (aus einer einfaktoriellen Varianzanalyse) zu konstruieren.
+Neben der Tatsache dass diese für sich gesehen informativ sind, stellen sie auch eine Art Baseline zur Bewertung der between-method variability dar.
+
+<!-- Hahn & Meeker ... Statistical intervals -->
+<!-- https://ebookcentral.proquest.com/lib/tuchemnitz/detail.action?docID=4822515 -->
+<!-- In the best of situations, one can rely on physical understanding, or information from outside the study, to justify the practical assumptions.
+Such evaluations, however, are principally the responsibility of the subject-matter expert.
+Often, the assessment of such assumptions is far from clear-cut.
+-->
+Hier zeigt sich dass für die Konstruktion geeigneter Fehlerbänder keine universal gültigen Antworten gibt, sondern die Wahl der statistischen Intervalle - wenn möglich -  vom Kontext der praktischen Anwendung des Messsystems abhängig gemacht werden sollte.
+Realistisch betrachtet ist der spätere Awendungsfall zum Zeitpunkt der Validierung nicht immer klar definiert.
+In solchen Fällen empfehlen wir verschiedene Bänder für within und between-subject Designs zu konstruieren.
+Falls das für Nich-Statistiker als zu kompliziert empfunden wird scheinen die untersuchten POINT-Bänder (mit messwiederholten Werten!) einen brauchbaren Kompromiss zwischen zu breiten und zu engen Intervallen darzustellen.
+Zudem sollten berechnete Fehlerintervalle stets grafisch anhand der Messdaten (Originalkurven und/oder Differenzkurven) evaluiert werden.
+
+In jedem Fall sollten aber aus den in der Einleitung genannten Gründen auf kontinuierliche Methoden gesetzt werden.
+Wir spekulieren dass der seltene Einsatz kontinuierlicher Verfahren in direktem Zusammenhang mit der Komplexität der Verfahren steht, die eine hohe Hürde für viele Praktiker mit weniger ausgeprägten Statistik- und Programmierkenntnissen darstellt.
+Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code frei zur Verfügung (GPL).
+
+<!-- Die Erkenntnisse der vorliegenden Studie können nicht 1:1 auf andere Arten von Daten (z. B. Bodenreaktionskraftkurven, EMG Signale) übertragen, sondern gelten primär für glatte Kurven. -->
+Was im vorliegenden Paper nicht untersucht wurde ist der Fall, wenn keine glatten Kurven Kurven (z.
+B.
+verrauschte Signale wie EMG) untersucht werden.
+Diese Betrachtungen sprengen aber den Rahmen des vorliegenden Papers.
+Zudem fallen Differenzenkurven i.
 d.
 R.
-ähnlicher (d.
-h.
-sie streuen weniger) als die Kurven verschiedener Probanden.
-Entsprechend besteht die Gefahr, die Streuung der Differenzen zwischen zwei Messmethoden zu unterschätzen.
-Um dem entgegenzuwirken, verwenden Roislien nur jeweils eine Kurve je Proband, was wiederum zu Lasten einer erheblichen Unsicherheit bezüglich der Streuung der berechneten LoA erkauft wird.
+deutlich glatter aus als die verrauschten Originalkurven und das Problem ist womöglich deutlich geringer als momentan zu erwarten wäre.
+<!-- Dann gibt es eine Reihe zusätzlicher Probleme, z.
+B.
+dass die Approximation mithilfe mathematischer Funktionen andere Funktionen (B-Splines anstatt Fourier) erfordert oder gänzlich unmöglich wird.
+-->
 
-Stichprobengröße: Es kann natürlich gut sein, dass die punnktweisen Verfahren bei höheren Stichprobenumfängen besser performen.
+!-- Ludbrook (2010)
+<!-- First, you should decide which set of limits arebest suited to your goal: 95% confidence limits indicate the range within which 95%of differences in the population from which you drew your sample should lie, so, in a sense, confidence limits are concerned with the present; 95%tolerance limits (shorthand for 95%tol-erance limits with 95%confidence) indicate the range within which asingle, new, observation should lie if it is drawn from the same popu-lation as was the sample you have studied and so, in a sense, toler-ance limits are concerned with the future.
+Which of these two types of limits should you estimate?
+The confidence limits can be used to decide whether one method can safely be substituted for another; hence, Altman and Bland’s description of them as ‘limits of agree-ment’.4,7,12 It is up to you to argue that your 95%confidence limits are too wide to safely allow one method to be substituted for the other or, conversely, that they are narrow enough to allow substitution.
+However, if you imagine that on some future occasion you obtain a certain value by Method A and want to know just how wide the range of values is that you would have obtained by Method B, then the 95%tolerance limits are appropriate.
+-->
 
-Die Berechnung der LoA ist mit bestimmten (parametrischen) Modellvoraussetzungen verbunden, dass ist aber am Ende eigentlich egal, weil: Was mache ich denn wenn die verletzt sind?
+<!-- Anforderungen an die Fehlerintervalle gestellt werden?
+Im Sinne einer möglichst hohen coverage liefern breitere Prädiktionsintervalle eine konservative Schätzung des zu zufälligen Messfehlers.
+Im Kontext von Messsystemvalidierungen kann es aber durchaus sinnvoll sein einen Teil der coverage zugunsten weniger konservativer Intervalle zu opfern solange diese die Streuung der Differenzen adäquat charakterisieren.
+Die POINT-Intervalle scheinen hier einen guten Kompromiss darzustellen.
+-->
+
+<!-- Eine mögliche Alternative zu den untersuchten Verfahren -->
+<!-- Ein möglicher Kompromiss zwischen den gebootstrappten, funktionalen Intervallen in BOOT und ... in POINT sind die von Francq vorgeschlagenen Toleranzintervalle. -->
+Francq et al.
+(2020) schlagen vor stattdessen Toleranzintervalle zu verwenden.
+Toleranzintervalle (auch beta-expectation tolerance interval ($\beta$ TI)) sind statistische Intervalle, die mit einem gewissen Konfidenzniveau einen bestimmten Anteil einer Stichprobenpopulation (z.
+B.
+95% der Differenzkurven) enthalten.
+
+<!-- In terms of terminology, tolerance means, in this context, that some difference between the methods is tolerated (the measurements are still comparable in practice).
+Furthermore, the tolerance interval is exact and therefore more appropriate than the agreement interval, as the next sections will show.
+-->
+
+Toleranz in diesem Zusammenhang, dass ein gewisser Unterschied zwischen den Methoden toleriert wird (die Messungen sind in der Praxis noch vergleichbar).
+Außerdem ist das Toleranzintervall exakter und daher besser geeignet als das Übereinstimmungsintervall, wie die nächsten Abschnitte zeigen werden.
+
+Solche Toleranzintervalle sind mathematisch äquivalent zu Prädiktionsintervallen, auch wenn sich die Interpretationen unterscheiden.
+
+<!-- Francq -->
+<!-- https://onlinelibrary.wiley.com/doi/full/10.1002/sim.8709 -->
+<!-- However, this interval is approximate (too narrow) and several authors propose calculating a confidence interval around each bound.
+This article demonstrates that this approach is misleading, awkward, and confusing.
+On the other hand, tolerance intervals are exact and can include a confidence level if needed.
+Tolerance intervals are also easier to calculate and to interpret.
+-->
+
+<!-- The agreement interval is approximate and therefore, it should be interpreted with caution (especially for small sample sizes).
+Calculating a confidence interval for each bound of the agreement interval is awkward and brings even more confusion to its interpretation.
+While Hamilton and Stamey recommend calculating the AIs with its confidence intervals in their paper entitled “Using Bland-Altman to assess agreement between two medical devices-don't forget the confidence intervals!,”26 we have shown in this article that tolerance intervals are better and should be encouraged.
+-->
+
+<!-- In terms of terminology, tolerance means, in this context, that some difference between the methods is tolerated (the measurements are still comparable in practice).
+Furthermore, the tolerance interval is exact and therefore more appropriate than the agreement interval, as the next sections will show.
+-->
+
+<!-- Ludbrook already pointed out the complexity of the agreement interval with its confidence intervals while the tolerance interval is the appropriate solution.5  -->
+´
+
+<!-- , die beim ROISLIEN-Verfahren höher als bei POINT ist (und um ca. Faktor 3 höher als bei BOOT)  -->
+
+<!-- https://onlinelibrary.wiley.com/doi/full/10.1002/sim.8709 -->
+
+<!-- Ludbrook already pointed out the complexity of the agreement interval with its confidence intervals while the tolerance interval is the appropriate solution.5  -->
+
+<!-- The prediction interval here presented is equivalent to the beta-expectation Tolerance interval (also called “tolerance interval type I,” see Table 1).5, 18, 34 Their definition and interpretation are, however, different.
+A prediction interval is an interval where a future measurement (eg, a future difference between two medical devices) is expected to lie with a given confidence level.
+A beta-expectation tolerance interval is an interval where a given proportion of the population should lie, on average (eg, 95% of the differences).
+These two definitions lead to the same mathematical formula (3).
+-->
+
+#### Datensätze
 
@@
 Gerade bei Signalen mit sprunghaften Änderungen ist denkbar dass punktweise Intervalle zu plausibleren Ergebnissen führen können.
-<!-- Lenhoff geht z. B. nicht für nichtstionäre Signale! -->
+<!-- Nur glatte Kurven bzw. glatte Differenzkurven (was in der Praxis wohl auch meistens der Fall ist) ... vllt. haben -->
 
-KONFIDENZINTERVALLE VS. PRÄDIKTIONSINTERVALLE
-<!-- Prädiktionsbänder sind kontinuierliche Intervalle, in denen eine zukünftige Beobachtung einer Zufallsvariablen (Messkurve) mit einem gegebenen Konfidenzniveau enthalten ist.
-Anhand dieser Prädiktionsbänder werden Entscheidungen über die Form und den Effekt von Behandlungen für einzelne Probanden getroffen.
-Zu enge Prädiktionsbänder implizieren im beschriebenen Anwendungsfall, dass der Gang eines Probanden ein von der Norm abweichendes Muster aufweist, dass behandelt werden muss.
--->
-Der hier beschriebene Anwendungsfall (Prädiktionsbänder) deckt natürlich nur ein gewisses Spektrum relevanter Fragestellungen ab.
-In vielen weit verbreiteten Studiendesigns, z.
-B.
-Interventionsstudien, ist mit den breiten Prädiktionsbändern oder den noch breiteren Toleranzbändern nicht viel zu holen.
-Mit den im Paper kontruierten Bändern würde nahezu jeder Effekt im Messfehlintervall untergehen.
-In solchen Fällen Studien sind aber ohnehin eher Mittelwerte von Interesse, die in Signifikanztests auf Unterschiede untersucht werden.
-In diesem Fall sind die schmaleren Konfidenzintervalle eher geeignet um die Abweichung/Übereinstimmung zwischen den gemessenen Mittelwertkurven zweier Messsysteme zu beschreiben.
-<!-- Für Details zu Unterschieden in den coverage properties von LoA und herkömmlichen Prädiktionsintervallen sei an dieser Stelle auf Francq et al. (2020) verwiesen. -->
+<!-- Joch et al. (2019) -->
+<!-- https://link.springer.com/article/10.3758/s13428-018-1060-5 -->
 
-<!-- LENHOFF -->
-<!-- In some research studies it is important to determine confidence bands for the mean curve of a given population.
-For example, a researcher might want to characterize the effect upon the mean external varus moment of a group of subjects treated with a given type of knee brace.
-In this example, the researcher would wish to determine a confidence band for the mean difference in the external varus moment for subjects while wearing the brace and the external varus moment for the same subjects while not wearing the brace.
-If the confidence -->
-<!-- band about the difference of the braced and un-braced curves of each subject contains the horizontal line at value zero, then one could state that there was no significant difference between the two conditions.
-This application is illustrated in Fig.
-1.
+A more critical limitation of the FBRT procedure, as the authors described it, is that it cannot be applied to all kinds of continuous data.
+Lenhoff and colleagues fitted gait data with mathematical functions representing the data as closely as possible.
+However, recorded data cannot always be represented by adequate fits—for instance, when no function family exists that is capable of fitting all the entities (see Appendix 2).
+EEG data, for example, are characterized by rapid changes in voltage polarity that cannot reasonably be represented by a single mathematical function prototype.
+Another limitation of the FBRT, as the authors describe it, is that the C factor is calculated using the data from only one sample (mostly representing the data of a null/control condition).
+By looking at a single data point analysis, for example the t test, the estimation of the variance parameter would integrate the data from all measured samples, which should in consequence lead to a more reliable estimation.
 -->
 
-https://statmodeling.stat.columbia.edu/2010/12/05/what_do_practit/
-The key assumptions of a regression model are validity and additivity.
-Except when you’re focused on predictions, don’t spend one minute worrying about distributional issues such as normality or equal variance of the errors.
+<!-- In conclusion, all of the described methods to statistically analyze continuous data have inherent problems that limit their scopes of application.
+The FBRT of Lenhoff et al.
+(1999) seems to be the most appropriate and useful approach with respect to continuous data that are not normally distributed and that can be represented by mathematical functions.
+For cases in which the data cannot be fitted, however, and therefore are not suited for the FBRT procedure, we developed a method combining a point-by-point variance estimate with a resampling technique in order to calculate confidence bands according to the FBRT.
+In addition, the new method includes the calculation of a C factor that integrates the data of all samples.
+We call this method the point-based resampling technique (PBRT); and it will be explained in the following sections.
+For illustration purposes, we will demonstrate the application of this method using exemplary EEG data from an event-related potential (ERP) analysis.
+Although we demonstrate the PBRT in a between-groups comparison, this procedure can also be used with other continuous data and different experimental designs (e.g., within-subjects comparisons, as is described in Appendix 3).
+-->
 
+#### Grundsaetzliches / Sonstiges
 Wir empfehlen grundsätzlich die Daten zu plotten.
@@
 
-Die Erkenntnisse der vorliegenden Studie können nicht 1:1 auf andere Arten von Daten (z. B. Bodenreaktionskraftkurven, EMG Signale) übertragen, sondern gelten primär für glatte Kurven.
+A word on sample size in validation studies ...
 
@@
 
-Trotz der theoretisch und praktisch demonstrierten Überlegenheit funktionaler Verfahren konnten sich deren Einsatz im Kontext der Validierung von Messinstrumenten noch nicht recht durchsetzen.
-Wir spekulieren, dass dies in direktem Zusammenhang mit der Komplexität der Verfahren steht, die eine hohe Hürde für viele Praktiker mit weniger ausgeprägten Statistik- und Programmierkenntnissen darstellt.
-Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code frei zur Verfügung (GPL).
+Als Ansatz für die Konstruktion besserer Bänder: Nichtparametrischer Ansatz mit Perzentilen (asymmetrische Bänder)
 
-Angesichts der überragenden Bedeutung der Kenntniss des Messfehlers ist es erstaunlich wie wenig sich viele Studien mit der Frage nach der Größe des Messfehlers beschäftigen.
-Gut demonstrieren lässt sich das am Beispiel von IMU's ...
+<!-- Thus, this function-based resampling technique (FBRT), as we call it, is an advanced tool that can be applied even to data with asymmetric distributions. -->
 
-Als Ansatz für die Konstruktion besserer Bänder:
-Nichtparametrischer Ansatz mit Perzentilen (asymmetrische Bänder)
-
-Wir glauben dass die geringe Verbreitung kontinuierlicher Methoden trotz erwiesenermaßen gut funktionierender Methoden in der Praxis an der programmiertechnischen Umsetzung scheitert.
-Wir hoffen mit diesem Paper und dem Code im Anhang einen Beitrag dazu zu leisten, dass mehr Forscher diese Methoden adaptieren und veralteten/ungeeigneten Methoden (Korrelationskoeffizienten etc.) den Rücken kehren.
-
-<!-- Es gibt auch immer mehr readily available (und kostenlose) Lösungen, siehe https://mjskay.github.io/ggdist/articles/lineribbon.html -->
-
-<!-- Limitations of curvewise intervals -->
-<!-- https://mjskay.github.io/ggdist/articles/lineribbon.html -->
-<!-- One challenge with curvewise intervals is that they can tend to be very conservative, especially at moderate-to-large intervals widths. -->
-<!-- Notice how noisy the curvewise intervals are.
-In addition, because a number of curves tend to start low and end high (or vice versa), above 50%, the bands rapidly expand to cover almost all of the curves in the sample, regardless of coverage level.
--->
-<!-- In general I have found that there is no one method that consistently works well on all datasets.
-No matter the method, intervals often become problematic above 50%, hence the default .width for curve_interval() is 0.5 (unlike the default for point_interval(), which is 0.95).
-In any case, caution when using these intervals is advised.
--->
-
-<!-- Welcher coverage Strategie ist besser (Entire curve (all-or-nothing) vs. percent-of-curve)? -->
-<!-- Denkansatz aus Juul et al. (2020) -->
-<!-- At the same time, however, it is clear that there is still important research to be done. We have suggested different ways of ranking centrality of ensemble curves. Each rank- -->
-<!-- ing has its merits: For example, the all-or-nothing ranking penalizes curves which are shaped differently than other ensemble curves and a weighted ranking can reward curves that are central in the very near future.
-A thorough investigation of these merits and trade-offs is needed.
-Until this has been investigated, we encourage researchers to be creative and mindful about the problems that each statistical method could have.
-And to communicate this uncertainty openly to decision makers.
--->
-
-<!-- Für die hier untersuchten biomechanischen Signale nicht so relevant, aber an sich ein sehr interessantes Feature ist das hier: -->
-<!-- Leicht abgewandelt nach Juul et al. (2020): -->
-<!-- In addition to the ranking methods (for finding quantiles/percentiles) mentioned above, one can rank the curves according to some feature of interest.
-In Fig.
-2E we show the curve box plots obtained when we rank curves according to their projected maximum values of newly hospitalized cases in a single day; in other words, the median projected peak value received -->
-<!-- the highest centrality. -->
-
-<!-- Interessant beim Juul Paper ist auch der Umgang mit ("abgeschnittenen") Extremwerten: -->
-<!-- Here we have argued that computing confidence intervals using fixed-time descriptive systematically suppresses trajectory extremes. This is natural. -->
-<!-- Fixed-time descriptive statistics are designed to show the least extreme predictions on a given date, not to take entire curves into account. -->
-
-<!-- <!-- Es gibt eine Reihe von Ansätzen aus anderen Bereichen (z. B. Covid Prädiktion), bei denen sog. Function oder Curve Boxplots eingesetzt werden. -->
-<!-- <!-- https://mjskay.github.io/ggdist/reference/curve_interval.html -->
-<!-- <!-- Hier referenzierte Paper u. a.  -->
-<!-- <!-- - Juul Jonas, Kaare Græsbøll, Lasse Engbo Christiansen, and Sune Lehmann.
-(2020).
-"Fixed-time descriptive statistics underestimate extremes of epidemic curve ensembles".
-arXiv e-print.
-arXiv:2007.05035 -->
-<!-- <!-- - Mirzargar, Mahsa, Ross T Whitaker, and Robert M Kirby.
-(2014).
-"Curve Boxplot: Generalization of Boxplot for Ensembles of Curves".
-IEEE Transactions on Visualization and Computer Graphics.
-20(12): 2654-2663.
-doi: 10.1109/TVCG.2014.2346455 -->
-
-<!-- <!-- See Mirzargar et al. (2014) or Juul et al. (2020) for an accessible introduction to data depth and curve boxplots / functional boxplots. -->
+Wichtig ist zudem dass die ROISLIEN-Methode - im Gegensatz zu BOOT - kein vollständig funktionales Verfahren ist, da im Anschluss an die Approximation der Kurven mithilfe von Fourrierreihen eine punktweise Berechnung der Prädiktionsintervallgrenzen stattfindet.
+Wir raten angesichts der höheren Unsicherheit, der schlechteren coverage properties und der höheren Komplexität im Vergleich mit dem POINT-Ansatz vom Einsatz des ROISLIEN-Verfahrens ab.
 
 # Conclusion
+
+Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
+In diesem Sinne stellen wir auch den zugehörigen R Code im Anhang zur freien Verfügung.
+Für die Zukunft ist auch ein R Paket geplant, dass dann womöglch auch andere Intervalle (Toleranzintervalle, Konfidenzintervalle) enthält.
+
+Verwendet messwiederholte Daten!
 

### 2021-11-12 R/distance_2_floa.R
Update

@@ -1,51 +1,0 @@
-# Average distance to the outer point of the curve set
-distance_2_floa <- function (data, floa) {
-
-  # floa variables need to be arranged in columns
-  if (dim(floa)[1] < dim(floa)[2]) {
-    floa <- t(floa)
-  }
-
-  if (!is.data.frame(floa)) {
-    floa <- as.data.frame(floa)
-  }
-
-  device1 <- data.frame(subset(data, device == "IMU")$value)
-  device2 <- data.frame(subset(data, device == "MC")$value)
-  device.diff <- device1 - device2
-  colnames(device.diff)[1] <- "value"
-
-  device.diff$frame <- seq(0, 100)
-  n.frames <- length(unique(data$frame))
-  n.strides <- length(unique(data$strideID))
-  n.subjects <- length(unique(data$subjectID))
-  strides.per.subject <- length(unique(data$strideID)) / n.subjects
-  device.diff$strideID <- as.factor(rep(1:n.strides, each = 101))
-  device.diff$subjectID <- as.factor(rep(1:n.subjects, each = strides.per.subject * n.frames))
-
-  upr.dist <- vector(mode = "list", length = n.frames)
-  lwr.dist <- vector(mode = "list", length = n.frames)
-  for (frame.idx in seq(0, 100)) {
-    data.per.frame <- subset(device.diff, subset = frame == frame.idx)
-
-    curve.extrema <- range(data.per.frame$value)
-    curve.max <- max(curve.extrema)
-    curve.min <- min(curve.extrema)
-
-    # Account for the loop index starting at 0
-    frame.idx.correct <- ifelse(frame.idx == 0, frame.idx + 1, frame.idx)
-
-    floa.upr <- floa$upper[frame.idx + 1]
-    floa.lwr <- floa$lower[frame.idx + 1]
-
-    upr.dist[[frame.idx.correct]] <- floa.upr - curve.max
-    lwr.dist[[frame.idx.correct]] <- floa.lwr - curve.min
-  }
-
-  distance <- data.frame(unlist(upr.dist),
-                         unlist(lwr.dist))
-
-  colnames(distance) <- c("upr.dist", "lwr.dist")
-
-  return(distance)
-}

### 2021-11-12 R/estimate_uncertainty_loa.R

@@ -22,7 +22,6 @@ estimate_uncertainty_loa <- function (data, n.boot) {
   }
 
   # Pointwise LoA return the same estimate every time (all curves are included)
-  # (Therefore, they are outside the loop)
   floa.point <- data.frame(t(floa_point(data)))
 
   floa.roislien.upper.unlist <- matrix(unlist(floa.roislien.upper), 101)
@@ -132,15 +131,15 @@ estimate_uncertainty_loa <- function (data, n.boot) {
 
 
   # (Uncertainty) Area between 2.5 and 97.5 borders
-  uncert.area.limits <- c(sum(pi95$floa.roislien.upper.97.5) - sum(pi95$floa.roislien.upper.2.5),
-                          sum(pi95$floa.roislien.lower.97.5) - sum(pi95$floa.roislien.lower.2.5),
-                          sum(pi95$floa.lenhoff.upper.97.5) - sum(pi95$floa.lenhoff.upper.2.5),
-                          sum(pi95$floa.lenhoff.lower.97.5) - sum(pi95$floa.lenhoff.lower.2.5),
-                          sum(pi95$floa.point.upper.97.5) - sum(pi95$floa.point.upper.2.5),
-                          sum(pi95$floa.point.lower.97.5) - sum(pi95$floa.point.lower.2.5)
+  # POINT values are identical for the upper and lower limits as symmetric confidence intervals were calculated.
+  uncert.area.limits <- c(round(sum(pi95$floa.point.lower.97.5) - sum(pi95$floa.point.lower.2.5), 2),
+                          round(sum(pi95$floa.roislien.upper.97.5) - sum(pi95$floa.roislien.upper.2.5), 2),
+                          round(sum(pi95$floa.roislien.lower.97.5) - sum(pi95$floa.roislien.lower.2.5), 2),
+                          round(sum(pi95$floa.lenhoff.upper.97.5) - sum(pi95$floa.lenhoff.upper.2.5), 2),
+                          round(sum(pi95$floa.lenhoff.lower.97.5) - sum(pi95$floa.lenhoff.lower.2.5), 2)
                           )
 
-  names(uncert.area.limits) <- c("au.roislien", "al.roislien", "au.lenhoff", "al.lenhoff", "au.point", "al.point")
+  names(uncert.area.limits) <- c("POINT", "ROISLIEN (U)", "ROISLIEN (L)", "BOOT (U)", "BOOT (L)")
 
   return(uncert.area.limits)
 }

### 2021-11-12 R/floa_point.R

@@ -51,13 +51,12 @@ floa_point <- function(data) {
                           )
 
   # PRECISION OF ESTIMATED LIMITS OF AGREEMENT
-  # Confidence intervals for the 95% limits of agreement (Bland & Altman, 1986)
-  # The standard error of the limits is approximately sqrt(3*s^2 / n)
+  # Confidence intervals for the 95% limits of agreement (Bland & Altman, 1999)
   n.curves <- length(unique(data$strideID))
-  # standard.error.d2s <- sqrt(3*floa.point$total.sd^2 / n.curves)
+  # Formula from Francq et al. (2020)
   standard.error.d2s <- sqrt(1/n.curves + (z0.975^2 / (2*(n.curves-1)))) * floa.point$total.sd
 
-  # t-score of the 95th quantile of the Student t distribution with df = (n.curves - 1)
+  # t-score of the 95th quantile of the Student t distribution with df = (n - 1)
   t <- qt(.975, df = n.curves - 1)
 
   floa.point.bnd <- rbind(floa.point.bnd,

### 2021-11-12 R/get_coverage_fraction.R

@@ -2,17 +2,10 @@ get_coverage_fraction <- function (data, floa) {
 
   # Get difference curves
   # ----------------------------------------------------------------------------
-  device1 <- data.frame(
-    subset(data, device == "IMU")
-  )
+  device1 <- data.frame(subset(data, device == "IMU"))
+  device2 <- data.frame(subset(data, device == "MC"))
 
-  device2 <- data.frame(
-    subset(data, device == "MC")
-  )
-
-  device.diff <- subset(device1,
-                        select = c(subjectID, strideID, frame))
-
+  device.diff <- subset(device1, select = c(subjectID, strideID, frame))
   device.diff$value <- device1$value - device2$value
 
 
@@ -27,19 +20,19 @@ get_coverage_fraction <- function (data, floa) {
   stride.indices <- unique(data$strideID)
 
   for (stride.idx in stride.indices){
-
     curve <- subset(device.diff, strideID == stride.idx)
 
     # Compare difference curves with upper and lower boundaries
     below.thresh <- curve$value < lwr.bnd
     above.thresh <- curve$value > upr.bnd
 
-    # Mean percentage of points outside the limits
+    # Percentage of points outside the limits
     inside <- below.thresh == above.thresh
     inside.thresh.perc <- c(inside.thresh.perc,
                             sum(inside[TRUE]) / length(curve$value))
   }
 
+  # Mean percentage of points outside the limits
   coverage <- round(mean(inside.thresh.perc), 3)
 
   return(coverage)

### 2021-11-12 R/plot_loa.R

@@ -24,9 +24,10 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
 
   device.diff$frame <- seq(0, 100)
   n.frames <- length(unique(data$frame))
-  n.strides <- length(unique(data$strideID))
+  n.strides <- 110 # length(unique(data$strideID))
   n.subjects <- length(unique(data$subjectID))
-  strides.per.subject <- length(unique(data$strideID)) / n.subjects
+  strides.per.subject <- round(length(unique(data$strideID)) / n.subjects) # round() needed because one of the data sets ("imu_mc") is unbalanced and will result in error later on
+  # strides.per.subject <- length(unique(data$strideID)) / n.subjects
   device.diff$strideID <- as.factor(rep(1:n.strides, each = 101))
   device.diff$subjectID <- as.factor(rep(1:n.subjects, each = strides.per.subject * n.frames))
 
@@ -69,13 +70,13 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
               size = 1.5,
               colour = "#009E73") +
     scale_y_continuous(limits = c(ylim[1], ylim[2])) +
-    labs(x = "Time-normalized signal [%]", y = "Difference") +
-    theme_minimal() +
-    theme(axis.text.x = element_text(size = 20),
-          axis.title.x = element_text(size = 22),
-          axis.text.y = element_text(size = 20),
-          axis.title.y = element_text(size = 22),
-          legend.position = "none")
+    labs(x = "Time-normalized signal [%]", y = "Difference")
+    # theme_minimal() +
+    # theme(axis.text.x = element_text(size = 20),
+    #       axis.title.x = element_text(size = 22),
+    #       axis.text.y = element_text(size = 20),
+    #       axis.title.y = element_text(size = 22),
+    #       legend.position = "none")
 
   PLOT.DIFF
 

### 2021-11-12 R/simulate_data.R

@@ -23,10 +23,11 @@
 imu.mc <- readRDS(paste0(dir.data, "/", "data.rds"))
 
 # Pick 10 random strides from each of the 11 subjects
-stride.idx <- imu.mc %>% dplyr:::group_by(subjectID)  %>% dplyr::summarize(sample_n(data.frame(strideID), 10, replace = FALSE)) # sample(strideID, size = 10, replace = FALSE)
+stride.idx <- imu.mc %>% dplyr:::group_by(subjectID) %>% dplyr::summarize(sample_n(data.frame(strideID), 10, replace = FALSE))
 stride.idx <- stride.idx$strideID
 stride.idx <- as.numeric(stride.idx)
 
+# Detect and change all duplicated strides
 double.idx <- which(duplicated(stride.idx))
 stride.idx[double.idx] <- double.idx + 1
 imu.mc.balanced <- imu.mc[imu.mc$strideID %in% stride.idx, ]
@@ -81,9 +82,6 @@ for (subj.idx in 1:n.subj) {
 
     mc <- rnorm(1, offset.mean, 0.05) + sine.1 + sine.2
     imu <- rnorm(1, offset.mean, 0.05) + sine.3 + sine.4
-
-    # plot(mc, type = "l")
-    # lines(imu, col = "red")
 
     value.subj <- c(value.subj, c(imu, mc))
 
@@ -153,9 +151,6 @@ for (subj.idx in 1:n.subj) {
     mc <- sine.1 + sine.2
     imu <- offset + sine.3 + sine.4
 
-    # plot(mc, type = "l")
-    # lines(imu, col = "red")
-
     value.subj <- c(value.subj, c(imu, mc))
 
     device.imu <- rep("IMU", n.frames)
@@ -223,9 +218,6 @@ for (subj.idx in 1:n.subj) {
     mc <- sine.1 + sine.2
     imu <- offset + sine.3 + sine.4
 
-    # plot(mc, type = "l")
-    # lines(imu, col = "red")
-
     value.subj <- c(value.subj, c(imu, mc))
 
     device.imu <- rep("IMU", n.frames)
@@ -292,9 +284,6 @@ for (subj.idx in 1:n.subj) {
 
     mc <- rnorm(1, offset.mean, 0.05) + sine.1 + sine.2
     imu <- rnorm(1, offset.mean, 0.05) + sine.3 + sine.4 + trend
-
[…]

### 2021-11-12 R/singlecurve_coverage.R

@@ -16,9 +16,6 @@ singlecurve_coverage <- function (data, n.boot) {
   # ####################################################################
   n.curves <- unique(data$strideID)
 
-  # cover.cross.rcb.v1   <- vector(mode = "list", length = length(n.curves))
-  # cover.cross.rcb.v2   <- vector(mode = "list", length = length(n.curves))
-  # cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
   cover.cross.lenhoff <- vector(mode = "list", length = length(n.curves))
@@ -28,9 +25,6 @@ singlecurve_coverage <- function (data, n.boot) {
     data.one.out <- subset(data, strideID != curve.idx)
 
     floa.point    <- floa_point(data.one.out)
-    # floa.v1       <- floa_rcb(data.one.out, n.boot, ver = "v1")
-    # floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
-    # floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out)
     floa.lenhoff  <- floa_lenhoff(data,
                                   k_reihe = 50,
@@ -47,18 +41,12 @@ singlecurve_coverage <- function (data, n.boot) {
     device.diff <- device1$value - device2$value
 
     # Get coverage for the left out (difference) curve -------------------
-    # cover.cross.rcb.v1[curve.idx]   <- get_coverage_singlecurve(device.diff, floa.v1)
-    # cover.cross.rcb.v2[curve.idx]   <- get_coverage_singlecurve(device.diff, floa.v2)
-    # cover.cross.rcb.v3[curve.idx]   <- get_coverage_singlecurve(device.diff, floa.v3)
     cover.cross.point[curve.idx]      <- get_coverage_singlecurve(device.diff, floa.point)
     cover.cross.roislien[curve.idx]   <- get_coverage_singlecurve(device.diff, floa.roislien)
     cover.cross.lenhoff[curve.idx]    <- get_coverage_singlecurve(device.diff, floa.lenhoff)
   }
 
   cover.cross <- cbind(
-                      # unlist(cover.cross.rcb.v1),
-                      # unlist(cover.cross.rcb.v2),
-                      # unlist(cover.cross.rcb.v3),
[…]

### 2021-11-12 paper.Rmd
Update paper draft

@@
 library(ggplot2)
-# library(boot)
 
@@
 source("plot_cov_ver.R")
-# source("distance_2_floa.R")
 source("estimate_uncertainty_loa.R")

### 2021-11-23 paper.Rmd
Continue paper draft

@@
 ---
-title: Comparison of continuous prediction bands for the validation of biomechanical curve data
+title: Comparison of continuous prediction intervals for the validation of biomechanical curve data
 author:
@@
 Lenhoff et al.
-(1999) haben das Problem punktweise konstruierter Prädiktionsbänder am Beispiel von 1D Gelenkwinkelverläufen (Kurven) demonstriert.
+(1999) haben das Problem punktweise konstruierter Bänder am Beispiel von 1D Gelenkwinkelverläufen (Kurven) demonstriert.
 Dazu verglichen sie punktweise berechnete Prädiktionsbänder gegenüber einer funktionalen Methode bei der Kurvendaten zunächst mithilfe von Fourierfunktionen approximiert werden und aus den so gebildeten funktionalen Kurven durch Bootstrapping Prädiktionsbänder gebildet werden (Sutherland, Olshen, Biden & Wyatt (1988, 1989)).
@@
 Limits of Agreement (LoA: 95% der Differenzen zwischen zwei Messsystemen sind im Intervall mittlere Differenz $\pm$ 1.96 Standardabweichungen der Differenzen enthalten), für Kurvendaten erweitert wird.
-LoA sind das am weitesten verbreitete verfahren in Methodenvergleichen (Ludbrook, 2010) und stellen im Wesentlichen Prädiktionsintervalle dar (obwohl das tatsächliche Prädiktionslevel der LoA stets niedriger ist als das nomale Niveau und für kleine Stichproben kollabiert (Francq, 2020)).
+LoA sind das am weitesten verbreitete Verfahren in Methodenvergleichen (Ludbrook, 2010) und stellen per Deinfition Prädiktionsintervalle dar (obwohl das tatsächliche Prädiktionslevel der LoA stets niedriger ist als das nomale Niveau und für kleine Stichproben kollabiert (Francq, 2020)).
 
@@
 # Results
-
 Figure 2 enthält die qualitative Verläufe der Prädiktionsbänder der drei Methoden gegenüber den Differenzkurven.
@@
 Die Prädiktionsbänder der punktweisen Verfahren (POINT, ROISLIEN) fallen zum Teil deutlich schmaler aus und unterrepräsentieren den Messfehler in Bezug auf die Abdeckung ganzer Kurven.
-Es zeigt sich jedoch dass die Performance insgesamt sehr gut ist und die Differenzkurven nur an wenigen Stellen außerhalb der Prädiktionsgrenzen liegen.
+Bezüglich des Anteils der abgedeckten Punkte pro Kurve erzielen die punktweise konstruierten Bänder sehr gute coverage Werte (Median 100% über alle $D = 110$ Iterationen der leave-one-out Kreuzvalidierung).
 
@@
 Dabei ist die Unsicherheit der Intervallgrenzen beim ROISLIEN-Verfahren etwas höher als beim POINT-Verfahren und nahezu 3x so hoch wie beim BOOT Verfahren.
-
 Die BOOT-Bänder weisen eine deutlich geringere Streuung der Intervallgrenzen auf als die beiden übrigen Methoden.
@@
 # Discussion
-Die Tatsache dass ein erheblicher Teil biomechanischer Messsysteme höherdimensionale Daten liefert, spiegelt sich bislang kaum in der Methodik von Validierungsstudien wieder.
-Um die Aussagekraft von Validierungsstudien diesbezüglich zu erhöhen braucht es Fehlerintervalle die alle Punkte des Messsignals.
-Zu diesem Zweck wurden im vorliegenden Paper drei relevante Verfahren (POINT, ROISLIEN, BOOT) zur Berechnung kontinuierlicher Prädiktionsintervalle identifiziert und für ausgewählte Messfehlerszenarien am Beispiel von 1D Kurvendaten empirisch analysiert.
+Die Tatsache, dass ein erheblicher Teil biomechanischer Messsysteme höherdimensionale Daten (> 0D) liefert, spiegelt sich bislang kaum in der Methodik von Validierungsstudien wieder.
+Um die Aussagekraft von Validierungsuntersuchungen diesbezüglich zu erhöhen braucht es Fehlerintervalle, die alle Punkte eines Messsignals berücksichtigen.
+Zu diesem Zweck wurden im vorliegenden Paper drei relevante Verfahren zur Berechnung kontinuierlicher Prädiktionsintervalle identifiziert (POINT, ROISLIEN, BOOT) und für ausgewählte Messfehlerszenarien am Beispiel von 1D Kurvendaten analysiert.
+BOOT ist ein funktionales Verfahren in dem diskrete Kurven mit mathematischen Funktionen approximiert werden und alle weiteren Berechnungen anhand dieser funktionalen Objekten vorgenommen werden.
+Die ROISLIEN-Methode wurde als funktionales Verfahren entwickelt, muss aber strenggenommen (wie POINT) zu den punktweisen Verfahren gezählt werden, da die eigentliche Berechnung der Intervallgrenzen separat für jeden Messzeitpunkt erfolgt.
 
 Die Intervalle der BOOT-Methode erreichen in allen untersuchten Szenarien die nominale coverage und sind mit weniger estimation uncertainty belastet als die beiden übrigen Verfahren (POINT, ROISLIEN).
-Diese Performance ist insofern zu erwarten, dass die BOOT-Bänder anhand eines Korrekturfaktors auf das nominale Niveau hin optimiert werden.
-Dies ist auch ein wesentlicher Unterschied zu den parametrischen POINT und ROISLIEN Methoden, bei denen die Intervallgrenzen auf Grundlage der Verteilung der Differenzen in der Stichprobe geschätzt werden.
+Dies ist insofern zu erwarten, als dass die BOOT-Bänder anhand eines Korrekturfaktors auf das nominale Niveau optimiert werden.
+Dies ist ein wesentlicher Unterschied zu den parametrischen POINT und ROISLIEN Methoden, bei denen die Intervallgrenzen auf Grundlage der Verteilung der Differenzen in der Stichprobe geschätzt werden.
 
 Die punktweise berechneten Bänder fallen erwartungsgemäß enger aus als die funktionalen BOOT-Bänder (Figure 2, Table 1 & 2).
-Die Prädiktionsintervalle der punktweisen Verfahren performen jedoch durchaus sehr gut wenn man den Anteil der abgedeckten Punkte pro Kurve über alle $D = 110$ Iterationen der leave-one-out Validierung als Maß für die Performance verwendet.
+Die coverage der punktweise konstruierten Bänder ist dennoch ausgezeichnet wenn man den Anteil der abgedeckten Punkte pro Kurve über alle $D = 110$ Iterationen der leave-one-out Kreuzvalidierung als Bewertungsgrundlage nimmt.
 Die zugehörige Verteilung ist extrem linksschief, d.
 h.
-bis auf wenige Kurven liegen stets die allermeisten Kurvenpunkte im Prädiktionsintervall.
-Die ROISLIEN-Methode wird hier im Übrigen nicht als funktionales, sondern als punktweises Verfahren gezählt, da im Anschluss an die Approximation der Kurven mithilfe von Fourrierreihen eine punktweise Berechnung der Prädiktionsintervallgrenzen stattfindet.
+bis auf wenige Ausnahmen liegen stets die allermeisten Punkte der Testkurven im trainierten Prädiktionsintervall.
 
-Die Verläufe der Intervallgrenzen im POINT und BOOT-Verfahren Verfahren sind qualitativ ähnlich, die coverage (Table 1 & 2) ist beim ROISLIEN-Verfahren jedoch etwas niedriger.
-Die Ursache hierfür liegt in der höheren Streuung der ROISLIEN-Intervallgrenzen über $D = 110$ Iterationen des leave-one-out Validierungsverfahrens, da hier nur eine Kurve je Proband verwendet wird und kleinere Stichproben generell einer größeren Unsicherheit unterliegen.
-Bei der POINT-Methode hingegen werden stets alle Kurven im Datensatz berücksichtigt, womit es für die Daten der Stichprobe keine Unsicherheit bezüglich der Intervallgrenzen gibt.
-Um die inhärente Unsicherheit des Samplingprozesses zu berücksichtigen schlagen Bland & Altman allerdings vor die LoA um Konfidenzbänder zu erweitern.
-Der Einsatz zusätzlicher Konfidenzbänder führt die ursprüngliche Idee einfach zu berechnender, intuitiver Unsicherheitsintervalle jedoch ein Stück weit ad absurdum (Francq et al., 2020), weshalb in dieser Arbeit auf eine Verbreiterung der LoA durch Konfidenzbänder verzichtete wurde.
-Ohnehin ist die korrekte Umsetzung der im vorliegenden Paper umgesetzten POINT-Methode nach Bland & Altman (1999) (mit Varianzanteilen aus einer ANOVA mit Messwiederholung etc.) deutlich komplexer als das im ROISLEN-Verfahren eingesetzte minimale Bland & Altman Modell ($\overline{d} \pm 1.96SD$).
-<!-- "would complicate the analysis, makes it awkward and confusing, and is useless as the exact CI with the t-distribution is available"  -->
-
-Die punktweisen Verfahren liefern zudem auch für die Datensätze plausible Verläufe, in denen erhebliche Verletzungen der Modellannahmen vorliegen (nicht-gaussverteilten Differenzen, phase shift in x-axis direction).
-Für die Praxis ist das eine wichtige Erkenntnis, da diese Annahmen beim Vergleich verschiedener Messmethoden häufig verletzt sind.
-Die Verteilung der Differenzen zwischen zwei Systemen wird z.
+Die punktweisen Verfahren liefern auch für Datensätze plausible Verläufe, in denen erhebliche Verletzungen der parametrischen Modellannahmen vorliegen (nicht-gaussverteilten Differenzen, phase shift in x-axis direction).
+Für die Praxis ist das eine wichtige Erkenntnis, da diese Annahmen häufig verletzt sind, z.
 B.
-erheblich verzerrt, wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder zu geringer Auflösung abgeschnitten werden, oder wegen (nichtlinearem) Drift.
+wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder zu geringer Auflösung abgeschnitten werden, oder in Gegenwart von nichtlinearem Drift.
 Speziell im der letzten Fall sind Alternativen zum BOOT-Verfahren von Interesse, da die Methode nur für stationäre Signale zuverlässig funktioniert (Lenhoff et al., 1999).
 
-Die BOOT-Intervalle fallen stellenweise deutlich breiter aus als die punktweisen.
-Das resultiert in hervorragenden coverage Werten, produziert aber auch stark konservative Fehlerbänder.
-Fairerweise muss man sagen das die BOOT-Intervalle nie für diesen Einsatzzweck entwickelt wurden.
-Im Allgemeinen sind Prädiktionsintervalle überall dort von Interesse, wo neue Messwerte gegen normative Daten (z.
+Die Verläufe der Intervallgrenzen im POINT und ROISLIEN-Verfahren Verfahren sind qualitativ ähnlich, die coverage (Table 1 & 2) ist beim ROISLIEN-Verfahren jedoch etwas niedriger.
+Die Ursache hierfür liegt in der höheren Streuung der ROISLIEN-Intervallgrenzen über $D = 110$ Iterationen des leave-one-out Validierungsverfahrens, da hier nur jeweils eine Kurve pro Proband gezogen wird und kleinere Stichproben einer größeren Unsicherheit unterliegen.
+Bei der POINT-Methode hingegen werden stets alle Kurven im Datensatz berücksichtigt.
+Um der prinzipiellen Unsicherheit von Samplingprozessen Rechnung zu tragen, empfehlen Bland & Altman jedoch die Erweiterung der LoA um Konfidenzbänder.
+Dieser Vorschlag führt die ursprüngliche Idee einfach zu berechnender, intuitiver Unsicherheitsintervalle allerdings ein Stück weit ad absurdum (Francq et al., 2020), weshalb in der vorliegenden Arbeit auf eine Verbreiterung der LoA durch Konfidenzbänder verzichtet wurde.
+Ohnehin ist die korrekte Umsetzung der POINT-Methode nach Bland & Altman (1999) (mit Varianzanteilen aus einer ANOVA mit Messwiederholung etc.) deutlich komplexer als das im ROISLEN-Verfahren eingesetzte minimale Bland & Altman Modell ($\overline{d} \pm 1.96SD$).
+Dadurch fällt auch der eigentliche Vorteil punktweiser Intervalle gegenüber funktionalen Verfahren - die Einfachheit der Berechnung - erheblich geringer aus.
+<!-- Ludbrook already pointed out the complexity of the agreement interval with its confidence intervals while the tolerance interval is the appropriate solution.5  -->
+
+Das punktweise Vorgehen hat noch einen anderen möglichen Vorteil gegenüber funktionalen Methoden: Sie können auf alle Arten kontinuierlicher Daten angewendet werden.
+Im Gegensatz zu den hier untersuchten glatten Kurven lassen sich mathematische Funktionen auf Signale mit vielen sprunghaften Veränderungen im Signalverlauf wie z.
+B.
+EMG-Signale nicht fitten.
+Die Untersuchung dieser Art von Kurven war allerdings nicht Gegenstand des vorliegenden Papers.
+Die hier vorgestellten Datensätze decken dennoch ein breites Spektrum biomechanischer Signale ab.
+
+#### Auswahl geeigneter Fehlerbänder
+Abgesehen von Unterschieden in der Berechnung und den coverage-Eigenschaften ist allen drei untersuchten Methoden gemein, dass Prädiktionsintervalle konstruiert werden.
+Um eine möglichst hohe coverage zu erzielen fallen diese Intervalle daher gezielt breit aus.
+Das ist im Sinne von Bland & Altman erwünscht um einen möglichst hohen Anteil von Einzelbeobachtungen in den Intervallen zu enthalten und so eine gewisse Sicherheit zu haben, dass auch die nächste Einzelmessung im Intervall liegt.
+Im biomechanischen Kontext ist das z.
+B.
+dann relevant, wenn neue Messwerte gegen normative Daten (z.
 B.
 aus einer Referenzdatenbank) verglichen werden.
-Wenn die Abweichung ein bestimmtes Maß überschreitet, d.
+Wenn hier die Abweichung ein bestimmtes Maß überschreitet, d.
 h.
-der Proband nicht zur Referenzpopulation gehört, dann werden - wie bei Olshen et al.
-(1989) - bestimmte pathologische Muster angenommen und eine - wie auch immer geartete - Therapie vorgeschlagen.
-Hier ist es also wichtig dass die Bänder nicht zu eng ausfallen, weil die Konsequenzen möglicher $\beta$ Fehler in diesem Kontext schwerwiegend sind.
+der Proband nicht zur Referenzpopulation gehört, werden bestimmte pathologische Muster angenommen und eine geeignete Therapie vorgeschlagen (Olshen et al., 1989).
+Entsprechend ist in diesem Kontext von Bedeutung, dass die Bänder nicht zu eng ausfallen, weil die Konsequenzen falsch positiver Schlüsse schwerwiegend sein können.
 
-Generell vermitteln die punktweisen Bänder (insbesondere POINT) ein deutlich realistischeres Bild des zufälligen Messfehlers.
-Für Differenzkurven zwischen zwei biomechanischen Messsystemen braucht es eigentlich keine Prädiktionsbänder die deutlich außerhalb der äußersten Kurve liegen wie bei BOOT da hier keine große Sprünge im Kurvenverlauf oder Ausreißer zu erwarten zu erwarten sind (vorausgesetzt die Stichprobe ist ausreichend groß und repräsentativ).
+Für viele andere Fragestellungen sind die breiten Prädiktionsintervalle allerdings nicht geeignet weil breitere Intervalle ein höheres Risiko für falsch negative Schlüsse bedeuten.
+Das ist z.
+B.
+der Fall wenn die Messungen im Kontext eines Vergleichs verschiedener Probandenkollektive erhoben werden (z.
+B.
+in Interventionsstudien oder case-control Studien).
+Gut demonstrieren lässt sich das am Beispiel der Differenzen in Fig.
+2 D (Differenzen zwischen einem optischen motion capture System und einer IMU).
+Eine mittels IMU gemessene Differenz zwischen den sagittalen Hüftwinkeln z.
+B.
+einer Kontroll- und Interventionsgruppe kann hier nur dann als tatsächlicher Effekt angenommen werden, wenn sie mehr als die (je nach Intervall ungefähr) $\pm$ 10° des zufälligen Messfehlers beträgt.
+Derartig große Effekte sind in biomechanischen Untersuchungen allerdings nur selten zu erwarten.
+<!-- Dies gilt umso mehr vor dem Hintergrund häufig eher geringer Effektstärken in der Biomechanik (Quelle ... Knudson? Metaanalysen!?).  -->
+<!-- Typische Effektgrößen liegen hier eher bei ... (Quelle). -->
+Die hier beobachtete Streuung der Differenzen ist dabei längst kein Szenario das weit von der Praxis entfernt ist.
+Gerade bei markerbasierten motion capture Systemen, die häufig als ground truth eingesetzt werden, ist die Reliabilität aufgrund von u.
+a.
+Weichteilbewegungen, Markerplatzierungsartefakten etc.
+fragwürdig, so dass hier leicht sehr breite (Prädiktions)Intervalle entstehen können.
+<!-- Werte/Studie zur Reliabilität von MoCap Systemen? -->
 
-Apropos unrealistisch weite Intervalle: In bestimmten Fällen, wie z.
-B.
-dem realweltlichen Datensatz fallen die Bänder aller drei Methoden so breit aus, dass mit dem neuen Messverfahren im Rahmen eines between-subject Designs kaum ein Effekt detektiert werden kann, der über den zufälligen Messfehler hinausgeht (Figure 2,  D).
-Anhand der Kurvenverläufe ist aber auch erkennbar, dass die Kurven geclustert auftreten weil die Streuung innerhalb der Probanden geringer ausfällt.
+Eine Alternative zu den breiten Prädiktionsintervallen sind Intervalle auf Grundlage der within-subject Standardabweichung (aus einer einfaktoriellen Varianzanalyse) (Bland & Altman, 1999).
+Wiederum am Beispiel Hüftgelenksdifferenzen (Fig.
+2, D) lässt sich erkennen, dass die Streuung innerhalb einzelner Probanden erheblich geringer ausfällt (erkennbar an den probandenweisen Kurvenclustern).
 Diese Eigenschaft kann ausgenutzt werden um engere Intervalle auf Grundlage der intraindividuellen Streuung zu konstruieren wenn primär die Reproduzierbarkeit der neuen Methode von Bedeutung ist.
@@
 bei prä-post-Vergleichen in Interventionsstudien).
-Bland & Altman (1999) schlagen hierfür vor Intervalle auf Grundlage der within-subject Standardabweichung (aus einer einfaktoriellen Varianzanalyse) zu konstruieren.
-Neben der Tatsache dass diese für sich gesehen informativ sind, stellen sie auch eine Art Baseline zur Bewertung der between-method variability dar.
+Neben der Tatsache dass diese Intervalle bereits für sich gesehen informativ sind, stellen sie auch eine Art Baseline zur Bewertung der between-method variability dar.
 
-<!-- Hahn & Meeker ... Statistical intervals -->
-<!-- https://ebookcentral.proquest.com/lib/tuchemnitz/detail.action?docID=4822515 -->
-<!-- In the best of situations, one can rely on physical understanding, or information from outside the study, to justify the practical assumptions.
-Such evaluations, however, are principally the responsibility of the subject-matter expert.
-Often, the assessment of such assumptions is far from clear-cut.
--->
-Hier zeigt sich dass für die Konstruktion geeigneter Fehlerbänder keine universal gültigen Antworten gibt, sondern die Wahl der statistischen Intervalle - wenn möglich -  vom Kontext der praktischen Anwendung des Messsystems abhängig gemacht werden sollte.
-Realistisch betrachtet ist der spätere Awendungsfall zum Zeitpunkt der Validierung nicht immer klar definiert.
-In solchen Fällen empfehlen wir verschiedene Bänder für within und between-subject Designs zu konstruieren.
-Falls das für Nich-Statistiker als zu kompliziert empfunden wird scheinen die untersuchten POINT-Bänder (mit messwiederholten Werten!) einen brauchbaren Kompromiss zwischen zu breiten und zu engen Intervallen darzustellen.
+In den Fällen wo Messsysteme zunächst validiert werden ohne dass unmittelbar klar ist für welche Anwendung das Messsystem später eingesetzt werden soll empfehlen wir verschiedene Bänder für within und between-subject Designs zu konstruieren.
+
+Falls in der Praxis als zu aufwändig bewertet wird, scheinen die untersuchten POINT-Bänder (mit messwiederholten Werten!) einen brauchbaren Kompromiss zwischen zu breiten und zu engen Intervallen darzustellen.
 Zudem sollten berechnete Fehlerintervalle stets grafisch anhand der Messdaten (Originalkurven und/oder Differenzkurven) evaluiert werden.
@@
 In jedem Fall sollten aber aus den in der Einleitung genannten Gründen auf kontinuierliche Methoden gesetzt werden.
+Diese müssen nicht zwingend mithilfe funktionaler Verfahren berechnet werden, sondern punktweise konstruierte Bänder können - je nach Anwendungsfall - durchaus gute Ergebnisse liefern.
+
 Wir spekulieren dass der seltene Einsatz kontinuierlicher Verfahren in direktem Zusammenhang mit der Komplexität der Verfahren steht, die eine hohe Hürde für viele Praktiker mit weniger ausgeprägten Statistik- und Programmierkenntnissen darstellt.
 Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code frei zur Verfügung (GPL).
-
-<!-- Die Erkenntnisse der vorliegenden Studie können nicht 1:1 auf andere Arten von Daten (z. B. Bodenreaktionskraftkurven, EMG Signale) übertragen, sondern gelten primär für glatte Kurven. -->
-Was im vorliegenden Paper nicht untersucht wurde ist der Fall, wenn keine glatten Kurven Kurven (z.
-B.
-verrauschte Signale wie EMG) untersucht werden.
-Diese Betrachtungen sprengen aber den Rahmen des vorliegenden Papers.
-Zudem fallen Differenzenkurven i.
-d.
-R.
-deutlich glatter aus als die verrauschten Originalkurven und das Problem ist womöglich deutlich geringer als momentan zu erwarten wäre.
-<!-- Dann gibt es eine Reihe zusätzlicher Probleme, z.
-B.
-dass die Approximation mithilfe mathematischer Funktionen andere Funktionen (B-Splines anstatt Fourier) erfordert oder gänzlich unmöglich wird.
--->
-
-!-- Ludbrook (2010)
-<!-- First, you should decide which set of limits arebest suited to your goal: 95% confidence limits indicate the range within which 95%of differences in the population from which you drew your sample should lie, so, in a sense, confidence limits are concerned with the present; 95%tolerance limits (shorthand for 95%tol-erance limits with 95%confidence) indicate the range within which asingle, new, observation should lie if it is drawn from the same popu-lation as was the sample you have studied and so, in a sense, toler-ance limits are concerned with the future.
-Which of these two types of limits should you estimate?
-The confidence limits can be used to decide whether one method can safely be substituted for another; hence, Altman and Bland’s description of them as ‘limits of agree-ment’.4,7,12 It is up to you to argue that your 95%confidence limits are too wide to safely allow one method to be substituted for the other or, conversely, that they are narrow enough to allow substitution.
-However, if you imagine that on some future occasion you obtain a certain value by Method A and want to know just how wide the range of values is that you would have obtained by Method B, then the 95%tolerance limits are appropriate.
--->
-
-<!-- Anforderungen an die Fehlerintervalle gestellt werden?
-Im Sinne einer möglichst hohen coverage liefern breitere Prädiktionsintervalle eine konservative Schätzung des zu zufälligen Messfehlers.
-Im Kontext von Messsystemvalidierungen kann es aber durchaus sinnvoll sein einen Teil der coverage zugunsten weniger konservativer Intervalle zu opfern solange diese die Streuung der Differenzen adäquat charakterisieren.
-Die POINT-Intervalle scheinen hier einen guten Kompromiss darzustellen.
--->
-
-<!-- Eine mögliche Alternative zu den untersuchten Verfahren -->
-<!-- Ein möglicher Kompromiss zwischen den gebootstrappten, funktionalen Intervallen in BOOT und ... in POINT sind die von Francq vorgeschlagenen Toleranzintervalle. -->
-Francq et al.
-(2020) schlagen vor stattdessen Toleranzintervalle zu verwenden.
-Toleranzintervalle (auch beta-expectation tolerance interval ($\beta$ TI)) sind statistische Intervalle, die mit einem gewissen Konfidenzniveau einen bestimmten Anteil einer Stichprobenpopulation (z.
-B.
-95% der Differenzkurven) enthalten.
-
-<!-- In terms of terminology, tolerance means, in this context, that some difference between the methods is tolerated (the measurements are still comparable in practice).
-Furthermore, the tolerance interval is exact and therefore more appropriate than the agreement interval, as the next sections will show.
--->
-
-Toleranz in diesem Zusammenhang, dass ein gewisser Unterschied zwischen den Methoden toleriert wird (die Messungen sind in der Praxis noch vergleichbar).
-Außerdem ist das Toleranzintervall exakter und daher besser geeignet als das Übereinstimmungsintervall, wie die nächsten Abschnitte zeigen werden.
-
-Solche Toleranzintervalle sind mathematisch äquivalent zu Prädiktionsintervallen, auch wenn sich die Interpretationen unterscheiden.
-
-<!-- Francq -->
-<!-- https://onlinelibrary.wiley.com/doi/full/10.1002/sim.8709 -->
-<!-- However, this interval is approximate (too narrow) and several authors propose calculating a confidence interval around each bound.
-This article demonstrates that this approach is misleading, awkward, and confusing.
-On the other hand, tolerance intervals are exact and can include a confidence level if needed.
-Tolerance intervals are also easier to calculate and to interpret.
--->
-
-<!-- The agreement interval is approximate and therefore, it should be interpreted with caution (especially for small sample sizes).
-Calculating a confidence interval for each bound of the agreement interval is awkward and brings even more confusion to its interpretation.
-While Hamilton and Stamey recommend calculating the AIs with its confidence intervals in their paper entitled “Using Bland-Altman to assess agreement between two medical devices-don't forget the confidence intervals!,”26 we have shown in this article that tolerance intervals are better and should be encouraged.
--->
-
-<!-- In terms of terminology, tolerance means, in this context, that some difference between the methods is tolerated (the measurements are still comparable in practice).
-Furthermore, the tolerance interval is exact and therefore more appropriate than the agreement interval, as the next sections will show.
--->
-
-<!-- Ludbrook already pointed out the complexity of the agreement interval with its confidence intervals while the tolerance interval is the appropriate solution.5  -->
-´
-
-<!-- , die beim ROISLIEN-Verfahren höher als bei POINT ist (und um ca. Faktor 3 höher als bei BOOT)  -->
-
-<!-- https://onlinelibrary.wiley.com/doi/full/10.1002/sim.8709 -->
-
-<!-- Ludbrook already pointed out the complexity of the agreement interval with its confidence intervals while the tolerance interval is the appropriate solution.5  -->
-
-<!-- The prediction interval here presented is equivalent to the beta-expectation Tolerance interval (also called “tolerance interval type I,” see Table 1).5, 18, 34 Their definition and interpretation are, however, different.
-A prediction interval is an interval where a future measurement (eg, a future difference between two medical devices) is expected to lie with a given confidence level.
-A beta-expectation tolerance interval is an interval where a given proportion of the population should lie, on average (eg, 95% of the differences).
-These two definitions lead to the same mathematical formula (3).
--->
-
-#### Datensätze
-
-Wir können natürlich nicht mit vier Datensätzen die gesamte Bandbreite abdecken, obwohl damit schon sehr viele Anwendungsfälle abgedeckt werden.
-Weiterhin ist der Anwendungsbereich auf eine spezielle Art von Kurven (glatte Kurve) beschränkt.
-Kurven mit vergleichbaren Eigenschaften werden von vielen biomechanischen Messsystemen erzeugt:  Bestimmte Bodenreaktionskräfte (horizontal), tiefpassgefilterte Signale, COP Verläufe etc..
-Es gibt aber auch Kurven die ganz andere Charkteristiken wie z.
-B.
-Sprünge/Unstetigkeiten im Verlauf (vertikale Bodenreaktionskurven), hohe Rauschanteile (z.
-B.
-EMG, ungefilterte Beschleunigungssignale), nichtstationäres Verhalten (z.
-B.
-Drift bei intergrierten Gyrosignalen) aufweisen.
-Es ist nicht klar inwiefern die untersuchten Methoden auch bei diesen Signalen brauchbare Ergebnisse liefern.
-Gerade bei Signalen mit sprunghaften Änderungen ist denkbar dass punktweise Intervalle zu plausibleren Ergebnissen führen können.
-<!-- Nur glatte Kurven bzw. glatte Differenzkurven (was in der Praxis wohl auch meistens der Fall ist) ... vllt. haben -->
-
-<!-- Joch et al. (2019) -->
-<!-- https://link.springer.com/article/10.3758/s13428-018-1060-5 -->
-
-A more critical limitation of the FBRT procedure, as the authors described it, is that it cannot be applied to all kinds of continuous data.
-Lenhoff and colleagues fitted gait data with mathematical functions representing the data as closely as possible.
-However, recorded data cannot always be represented by adequate fits—for instance, when no function family exists that is capable of fitting all the entities (see Appendix 2).
-EEG data, for example, are characterized by rapid changes in voltage polarity that cannot reasonably be represented by a single mathematical function prototype.
-Another limitation of the FBRT, as the authors describe it, is that the C factor is calculated using the data from only one sample (mostly representing the data of a null/control condition).
-By looking at a single data point analysis, for example the t test, the estimation of the variance parameter would integrate the data from all measured samples, which should in consequence lead to a more reliable estimation.
--->
-
-<!-- In conclusion, all of the described methods to statistically analyze continuous data have inherent problems that limit their scopes of application.
-The FBRT of Lenhoff et al.
-(1999) seems to be the most appropriate and useful approach with respect to continuous data that are not normally distributed and that can be represented by mathematical functions.
-For cases in which the data cannot be fitted, however, and therefore are not suited for the FBRT procedure, we developed a method combining a point-by-point variance estimate with a resampling technique in order to calculate confidence bands according to the FBRT.
-In addition, the new method includes the calculation of a C factor that integrates the data of all samples.
-We call this method the point-based resampling technique (PBRT); and it will be explained in the following sections.
-For illustration purposes, we will demonstrate the application of this method using exemplary EEG data from an event-related potential (ERP) analysis.
-Although we demonstrate the PBRT in a between-groups comparison, this procedure can also be used with other continuous data and different experimental designs (e.g., within-subjects comparisons, as is described in Appendix 3).
--->
 
@@
 
-<!-- Thus, this function-based resampling technique (FBRT), as we call it, is an advanced tool that can be applied even to data with asymmetric distributions. -->
-
-Wichtig ist zudem dass die ROISLIEN-Methode - im Gegensatz zu BOOT - kein vollständig funktionales Verfahren ist, da im Anschluss an die Approximation der Kurven mithilfe von Fourrierreihen eine punktweise Berechnung der Prädiktionsintervallgrenzen stattfindet.
-Wir raten angesichts der höheren Unsicherheit, der schlechteren coverage properties und der höheren Komplexität im Vergleich mit dem POINT-Ansatz vom Einsatz des ROISLIEN-Verfahrens ab.
-
 # Conclusion
 
-Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
+<!-- Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
 In diesem Sinne stellen wir auch den zugehörigen R Code im Anhang zur freien Verfügung.
 Für die Zukunft ist auch ein R Paket geplant, dass dann womöglch auch andere Intervalle (Toleranzintervalle, Konfidenzintervalle) enthält.
+-->
 
-Verwendet messwiederholte Daten!
+Punktweise Intervalle auf Grundlage parameterischer Modelle können durchaus brauchbare Ergebnisse liefern, auch bei Daten in denen die Voraussetzungen mitunter eheblich verletzt sind.
+...
+Voraussetzung ist das sie in geeigneter Weise berechnet werden.
+Dazu zählt u.
+a.
+auch die Berücksichtigung von messwiederholte Daten.
 

### 2021-11-24 paper.Rmd
Continue paper draft

@@
 ### TODO
+- Rolle von Messwiederholungen in der Fragestellung definieren!?
 - Tabellen- und Abbildungsbeschriftungen
@@
 - .bib file anlegen
-<!-- - Begriff spatiotemporal überdenken (wie ist das im Kontext FDA definiert) -->
+- Begriff concurrent validity reinbringen
+- Github-Link reinbringe
 
@@
 <!-- - Ratkowsky 1990 Handbook of nonlinear regression models -->
+<!-- - Meeker et al. Statistical intervals -->
+<!-- - Juul et al. (2020) lesen -->
 
@@
 Dazu verglichen sie punktweise berechnete Prädiktionsbänder gegenüber einer funktionalen Methode bei der Kurvendaten zunächst mithilfe von Fourierfunktionen approximiert werden und aus den so gebildeten funktionalen Kurven durch Bootstrapping Prädiktionsbänder gebildet werden (Sutherland, Olshen, Biden & Wyatt (1988, 1989)).
-They found that the Sutherland prediction bands provided appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands were shown to provide inadequate coverage (54%).
-
-<!-- Sutherland, Olshen, Biden & Wyatt (1988, 1989) (BOOT) haben für Gelenkwinkeldaten im Kontext von Ganganalysen bereits Ende der 1980er Jahre eine Methode vorgestellt, bei der Kurvendaten mithilfe von Fourierfunktionen approximiert werden und aus den so gebildeten funktionalen Kurven durch Bootstrapping Prädiktionsbänder zur Beschreibung eines normativen Gangbildes gebildet werden.
-Lenhoff et al.
-(1999) illustrated the same method on a generic gait analysis data set (comprised of joint angle curves) and compared it to pointwise Gaussian intervals.
-They found that the Sutherland prediction bands provided appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands were shown to provide inadequate coverage (54%).
--->
+They found that these prediction bands provided appropriate coverage for continuous curve gait data (86% coverage at 90% nominal coverage level) while pointwise Gaussian bands were shown to provide inadequate coverage (54%).
 
@@
 Limits of Agreement (LoA: 95% der Differenzen zwischen zwei Messsystemen sind im Intervall mittlere Differenz $\pm$ 1.96 Standardabweichungen der Differenzen enthalten), für Kurvendaten erweitert wird.
-LoA sind das am weitesten verbreitete Verfahren in Methodenvergleichen (Ludbrook, 2010) und stellen per Deinfition Prädiktionsintervalle dar (obwohl das tatsächliche Prädiktionslevel der LoA stets niedriger ist als das nomale Niveau und für kleine Stichproben kollabiert (Francq, 2020)).
+LoA sind das am weitesten verbreitete Verfahren in Methodenvergleichen (Ludbrook, 2010) und stellen per Definfition Prädiktionsintervalle dar (obwohl das tatsächliche Prädiktionslevel der LoA stets niedriger ist als das nominale Niveau und für kleine Stichproben kollabiert (Francq, 2020)).
 
-Ausgehend von diesen Vorarbeiten soll im vorliegenden Paper eine systematische Analyse von Verfahren zur Konstruktion kontinuierlicher Prädiktionsbändern für 1D Differenzkurven vorgenommen werden (im Kontext der Validierung von Messsystemen).
-Dazu werden drei verschiedene Verfahren anhand simulierter und realer Datensätze hinsichtlich ihrer Abdeckungseigenschaften verglichen: (1) punktweise Bland and Altman LoA (2) LoA nach Roislien et al.
+Das primäre Anliegen des vorliegenden Papers ist die Gegenüberstellung punktweiser gegen funktionale Methoden zur Konstruktion kontinuierlicher Fehlerintervalle für 1D Kurvendaten im Rahmen einer Messsystemvalidierung.
+Demonstriert werden soll das mithilfe simulierter und realer Datensätze exemplarisch für Prädiktionsbändern die mit drei verschiedenen Methoden konstruiert werden: (1) punktweise Bland and Altman LoA (2) LoA nach Roislien et al.
 (2012) und (3) gebootstrappte funktionale Prädiktionsbänder (Sutherland, Olshen, Biden & Wyatt (1988, 1989), Lenhoff et al.
 (1999)).
-
-Wir hoffen mit dieser Studie einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien zu setzen und Forschern praktische Hinweise für die Auswahl und Implementierung geeigneter Verfahren zur Verfügung zu stellen.
-In diesem Sinne ist auch der zugehörige R Code im Anhang zur Verfügung gestellt.
+Zudem soll praktische Hinweise zur Konstruktion und zum Einsatz unterschiedlicher Fehlerbänder für verschiedene Anwendungsszenarien (Untersuchungsdesigns) abgeleitet werden.
 
 # Methods
+
+Um die Vergleichbarkeit der verschiedenen Methoden zu gewährleisten, erfolgt die Konstruktion der Fehlerintervalle in allen drei Methoden anhand von Differenzen(kurven).
 
@@
 Die Tatsache, dass ein erheblicher Teil biomechanischer Messsysteme höherdimensionale Daten (> 0D) liefert, spiegelt sich bislang kaum in der Methodik von Validierungsstudien wieder.
-Um die Aussagekraft von Validierungsuntersuchungen diesbezüglich zu erhöhen braucht es Fehlerintervalle, die alle Punkte eines Messsignals berücksichtigen.
-Zu diesem Zweck wurden im vorliegenden Paper drei relevante Verfahren zur Berechnung kontinuierlicher Prädiktionsintervalle identifiziert (POINT, ROISLIEN, BOOT) und für ausgewählte Messfehlerszenarien am Beispiel von 1D Kurvendaten analysiert.
+Um die Aussagekraft von Validierungsuntersuchungen diesbezüglich zu erhöhen, braucht es Fehlerintervalle, die alle Punkte eines Messsignals berücksichtigen.
+Zu diesem Zweck wurden im vorliegenden Paper drei relevante Verfahren zur Berechnung kontinuierlicher Fehlerbänder identifiziert (POINT, ROISLIEN, BOOT) und für ausgewählte Messfehlerszenarien am Beispiel von 1D Kurvendaten analysiert.
 BOOT ist ein funktionales Verfahren in dem diskrete Kurven mit mathematischen Funktionen approximiert werden und alle weiteren Berechnungen anhand dieser funktionalen Objekten vorgenommen werden.
-Die ROISLIEN-Methode wurde als funktionales Verfahren entwickelt, muss aber strenggenommen (wie POINT) zu den punktweisen Verfahren gezählt werden, da die eigentliche Berechnung der Intervallgrenzen separat für jeden Messzeitpunkt erfolgt.
+Die ROISLIEN-Methode wurde als funktionales Verfahren entwickelt, muss aber strenggenommen zu den punktweisen Verfahren gezählt werden, da die eigentliche Berechnung der Intervallgrenzen separat für jeden Messzeitpunkt erfolgt.
 
+# Punktweise vs. funktionale Prädiktionsbänder
 Die Intervalle der BOOT-Methode erreichen in allen untersuchten Szenarien die nominale coverage und sind mit weniger estimation uncertainty belastet als die beiden übrigen Verfahren (POINT, ROISLIEN).
@@
 h.
-bis auf wenige Ausnahmen liegen stets die allermeisten Punkte der Testkurven im trainierten Prädiktionsintervall.
-
-Die punktweisen Verfahren liefern auch für Datensätze plausible Verläufe, in denen erhebliche Verletzungen der parametrischen Modellannahmen vorliegen (nicht-gaussverteilten Differenzen, phase shift in x-axis direction).
-Für die Praxis ist das eine wichtige Erkenntnis, da diese Annahmen häufig verletzt sind, z.
-B.
-wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder zu geringer Auflösung abgeschnitten werden, oder in Gegenwart von nichtlinearem Drift.
-Speziell im der letzten Fall sind Alternativen zum BOOT-Verfahren von Interesse, da die Methode nur für stationäre Signale zuverlässig funktioniert (Lenhoff et al., 1999).
+bis auf wenige Ausnahmen liegen stets nahezu alle Punkte der Testkurven im trainierten Prädiktionsintervall.
 
 Die Verläufe der Intervallgrenzen im POINT und ROISLIEN-Verfahren Verfahren sind qualitativ ähnlich, die coverage (Table 1 & 2) ist beim ROISLIEN-Verfahren jedoch etwas niedriger.
-Die Ursache hierfür liegt in der höheren Streuung der ROISLIEN-Intervallgrenzen über $D = 110$ Iterationen des leave-one-out Validierungsverfahrens, da hier nur jeweils eine Kurve pro Proband gezogen wird und kleinere Stichproben einer größeren Unsicherheit unterliegen.
+Die Ursache hierfür liegt in der höheren Streuung der ROISLIEN-Intervallgrenzen, da hier nur jeweils eine Kurve pro Proband gezogen wird und kleinere Stichproben generell einer größeren Unsicherheit unterliegen.
 Bei der POINT-Methode hingegen werden stets alle Kurven im Datensatz berücksichtigt.
-Um der prinzipiellen Unsicherheit von Samplingprozessen Rechnung zu tragen, empfehlen Bland & Altman jedoch die Erweiterung der LoA um Konfidenzbänder.
+Um der prinzipiellen Unsicherheit von Samplingprozessen Rechnung zu tragen, empfehlen Bland & Altman die Erweiterung der LoA um Konfidenzbänder.
 Dieser Vorschlag führt die ursprüngliche Idee einfach zu berechnender, intuitiver Unsicherheitsintervalle allerdings ein Stück weit ad absurdum (Francq et al., 2020), weshalb in der vorliegenden Arbeit auf eine Verbreiterung der LoA durch Konfidenzbänder verzichtet wurde.
@@
 Dadurch fällt auch der eigentliche Vorteil punktweiser Intervalle gegenüber funktionalen Verfahren - die Einfachheit der Berechnung - erheblich geringer aus.
-<!-- Ludbrook already pointed out the complexity of the agreement interval with its confidence intervals while the tolerance interval is the appropriate solution.5  -->
 
-Das punktweise Vorgehen hat noch einen anderen möglichen Vorteil gegenüber funktionalen Methoden: Sie können auf alle Arten kontinuierlicher Daten angewendet werden.
-Im Gegensatz zu den hier untersuchten glatten Kurven lassen sich mathematische Funktionen auf Signale mit vielen sprunghaften Veränderungen im Signalverlauf wie z.
+Die punktweisen Verfahren liefern auch für die Datensätze weitgehend plausible Verläufe, in denen erhebliche Verletzungen der parametrischen Modellannahmen vorliegen (nicht-gaussverteilten Differenzen, phase shift in x-axis direction).
+Für die Praxis ist das eine wichtige Erkenntnis, da diese Annahmen in vielen Fällen verletzt sind, z.
 B.
-EMG-Signale nicht fitten.
-Die Untersuchung dieser Art von Kurven war allerdings nicht Gegenstand des vorliegenden Papers.
-Die hier vorgestellten Datensätze decken dennoch ein breites Spektrum biomechanischer Signale ab.
+wenn Sensorwerte wegen einer Begrenzung des Messbereichs oder zu geringer Auflösung abgeschnitten werden, oder in Gegenwart von nichtlinearem Drift.
+Speziell im letzteren Fall sind Alternativen zum BOOT-Verfahren von Interesse, da die Methode nur für stationäre Signale zuverlässig funktioniert (Lenhoff et al., 1999).
+Das punktweise Vorgehen hat darüber hinaus einen Vorteil gegenüber funktionalen Verfahren wenn die untersuchten Kurven weniger glatt sind als in den hier untersuchten Beispieldatensätzen (z.
+B.
+EMG-Signale) weil das Fitten mathematischer Funktionen für diese Art Kurven nicht immer möglich ist.
+Praktisch gibt es aber nur wenige Signale bei denen das wirklich zum Tragen kommt: In den meisten Fällen werden verrauschte Signale (z.
+B.
+Beschleunigungssignale) ohnehin tiefpassgefiltert, so dass man auch im Rahmen der Validierung von glatten Kurven ausgehen kann.
+Eine genauere Analyse derartiger Kurven sprengt zudem den Rahmen dieser Arbeit.
 
-#### Auswahl geeigneter Fehlerbänder
-Abgesehen von Unterschieden in der Berechnung und den coverage-Eigenschaften ist allen drei untersuchten Methoden gemein, dass Prädiktionsintervalle konstruiert werden.
-Um eine möglichst hohe coverage zu erzielen fallen diese Intervalle daher gezielt breit aus.
-Das ist im Sinne von Bland & Altman erwünscht um einen möglichst hohen Anteil von Einzelbeobachtungen in den Intervallen zu enthalten und so eine gewisse Sicherheit zu haben, dass auch die nächste Einzelmessung im Intervall liegt.
+#### Konstruktion und Auswahl geeigneter Fehlerbänder
+Primäres Ziel des vorliegenden Papers war es Unterschiede zwischen kontinuierlichen und punktweisen Verfahren mit Blick auf die Konstruktion von Fehlerintervallen in Methodenvergleichen zu analysieren.
+Das am häufigsten genutze Tool in Methodenvergleichen ist das im medizinischen Kontext entwickelte Bland & Altman Verfahren, mit dem der Grad der Übereinstimmung anhand von Differenzen zwischen zwei Messsystemen bewertet wird.
+Die entsprechenden LoA fallen verhältnismäßig breit aus weil sie einen hohen Anteil (95%) der Einzelbeobachtungen enthalten sollten um eine gewisse Sicherheit zu haben, dass auch die nächste Einzelmessung im Intervall liegt.
 Im biomechanischen Kontext ist das z.
 B.
-dann relevant, wenn neue Messwerte gegen normative Daten (z.
+dann relevant, wenn neue Messwerte mit normativen Daten (z.
 B.
 aus einer Referenzdatenbank) verglichen werden.
-Wenn hier die Abweichung ein bestimmtes Maß überschreitet, d.
+Wenn die Abweichung ein bestimmtes Maß überschreitet, d.
 h.
-der Proband nicht zur Referenzpopulation gehört, werden bestimmte pathologische Muster angenommen und eine geeignete Therapie vorgeschlagen (Olshen et al., 1989).
-Entsprechend ist in diesem Kontext von Bedeutung, dass die Bänder nicht zu eng ausfallen, weil die Konsequenzen falsch positiver Schlüsse schwerwiegend sein können.
+der Proband nicht zur Referenzpopulation gehört, können z.
+B.
+pathologische Muster angenommen werden (Olshen et al., 1989).
+Dabei ist von Bedeutung, dass die Bänder nicht zu eng ausfallen, weil die Konsequenzen falsch positiver Schlüsse schwerwiegend sein können.
 
-Für viele andere Fragestellungen sind die breiten Prädiktionsintervalle allerdings nicht geeignet weil breitere Intervalle ein höheres Risiko für falsch negative Schlüsse bedeuten.
-Das ist z.
-B.
-der Fall wenn die Messungen im Kontext eines Vergleichs verschiedener Probandenkollektive erhoben werden (z.
+Für viele andere Fragestellungen sind Prädiktionsintervalle allerdings eher ungeeignet, weil breitere Intervalle ein höheres Risiko für falsch negative Schlüsse bedeuten.
+Das ist u.
+a.
+dann der Fall, wenn die Messwerte im Kontext eines Vergleichs verschiedener Probandenkollektive erhoben werden (z.
 B.
 in Interventionsstudien oder case-control Studien).
-Gut demonstrieren lässt sich das am Beispiel der Differenzen in Fig.
-2 D (Differenzen zwischen einem optischen motion capture System und einer IMU).
-Eine mittels IMU gemessene Differenz zwischen den sagittalen Hüftwinkeln z.
+Gut demonstrieren lässt sich das am Beispiel des realweltlichen Datensatzes in Fig.
+2 D (Differenzen (IMU - MC) sagittaler Hüftwinkel): Ein mittels IMU gemessener Unterschied zwischen den sagittalen Hüftwinkeln z.
 B.
-einer Kontroll- und Interventionsgruppe kann hier nur dann als tatsächlicher Effekt angenommen werden, wenn sie mehr als die (je nach Intervall ungefähr) $\pm$ 10° des zufälligen Messfehlers beträgt.
-Derartig große Effekte sind in biomechanischen Untersuchungen allerdings nur selten zu erwarten.
-<!-- Dies gilt umso mehr vor dem Hintergrund häufig eher geringer Effektstärken in der Biomechanik (Quelle ... Knudson? Metaanalysen!?).  -->
-<!-- Typische Effektgrößen liegen hier eher bei ... (Quelle). -->
-Die hier beobachtete Streuung der Differenzen ist dabei längst kein Szenario das weit von der Praxis entfernt ist.
-Gerade bei markerbasierten motion capture Systemen, die häufig als ground truth eingesetzt werden, ist die Reliabilität aufgrund von u.
-a.
-Weichteilbewegungen, Markerplatzierungsartefakten etc.
-fragwürdig, so dass hier leicht sehr breite (Prädiktions)Intervalle entstehen können.
-<!-- Werte/Studie zur Reliabilität von MoCap Systemen? -->
+einer Kontroll- und Interventionsgruppe kann hier nur dann als tatsächlicher Effekt angenommen werden, wenn die Differenz außerhalb der ermittelten Intervallgrenzen von ca.
+$\pm$ 10° liegt.
+Derartig große Effekte sind in biomechanischen Untersuchungen allerdings nur selten zu erwarten (Knudson, 2017).
+Die Streuung der Differenzen im Beispiel ist dabei längst kein unrealistisches Szenario.
+Gerade bei markerbasierten motion capture Systemen, die häufig als Referenz in Methodenvergleichen eingesetzt werden, ist die Reliabilität aufgrund von z.
+B.
+Weichteilbewegungen oder Markerplatzierungsartefakten limitiert (McGinley et al., 2009), so dass hier ohne Weiteres sehr breite (Prädiktions)Intervalle entstehen können.
 
-Eine Alternative zu den breiten Prädiktionsintervallen sind Intervalle auf Grundlage der within-subject Standardabweichung (aus einer einfaktoriellen Varianzanalyse) (Bland & Altman, 1999).
-Wiederum am Beispiel Hüftgelenksdifferenzen (Fig.
+Eine Alternative zu den breiten Prädiktionsintervallen sind Intervalle auf Grundlage der within-subject Standardabweichung (Bland & Altman, 1999).
+Bei den Hüftgelenksdifferenzen (Fig.
 2, D) lässt sich erkennen, dass die Streuung innerhalb einzelner Probanden erheblich geringer ausfällt (erkennbar an den probandenweisen Kurvenclustern).
@@
 B.
-der Fall wenn Untersuchungen mit einem reinen within-subjects Design durchgeführt werden (z.
+von Interesse wenn Untersuchungen mit einem reinen within-subjects Design durchgeführt werden (z.
 B.
-bei prä-post-Vergleichen in Interventionsstudien).
-Neben der Tatsache dass diese Intervalle bereits für sich gesehen informativ sind, stellen sie auch eine Art Baseline zur Bewertung der between-method variability dar.
+intraindividuelle bei prä-post-Vergleiche).
+Neben der Tatsache dass diese Intervalle bereits für sich gesehen informativ sind, stellen sie auch eine Art Baseline zur Bewertung der between-method variability dar (Bland & Altman, 1999).
+Was leider keines der vorgestellten Verfahren vorsieht ist die Möglichkeit asymmetrischer Bänder um Unterschiede in der Ausprägung des zufälligen Messfehlers abzubilden.
+Vorstellbar wäre z.
+B.
+eine Methode in der anstatt der symmetrischen um einen zentralen Wert Perzentile als Intervallgrenzen definiert werden.
+Interessant erscheint auch der Ansatz diese Intervallgrenzen (ähnlich wie in BOOT) auf eine bestimmte coverage probability hin zu korrigieren.
 
-In den Fällen wo Messsysteme zunächst validiert werden ohne dass unmittelbar klar ist für welche Anwendung das Messsystem später eingesetzt werden soll empfehlen wir verschiedene Bänder für within und between-subject Designs zu konstruieren.
-
-Falls in der Praxis als zu aufwändig bewertet wird, scheinen die untersuchten POINT-Bänder (mit messwiederholten Werten!) einen brauchbaren Kompromiss zwischen zu breiten und zu engen Intervallen darzustellen.
-Zudem sollten berechnete Fehlerintervalle stets grafisch anhand der Messdaten (Originalkurven und/oder Differenzkurven) evaluiert werden.
-
-In jedem Fall sollten aber aus den in der Einleitung genannten Gründen auf kontinuierliche Methoden gesetzt werden.
-Diese müssen nicht zwingend mithilfe funktionaler Verfahren berechnet werden, sondern punktweise konstruierte Bänder können - je nach Anwendungsfall - durchaus gute Ergebnisse liefern.
-
-Wir spekulieren dass der seltene Einsatz kontinuierlicher Verfahren in direktem Zusammenhang mit der Komplexität der Verfahren steht, die eine hohe Hürde für viele Praktiker mit weniger ausgeprägten Statistik- und Programmierkenntnissen darstellt.
-Um die Verbreitung funktionaler Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code frei zur Verfügung (GPL).
-
-#### Grundsaetzliches / Sonstiges
-Wir empfehlen grundsätzlich die Daten zu plotten.
-Am Ende ist jeder Datensatz bzw.
-jede Fragestellung unterschiedlich und ein pauschaler coverage value wird der Qualität des Messgeräts (in Bezug auf die jeweilige Fragestellung möglicherweise nicht gerecht).
-
-A word on sample size in validation studies ...
-
-SPM
-Pataky: A methodology called Statistical Parametric Mapping (SPM) (Friston et al., 2007) can partially offset these limitations by providing a framework for the continuous statistical analysis of smooth bounded nD fields.
-
-Als Ansatz für die Konstruktion besserer Bänder: Nichtparametrischer Ansatz mit Perzentilen (asymmetrische Bänder)
+In den Fällen wo Messsysteme zunächst ohne klar definiertes Anwendungsgebiet validiert werden, empfehlen wir verschiedene Bänder für within und between-subject Designs zu konstruieren.
+Punktweise konstruierte Bänder stellen hierbei eine brauchbare Alternative dar.
+Hier sollten aber das in dieser Arbeit implementierte, komplexere Modell mit Messwiederholungen berücksichtigt werden um keine zu engen Intervallgrenzen zu erhalten.
+In jedem Fall aber sollten aus den in der Einleitung aufgeführten Gründen kontinuierliche Methoden den Vorzug gegenüber diskreten Verfahren erhalten.
+Wir spekulieren dass der spärliche Einsatz kontinuierlicher Verfahren in direktem Zusammenhang mit der Komplexität der Verfahren steht, die vermutlich eine hohe Hürde für viele Forscher mit weniger ausgeprägten Statistik- und Programmierkenntnissen darstellt.
+Um die Verbreitung kontinuierlicher Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code im Anhang und zusätzlich (inkl.
+aller Datensätze) auf Github zur Verfügung ().
 
 # Conclusion
-
 <!-- Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
@@
 ...
-Voraussetzung ist das sie in geeigneter Weise berechnet werden.
+Voraussetzung ist das sie in geeigneter Weise berechnet werden ...
 Dazu zählt u.
 a.
-auch die Berücksichtigung von messwiederholte Daten.
-
-<!-- Aus Juul et al. (2020) -->
-<!-- In summary, when making forecasts of epidemic trajectories, it is important to represent the resulting curve ensembles in a way that captures the quantities of interest -->
-<!-- in an intuitive way. -->
+die Berücksichtigung von messwiederholten Daten ...
+man könnte auch wie bei Lenhoff einen Korrekturfaktor reinbringen

### 2021-11-29 paper.Rmd
Fix uncertainty band calculation

@@
 - Rolle von Messwiederholungen in der Fragestellung definieren!?
-- Tabellen- und Abbildungsbeschriftungen
-- Erklarende Abbildung Methodenkapitel uncertainty estimation (eahnlich Diss Doris)
+- Erklarende Abbildung Methodenkapitel uncertainty estimation (aehnlich Diss Doris)
 - Formelnummerierung
 <!-- - Exemplarische Paper für die jeweiligen simulierten Datensätze raussuchen -->
-- .bib file anlegen
 - Begriff concurrent validity reinbringen
-- Github-Link reinbringe
 
@@
 source("floa_rcb.R")
-source("lenhoff_doris.R")
+source("floa_boot.R")
 source("floa_point.R")
@@
 # Introduction
-<!-- Wie in allen empirischen Wissenschaften basiert der Erkenntnisgewinn auf der Qualität experimenteller Daten und hängt damit direkt von der Güte der eingesetzten Messverfahren ab. -->
 Die Biomechanik ist eine empirische Wissenschaft, d.
@@
 e.
-phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) wenn inferenzstatistische Verfahren wie Konfidenz- oder Prädiktionsintervalle simultan auf mehrere Samplingpunkte angewandt werden um kontinuierliche Fehlerbänder zu konstruieren, ist zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird weil die (nonzero) Kovarianz benachbarter Samplingpunkte nicht berücksichtigt wird (Olea & Plagborg-Moeller, 2018; Lenhoff et al., 1999).
+phase shift) and the risk of examining points that have little relevance for the system under investigation (Donoghue et al., 2008; Pataky et al., 2008; Richter et al., 2014; Park et al., 2017) (iii) wenn inferenzstatistische Verfahren wie Konfidenz- oder Prädiktionsintervalle simultan auf mehrere Samplingpunkte angewandt werden um kontinuierliche Fehlerbänder zu konstruieren, ist zu erwarten, dass die zufällige Fehlerkomponente unterschätzt wird weil die (nonzero) Kovarianz benachbarter Samplingpunkte nicht berücksichtigt wird (Lenhoff et al., 1999; Olea & Plagborg-Moeller, 2018;).
 Die üblicherweise eingesetzten Korrekturverfahren zur Anpassung des alpha levels (Hochberg & , 1987) führen aufgrund der hohen Anzahl an Samplingpunkten zu extrem konservativen Schätzungen des zufälligen Fehlers.
@@
   geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Value") +
+  labs(x = "Time-normalized signal [%]", y = "Angle [°]") +
   theme_minimal() +
@@
   geom_line(data = data.single.imu, aes(x = frame, y = value, group = strideID, colour = as.factor(subjectID)), alpha = 0.3) +
-  labs(x = "Time-normalized signal [%]", y = "Value") +
+  labs(x = "Time-normalized signal [%]", y = "Angle [°]") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(axis.text.x = element_text(size = 14),
+        axis.title.x = element_text(size = 14),
+        axis.text.y = element_text(size = 14),
+        axis.title.y = element_text(size = 14),
+        legend.position = "none")
 ```
@@
 PLOT
+
+ggsave("~/Nextcloud/project-fab-forschung/Publikationen/FLOA/tex/Grafiken/original_curves.png", device = "png", dpi = 300)
 ```
@@
 
-Der Code zum Erstellen der Kurven findet sich im Anhang und auf Github als R Code und txt Dateien.
+<!-- Der Code zum Erstellen der Kurven findet sich im Anhang und auf Github als R Code und txt Dateien. -->
 
@@
 ### Uncertainty estimation
-
 Die Grenzen der Prädiktionsintervalle unterliegen einem stochastischen Prozess und variieren in Abhängigkeit vom Samplingprozess in den jeweiligen Verfahren (POINT, ROISLIEN, BOOT).
@@
 floa.roislien <- floa_roislien(data)
-floa.lenhoff.pred  <- floa_lenhoff(data,
+floa.boot  <- floa_boot(data,
                         k_reihe = 50,
@@
 
-PLOT.1 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-3, 3))
+PLOT.1 <- plot_loa(data, floa.point, floa.roislien, floa.boot, ylim = c(-3, 3))
 
@@
 
-PLOT.2 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-8, 8))
+PLOT.2 <- plot_loa(data, floa.point, floa.roislien, floa.boot, ylim = c(-8, 8))
 
@@
 
-PLOT.3 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-5, 5))
+PLOT.3 <- plot_loa(data, floa.point, floa.roislien, floa.boot, ylim = c(-5, 5))
 
@@
 
-PLOT.4 <- plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-20, 20))
+PLOT.4 <- plot_loa(data, floa.point, floa.roislien, floa.boot, ylim = c(-20, 20))
 ```
@@
 PLOT
+
+ggsave("~/Nextcloud/project-fab-forschung/Publikationen/FLOA/tex/Grafiken/difference_intervals.png", device = "png", dpi = 300)
 ```
@@
 rownames(covered.curves.percent) <- c("A. GAUSS", "B. NONGAUSS", "C. XSHIFT", "D. REAL")
-knitr::kable(covered.curves.percent, caption = "Table 1: Percentages of entirely covered curves for methods POINT, ROISLIEN and BOOT across data sets A - D (A - Gaussian error model, B - Non-Gaussian errors and heteroskedasticity, C - Gaussian error model and phase shift in x-axis direction, D - Real-world hip joint angle curves).
+
+table.tmp <- knitr::kable(covered.curves.percent, caption = "Table 1: Percentages of entirely covered curves for methods POINT, ROISLIEN and BOOT across data sets A - D (A - Gaussian error model, B - Non-Gaussian errors and heteroskedasticity, C - Gaussian error model and phase shift in x-axis direction, D - Real-world hip joint angle curves).
 Coverage values are presented as mean values across $D = 110$ leave-one-out test curves.")
+
+table.tmp
+
+# writeLines(table.tmp, 'table1.tex')
 ```
@@
 Um die Verbreitung kontinuierlicher Methoden zu katalysieren stellen wir daher die bei der Erstellung des Papers implementierten Methoden als R Code im Anhang und zusätzlich (inkl.
-aller Datensätze) auf Github zur Verfügung ().
+aller Datensätze) auf Github zur Verfügung (https://github.com/koda86/floa).
 
 # Conclusion
-<!-- Obwohl wir in der Arbeit bei Weitem nicht alle Anwendungsfälle (Methoden, Intervalle, Datensätze) abdecken können, hoffen wir einen Impuls für die breitere Anwendung kontinuierlicher Methoden im Rahmen biomechanischer Validierungsstudien setzen zu können.
-In diesem Sinne stellen wir auch den zugehörigen R Code im Anhang zur freien Verfügung.
-Für die Zukunft ist auch ein R Paket geplant, dass dann womöglch auch andere Intervalle (Toleranzintervalle, Konfidenzintervalle) enthält.
--->
-
 Punktweise Intervalle auf Grundlage parameterischer Modelle können durchaus brauchbare Ergebnisse liefern, auch bei Daten in denen die Voraussetzungen mitunter eheblich verletzt sind.

### 2021-11-29 R/estimate_uncertainty_loa.R

@@ -1,146 +1,167 @@
 # Estimate uncertainty in different LoA approaches (95% percentile intervals for the LoA)
+
 estimate_uncertainty_loa <- function (data, n.boot) {
 
   floa.roislien.upper <- vector(mode = "list", length = n.boot)
   floa.roislien.lower <- vector(mode = "list", length = n.boot)
-  floa.lenhoff.upper <- vector(mode = "list", length = n.boot)
-  floa.lenhoff.lower <- vector(mode = "list", length = n.boot)
+  floa.boot.upper <- vector(mode = "list", length = n.boot)
+  floa.boot.lower <- vector(mode = "list", length = n.boot)
 
   for (i in 1:n.boot) {
     floa.roislien <- floa_roislien(data)
-    floa.lenhoff  <- floa_lenhoff(data,
-                                  k_reihe = 50,
-                                  n.boot = n.boot,
-                                  band = "prediction",
-                                  cp.begin = 0,
-                                  alpha = 0.05)
+    floa.boot  <- floa_boot(data,
+                            k_reihe = 50,
+                            n.boot = n.boot,
+                            band = "prediction",
+                            cp.begin = 0,
+                            alpha = 0.05)
 
-    floa.roislien.upper[[i]] <- floa.roislien["upper", ]
-    floa.roislien.lower[[i]] <- floa.roislien["lower", ]
-    floa.lenhoff.upper[[i]]  <- floa.lenhoff["upper", ]
-    floa.lenhoff.lower[[i]]  <- floa.lenhoff["lower", ]
+    floa.roislien.upper[[i]] <- floa.roislien["upper.loa", ]
+    floa.roislien.lower[[i]] <- floa.roislien["lower.loa", ]
+    floa.boot.upper[[i]]  <- floa.boot["upper.loa", ]
+    floa.boot.lower[[i]]  <- floa.boot["lower.loa", ]
   }
-
-  # Pointwise LoA return the same estimate every time (all curves are included)
-  floa.point <- data.frame(t(floa_point(data)))
 
   floa.roislien.upper.unlist <- matrix(unlist(floa.roislien.upper), 101)
   floa.roislien.lower.unlist <- matrix(unlist(floa.roislien.lower), 101)
[…]

### 2021-11-29 R/floa_boot.R

@@ -1,20 +1,19 @@
-floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
+floa_boot <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
 
   # Limit of Agreement calculation adapted from Lenhoff et al. (1999)
   # ----------------------------------------------------------------------------
   #
   # Function arguments:
-  # k_reihe  : number of coefficients
+  # k_reihe  : Number of coefficients
   # n.boot   : Number of bootstrap iterations
   # band     : Type of interval (prediction or confidence)
-  # cp.begin : Anfangswert Quantil
-  # alpha    : Irrtumswahrscheinlichkeit
+  # cp.begin : Initial value quantile
+  # alpha    : Significance level
   # ----------------------------------------------------------------------------
 
   # ----------------------------------------------------------------------------
   # Get difference curves
   # ----------------------------------------------------------------------------
-
   devices <- unique(data$device)
 
   device.1 <- subset(data, device  == devices[1])$value
@@ -32,7 +31,6 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
   # ----------------------------------------------------------------------------
   # Approximate time series (differences) using Fourier functions
   # ----------------------------------------------------------------------------
-
   fourier.koeffi    = matlab::zeros(c(k_reihe*2 + 1, n_kurven))
   fourier.real      = matlab::zeros(n_time, n_kurven)
   fourier.mean      = matlab::zeros(k_reihe*2 + 1)
@@ -42,7 +40,7 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
   fourier.std_all   = matlab::zeros(n_time, n_time)
   fourier.std       = matlab::zeros(n_time, 1)
 
-  # Fourierreihe aufstellen
+  # Set up a Fourier series
   # General: f(t) = mu + sum(alpha cos(2pi*k*t/T) + beta sin(2pi*k*t/T))
   fourier.s = rep(1, times = n_time)
   for (k in seq(1, k_reihe*2, 2)) {
@@ -72,7 +70,7 @@ floa_lenhoff <- function(data, k_reihe, n.boot, band, cp.begin, alpha) {
[…]

### 2021-11-29 R/floa_point.R

@@ -39,15 +39,13 @@ floa_point <- function(data) {
   floa.point <- data.frame(unlist(mean.diff), unlist(total.sd))
   names(floa.point) <- c("mean.diff", "total.sd")
 
-  # Calculate mean, upper and lower limits of agreement (boundaries) to get the
-  # same structure as returned by the other methods (i. e. floa_rcb)
+  # Calculate mean, upper and lower limits of agreement (floa.point.loa) to
+  # get the same structure as returned by the other methods (i. e. floa_rcb)
+  z0.975 <- qnorm(0.975, mean = 0, sd = 1) # Quantile of the standard normal distribution
 
-  # Quantile of the standard normal distribution
-  z0.975 <- qnorm(0.975, mean = 0, sd = 1)
-
-  floa.point.bnd <- rbind(floa.point$mean.diff + z0.975*floa.point$total.sd,
+  floa.point.loa <- rbind(floa.point$mean.diff + z0.975 * floa.point$total.sd,
                           floa.point$mean.diff,
-                          floa.point$mean.diff - z0.975*floa.point$total.sd
+                          floa.point$mean.diff - z0.975 * floa.point$total.sd
                           )
 
   # PRECISION OF ESTIMATED LIMITS OF AGREEMENT
@@ -59,14 +57,15 @@ floa_point <- function(data) {
   # t-score of the 95th quantile of the Student t distribution with df = (n - 1)
   t <- qt(.975, df = n.curves - 1)
 
-  floa.point.bnd <- rbind(floa.point.bnd,
-                     floa.point$mean.diff + z0.975*floa.point$total.sd + t*standard.error.d2s,
-                     floa.point$mean.diff + z0.975*floa.point$total.sd - t*standard.error.d2s,
-                     floa.point$mean.diff - z0.975*floa.point$total.sd + t*standard.error.d2s,
-                     floa.point$mean.diff - z0.975*floa.point$total.sd - t*standard.error.d2s)
+  # Upper and lower uncertainty limits
+  upper.ci.upper <- floa.point$mean.diff + z0.975*floa.point$total.sd + t*standard.error.d2s
+  upper.ci.lower <- floa.point$mean.diff + z0.975*floa.point$total.sd - t*standard.error.d2s
+  lower.ci.upper <- floa.point$mean.diff - z0.975*floa.point$total.sd + t*standard.error.d2s
[…]

### 2021-11-29 R/floa_roislien.R

@@ -16,7 +16,7 @@ floa_roislien <- function(data) {
                          func.mean,
                          func.mean - 1.96 * func.sd)
 
-  rownames(floa.roislien) <- c("upper", "mean", "lower")
+  rownames(floa.roislien) <- c("upper.loa", "mean", "lower.loa")
 
   return(floa.roislien)
 }

### 2021-11-29 R/get_coverage_singlecurve.R

@@ -4,8 +4,8 @@ get_coverage_singlecurve <- function (curve, floa) {
 
   # Calculate coverage (entire curves within the percentile boundaries)
   # ----------------------------------------------------------------------------
-  lwr.bnd <- floa["lower", ]
-  upr.bnd <- floa["upper", ]
+  lwr.bnd <- floa[grep("low", rownames(floa)), ]
+  upr.bnd <- floa[grep("up", rownames(floa)), ]
 
   outside <- 0
 

### 2021-11-29 R/get_coverage_singlecurve_fraction.R

@@ -4,8 +4,8 @@ get_coverage_singlecurve_fraction <- function (curve, floa) {
 
   # Calculate coverage (curve points within the percentile boundaries)
   # ----------------------------------------------------------------------------
-  lwr.bnd <- floa["lower", ]
-  upr.bnd <- floa["upper", ]
+  lwr.bnd <- floa[grep("low", rownames(floa)), ]
+  upr.bnd <- floa[grep("up", rownames(floa)), ]
 
   outside <- 0
 

### 2021-11-29 R/main.R

@@ -41,7 +41,7 @@ source("fdaDelta.R")
 source("pick_subwise_curves.R")
 source("draw_clusters.R")
 source("floa_rcb.R")
-source("lenhoff_doris.R")
+source("floa_boot.R")
 source("floa_point.R")
 source("floa_roislien.R")
 source("plot_loa.R")
@@ -54,7 +54,6 @@ source("crossval_coverage_fraction.R")
 source("singlecurve_coverage.R")
 source("singlecurve_coverage_fraction.R")
 source("plot_cov_ver.R")
-source("distance_2_floa.R")
 source("estimate_uncertainty_loa.R")
 
 # ********************************* Data sets **********************************
@@ -108,26 +107,25 @@ n.boot <- 400
 # v4 : Roislien approach (Get one random stride from each subject ONCE and boot-
 #      strap the resulting sample (of length (n=length(subjects))
 # floa <- floa_rcb(data, n.boot, ver = "v2")
-
 floa.point    <- floa_point(data)
 floa.roislien <- floa_roislien(data)
-floa.lenhoff.pred  <- floa_lenhoff(data,
-                                   k_reihe = 50,
-                                   n.boot = n.boot,
-                                   band = "prediction",
-                                   cp.begin = 0,
-                                   alpha = 0.05)
+floa.boot.pred  <- floa_boot(data,
+                             k_reihe = 50,
+                             n.boot = n.boot,
+                             band = "prediction",
+                             cp.begin = 0,
+                             alpha = 0.05)
 
-floa.lenhoff.conf  <- floa_lenhoff(data,
-                                   k_reihe = 50,
-                                   n.boot = n.boot,
-                                   band = "confidence",
-                                   cp.begin = 0,
-                                   alpha = 0.05)
+floa.boot.conf  <- floa_boot(data,
+                             k_reihe = 50,
+                             n.boot = n.boot,
+                             band = "confidence",
+                             cp.begin = 0,
+                             alpha = 0.05)
 
 
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff.pred, ylim = c(-15, 15))
-plot_loa(data, floa.point, floa.roislien, floa.lenhoff.conf, ylim = c(-15, 15))
+plot_loa(data, floa.point, floa.roislien, floa.boot.pred, ylim = c(-15, 15))
+plot_loa(data, floa.point, floa.roislien, floa.boot.conf, ylim = c(-15, 15))
 
 
 # ********************************* Coverage ***********************************

### 2021-11-29 R/plot_loa.R

@@ -1,4 +1,4 @@
-plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
+plot_loa <- function (data, floa.point, floa.roislien, floa.boot, ylim) {
 
   # Plots the limits across the respective methods against all difference curves
   # in the data set
@@ -7,13 +7,13 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
   # Arguments:
   # floa.point: point-wise limits of agreement (Bland & Altman, 2007)
   # floa.roislien: limits of agreement as in Roislien et al. (2012)
-  # floa.lenhoff: limits of agreement as in Lenhoff et al. (1999)
+  # floa.boot: limits of agreement as in Lenhoff et al. (1999)
   # ylim: y-axis limits (2 element numeric object)
   # ----------------------------------------------------------------------------
 
   floa.point <- data.frame(t(floa.point))
   floa.roislien <- data.frame(t(floa.roislien))
-  floa.lenhoff <- data.frame(t(floa.lenhoff))
+  floa.boot <- data.frame(t(floa.boot))
 
   # Difference curves -------------------------------------------------
   device1 <- subset(data, device == "IMU", select = value)
@@ -39,37 +39,37 @@ plot_loa <- function (data, floa.point, floa.roislien, floa.lenhoff, ylim) {
               alpha = 0.15) +
     scale_color_grey(start = 0.8, end = 0.2) +
     # Plot lower limit of agreement
-    geom_line(data = floa.lenhoff,
-              aes(x = seq(0, 100), y = upper, col = "red", group = 1),
+    geom_line(data = floa.boot,
+              aes(x = seq(0, 100), y = upper.loa, col = "red", group = 1),
               linetype = "solid",
               size = 1.5,
               colour = "#D55E00") +
-    geom_line(data = floa.lenhoff,
-              aes(x = seq(0, 100), y = lower, col = "red", group = 1),
+    geom_line(data = floa.boot,
+              aes(x = seq(0, 100), y = lower.loa, col = "red", group = 1),
               linetype = "solid",
               size = 1.5,
               colour = "#D55E00") +
     geom_line(data = floa.roislien,
-              aes(x = seq(0, 100), y = upper),
+              aes(x = seq(0, 100), y = upper.loa),
               linetype = "solid",
               size = 1.5,
               colour = "#56B4E9") +
     geom_line(data = floa.roislien,
[…]

### 2021-11-29 R/singlecurve_coverage.R

@@ -1,24 +1,13 @@
 singlecurve_coverage <- function (data, n.boot) {
-  # ####################################################################
+  # --------------------------------------------------------------------
   # Leave-one (curve) out method to estimate the uncertainty in the
   # achieved coverage (see Lenhoff et al. (1999))
   # --------------------------------------------------------------------
-  #
-  # Currently, different versions of the sampling process in draw_clusters()
-  # (nested in floa_rcb()) are implemented:
-  #
-  # v1 : n = length(subjects) random strides from all strides
-  # v2 : One random stride per subject
-  # v3 : Fetch a SINGLE random stride from all strides
-  # v4 : Roislien approach (Get one random stride from each subject ONCE
-  #      and bootstrap the resulting sample
-  # v5 : Pointwise B&A limits
-  # ####################################################################
   n.curves <- unique(data$strideID)
 
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
-  cover.cross.lenhoff <- vector(mode = "list", length = length(n.curves))
+  cover.cross.boot <- vector(mode = "list", length = length(n.curves))
 
   for (curve.idx in n.curves) {
     # Calculate FLoA with one curve left out -----------------------------
@@ -26,7 +15,7 @@ singlecurve_coverage <- function (data, n.boot) {
 
     floa.point    <- floa_point(data.one.out)
     floa.roislien <- floa_roislien(data.one.out)
-    floa.lenhoff  <- floa_lenhoff(data,
+    floa.boot  <- floa_boot(data,
                                   k_reihe = 50,
                                   n.boot = n.boot,
                                   band = "prediction",
@@ -43,13 +32,13 @@ singlecurve_coverage <- function (data, n.boot) {
     # Get coverage for the left out (difference) curve -------------------
     cover.cross.point[curve.idx]      <- get_coverage_singlecurve(device.diff, floa.point)
     cover.cross.roislien[curve.idx]   <- get_coverage_singlecurve(device.diff, floa.roislien)
-    cover.cross.lenhoff[curve.idx]    <- get_coverage_singlecurve(device.diff, floa.lenhoff)
+    cover.cross.boot[curve.idx]    <- get_coverage_singlecurve(device.diff, floa.boot)
   }
 
   cover.cross <- cbind(
                       unlist(cover.cross.point),
                       unlist(cover.cross.roislien),
[…]

### 2021-11-29 R/singlecurve_coverage_fraction.R

@@ -31,7 +31,7 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
   # cover.cross.rcb.v3   <- vector(mode = "list", length = length(n.curves))
   cover.cross.point    <- vector(mode = "list", length = length(n.curves))
   cover.cross.roislien <- vector(mode = "list", length = length(n.curves))
-  cover.cross.lenhoff <- vector(mode = "list", length = length(n.curves))
+  cover.cross.boot <- vector(mode = "list", length = length(n.curves))
 
   for (curve.idx in n.curves) {
     # Calculate FLoA with one curve left out -----------------------------
@@ -42,7 +42,7 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
     # floa.v2       <- floa_rcb(data.one.out, n.boot, ver = "v2")
     # floa.v3       <- floa_rcb(data.one.out, n.boot, ver = "v3")
     floa.roislien <- floa_roislien(data.one.out)
-    floa.lenhoff  <- floa_lenhoff(data,
+    floa.boot  <- floa_boot(data,
                                   k_reihe = 50,
                                   n.boot = n.boot,
                                   band = "prediction",
@@ -63,7 +63,7 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
     # cover.cross.rcb.v3[curve.idx]   <- get_coverage_singlecurve_fraction(device.diff, floa.v3)
     cover.cross.point[curve.idx]      <- get_coverage_singlecurve_fraction(device.diff, floa.point)
     cover.cross.roislien[curve.idx]   <- get_coverage_singlecurve_fraction(device.diff, floa.roislien)
-    cover.cross.lenhoff[curve.idx]    <- get_coverage_singlecurve_fraction(device.diff, floa.lenhoff)
+    cover.cross.boot[curve.idx]    <- get_coverage_singlecurve_fraction(device.diff, floa.boot)
   }
 
   cover.cross <- cbind(
@@ -72,7 +72,7 @@ singlecurve_coverage_fraction <- function (data, n.boot) {
                       # unlist(cover.cross.rcb.v3),
                       unlist(cover.cross.point),
                       unlist(cover.cross.roislien),
-                      unlist(cover.cross.lenhoff)
+                      unlist(cover.cross.boot)
                       )
 
   return(cover.cross)

### 2021-12-07 man/floa_doc.Rmd
Move files to man folder

--- 2021-06-10 floa_doc.Rmd
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: github_document
-# output:
-#   html_document: 
-#     keep_md: yes
-#     theme: united
-#     toc: yes
-#     toc_float: yes
+# output: github_document
+output:
+  html_document:
+    keep_md: yes
+    theme: united
+    toc: yes
+    toc_float: yes
 ---
